### 2019-01-20 ext_risk_copula.Rmd
modified

---
title: "ext_risk_copula"
author: "Shyamolina Ghosh"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
source("mtime.R")
source("mydispersal.R")
```

#Ricker model :
```{r fig_ricker, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)

rlist<-c(0,0.1,0.3,0.5,0.7,0.9,1.0,1.1,1.3,1.5,1.7,1.9) #rc = 1
numsims<-10000
numsteps<-25
numlocs<-5
a<-NA
b<-NA
K<-50
scl<-1
ext_thrs<-K/10
model<-"ricker"

tempo<-paste("./Results/",model,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")

ricker_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
colnames(ricker_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")

# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
for(i in c(1:length(rlist))){
    r<- rlist[i]
    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
                 scl=scl,model=model,disp_everywhere=F,ploton = F)
    ricker_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
    ricker_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
    mtext(paste0("r = ", r))
}
par(op)
dev.off()

saveRDS(ricker_extrisk_d0,paste(resloc,"ricker_extrisk_d0.RDS",sep=""))

# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
for(r in rlist){
    #cat("r=",r,"\n")
    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
                 scl=scl,model=model,disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
    mtext(paste0("r = ", r))
}
par(op)
dev.off()
```

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewinton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ricker/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewinton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }

<!--
In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the lewinton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;

\begin{equation}\label{eq.extrisk_model}
P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
\end{equation}

$r$ is the liner growth rate and $K$ is the carrying capacity of the environment. $\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence. We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion). We simulated Eq. \ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $10$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;

- LC model (r=0) :

    * For LC model, as expected, when extinction risk calculated after $10$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one. [see first subplot of Fig. \ref{fig_ricker_risk_local_disp} and Fig. \ref{fig_ricker_risk_global_disp}]

- Ricker model (r $\neq$ 0):

    * For ricker model when extinction risk calculated after $10$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations (rc =1, under- to over-compensatory dynamics). [see Fig. \ref{fig_rikcer_risk_local_disp} - \ref{fig_ricker_risk_global_disp}]

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. (produce the figs if not done yet)

With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig. \ref{fig_risk_numlocs_effect_local_disp} - Fig. \ref{fig_risk_numlocs_effect_global_disp}] for both model. (produce the figs if not done yet)
-->

#Hassell Model :

```{r fig_hassell, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)

model<-"hassell"

tempo<-paste("./Results/",model,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")

rlist<-seq(from=1.5,to=7.5,by=0.5)
numsims<-10000
numsteps<-25
numlocs<-5
a<-0.5
b<-100
scl<-1
rc<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition 
rlist<-sort(c(rlist,rc))

hassell_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
colnames(hassell_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")

# plot for local dispersal in D matrix 
pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,1.2))
for(i in c(1:length(rlist))){
    r<- rlist[i]
    K_e<-((r^(1/b))-1)/a
    K<-K_e
    ext_thrs<-K/10
    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
                 scl=scl,model=model,disp_everywhere=F,ploton = F)
    hassell_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
    hassell_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
    mtext(paste0("r = ", r))
}
par(op)
dev.off()

saveRDS(hassell_extrisk_d0,paste(resloc,"hassell_extrisk_d0.RDS",sep=""))

# plot for local dispersal in D matrix 
pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,1.2))
for(r in rlist){
    #cat("r=",r,"\n")
    K_e<-((r^(1/b))-1)/a
    K<-K_e
    ext_thrs<-K/10
    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
                 scl=scl,model=model,disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
    mtext(paste0("r = ", r))
}
par(op)
dev.off()
```

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/hassell/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/hassell/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }

#Maynard Smith Model :

```{r fig_msmith_r_vary_localdisp, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)

model<-"msmith"

tempo<-paste("./Results/",model,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")

rlist<-seq(from=1.2,to=2.8,by=0.2)
numsims<-10000
numsteps<-25
numlocs<-5
a<-1
b<-3
scl<-1

#msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
#colnames(msmith_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")

# plot for local dispersal in D matrix 
pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
for(r in rlist){
    K_e<-((r-1)^(1/b))/a
    K<-K_e
    ext_thrs<-K/10
    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
                 scl=scl,model=model,disp_everywhere=F,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
    mtext(paste0("r = ", r))
}
par(op)
dev.off()

# plot for global dispersal in D matrix 
pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
for(r in rlist){
    K_e<-((r-1)^(1/b))/a
    K<-K_e
    ext_thrs<-K/10
    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
                 scl=scl,model=model,disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
    mtext(paste0("r = ", r))
}
par(op)
dev.off()
```

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }

### 2019-01-20 ext_risk_copula.Rmd
adding ext_risk for d=0 plot : ricker and hassell : CI need to confirm

@@
 
+```{r fig_ricker_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+
+ricker_extrisk_d0<-readRDS("./Results/ricker/ricker_extrisk_d0.RDS")
+numsims<-10000
+ricker_extrisk_d0$CI0.025left<-qbinom(ricker_extrisk_d0$risk_left_d0,size=numsims,prob=0.025)/numsims
+ricker_extrisk_d0$CI0.975left<-qbinom(ricker_extrisk_d0$risk_left_d0,size=numsims,prob=0.975)/numsims
+
+ricker_extrisk_d0$CI0.025right<-qbinom(ricker_extrisk_d0$risk_right_d0,size=numsims,prob=0.025)/numsims
+ricker_extrisk_d0$CI0.975right<-qbinom(ricker_extrisk_d0$risk_right_d0,size=numsims,prob=0.975)/numsims
+
+pdf("./Results/ricker/ricker_extrisk_d0.pdf",height = 6,width=6)
+
+plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(0,2),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,angle=90,length=0.05,col=rgb(1,0,0,0.5),code=3)
+
+lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025right,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975right,angle=90,length=0.05,col=rgb(0,0,1,0.5),code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(0.2,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+
+dev.off()
+```
+
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewinton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ricker/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewinton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
+```{r fig_hassell_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+hassell_extrisk_d0<-readRDS("./Results/hassell/hassell_extrisk_d0.RDS")
+
+hassell_extrisk_d0$CI0.025left<-qbinom(hassell_extrisk_d0$risk_left_d0,size=numsims,prob=0.025)/numsims
+hassell_extrisk_d0$CI0.975left<-qbinom(hassell_extrisk_d0$risk_left_d0,size=numsims,prob=0.975)/numsims
+
+hassell_extrisk_d0$CI0.025right<-qbinom(hassell_extrisk_d0$risk_right_d0,size=numsims,prob=0.025)/numsims
+hassell_extrisk_d0$CI0.975right<-qbinom(hassell_extrisk_d0$risk_right_d0,size=numsims,prob=0.975)/numsims
+
+pdf("./Results/hassell/hassell_extrisk_d0.pdf",height = 6,width=6)
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.4,7.6),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.05,col=rgb(1,0,0,0.5),code=3)
+
+lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025right,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975right,angle=90,length=0.05,col=rgb(0,0,1,0.5),code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(2,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+
+dev.off()
+```
+
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/hassell/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/hassell/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/hassell/hassell_extrisk_d0.pdf){ width=40%, height=40% }
 

### 2019-01-21 1var_pop_model_bifurcation.R
bifurcation plot

@@ -1,0 +1,96 @@
+# Bifurcation diagram for 1variable model
+#                    Ricker model : P(t+1)=P(t)*exp(r(1-(P(t)/K))) # rc=1
+#                    Hassell model : P(t+1)=r*P(t)/(1+(a*P(t)))^b  # rc=(b/(b-1))^b
+#                    Maynard-smith model : P(t+1)=(r*P(t))/(1+((a*P(t))^b)) # rc=b/(b-1)
+
+
+fn_d1_eqm<-function(params,model){
+  if(model=="ricker"){
+    fn<-expression(p*exp(r*(1-(p/K))))
+    fn_d1<-D(fn,"p")
+    r<-params[1]
+    K<-params[2]
+    p_e<-K
+    res<-eval(fn_d1,list(r=r,K=K,p=p_e))
+  }
+  
+  if(model=="hassell"){
+    fn<-expression(r*p/(1+(a*p))^b)
+    fn_d1<-D(fn,"p")
+    r<-params[1]
+    a<-params[2]
+    b<-params[3]
+    p_e<-((r^(1/b))-1)/a
+    res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
+  }
+  
+  if(model=="msmith"){
+    fn<-expression((r*p)/(1+((a*p)^b)))
+    fn_d1<-D(fn,"p")
+    r<-params[1]
+    a<-params[2]
+    b<-params[3]
+    p_e<-((r-1)^(1/b))/a
+    res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
+  }
+  
+  return(res)
+}
+
+
+# calling for ricker
+rlist<-sort(c(0,seq(from=0.1,to=1.9,by=0.2),1))
+K<-50 
+res_ricker<-c() 
+for(r in rlist){
+  ans<-fn_d1_eqm(params=c(r,K),model="ricker")
+  res_ricker<-c(res_ricker,ans)
+}
+
+plot(rlist,res_ricker,col="red",type="b",xlab="r",ylab="ricker_1st_der_at_eqm")
+abline(h=0)
+
+formatC(res_ricker[7],digits = 60, format="f") #formatting upto 60th decimal : eqm point
+
+# calling for hassell
+rlist<-c(seq(from=1.5,to=2.5,by=0.5),seq(from=2.731,to=2.732,by=0.00000001),seq(from=3,to=7.5,by=0.5))
+a<-0.5
+b<-100
[…]

### 2019-01-21 ext_risk_copula.Rmd
adding msmith d0_extrisk_vs_r plot

@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_d0}](./Results/ricker/ricker_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/hassell/hassell_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0}](./Results/hassell/hassell_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-rlist<-seq(from=1.2,to=2.8,by=0.2)
 numsims<-10000
@@
 numlocs<-5
-a<-1
-b<-3
+a<-0.5
+b<-4
 scl<-1
 
-#msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-#colnames(msmith_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+rc<-b/(b-1)
+rlist<-sort(c(seq(from=1.2,to=2.8,by=0.2),rc))
+
+msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+colnames(msmith_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
 
 # plot for local dispersal in D matrix 
-pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
-for(r in rlist){
+pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
     K_e<-((r-1)^(1/b))/a
@@
                  scl=scl,model=model,disp_everywhere=F,ploton = F)
+    msmith_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    msmith_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
@@
 
+saveRDS(msmith_extrisk_d0,paste(resloc,"msmith_extrisk_d0.RDS",sep=""))
+
 # plot for global dispersal in D matrix 
-pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
 for(r in rlist){
@@
 
+```{r fig_msmith_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+msmith_extrisk_d0<-readRDS("./Results/msmith/msmith_extrisk_d0.RDS")
+
+msmith_extrisk_d0$CI0.025left<-qbinom(msmith_extrisk_d0$risk_left_d0,size=numsims,prob=0.025)/numsims
+msmith_extrisk_d0$CI0.975left<-qbinom(msmith_extrisk_d0$risk_left_d0,size=numsims,prob=0.975)/numsims
+
+msmith_extrisk_d0$CI0.025right<-qbinom(msmith_extrisk_d0$risk_right_d0,size=numsims,prob=0.025)/numsims
+msmith_extrisk_d0$CI0.975right<-qbinom(msmith_extrisk_d0$risk_right_d0,size=numsims,prob=0.975)/numsims
+
+pdf("./Results/msmith/msmith_extrisk_d0.pdf",height = 6,width=6)
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.05,col=rgb(1,0,0,0.5),code=3)
+
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.05,col=rgb(0,0,1,0.5),code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(1.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+
+dev.off()
+```
+
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=1.33); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith/msmith_extrisk_d0.pdf){ width=40%, height=40% }

### 2019-01-21 1var_pop_models.R

@@ -76,9 +76,9 @@ Msmith<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-r<-1.8
-a<-1
-b<-3
+r<-1.2
+a<-0.5
+b<-4
 K_e<-((r-1)^(1/b))/a
 K_e
 Msmith(r=r,a=a,b=b,p0=K_e-0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4

### 2019-01-22 ext_risk_copula.Rmd
adding Pennycuick model results

@@
 
-```{r fig_msmith_r_vary_localdisp, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r fig_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
 
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=1.33); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith/msmith_extrisk_d0.pdf){ width=40%, height=40% }
+
+# Pennycuick model :
+
+```{r get_rc_pennycuick,echo=F,cache=T,cache.extra=list(seed,mtime("1var_pop_model_bifurcation.R"))}
+source("./1var_pop_model_bifurcation.R")
+a_pennycuick<-0.1
+b_pennycuick<-0.1
+eps<-1e-10 #you can change it upto which decimal point you want it to be exact
+rlist<-seq(from=4.3231657,to=4.3231658,by=eps)
+res_pennycuick<-c() 
+for(r in rlist){
+  ans<-fn_d1_eqm(params=c(r,a,b),model="pennycuick")
+  res_pennycuick<-c(res_pennycuick,ans)
+}
+
+#plot(rlist,res_pennycuick,col="red",type="l",xlab="r",ylab="pennycuick_1st_der_at_eqm")
+#abline(h=0)
+
+rc_pennycuick<-rlist[which.min(abs(res_pennycuick-0))]
+formatC(rc_pennycuick,digits = 10, format="f") # accurate upto 10th decimal
+```
+
+```{r fig_pennycuick, echo=F,  results="hide", cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+model<-"pennycuick"
+
+tempo<-paste("./Results/",model,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+numsims<-10000
+numsteps<-25
+numlocs<-5
+a<-a_pennycuick
+b<-b_pennycuick
+scl<-1
+
+rlist<-sort(c(seq(from=2,to=7,by=0.5),rc_pennycuick))
+
+pennycuick_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) # a dataframe to store ext risk at dispersal d=0 for all r
+colnames(pennycuick_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+# plot for local dispersal in D matrix 
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
+    K_e<-(b/a)*(a+log(r-1))
+    K<-K_e
+    ext_thrs<-K/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+    pennycuick_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    pennycuick_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+saveRDS(pennycuick_extrisk_d0,paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
+
+# plot for global dispersal in D matrix 
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+for(r in rlist){
+    K_e<-(b/a)*(a+log(r-1))
+    K<-K_e
+    ext_thrs<-K/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,ploton = F)
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 1, b = 3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }

### 2019-01-22 1var_pop_model_bifurcation.R

@@ -1,7 +1,8 @@
 # Bifurcation diagram for 1variable model
-#                    Ricker model : P(t+1)=P(t)*exp(r(1-(P(t)/K))) # rc=1
-#                    Hassell model : P(t+1)=r*P(t)/(1+(a*P(t)))^b  # rc=(b/(b-1))^b
-#                    Maynard-smith model : P(t+1)=(r*P(t))/(1+((a*P(t))^b)) # rc=b/(b-1)
+#                    Ricker model : p(t+1)=p(t)*exp(r(1-(p(t)/K))) # rc=1
+#                    Hassell model : p(t+1)=r*p(t)/(1+(a*p(t)))^b  # rc=(b/(b-1))^b
+#                    Maynard-smith model : p(t+1)=(r*p(t))/(1+((a*p(t))^b)) # rc=b/(b-1)
+#                    Pennycuick model : p(t+1)=(r*p(t))/(1+exp(-a*(1-(p(t)/b)))) # rc calculate numerically
 
 
 fn_d1_eqm<-function(params,model){
@@ -34,59 +35,85 @@ fn_d1_eqm<-function(params,model){
     res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
   }
   
+  if(model=="pennycuick"){
+    fn<-expression((r*p)/(1+exp(-a*(1-(p/b)))))
+    fn_d1<-D(fn,"p")
+    r<-params[1]
+    a<-params[2]
+    b<-params[3]
+    p_e<-(b/a)*(a+log(r-1))
+    res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
+  }
+  
   return(res)
 }
 
-
+#----------------------------------------------------------------
 # calling for ricker
-rlist<-sort(c(0,seq(from=0.1,to=1.9,by=0.2),1))
-K<-50 
-res_ricker<-c() 
-for(r in rlist){
-  ans<-fn_d1_eqm(params=c(r,K),model="ricker")
-  res_ricker<-c(res_ricker,ans)
-}
+#rlist<-sort(c(0,seq(from=0.1,to=1.9,by=0.2),1))
+#K<-50 
+#res_ricker<-c() 
+#for(r in rlist){
+#  ans<-fn_d1_eqm(params=c(r,K),model="ricker")
+#  res_ricker<-c(res_ricker,ans)
+#}
 
-plot(rlist,res_ricker,col="red",type="b",xlab="r",ylab="ricker_1st_der_at_eqm")
-abline(h=0)
[…]

### 2019-01-22 1var_pop_models.R

@@ -103,23 +103,32 @@ Verhulst<-function(r,p0,lensim){
 }
 
 #call the function
-Verhulst(r=1.2,p0=0.4,lensim=400) # eqm. point =1
+Verhulst(r=1.4,p0=0.4,lensim=400) # eqm. point =1
 
+#-------------------------------------------------------------
+# Exploring pennycuick model
+Pennycuick<-function(r,a,b,p0,lensim){
+  pop<-c(p0)
+  time<-c(0)
+  
+  for(it in c(1:lensim)){
+    pt<-(r*p0)/(1+exp(-a*(1-(p0/b))))
+    time<-c(time,it)
+    pop<-c(pop,pt)
+    p0<-pt
+  }
+  
+  plot(time,pop,type="b")
+  
+}
 
+#call the function
+a<-0.1
+b<-0.1
+r<-2
+K_e<-(b/a)*(a+log(r-1))
+K_e
+Pennycuick(r=r,a=a,b=b,p0=K_e-0.01,lensim=200) 
 
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
+#-----------------------------------------
 

### 2019-01-22 mydispersal.R

@@ -74,6 +74,7 @@ popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
     lam_verhulst<-1+(r*(1-res[,,tct]))
     lam_hassell<-r/((1+(a*res[,,tct]))^b)
     lam_msmith<-r/(1+((a*res[,,tct])^b))
+    lam_pennycuick<-r/(1+exp(-a*(1-(res[,,tct]/b))))
     
     #growth prior to dispersal
     if(model=="ricker"){
@@ -84,7 +85,9 @@ popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
       given_model<-lam_hassell
     }else if(model=="msmith"){
       given_model<-lam_msmith
-    }else{
+    }else if(model=="pennycuick"){
+      given_model<-lam_pennycuick
+    }else{ 
       warning("model not specified",immediate.=T,call.=T)
     }
     

### 2019-01-22 ext_risk_copula.Rmd
error fixed in qbinom function

@@
 numsims<-10000
-ricker_extrisk_d0$CI0.025left<-qbinom(ricker_extrisk_d0$risk_left_d0,size=numsims,prob=0.025)/numsims
-ricker_extrisk_d0$CI0.975left<-qbinom(ricker_extrisk_d0$risk_left_d0,size=numsims,prob=0.975)/numsims
+ricker_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_left_d0)/numsims
+ricker_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_left_d0)/numsims
 
-ricker_extrisk_d0$CI0.025right<-qbinom(ricker_extrisk_d0$risk_right_d0,size=numsims,prob=0.025)/numsims
-ricker_extrisk_d0$CI0.975right<-qbinom(ricker_extrisk_d0$risk_right_d0,size=numsims,prob=0.975)/numsims
+ricker_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_right_d0)/numsims
+ricker_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_right_d0)/numsims
 
@@
 plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(0,2),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,angle=90,length=0.05,col=rgb(1,0,0,0.5),code=3)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
 
 lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025right,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975right,angle=90,length=0.05,col=rgb(0,0,1,0.5),code=3)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025right,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
 mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
@@
 
-hassell_extrisk_d0$CI0.025left<-qbinom(hassell_extrisk_d0$risk_left_d0,size=numsims,prob=0.025)/numsims
-hassell_extrisk_d0$CI0.975left<-qbinom(hassell_extrisk_d0$risk_left_d0,size=numsims,prob=0.975)/numsims
+hassell_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_left_d0)/numsims
+hassell_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_left_d0)/numsims
 
-hassell_extrisk_d0$CI0.025right<-qbinom(hassell_extrisk_d0$risk_right_d0,size=numsims,prob=0.025)/numsims
-hassell_extrisk_d0$CI0.975right<-qbinom(hassell_extrisk_d0$risk_right_d0,size=numsims,prob=0.975)/numsims
+hassell_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_right_d0)/numsims
+hassell_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_right_d0)/numsims
 
@@
 plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.4,7.6),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.05,col=rgb(1,0,0,0.5),code=3)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
 
 lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025right,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975right,angle=90,length=0.05,col=rgb(0,0,1,0.5),code=3)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025right,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
 mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
@@
 
-msmith_extrisk_d0$CI0.025left<-qbinom(msmith_extrisk_d0$risk_left_d0,size=numsims,prob=0.025)/numsims
-msmith_extrisk_d0$CI0.975left<-qbinom(msmith_extrisk_d0$risk_left_d0,size=numsims,prob=0.975)/numsims
+msmith_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
+msmith_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
 
-msmith_extrisk_d0$CI0.025right<-qbinom(msmith_extrisk_d0$risk_right_d0,size=numsims,prob=0.025)/numsims
-msmith_extrisk_d0$CI0.975right<-qbinom(msmith_extrisk_d0$risk_right_d0,size=numsims,prob=0.975)/numsims
+msmith_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
+msmith_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
 
 pdf("./Results/msmith/msmith_extrisk_d0.pdf",height = 6,width=6)
-plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.05,col=rgb(1,0,0,0.5),code=3)
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0.8,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
 
 lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.05,col=rgb(0,0,1,0.5),code=3)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
 mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=1.33); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith/msmith_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith/msmith_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
+```{r fig_pennycuick_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+pennycuick_extrisk_d0<-readRDS("./Results/pennycuick/pennycuick_extrisk_d0.RDS")
+
+pennycuick_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
+pennycuick_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
+
+pennycuick_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
+pennycuick_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
+
+pdf("./Results/pennycuick/pennycuick_extrisk_d0.pdf",height = 6,width=6)
+plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+     xlim = c(2,7),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025right,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(3,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+
+dev.off()
+```
+
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 1, b = 3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000\label{fig_pennycuick_risk_d0}](./Results/pennycuick/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }

### 2019-01-25 ext_risk_copula.Rmd
try to add malthus model

@@
 numlocs<-5
-a<-NA
-b<-NA
 K<-50
@@
     r<- rlist[i]
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    params<-c(r,K)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=F,ploton = F)
@@
     #cat("r=",r,"\n")
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    params<-c(r,K)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=T,ploton = F)
@@
     K_e<-((r^(1/b))-1)/a
-    K<-K_e
-    ext_thrs<-K/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    ext_thrs<-K_e/10
+    params<-c(r,a,b)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=F,ploton = F)
@@
     K_e<-((r^(1/b))-1)/a
-    K<-K_e
-    ext_thrs<-K/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    ext_thrs<-K_e/10
+    params<-c(r,a,b)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=T,ploton = F)
@@
     r<-rlist[i]
+    params<-c(r,a,b)
     K_e<-((r-1)^(1/b))/a
-    K<-K_e
-    ext_thrs<-K/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=F,ploton = F)
@@
 for(r in rlist){
+    params<-c(r,a,b)
     K_e<-((r-1)^(1/b))/a
-    K<-K_e
-    ext_thrs<-K/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=T,ploton = F)
@@
     r<-rlist[i]
+    params<-c(r,a,b)
     K_e<-(b/a)*(a+log(r-1))
-    K<-K_e
-    ext_thrs<-K/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=F,ploton = F)
@@
 for(r in rlist){
+    params<-c(r,a,b)
     K_e<-(b/a)*(a+log(r-1))
-    K<-K_e
-    ext_thrs<-K/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                  scl=scl,model=model,disp_everywhere=T,ploton = F)

### 2019-01-25 1var_pop_model_bifurcation.R

@@ -45,6 +45,16 @@ fn_d1_eqm<-function(params,model){
     res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
   }
   
+  if(model=="malthus"){
+    fn<-expression(r*p*(K-p+(L*log(p))))
+    fn_d1<-D(fn,"p")
+    r<-params[1]
+    K<-params[2]
+    L<-params[3]
+    p_e<-params[4]
+    res<-eval(fn_d1,list(r=r,K=K,L=L,p=p_e))
+  }
+  
   return(res)
 }
 
@@ -115,9 +125,33 @@ fn_d1_eqm<-function(params,model){
 #rc_numerics<-rlist[which.min(abs(res_pennycuick-0))]
 #formatC(rc_numerics,digits = 10, format="f") # compare with rc_analytical
 
+#------------------------------------------
+# calling for Malthus
+#library(nleqslv)
+#eqs<-function(z,r,K,L){
+#  x<-z^2
+#  y<-numeric(1)
+#  y[1]<-(L*log(x))-x+K-(1/r)
+#  return(y)
+#}
 
 
+#zstart <- 100
+#eps<-1e-10
+#rlist<-seq(from=0.003230,to=0.003232,by=eps)
+#K<-600
+#L<-4
+#res_malthus<-c()
+#for(r in rlist){
+#  geteqm<-nleqslv(zstart, eqs, method="Newton",r=r,K=K,L=L)
+#  p_e<-(geteqm$x)^2
+#  ans<-fn_d1_eqm(params=c(r,K,L,p_e),model="malthus")
+#  res_malthus<-c(res_malthus,ans)
+#} 
 
+#plot(rlist,res_malthus,col="red",type="l",xlab="r",ylab="malthus_1st_der_at_eqm")
+#abline(h=0)
 
-
+#rc_numerics<-rlist[which.min(abs(res_malthus-0))]
+#formatC(rc_numerics,digits = 10, format="f") # compare with rc_analytical
 

### 2019-01-25 1var_pop_models.R

@@ -131,4 +131,38 @@ K_e
 Pennycuick(r=r,a=a,b=b,p0=K_e-0.01,lensim=200) 
 
 #-----------------------------------------
+# Exploring Malthus model
+Malthus<-function(r,K,L,p0,lensim){
+  pop<-c(p0)
+  time<-c(0)
+  
+  for(it in c(1:lensim)){
+    pt<-r*p0*(K-p0+(L*log(p0)))
+    time<-c(time,it)
+    pop<-c(pop,pt)
+    p0<-pt
+  }
+  
+  plot(time,pop,type="b")
+  
+}
+
+#call the function
+r<-0.0032
+K<-600
+L<-4
+
+Malthus(r=r,K=K,L=L,p0=0.1,lensim=200) 
+
+
+
+
+
+
+
+
+
+
+
+
 

### 2019-01-25 mydispersal.R

@@ -47,13 +47,13 @@ Dispersal_mat<-function(numlocs,d,disp_everywhere){
 #ns       An numsims by numlocs by numsteps array of the epsilons, 
 #           where numsteps is the number of time steps you want
 #D        Dispersal matrix
-#r,K,a,b  model parameters
-# ext_thrs  a threshold below which populations go extinct
+#params   a vector with model parameters in order as written in equation (e.g. r,a,b... or r,K,L)
+#ext_thrs  a threshold below which populations go extinct
 #model    a character specifying model name
 #--------Output------------------------------------------------------
 #A numsims by numlocs by numsteps+1 array of populations
 #----------------------------------------------------------------------
-popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
+popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
   numsims<-dim(ns)[1]
   numlocs<-dim(ns)[2]
   numsteps<-dim(ns)[3]
@@ -70,23 +70,40 @@ popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
     #growth rates based on the noise
     lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
     
-    lam_ricker<-exp(r*(1-(res[,,tct]/K)))
-    lam_verhulst<-1+(r*(1-res[,,tct]))
-    lam_hassell<-r/((1+(a*res[,,tct]))^b)
-    lam_msmith<-r/(1+((a*res[,,tct])^b))
-    lam_pennycuick<-r/(1+exp(-a*(1-(res[,,tct]/b))))
-    
     #growth prior to dispersal
     if(model=="ricker"){
+      r<-params[1]
+      K<-params[2]
+      lam_ricker<-exp(r*(1-(res[,,tct]/K)))
       given_model<-lam_ricker
     }else if(model=="verhulst"){
+      r<-params[1]
+      lam_verhulst<-1+(r*(1-res[,,tct]))
       given_model<-lam_verhulst
     }else if(model=="hassell"){
+      r<-params[1]
+      a<-params[2]
+      b<-params[3]
+      lam_hassell<-r/((1+(a*res[,,tct]))^b)
       given_model<-lam_hassell
     }else if(model=="msmith"){
+      r<-params[1]
+      a<-params[2]
+      b<-params[3]
+      lam_msmith<-r/(1+((a*res[,,tct])^b))
[…]

### 2019-01-27 ext_risk_copula.Rmd
try to add austin-brewer model

@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 1, b = 3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
 

### 2019-01-27 1var_pop_models.R

@@ -154,6 +154,55 @@ L<-4
 
 Malthus(r=r,K=K,L=L,p0=0.1,lensim=200) 
 
+#-----------------------------------------
+# Exploring Austin-Brewer model
+AustinBrewer<-function(r,K,s,p0,lensim){
+  pop<-c(p0)
+  time<-c(0)
+  
+  for(it in c(1:lensim)){
+    pt<-p0*(1+(r*(K-p0)*(1-exp(-s*p0)))) 
+    time<-c(time,it)
+    pop<-c(pop,pt)
+    p0<-pt
+  }
+  
+  plot(time,pop,type="b")
+  
+}
+
+#call the function
+s<-0.13
+K<-45
+rc<-1/(K*(1-exp(-s*K)))
+rc
+r<-0.01
+AustinBrewer(r=r,K=K,s=s,p0=1,lensim=200) 
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
 
 
 

### 2019-01-27 mydispersal.R

@@ -104,6 +104,12 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
       L<-params[3]
       lam_malthus<-r*(K-res[,,tct]+(L*log(res[,,tct])))
       given_model<-lam_malthus
+    }else if(model=="austinbrewer"){
+      r<-params[1]
+      K<-params[2]
+      s<-params[3]
+      lam_austin<-(1+(r*(K-res[,,tct])*(1-exp(-s*res[,,tct])))) 
+      given_model<-lam_austin
     }else{ 
       warning("model not specified",immediate.=T,call.=T)
     }

### 2019-01-29 ext_risk_copula.Rmd
adding msmith result with scl=0.8

@@
 
-```{r fig_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r fig_msmith_scl_1, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 model<-"msmith"
+scl<-1
 
-tempo<-paste("./Results/",model,sep="")
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
 if (!dir.exists(tempo)){
@@
 b<-4
-scl<-1
 
@@
 dev.off()
-```
 
-```{r fig_msmith_extrisk_d0, echo=F,results="hide"}
-set.seed(seed)
-numsims<-10000
-msmith_extrisk_d0<-readRDS("./Results/msmith/msmith_extrisk_d0.RDS")
-
+# Plot ext_risk_d0
 msmith_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
@@
 
-pdf("./Results/msmith/msmith_extrisk_d0.pdf",height = 6,width=6)
-plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0.8,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
 arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
 
-lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
 arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
@@
 legend(1.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-
 dev.off()
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith_scl_1/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith_scl_1/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith/msmith_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith_scl_1/msmith_extrisk_d0.pdf){ width=40%, height=40% }
+
+```{r fig_msmith_scl_0.8, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+
+model<-"msmith"
+scl<-0.8
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+numsims<-10000
+numsteps<-25
+numlocs<-5
+a<-0.5
+b<-4
+
+rc<-b/(b-1)
+rlist<-sort(c(seq(from=1.2,to=2.8,by=0.2),rc))
+
+msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+colnames(msmith_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+# plot for local dispersal in D matrix 
+pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
+    params<-c(r,a,b)
+    K_e<-((r-1)^(1/b))/a
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+    msmith_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    msmith_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+saveRDS(msmith_extrisk_d0,paste(resloc,"msmith_extrisk_d0.RDS",sep=""))
+
+# plot for global dispersal in D matrix 
+pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+for(r in rlist){
+    params<-c(r,a,b)
+    K_e<-((r-1)^(1/b))/a
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,ploton = F)
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# Plot ext_risk_d0
+msmith_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
+msmith_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
+
+msmith_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
+msmith_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
+
+pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(1.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0_scl_0.8}](./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000\label{fig_pennycuick_risk_d0}](./Results/pennycuick/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
+
+# Austin Brewer model :

### 2019-01-29 1var_pop_models.R

@@ -173,14 +173,41 @@ AustinBrewer<-function(r,K,s,p0,lensim){
 
 #call the function
 s<-0.13
-K<-45
+K<-100
 rc<-1/(K*(1-exp(-s*K)))
 rc
-r<-0.01
+r<-0.014
 AustinBrewer(r=r,K=K,s=s,p0=1,lensim=200) 
 
+#----------------------------------------------------
+# Exploring Varley model
 
+Varley<-function(r,b,C,p0,lensim){
+  pop<-c(p0)
+  time<-c(0)
+  
+  for(it in c(1:lensim)){
+    if(p0<=C){
+      pt<-r*p0
+    }else{
+      pt<-r*(p0^(1-b))
+    }
+    time<-c(time,it)
+    pop<-c(pop,pt)
+    p0<-pt
+  }
+  
+  plot(time,pop,type="b")
+  
+}
 
+#call the function
+b<-1.4
+r<-4
+C<-1
+p0<-r^(1/b)
+p0
+Varley(r=r,b=b,C=C,p0=p0-1,lensim=200) 
 
 
 

### 2019-01-29 ExtremeTailDep.R

@@ -11,7 +11,7 @@
 #
 retd<-function(n,d,rl)
 {
-  m<-matrix(rnorm(n*(d+1)),n,d+1)
+  m<-matrix(rnorm(n*(d+1),mean=0,sd=1),n,d+1)
   flag<-(m[,1]>0) #a coin toss
   m<-m[,2:(d+1)] #an n by d matrix of iid standard normals
   

### 2019-01-29 mydispersal.R

@@ -74,47 +74,56 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
     if(model=="ricker"){
       r<-params[1]
       K<-params[2]
-      lam_ricker<-exp(r*(1-(res[,,tct]/K)))
-      given_model<-lam_ricker
+      lam_ricker<-exp(r*(1-(res[,,tct]/K)))*res[,,tct]
+      res[,,tct+1]<-lam_ricker*lam_sto #numsims by numlocs matrix
     }else if(model=="verhulst"){
       r<-params[1]
-      lam_verhulst<-1+(r*(1-res[,,tct]))
-      given_model<-lam_verhulst
+      lam_verhulst<-1+(r*(1-res[,,tct]))*res[,,tct]
+      res[,,tct+1]<-lam_verhulst*lam_sto #numsims by numlocs matrix
     }else if(model=="hassell"){
       r<-params[1]
       a<-params[2]
       b<-params[3]
-      lam_hassell<-r/((1+(a*res[,,tct]))^b)
-      given_model<-lam_hassell
+      lam_hassell<-(r*res[,,tct])/((1+(a*res[,,tct]))^b)
+      res[,,tct+1]<-lam_hassell*lam_sto #numsims by numlocs matrix
     }else if(model=="msmith"){
       r<-params[1]
       a<-params[2]
       b<-params[3]
-      lam_msmith<-r/(1+((a*res[,,tct])^b))
-      given_model<-lam_msmith
+      lam_msmith<-(r*res[,,tct])/(1+((a*res[,,tct])^b))
+      res[,,tct+1]<-lam_msmith*lam_sto #numsims by numlocs matrix
     }else if(model=="pennycuick"){
       r<-params[1]
       a<-params[2]
       b<-params[3]
-      lam_pennycuick<-r/(1+exp(-a*(1-(res[,,tct]/b))))
-      given_model<-lam_pennycuick
+      lam_pennycuick<-(r*res[,,tct])/(1+exp(-a*(1-(res[,,tct]/b))))
+      res[,,tct+1]<-lam_pennycuick*lam_sto #numsims by numlocs matrix
     }else if(model=="malthus"){
       r<-params[1]
       K<-params[2]
       L<-params[3]
-      lam_malthus<-r*(K-res[,,tct]+(L*log(res[,,tct])))
-      given_model<-lam_malthus
[…]

### 2019-01-31 ext_risk_copula.Rmd
adding austin brewer results with scl=0.8

@@
 # Austin Brewer model :
+
+```{r fig_abrewer_scl_0.8, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+
+model<-"austinbrewer"
+scl<-0.8
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+numsims<-10000
+numsteps<-25
+numlocs<-5
+s<-0.2
+K<-50
+p0<-K
+rc<-1/(K*(1-exp(-s*K)))
+
+rlist<-sort(c(0.002,0.004,0.006,0.008,0.01,0.015,rc,0.022,0.026))
+
+abrewer_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+colnames(abrewer_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+# plot for local dispersal in D matrix 
+pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
+    params<-c(r,K,s)
+    ext_thrs<-p0/10
+    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+    abrewer_extrisk_d0$risk_left_d0[i]<-s0$risk_left[1]
+    abrewer_extrisk_d0$risk_right_d0[i]<-s0$risk_right[1]
+    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+saveRDS(abrewer_extrisk_d0,paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
+
+# plot for global dispersal in D matrix 
+pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+for(r in rlist){
+    params<-c(r,K,s)
+    ext_thrs<-p0/10
+    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,ploton = F)
+    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# Plot ext_risk_d0
+abrewer_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
+abrewer_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
+
+abrewer_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
+abrewer_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
+
+pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+     xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025left,x1=abrewer_extrisk_d0$r,
+       y1=abrewer_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025right,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(1.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.8}](./Results/abrewer_scl_0.8/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_abrewer_r_vary_scl_0.8}](./Results/abrewer_scl_0.8/abrewer_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.8}](./Results/abrewer_scl_0.8/abrewer_extrisk_d0.pdf){ width=40%, height=40% }

### 2019-01-31 1var_pop_models.R

@@ -172,11 +172,11 @@ AustinBrewer<-function(r,K,s,p0,lensim){
 }
 
 #call the function
-s<-0.13
-K<-100
+s<-0.2
+K<-50
 rc<-1/(K*(1-exp(-s*K)))
 rc
-r<-0.014
+r<-0.026
 AustinBrewer(r=r,K=K,s=s,p0=1,lensim=200) 
 
 #----------------------------------------------------

### 2019-02-01 ext_risk_copula.Rmd
adding results : verhulst model

--- 2019-01-29 ext_risk_copula.Rmd
@@
 # plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
 for(r in rlist){
@@
 # Austin Brewer model :
+
+```{r fig_abrewer_scl_0.8, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+
+model<-"austinbrewer"
+scl<-0.8
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+numsims<-10000
+numsteps<-25
+numlocs<-5
+s<-0.2
+K<-50
+p0<-K
+rc<-1/(K*(1-exp(-s*K)))
+
+rlist<-sort(c(0.002,0.004,0.006,0.008,0.01,0.015,rc,0.022,0.026))
+
+abrewer_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+colnames(abrewer_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+# plot for local dispersal in D matrix 
+pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
+    params<-c(r,K,s)
+    ext_thrs<-p0/10
+    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+    abrewer_extrisk_d0$risk_left_d0[i]<-s0$risk_left[1]
+    abrewer_extrisk_d0$risk_right_d0[i]<-s0$risk_right[1]
+    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+saveRDS(abrewer_extrisk_d0,paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
+
+# plot for global dispersal in D matrix 
+pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+for(r in rlist){
+    params<-c(r,K,s)
+    ext_thrs<-p0/10
+    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,ploton = F)
+    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# Plot ext_risk_d0
+abrewer_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
+abrewer_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
+
+abrewer_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
+abrewer_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
+
+pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025left,x1=abrewer_extrisk_d0$r,
+       y1=abrewer_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025right,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(0.005,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_abrewer_r_vary_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
+
+# Verhulst model :
+
+```{r fig_verhulst_scl_2, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+
+model<-"verhulst"
+scl<-2.5
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+numsims<-10000
+numsteps<-25
+numlocs<-5
+p0<-1
+
+rlist<-seq(from=0.2,to=1.8,by=0.2)
+
+verhulst_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+colnames(verhulst_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+# plot for local dispersal in D matrix 
+pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
+    ext_thrs<-p0/10
+    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=r,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+    verhulst_extrisk_d0$risk_left_d0[i]<-s0$risk_left[1]
+    verhulst_extrisk_d0$risk_right_d0[i]<-s0$risk_right[1]
+    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+saveRDS(verhulst_extrisk_d0,paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
+
+# plot for global dispersal in D matrix 
+pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+for(r in rlist){
+    ext_thrs<-p0/10
+    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=r,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,ploton = F)
+    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# Plot ext_risk_d0
+verhulst_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
+verhulst_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
+
+verhulst_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
+verhulst_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
+
+pdf(paste(resloc,"verhulst_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025left,x1=verhulst_extrisk_d0$r,
+       y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025right,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(0.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=2.5); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_extrisk_d0.pdf){ width=40%, height=40% }

### 2019-02-01 1var_pop_models.R

@@ -99,11 +99,11 @@ Verhulst<-function(r,p0,lensim){
     p0<-pt
   }
   
-  plot(time,pop,type="b")
+  plot(time,pop,type="l")
 }
 
 #call the function
-Verhulst(r=1.4,p0=0.4,lensim=400) # eqm. point =1
+Verhulst(r=1.8,p0=0.4,lensim=400) # eqm. point =1
 
 #-------------------------------------------------------------
 # Exploring pennycuick model
@@ -202,9 +202,9 @@ Varley<-function(r,b,C,p0,lensim){
 }
 
 #call the function
-b<-1.4
+b<-0.2
 r<-4
-C<-1
+C<-2
 p0<-r^(1/b)
 p0
 Varley(r=r,b=b,C=C,p0=p0-1,lensim=200) 

### 2019-02-15 ext_risk_copula.Rmd
multiple changes : to be continued

@@
 set.seed(seed)
-
+library(abind)
 rlist<-c(0,0.1,0.3,0.5,0.7,0.9,1.0,1.1,1.3,1.5,1.7,1.9) #rc = 1
@@
 
+res_ricker_local<-c()
+
 # plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+#pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+#op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
 for(i in c(1:length(rlist))){
@@
                  scl=scl,model=model,disp_everywhere=F,ploton = F)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$avg_acf_left,s$avg_acf_right)
+
+    res_ricker_local<-abind(res_ricker_local,temp_mat)
+
     ricker_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
     ricker_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-    mtext(paste0("r = ", r))
+    #plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+    #     panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    #lines(s$d_seq,s$risk_right,type="b",col="blue")
+    #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+    #mtext(paste0("r = ", r))
 }
-par(op)
-dev.off()
+#par(op)
+#dev.off()
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","avgACF_left","avgACF_right")
+dimnames(res_ricker_local)<-list(NULL,cnm,anm)
 
 saveRDS(ricker_extrisk_d0,paste(resloc,"ricker_extrisk_d0.RDS",sep=""))
+saveRDS(res_ricker_local,paste(resloc,"res_ricker_local.RDS",sep=""))
 
-# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+res_ricker_global<-c()
+# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
+#pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+#op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
 for(r in rlist){
@@
                  scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$avg_acf_left,s$avg_acf_right)
+    colnames(temp_mat)<-c("d","risk_left","risk_right","avgACF_left","avgACF_right")
+
+    res_ricker_global<-abind(res_ricker_global,temp_mat)
+    #plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+    #     panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    #lines(s$d_seq,s$risk_right,type="b",col="blue")
+    #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    #mtext(paste0("r = ", r))
 }
-par(op)
-dev.off()
+#par(op)
+#dev.off()
 ```
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_d0}](./Results/ricker/ricker_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--
-In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the lewinton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
-
-\begin{equation}\label{eq.extrisk_model}
-P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
-\end{equation}
-
-$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
-$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
-We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
-We simulated Eq.
-\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $10$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
-
-- LC model (r=0) :
-
-    * For LC model, as expected, when extinction risk calculated after $10$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-[see first subplot of Fig.
-\ref{fig_ricker_risk_local_disp} and Fig.
-\ref{fig_ricker_risk_global_disp}]
-
-- Ricker model (r $\neq$ 0):
-
-    * For ricker model when extinction risk calculated after $10$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations (rc =1, under- to over-compensatory dynamics).
-[see Fig.
-\ref{fig_rikcer_risk_local_disp} - \ref{fig_ricker_risk_global_disp}]
-
-    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. (produce the figs if not done yet)
-
-With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
-\ref{fig_risk_numlocs_effect_local_disp} - Fig.
-\ref{fig_risk_numlocs_effect_global_disp}] for both model.
-(produce the figs if not done yet)
--->
 
@@
 #Maynard Smith Model :
-
-```{r fig_msmith_scl_1, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
-set.seed(seed)
-
-model<-"msmith"
-scl<-1
-
-tempo<-paste("./Results/",model,"_scl_",scl,sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-
-numsims<-10000
-numsteps<-25
-numlocs<-5
-a<-0.5
-b<-4
-
-rc<-b/(b-1)
-rlist<-sort(c(seq(from=1.2,to=2.8,by=0.2),rc))
-
-msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(msmith_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
-
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
-for(i in c(1:length(rlist))){
-    r<-rlist[i]
-    params<-c(r,a,b)
-    K_e<-((r-1)^(1/b))/a
-    ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
-    msmith_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    msmith_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
-
-saveRDS(msmith_extrisk_d0,paste(resloc,"msmith_extrisk_d0.RDS",sep=""))
-
-# plot for global dispersal in D matrix 
-pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
-for(r in rlist){
-    params<-c(r,a,b)
-    K_e<-((r-1)^(1/b))/a
-    ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
-
-# Plot ext_risk_d0
-msmith_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
-msmith_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
-
-msmith_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
-msmith_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=6,width=6)
-plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-
-lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(1.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-dev.off()
-```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/msmith_scl_1/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/msmith_scl_1/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0}](./Results/msmith_scl_1/msmith_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-```{r fig_verhulst_scl_2, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r fig_verhulst_scl_2.5, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)

### 2019-02-15 1var_pop_models.R

@@ -202,9 +202,9 @@ Varley<-function(r,b,C,p0,lensim){
 }
 
 #call the function
-b<-0.2
+b<-1.2
 r<-4
-C<-2
+C<-1
 p0<-r^(1/b)
 p0
 Varley(r=r,b=b,C=C,p0=p0-1,lensim=200) 

### 2019-02-15 mydispersal.R

@@ -144,6 +144,25 @@ extrisk<-function(sims){
   totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
   return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])
 }
+
+# function to calculate average ACF (lag=1) from all numlocs after specified time steps
+# input : 
+#        1) sims : an array (numsims by numlocs by numsteps+1)
+#        2) nts : numsteps after which you want the acf; generally equal to numsteps+1 as to compare with extinction risk
+get_avg_acf<-function(sims,nts){
+  m<-sims[,,nts]
+  macf<-apply(X=m,MARGIN = 2,FUN=acf,lag.max = 1,type="correlation",plot=F)
+  acfs<-c() # initialize to store acf with lag1 for each col of matrix m
+  for(i in c(1:ncol(m))){
+    acfs<-c(acfs,macf[[i]]$acf[2])
+  }
+  ans<-mean(acfs)
+  return(ans)
+}
+
+#get_avg_acf(sims=sims,nts=numsteps+1)
+
+
 #-------------------------------------------------------------------------------------------------
 # function to optionally plot extinction risk agsinst time and give you back extinction risk 
 #                                                            after specified numsteps 
@@ -158,8 +177,9 @@ extrisk<-function(sims){
 #       8) scl : a scaling factor which is multiplied with noise generated
 #       9) model : a character specifying model name
 #       10) ploton : logical(T or F) to get optional plot
+#       11) resloc : location to save plots
 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton){
+plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc){
  
   ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
   #                                                      numpoints=numsteps*numsims)
[…]

### 2019-02-15 ext_risk_copula.Rmd
fixed : ricker,hassell

@@
 #Ricker model :
-```{r fig_ricker, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_ricker, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-ricker_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+#initiate a dataframe to store ext risk at dispersal d=0 for all r
+ricker_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) 
 colnames(ricker_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
@@
 
-# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-#pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-#op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
 for(i in c(1:length(rlist))){
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
 
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$avg_acf_left,s$avg_acf_right)
-
-    res_ricker_local<-abind(res_ricker_local,temp_mat)
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    res_ricker_local<-abind(res_ricker_local,temp_mat,along = 3)
 
@@
     ricker_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-    #plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
-    #     panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    #lines(s$d_seq,s$risk_right,type="b",col="blue")
-    #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-    #mtext(paste0("r = ", r))
 }
-#par(op)
-#dev.off()
+
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","avgACF_left","avgACF_right")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
 dimnames(res_ricker_local)<-list(NULL,cnm,anm)
@@
 
+# For global dispersion
 res_ricker_global<-c()
-# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-#pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-#op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
 for(r in rlist){
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
 
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$avg_acf_left,s$avg_acf_right)
-    colnames(temp_mat)<-c("d","risk_left","risk_right","avgACF_left","avgACF_right")
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    res_ricker_global<-abind(res_ricker_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+dimnames(res_ricker_global)<-list(NULL,cnm,anm)
+saveRDS(res_ricker_global,paste(resloc,"res_ricker_global.RDS",sep=""))
+```
 
-    res_ricker_global<-abind(res_ricker_global,temp_mat)
-    #plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
-    #     panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    #lines(s$d_seq,s$risk_right,type="b",col="blue")
-    #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    #mtext(paste0("r = ", r))
+```{r fig_ricker, echo=F, results="hide",cache=T, cache.extra=list(seed,res_ricker_local,res_ricker_global)}
+pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_ricker_local)[3]){
+  s<-res_ricker_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  lines(s$d, s$noise_avg_acf_left, col="green")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
-#par(op)
-#dev.off()
+par(op)
+dev.off()
+
+pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_ricker_global)[3]){
+  s<-res_ricker_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
+
+  lines(s$d, s$noise_avg_acf_left, col="green")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
+}
+par(op)
+dev.off()
 ```
@@
 
-plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(0,2),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",
+     ylim=c(0,1),xlim = c(0,2),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,
+       angle=90,length=0.1,col="red",code=3)
+abline(v=1,lty="dotted") #rc=1
 
@@
 
-```{r fig_hassell, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_hassell, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-1
-rc<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition 
-rlist<-sort(c(rlist,rc))
+rc_hassell<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition 
+rlist<-sort(c(rlist,rc_hassell))
 
@@
 
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
-op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,1.2))
+res_hassell_local<-c()
+
 for(i in c(1:length(rlist))){
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    res_hassell_local<-abind(res_hassell_local,temp_mat,along = 3)
+
     hassell_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
     hassell_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
 }
-par(op)
-dev.off()
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+dimnames(res_hassell_local)<-list(NULL,cnm,anm)
 
 saveRDS(hassell_extrisk_d0,paste(resloc,"hassell_extrisk_d0.RDS",sep=""))
+saveRDS(res_hassell_local,paste(resloc,"res_hassell_local.RDS",sep=""))
 
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
-op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,1.2))
+# For global dispersion
+res_hassell_global<-c()
 for(r in rlist){
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    res_hassell_global<-abind(res_hassell_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+dimnames(res_hassell_global)<-list(NULL,cnm,anm)
+saveRDS(res_hassell_global,paste(resloc,"res_hassell_global.RDS",sep=""))
+```
+
+```{r fig_hassell, echo=F, results="hide",cache=T, cache.extra=list(seed,res_hassell_local,res_hassell_global)}
+pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=18)
+op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_hassell_local)[3]){
+  s<-res_hassell_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  lines(s$d, s$noise_avg_acf_left, col="green")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=18)
+op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_hassell_global)[3]){
+  s<-res_hassell_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  lines(s$d, s$noise_avg_acf_left, col="green")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_hassell,lty="dotted")
 

### 2019-02-15 mydispersal.R

@@ -147,7 +147,7 @@ extrisk<-function(sims){
 
 # function to calculate average ACF (lag=1) from all numlocs after specified time steps
 # input : 
-#        1) sims : an array (numsims by numlocs by numsteps+1)
+#        1) sims : an array (numsims by numlocs by numsteps)
 #        2) nts : numsteps after which you want the acf; generally equal to numsteps+1 as to compare with extinction risk
 get_avg_acf<-function(sims,nts){
   m<-sims[,,nts]
@@ -160,8 +160,7 @@ get_avg_acf<-function(sims,nts){
   return(ans)
 }
 
-#get_avg_acf(sims=sims,nts=numsteps+1)
-
+# I checked : ACF of noise (right tail) = ACF of noise (left tail)
 
 #-------------------------------------------------------------------------------------------------
 # function to optionally plot extinction risk agsinst time and give you back extinction risk 
@@ -190,12 +189,16 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
   ns1<-scl*ns1
   pops1<-popsim_ml_D(p0=rep(p0,numlocs),ns=ns1,D=D,params=params,ext_thrs=ext_thrs,model=model)
   risk_right<-extrisk(pops1) # a vector
-  acf_right<-get_avg_acf(sims=pops1,nts=numsteps+1) # a number
+  
+  noise_acf_right<-get_avg_acf(sims=ns1,nts=numsteps) # a number
+  pop_acf_right<-get_avg_acf(sims=pops1,nts=numsteps+1) # a number
   
   ns2<-(-ns1)
   pops2<-popsim_ml_D(p0=rep(p0,numlocs),ns=ns2,D=D,params=params,ext_thrs=ext_thrs,model=model)
   risk_left<-extrisk(pops2) # a vector
-  acf_left<-get_avg_acf(sims=pops2,nts=numsteps+1) # a number
+  
+  noise_acf_left<-get_avg_acf(sims=ns2,nts=numsteps) # a number
+  pop_acf_left<-get_avg_acf(sims=pops2,nts=numsteps+1) # a number
   
   if(ploton==T){
     # noise time series plot for last simulation (numsim=numsims) from each patches
@@ -259,8 +262,10 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
   
[…]

### 2019-02-16 ext_risk_copula.Rmd
upto msmith fixed

@@
 seed<-101
+library(abind)
 source("mtime.R")
-source("mydispersal.R")
 ```
@@
 set.seed(seed)
-library(abind)
+source("mydispersal.R")
 rlist<-c(0,0.1,0.3,0.5,0.7,0.9,1.0,1.1,1.3,1.5,1.7,1.9) #rc = 1
@@
 
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
     res_ricker_local<-abind(res_ricker_local,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
 dimnames(res_ricker_local)<-list(NULL,cnm,anm)
@@
 res_ricker_global<-c()
+
 for(r in rlist){
@@
 
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
     res_ricker_global<-abind(res_ricker_global,temp_mat,along = 3)
 }
+
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
 dimnames(res_ricker_global)<-list(NULL,cnm,anm)
@@
   lines(s$d, s$pop_avg_acf_right, col="skyblue")
-
-  lines(s$d, s$noise_avg_acf_left, col="green")
 
@@
   lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
-
-  lines(s$d, s$noise_avg_acf_left, col="green")
 
@@
 set.seed(seed)
+source("mydispersal.R")
 
@@
 
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
     res_hassell_local<-abind(res_hassell_local,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
 dimnames(res_hassell_local)<-list(NULL,cnm,anm)
@@
 
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right,s$noise_avg_acf_left,s$noise_avg_acf_right)
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
     res_hassell_global<-abind(res_hassell_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right","noise_avg_acf_left","noise_avg_acf_right")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
 dimnames(res_hassell_global)<-list(NULL,cnm,anm)
@@
 
-```{r fig_msmith_scl_0.8, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
+source("mydispersal.R")
 
@@
 
-rc<-b/(b-1)
-rlist<-sort(c(seq(from=1.2,to=2.8,by=0.2),rc))
+rc_msmith<-b/(b-1)
+rlist<-sort(c(seq(from=1.2,to=2.8,by=0.2),rc_msmith))
 
@@
 
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+#for local dispersal in D matrix 
+res_msmith_local<-c()
+
 for(i in c(1:length(rlist))){
-    r<-rlist[i]
+    r<- rlist[i]
     params<-c(r,a,b)
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_msmith_local<-abind(res_msmith_local,temp_mat,along = 3)
+
     msmith_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
     msmith_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+}
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_msmith_local)<-list(NULL,cnm,anm)
+
+saveRDS(msmith_extrisk_d0,paste(resloc,"msmith_extrisk_d0.RDS",sep=""))
+saveRDS(res_msmith_local,paste(resloc,"res_msmith_local.RDS",sep=""))
+
+# For global dispersal in D matrix 
+res_msmith_global<-c()
+for(r in rlist){
+    #cat("r=",r,"\n")
+    params<-c(r,a,b)
+    K_e<-((r-1)^(1/b))/a
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_msmith_global<-abind(res_msmith_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_msmith_global)<-list(NULL,cnm,anm)
+saveRDS(res_msmith_global,paste(resloc,"res_msmith_global.RDS",sep=""))
+```
+
+```{r fig_msmith, echo=F, results="hide",cache=T, cache.extra=list(seed,res_msmith_local,res_msmith_global)}
+pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_msmith_local)[3]){
+  s<-res_msmith_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 
-saveRDS(msmith_extrisk_d0,paste(resloc,"msmith_extrisk_d0.RDS",sep=""))
+pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
 
-# plot for global dispersal in D matrix 
-pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
-for(r in rlist){
-    params<-c(r,a,b)
-    K_e<-((r-1)^(1/b))/a
-    ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+for(i in 1:dim(res_msmith_global)[3]){
+  s<-res_msmith_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 dev.off()
+```
 
-# Plot ext_risk_d0
+```{r fig_msmith_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+
+msmith_extrisk_d0<-readRDS(paste(resloc,"msmith_extrisk_d0.RDS",sep=""))
+numsims<-10000
 msmith_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
@@
 arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_msmith,lty="dotted")
 

### 2019-02-16 mydispersal.R

@@ -148,18 +148,23 @@ extrisk<-function(sims){
 # function to calculate average ACF (lag=1) from all numlocs after specified time steps
 # input : 
 #        1) sims : an array (numsims by numlocs by numsteps)
-#        2) nts : numsteps after which you want the acf; generally equal to numsteps+1 as to compare with extinction risk
-get_avg_acf<-function(sims,nts){
-  m<-sims[,,nts]
-  macf<-apply(X=m,MARGIN = 2,FUN=acf,lag.max = 1,type="correlation",plot=F)
-  acfs<-c() # initialize to store acf with lag1 for each col of matrix m
-  for(i in c(1:ncol(m))){
-    acfs<-c(acfs,macf[[i]]$acf[2])
+
+get_avg_acf<-function(sims){
+  #x<-sims[,,-1] # to get rid off the initial p0 values
+  x<-sims # dim(x) = numsims x numlocs x numsteps
+  acfs<-c() # to store acf from timeseries of each location for each simulation
+  
+  for(i in c(1:dim(x)[1])){ # loop for each numsim
+    m<-x[i,,] # matrix of dim = numlocs by numsteps
+    macf<-apply(X=m,MARGIN = 1,FUN=acf,lag.max = 1,type="correlation",plot=F) # along row : time series from each location
+    s<-unlist(macf)
+    id<-which(names(s)=="acf2")
+    sl1<-as.numeric(s[id])
+    acfs<-c(acfs,sl1)
   }
   ans<-mean(acfs)
   return(ans)
 }
-
 # I checked : ACF of noise (right tail) = ACF of noise (left tail)
 
 #-------------------------------------------------------------------------------------------------
@@ -190,15 +195,20 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
   pops1<-popsim_ml_D(p0=rep(p0,numlocs),ns=ns1,D=D,params=params,ext_thrs=ext_thrs,model=model)
   risk_right<-extrisk(pops1) # a vector
   
-  noise_acf_right<-get_avg_acf(sims=ns1,nts=numsteps) # a number
-  pop_acf_right<-get_avg_acf(sims=pops1,nts=numsteps+1) # a number
+  
+  pop_acf_right<-get_avg_acf(sims=pops1[,,-1]) # a number
   
   ns2<-(-ns1)
[…]

### 2019-02-16 ext_risk_copula.Rmd
all cleaned code as of now with plotting avg acf for pop

@@
 pdf("./Results/hassell/hassell_extrisk_d0.pdf",height = 6,width=6)
-plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.4,7.6),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.4,7.6),
+     xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_hassell,lty="dotted")
@@
 pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=6,width=6)
-plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),
+     xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_msmith,lty="dotted")
@@
 
-```{r fig_pennycuick, echo=F,  results="hide", cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_pennycuick, echo=F,  results="hide", cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
+source("mydispersal.R")
+
 model<-"pennycuick"
@@
 
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
+#for local dispersal in D matrix 
+res_pennycuick_local<-c()
+
 for(i in c(1:length(rlist))){
-    r<-rlist[i]
+    r<- rlist[i]
     params<-c(r,a,b)
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_pennycuick_local<-abind(res_pennycuick_local,temp_mat,along = 3)
+
     pennycuick_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
     pennycuick_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+}
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_pennycuick_local)<-list(NULL,cnm,anm)
+
+saveRDS(pennycuick_extrisk_d0,paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
+saveRDS(res_pennycuick_local,paste(resloc,"res_pennycuick_local.RDS",sep=""))
+
+# For global dispersal in D matrix 
+res_pennycuick_global<-c()
+for(r in rlist){
+    #cat("r=",r,"\n")
+    params<-c(r,a,b)
+    K_e<-(b/a)*(a+log(r-1))
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_pennycuick_global<-abind(res_pennycuick_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_pennycuick_global)<-list(NULL,cnm,anm)
+saveRDS(res_pennycuick_global,paste(resloc,"res_pennycuick_global.RDS",sep=""))
+```
+
+```{r fig_pennycuick, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_pennycuick_local)[3]){
+  s<-res_pennycuick_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 
-saveRDS(pennycuick_extrisk_d0,paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
 
-# plot for global dispersal in D matrix 
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.2))
-for(r in rlist){
-    params<-c(r,a,b)
-    K_e<-(b/a)*(a+log(r-1))
-    ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+for(i in 1:dim(res_pennycuick_global)[3]){
+  s<-res_pennycuick_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = c(2,7),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
-arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+     xlim = c(2,7),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_pennycuick,lty="dotted")
 
@@
 
-```{r fig_abrewer_scl_0.8, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_abrewer, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
+source("mydispersal.R")
 
@@
 p0<-K
-rc<-1/(K*(1-exp(-s*K)))
+rc_abrewer<-1/(K*(1-exp(-s*K)))
 
-rlist<-sort(c(0.002,0.004,0.006,0.008,0.01,0.015,rc,0.022,0.026))
+rlist<-sort(c(0.002,0.004,0.006,0.008,0.01,0.015,rc_abrewer,0.022,0.026))
 
@@
 
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+#for local dispersal in D matrix 
+res_abrewer_local<-c()
+
 for(i in c(1:length(rlist))){
-    r<-rlist[i]
+    r<- rlist[i]
     params<-c(r,K,s)
     ext_thrs<-p0/10
-    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
-    abrewer_extrisk_d0$risk_left_d0[i]<-s0$risk_left[1]
-    abrewer_extrisk_d0$risk_right_d0[i]<-s0$risk_right[1]
-    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_abrewer_local<-abind(res_abrewer_local,temp_mat,along = 3)
+
+    abrewer_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    abrewer_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+}
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_abrewer_local)<-list(NULL,cnm,anm)
+
+saveRDS(abrewer_extrisk_d0,paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
+saveRDS(res_abrewer_local,paste(resloc,"res_abrewer_local.RDS",sep=""))
+
+# For global dispersal in D matrix 
+res_abrewer_global<-c()
+for(r in rlist){
+    #cat("r=",r,"\n")
+    params<-c(r,K,s)
+    ext_thrs<-p0/10
+
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_abrewer_global<-abind(res_abrewer_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_abrewer_global)<-list(NULL,cnm,anm)
+saveRDS(res_abrewer_global,paste(resloc,"res_abrewer_global.RDS",sep=""))
+```
+
+```{r fig_abrewer, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
+pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_abrewer_local)[3]){
+  s<-res_abrewer_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 
-saveRDS(abrewer_extrisk_d0,paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
+pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
 
-# plot for global dispersal in D matrix 
-pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
-for(r in rlist){
-    params<-c(r,K,s)
-    ext_thrs<-p0/10
-    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+for(i in 1:dim(res_abrewer_global)[3]){
+  s<-res_abrewer_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 dev.off()
+```
 
-# Plot ext_risk_d0
+```{r fig_abrewer_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+abrewer_extrisk_d0<-readRDS(paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
+
 abrewer_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
 abrewer_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
-
 abrewer_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
 abrewer_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
-
 pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=6,width=6)
 plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
 arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025left,x1=abrewer_extrisk_d0$r,
        y1=abrewer_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_abrewer,lty="dotted")
 
@@
 
-```{r fig_verhulst_scl_2.5, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_verhulst, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
+source("mydispersal.R")
 
@@
 
-rlist<-seq(from=0.2,to=1.8,by=0.2)
+rlist<-seq(from=0.2,to=1.8,by=0.2) #rc=1
 
@@
 
-# plot for local dispersal in D matrix 
-pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
+#for local dispersal in D matrix 
+res_verhulst_local<-c()
+
 for(i in c(1:length(rlist))){
     r<-rlist[i]
+    params<-r
     ext_thrs<-p0/10
-    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=r,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,ploton = F)
-    verhulst_extrisk_d0$risk_left_d0[i]<-s0$risk_left[1]
-    verhulst_extrisk_d0$risk_right_d0[i]<-s0$risk_right[1]
-    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_verhulst_local<-abind(res_verhulst_local,temp_mat,along = 3)
+
+    verhulst_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    verhulst_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+}
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_verhulst_local)<-list(NULL,cnm,anm)
+
+saveRDS(verhulst_extrisk_d0,paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
+saveRDS(res_verhulst_local,paste(resloc,"res_verhulst_local.RDS",sep=""))
+
+# For global dispersal in D matrix 
+res_verhulst_global<-c()
+for(r in rlist){
+    params<-r
+    ext_thrs<-p0/10
+
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_verhulst_global<-abind(res_verhulst_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_verhulst_global)<-list(NULL,cnm,anm)
+saveRDS(res_verhulst_global,paste(resloc,"res_verhulst_global.RDS",sep=""))
+```
+
+```{r fig_verhulst, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
+pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_verhulst_local)[3]){
+  s<-res_verhulst_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 
-saveRDS(verhulst_extrisk_d0,paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
+pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
 
-# plot for global dispersal in D matrix 
-pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.2))
-for(r in rlist){
-    ext_thrs<-p0/10
-    s0<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=r,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,ploton = F)
-    plot(s0$d_seq,s0$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-         xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s0$d_seq,s0$risk_right,type="b",col="blue")
-    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-    mtext(paste0("r = ", r))
+for(i in 1:dim(res_verhulst_global)[3]){
+  s<-res_verhulst_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 dev.off()
+```
 
-# Plot ext_risk_d0
+```{r fig_verhulst_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+verhulst_extrisk_d0<-readRDS(paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
+
 verhulst_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
@@
 plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="ext_risk",cex.lab=1.5,pch=1)
+     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
 arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025left,x1=verhulst_extrisk_d0$r,
        y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=1,lty="dotted")
 

### 2019-02-16 mydispersal.R

@@ -212,7 +212,7 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
   
   if(ploton==T){
     # noise time series plot for last simulation (numsim=numsims) from each patches
-    pdf(paste0(resloc,"params_",params,"_noise_lastsim_timeseries_all_locs.pdf",sep=""),height=5,width=10)
+    pdf(paste0(resloc,"r_",params[1],"_noise_lastsim_timeseries_all_locs.pdf",sep=""),height=5,width=10)
     op<-par(mfrow=c(1,2),mar=c(3.5,4.5,2,3.5),mgp=c(1.9,0.5,0))
     
     ns_sim1<-ns1[numsims,,]
@@ -234,7 +234,7 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
     par(op)
     dev.off()
     
-    pdf(paste0(resloc,"params_",params,"_pops_lastsim_timeseries_all_locs.pdf",sep=""),height=5,width=10)
+    pdf(paste0(resloc,"r_",params[1],"_pops_lastsim_timeseries_all_locs.pdf",sep=""),height=5,width=10)
     op<-par(mfrow=c(1,2),mar=c(3.5,4.5,2,3.5),mgp=c(1.9,0.5,0))
     
     pop_sim1<-pops1[numsims,,]
@@ -256,7 +256,7 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
     par(op)
     dev.off()
     
-    pdf(paste0(resloc,"params_",params,"_extrisk_vs_time.pdf",sep=""),height=5,width=10)
+    pdf(paste0(resloc,"r_",params[1],"_extrisk_vs_time.pdf",sep=""),height=5,width=10)
     op<-par(mfrow=c(1,2),mar=c(3.5,4.5,2,3.5),mgp=c(1.9,0.5,0))
     plot(1:length(risk_right),risk_right,type='b',
          xlab='Time',ylab='Risk',col='blue',panel.first = grid())

### 2019-02-17 ext_risk_copula.Rmd
running code with acf on 26 timesteps and omitting austin-brewer results

@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
 
@@
 
+# Verhulst model :
+
+```{r sim_verhulst, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+source("mydispersal.R")
+
+model<-"verhulst"
+scl<-2.5
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+numsims<-10000
+numsteps<-25
+numlocs<-5
+p0<-1
+
+rlist<-seq(from=0.2,to=1.8,by=0.2) #rc=1
+
+verhulst_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
+colnames(verhulst_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+#for local dispersal in D matrix 
+res_verhulst_local<-c()
+
+for(i in c(1:length(rlist))){
+    r<-rlist[i]
+    params<-r
+    ext_thrs<-p0/10
+
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_verhulst_local<-abind(res_verhulst_local,temp_mat,along = 3)
+
+    verhulst_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    verhulst_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+}
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_verhulst_local)<-list(NULL,cnm,anm)
+
+saveRDS(verhulst_extrisk_d0,paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
+saveRDS(res_verhulst_local,paste(resloc,"res_verhulst_local.RDS",sep=""))
+
+# For global dispersal in D matrix 
+res_verhulst_global<-c()
+for(r in rlist){
+    params<-r
+    ext_thrs<-p0/10
+
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_verhulst_global<-abind(res_verhulst_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_verhulst_global)<-list(NULL,cnm,anm)
+saveRDS(res_verhulst_global,paste(resloc,"res_verhulst_global.RDS",sep=""))
+```
+
+```{r fig_verhulst, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
+pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_verhulst_local)[3]){
+  s<-res_verhulst_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+
+for(i in 1:dim(res_verhulst_global)[3]){
+  s<-res_verhulst_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.04,0.04)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
+
+  ## a little farther out (line=3.2) to make room for labels
+  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
+}
+par(op)
+dev.off()
+```
+
+```{r fig_verhulst_extrisk_d0, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+verhulst_extrisk_d0<-readRDS(paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
+
+verhulst_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
+verhulst_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
+
+verhulst_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
+verhulst_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
+
+pdf(paste(resloc,"verhulst_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025left,x1=verhulst_extrisk_d0$r,
+       y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=1,lty="dotted")
+
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025right,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(0.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=2.5); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
+
+<!--
 # Austin Brewer model :
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
 
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
-
-# Verhulst model :
-
-```{r sim_verhulst, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
-set.seed(seed)
-source("mydispersal.R")
-
-model<-"verhulst"
-scl<-2.5
-
-tempo<-paste("./Results/",model,"_scl_",scl,sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-
-numsims<-10000
-numsteps<-25
-numlocs<-5
-p0<-1
-
-rlist<-seq(from=0.2,to=1.8,by=0.2) #rc=1
-
-verhulst_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(verhulst_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
-
-#for local dispersal in D matrix 
-res_verhulst_local<-c()
-
-for(i in c(1:length(rlist))){
-    r<-rlist[i]
-    params<-r
-    ext_thrs<-p0/10
-
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
-    res_verhulst_local<-abind(res_verhulst_local,temp_mat,along = 3)
-
-    verhulst_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    verhulst_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-}
-
-anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
-dimnames(res_verhulst_local)<-list(NULL,cnm,anm)
-
-saveRDS(verhulst_extrisk_d0,paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
-saveRDS(res_verhulst_local,paste(resloc,"res_verhulst_local.RDS",sep=""))
-
-# For global dispersal in D matrix 
-res_verhulst_global<-c()
-for(r in rlist){
-    params<-r
-    ext_thrs<-p0/10
-
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
-    res_verhulst_global<-abind(res_verhulst_global,temp_mat,along = 3)
-}
-anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
-dimnames(res_verhulst_global)<-list(NULL,cnm,anm)
-saveRDS(res_verhulst_global,paste(resloc,"res_verhulst_global.RDS",sep=""))
-```
-
-```{r fig_verhulst, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
-pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
-
-for(i in 1:dim(res_verhulst_local)[3]){
-  s<-res_verhulst_local[,,i]
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  lines(s$d, s$pop_avg_acf_right, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
-}
-par(op)
-dev.off()
-
-pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
-
-for(i in 1:dim(res_verhulst_global)[3]){
-  s<-res_verhulst_global[,,i]
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
-}
-par(op)
-dev.off()
-```
-
-```{r fig_verhulst_extrisk_d0, echo=F,results="hide"}
-set.seed(seed)
-numsims<-10000
-verhulst_extrisk_d0<-readRDS(paste(resloc,"verhulst_extrisk_d0.RDS",sep=""))
-
-verhulst_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
-verhulst_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
-
-verhulst_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
-verhulst_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"verhulst_extrisk_d0.pdf",sep=""),height=6,width=6)
-plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
-arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025left,x1=verhulst_extrisk_d0$r,
-       y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=1,lty="dotted")
-
-lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
-arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025right,
-       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(0.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-dev.off()
-```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=2.5); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
+-->

### 2019-02-17 mydispersal.R

@@ -53,7 +53,7 @@ Dispersal_mat<-function(numlocs,d,disp_everywhere){
 #--------Output------------------------------------------------------
 #A numsims by numlocs by numsteps+1 array of populations
 #----------------------------------------------------------------------
-popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
+popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model,checkon){
   numsims<-dim(ns)[1]
   numlocs<-dim(ns)[2]
   numsteps<-dim(ns)[3]
@@ -108,7 +108,16 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
       r<-params[1]
       K<-params[2]
       s<-params[3]
-      lam_austin<-(1+(r*(K-res[,,tct])*(1-exp(-s*res[,,tct]))))*res[,,tct] 
+      
+      xx<-res[,,tct]
+      
+      if(checkon==T){
+        cat("tct=",tct,"\n")
+        saveRDS(xx,"./res_tct_abrewer.RDS")
+        saveRDS(lam_sto,"./lam_sto_abrewer.RDS")
+      }
+      
+      lam_austin<-xx*(1+(r*(K-xx)*(1-exp(-(s*xx)))))
       res[,,tct+1]<-lam_austin*lam_sto #numsims by numlocs matrix
     }else if(model=="varley"){
       r<-params[1]
@@ -162,7 +171,9 @@ get_avg_acf<-function(sims){
     sl1<-as.numeric(s[id])
     acfs<-c(acfs,sl1)
   }
-  ans<-mean(acfs)
+  # which(is.na(acfs))
+  ans<-mean(acfs,na.rm=T) # This line to remove NaN which is produced due to pop = 0 throughout the time at any patch
+                          # you can see that in the row of matrix m (for a given i)
   return(ans)
 }
 # I checked : ACF of noise (right tail) = ACF of noise (left tail)
@@ -183,7 +194,7 @@ get_avg_acf<-function(sims){
 #       10) ploton : logical(T or F) to get optional plot
 #       11) resloc : location to save plots
 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc){
[…]

### 2019-02-17 ext_risk_copula.Rmd
models ok, excluding abrewer

@@
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
@@
 
-  ylm2<-c(-0.04,0.04)
+  ylm2<-c(-0.2,0.6)
   ## Plot the second plot and put axis scale on right
@@
        axes=FALSE, type="l", col="orange")
+  abline(h=0,lty="dotted",col="grey")
   lines(s$d, s$pop_avg_acf_right, col="skyblue")
@@
        axes=FALSE, type="l", col="orange")
+  abline(h=0,lty="dotted",col="grey")
   lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
@@
 
-pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
+pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
 op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
@@
 
-```{r get_rc_pennycuick,echo=F,cache=T,cache.extra=list(seed,mtime("1var_pop_model_bifurcation.R"))}
+```{r get_rc_pennycuick,echo=F,results="hide",cache=T,cache.extra=list(seed,mtime("1var_pop_model_bifurcation.R"))}
 source("./1var_pop_model_bifurcation.R")
@@
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
   lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
          col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=2.5); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--
-# Austin Brewer model :
-
-```{r sim_abrewer, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
-set.seed(seed)
-source("mydispersal.R")
-
-model<-"austinbrewer"
-scl<-0.8
-
-tempo<-paste("./Results/",model,"_scl_",scl,sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-
-numsims<-10000
-numsteps<-25
-numlocs<-5
-s<-0.2
-K<-50
-p0<-K
-rc_abrewer<-1/(K*(1-exp(-s*K)))
-
-rlist<-sort(c(0.002,0.004,0.006,0.008,0.01,0.015,rc_abrewer,0.022,0.026))
-
-abrewer_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(abrewer_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
-
-#for local dispersal in D matrix 
-res_abrewer_local<-c()
-
-for(i in c(1:length(rlist))){
-    r<- rlist[i]
-    params<-c(r,K,s)
-    ext_thrs<-p0/10
-
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
-    res_abrewer_local<-abind(res_abrewer_local,temp_mat,along = 3)
-
-    abrewer_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    abrewer_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-}
-
-anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
-dimnames(res_abrewer_local)<-list(NULL,cnm,anm)
-
-saveRDS(abrewer_extrisk_d0,paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
-saveRDS(res_abrewer_local,paste(resloc,"res_abrewer_local.RDS",sep=""))
-
-# For global dispersal in D matrix 
-res_abrewer_global<-c()
-for(r in rlist){
-    #cat("r=",r,"\n")
-    params<-c(r,K,s)
-    ext_thrs<-p0/10
-
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
-    res_abrewer_global<-abind(res_abrewer_global,temp_mat,along = 3)
-}
-anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
-dimnames(res_abrewer_global)<-list(NULL,cnm,anm)
-saveRDS(res_abrewer_global,paste(resloc,"res_abrewer_global.RDS",sep=""))
-```
-
-```{r fig_abrewer, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
-pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
-
-for(i in 1:dim(res_abrewer_local)[3]){
-  s<-res_abrewer_local[,,i]
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  lines(s$d, s$pop_avg_acf_right, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
-}
-par(op)
-dev.off()
-
-pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
-
-for(i in 1:dim(res_abrewer_global)[3]){
-  s<-res_abrewer_global[,,i]
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
-}
-par(op)
-dev.off()
-```
-
-```{r fig_abrewer_extrisk_d0, echo=F,results="hide"}
-set.seed(seed)
-numsims<-10000
-abrewer_extrisk_d0<-readRDS(paste(resloc,"abrewer_extrisk_d0.RDS",sep=""))
-
-abrewer_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
-abrewer_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
-abrewer_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
-abrewer_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
-pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=6,width=6)
-plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
-arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025left,x1=abrewer_extrisk_d0$r,
-       y1=abrewer_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_abrewer,lty="dotted")
-
-lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
-arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025right,
-       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(0.005,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-dev.off()
-```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_abrewer_r_vary_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
--->

### 2019-02-18 ext_risk_copula.Rmd
adding abrewer model

--- 2019-02-17 ext_risk_copula.Rmd
@@
 
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
+  ylm2<-c(-0.2,0.6)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  abline(h=0,lty="dotted",col="grey")
   lines(s$d, s$pop_avg_acf_right, col="skyblue")
@@
 
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
+  ylm2<-c(-0.2,0.6)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  abline(h=0,lty="dotted",col="grey")
   lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
@@
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.2,0.6)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  abline(h=0,lty="dotted",col="grey")
   lines(s$d, s$pop_avg_acf_right, col="skyblue")
@@
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.04,0.04)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", rlist[i]))
+
+  ## Allow a second plot on the same graph
+  par(new=TRUE)
+
+  ylm2<-c(-0.2,0.6)
+  ## Plot the second plot and put axis scale on right
+  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
+       axes=FALSE, type="l", col="orange")
+  abline(h=0,lty="dotted",col="grey")
   lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
@@
 
-<!--
 # Austin Brewer model :
@@
 K<-50
-p0<-K
+
 rc_abrewer<-1/(K*(1-exp(-s*K)))
@@
   r<- rlist[i]
+  K<-50
+  s<-0.2
+  p0<-K
   params<-c(r,K,s)
@@
   #cat("r=",r,"\n")
+  K<-50
+  s<-0.2
+  p0<-K
   params<-c(r,K,s)
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
--->

### 2019-02-20 ext_risk_copula.Rmd
omitting acf results; adding r upto bifurcation points : 95% interval

@@
 source("mydispersal.R")
-rlist<-c(0,0.1,0.3,0.5,0.7,0.9,1.0,1.1,1.3,1.5,1.7,1.9) #rc = 1
+
+rc_ricker<-1 # transition from under to overcompensatory dynamics
+r_bf_ricker<-2 # bifurcation parameter stable to unstable eqm
+
+qrng<-runif(10000,min=0,max=r_bf_ricker)
+qr<-quantile(qrng,probs=c(.025,.975))
+
+rlist<-unique(c(seq(from=range(qr)[1],to=rc_ricker,length.out = 5),c(seq(from=rc_ricker,to=range(qr)[2],length.out = 5))))
+
 numsims<-10000
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
 ```{r fig_ricker, echo=F, results="hide",cache=T, cache.extra=list(seed,res_ricker_local,res_ricker_global)}
-pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
 
@@
 
+  #---------------------------------------------------
   ## Allow a second plot on the same graph
-  par(new=TRUE)
+  #par(new=TRUE)
 
-  ylm2<-c(-0.2,0.6)
+  #ylm2<-c(-0.2,0.6)
   ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+  #plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
        #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  abline(h=0,lty="dotted",col="grey")
-  lines(s$d, s$pop_avg_acf_right, col="skyblue")
+  #     axes=FALSE, type="l", col="orange")
+ # abline(h=0,lty="dotted",col="grey")
+ # lines(s$d, s$pop_avg_acf_right, col="skyblue")
 
   ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
+ # mtext("Avg. ACF",side=4,line=3.2,col="green4") 
+ # axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
+  #---------------------------------------------------------
 }
@@
 
-pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
 
@@
   mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.2,0.6)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  abline(h=0,lty="dotted",col="grey")
-  lines(s$d, s$pop_avg_acf_right, pch=16, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
        angle=90,length=0.1,col="red",code=3)
-abline(v=1,lty="dotted") #rc=1
+abline(v=rc_ricker,lty="dotted") #rc=1
 
@@
 
-rlist<-seq(from=1.5,to=7.5,by=0.5)
 numsims<-10000
@@
 scl<-1
-rc_hassell<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition 
-rlist<-sort(c(rlist,rc_hassell))
+rc_hassell<-((b/(b-1)))^b # transition from under to overcompensatory dynamics
+r_bf_hassell<- ((b/(b-2)))^b # bifurcation parameter stable to unstable eqm
+
+qrng<-runif(10000,min=0,max=r_bf_hussell)
+qr<-quantile(qrng,probs=c(.025,.975))
+
+rlist<-unique(c(seq(from=range(qr)[1],to=rc_hassell,length.out = 5),c(seq(from=rc_hassell,to=range(qr)[2],length.out = 5))))
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
 ```{r fig_hassell, echo=F, results="hide",cache=T, cache.extra=list(seed,res_hassell_local,res_hassell_global)}
-pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=18)
-op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,6.2))
+pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
 
@@
   mtext(paste0("r = ", rlist[i]))
-
-  ## Allow a second plot on the same graph
-  par(new=TRUE)
-
-  ylm2<-c(-0.2,0.6)
-  ## Plot the second plot and put axis scale on right
-  plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-       axes=FALSE, type="l", col="orange")
-  abline(h=0,lty="dotted",col="grey")
-  lines(s$d, s$pop_avg_acf_right, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
-  mtext("Avg. ACF",side=4,line=3.2,col="green4") 
-  axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
 }
@@
 
-pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=18)
-op<-par(mfrow=c(3,5),mar=c(5.2,4.2,1.2,6.2))
+pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
 
@@
 rc_msmith<-b/(b-1)
-rlist<-sort(c(seq(from=1.2,to=2.8,by=0.2),rc_msmith))
+r_bf_msmith<-b/(b-2) # bifurcation parameter stable to unstable eqm
+
+qrng<-runif(10000,min=0,max=r_bf_msmith)
+qr<-quantile(qrng,probs=c(.025,.975))
+
+rlist<-unique(c(seq(from=range(qr)[1],to=rc_msmith,length.out = 5),c(seq(from=rc_msmith,to=range(qr)[2],length.out = 5))))
 
@@
 ```{r fig_msmith, echo=F, results="hide",cache=T, cache.extra=list(seed,res_msmith_local,res_msmith_global)}
-pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
+pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
@@
 
-pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
+pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
@@
 eps<-1e-10 #you can change it upto which decimal point you want it to be exact
-rlist<-seq(from=4.3231657,to=4.3231658,by=eps)
+rlist<-seq(from=4.3231657,to=4.3231659,by=eps)
 res_pennycuick<-c() 
 for(r in rlist){
-  ans<-fn_d1_eqm(params=c(r,a,b),model="pennycuick")
+  ans<-fn_d1_eqm(params=c(r,a_pennycuick,b_pennycuick),model="pennycuick")
   res_pennycuick<-c(res_pennycuick,ans)
@@
 formatC(rc_pennycuick,digits = 10, format="f") # accurate upto 10th decimal
+
+# This is to find the critical r value where stable to unstable eqm occurs
+rlist<-seq(from=9.4672389,to=9.46724,by=eps)
+#formatC(rlist,digits = 14, format="f")
+
+res_pennycuick<-c() 
+for(r in rlist){
+  ans<-fn_d1_eqm(params=c(r=r,a_pennycuick,b_pennycuick),model="pennycuick")
+  res_pennycuick<-c(res_pennycuick,ans)
+}
+
+#plot(rlist,res_pennycuick,col="red",type="l",xlab="r",ylab="pennycuick_1st_der_at_eqm")
+#abline(h=0)
+#fp<-0
+fp<--1
+r_bf_pennycuick<-rlist[which.min(abs(res_pennycuick-fp))]
+#formatC(rc_pen_uns,digits = 14, format="f") 
 ```
 
-```{r sim_pennycuick, echo=F,  results="hide", cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_pennycuick, echo=F,  results="hide", cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-rlist<-sort(c(seq(from=2,to=7,by=0.5),rc_pennycuick))
+qrng<-runif(10000,min=0,max=r_bf_pennycuick)
+qr<-quantile(qrng,probs=c(.025,.975))
+
+rlist<-unique(c(seq(from=range(qr)[1],to=rc_pennycuick,length.out = 5),c(seq(from=rc_pennycuick,to=range(qr)[2],length.out = 5))))
 
@@
 ```{r fig_pennycuick, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=15)
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
@@
 
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=15)
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
@@
 
-rlist<-seq(from=0.2,to=1.8,by=0.2) #rc=1
+rc_verhulst<-1 # transition from under to overcompensatory dynamics
+r_bf_verhulst<-2 # bifurcation parameter stable to unstable eqm
+
+qrng<-runif(10000,min=0,max=r_bf_verhulst)
+qr<-quantile(qrng,probs=c(.025,.975))
+
+rlist<-unique(c(seq(from=range(qr)[1],to=rc_verhulst,length.out = 5),c(seq(from=rc_verhulst,to=range(qr)[2],length.out = 5))))
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
 ```{r fig_verhulst, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
-pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
@@
 
-pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
@@
        y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=1,lty="dotted")
+abline(v=rc_verhulst,lty="dotted")
 
@@
 
-rc_abrewer<-1/(K*(1-exp(-s*K)))
+rc_abrewer<-1/(K*(1-exp(-s*K))) # transition from under to overcompensatory dynamics
+r_bf_abrewer<-2*rc_abrewer # bifurcation parameter stable to unstable eqm
 
-rlist<-sort(c(0.002,0.004,0.006,0.008,0.01,0.015,rc_abrewer,0.022,0.026))
+qrng<-runif(10000,min=0,max=r_bf_abrewer)
+qr<-quantile(qrng,probs=c(.025,.975))
+
+rlist<-unique(c(seq(from=range(qr)[1],to=rc_abrewer,length.out = 5),c(seq(from=rc_abrewer,to=range(qr)[2],length.out = 5))))
 
@@
   s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F)
+               scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
   s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F)
+               scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
 
@@
 ```{r fig_abrewer, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
-pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
@@
 
-pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))

### 2019-02-20 1var_pop_models.R

@@ -118,14 +118,14 @@ Pennycuick<-function(r,a,b,p0,lensim){
     p0<-pt
   }
   
-  plot(time,pop,type="b")
+  plot(time,pop,type="l")
   
 }
 
 #call the function
 a<-0.1
 b<-0.1
-r<-2
+r<-10
 K_e<-(b/a)*(a+log(r-1))
 K_e
 Pennycuick(r=r,a=a,b=b,p0=K_e-0.01,lensim=200) 
@@ -176,7 +176,7 @@ s<-0.2
 K<-50
 rc<-1/(K*(1-exp(-s*K)))
 rc
-r<-0.026
+r<-0.004
 AustinBrewer(r=r,K=K,s=s,p0=1,lensim=200) 
 
 #----------------------------------------------------

### 2019-02-20 mydispersal.R

@@ -109,15 +109,13 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model,checkon){
       K<-params[2]
       s<-params[3]
       
-      xx<-res[,,tct]
-      
       if(checkon==T){
         cat("tct=",tct,"\n")
-        saveRDS(xx,"./res_tct_abrewer.RDS")
+        saveRDS(res[,,tct],"./res_tct_abrewer.RDS")
         saveRDS(lam_sto,"./lam_sto_abrewer.RDS")
       }
       
-      lam_austin<-xx*(1+(r*(K-xx)*(1-exp(-(s*xx)))))
+      lam_austin<-(1+(r*(K-res[,,tct])*(1-exp(-(s*res[,,tct])))))*res[,,tct]
       res[,,tct+1]<-lam_austin*lam_sto #numsims by numlocs matrix
     }else if(model=="varley"){
       r<-params[1]
@@ -193,8 +191,9 @@ get_avg_acf<-function(sims){
 #       9) model : a character specifying model name
 #       10) ploton : logical(T or F) to get optional plot
 #       11) resloc : location to save plots
+#       checkon, getacf : these are tags
 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,checkon){
+plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,checkon,getacf){
  
   ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
   #                                                      numpoints=numsteps*numsims)
@@ -206,8 +205,6 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
   pops1<-popsim_ml_D(p0=rep(p0,numlocs),ns=ns1,D=D,params=params,ext_thrs=ext_thrs,model=model,checkon=checkon)
   risk_right<-extrisk(pops1) # a vector
   
-  pop_acf_right<-get_avg_acf(sims=pops1) # a number
-  
   ns2<-(-ns1)
   pops2<-popsim_ml_D(p0=rep(p0,numlocs),ns=ns2,D=D,params=params,ext_thrs=ext_thrs,model=model,checkon=checkon)
   risk_left<-extrisk(pops2) # a vector
[…]

### 2019-02-20 ext_risk_copula.Rmd
several changes

@@
 pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+#op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
          col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
+  mtext(paste0("r = ", round(rlist[i],5)))
 
@@
 pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
          col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", rlist[i]))
+  mtext(paste0("r = ", round(rlist[i],5)))
 }
@@
 
-qrng<-runif(10000,min=0,max=r_bf_hussell)
+qrng<-runif(10000,min=0,max=r_bf_hassell)
 qr<-quantile(qrng,probs=c(.025,.975))
@@
 pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 
@@
 pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
 

### 2019-02-20 ext_risk_copula.Rmd
bug fixed

@@
 
-qrng<-runif(10000,min=0,max=r_bf_hassell)
+qrng<-runif(10000,min=1,max=r_bf_hassell) # r must be > 1
 qr<-quantile(qrng,probs=c(.025,.975))
@@
 pdf("./Results/hassell/hassell_extrisk_d0.pdf",height = 6,width=6)
-plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.4,7.6),
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(rlist),
      xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
@@
 
-qrng<-runif(10000,min=0,max=r_bf_msmith)
+qrng<-runif(10000,min=1,max=r_bf_msmith) #r must be > 1 for +ve eqm pt.
 qr<-quantile(qrng,probs=c(.025,.975))

### 2019-02-20 ext_risk_copula.Rmd
bug fixed
minor changes

@@
 plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",
-     ylim=c(0,1),xlim = c(0,2),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     ylim=c(0,1),xlim = range(ricker_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
 arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,
@@
 pdf("./Results/hassell/hassell_extrisk_d0.pdf",height = 6,width=6)
-plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(rlist),
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
      xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
@@
 pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=6,width=6)
-plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = c(1.2,2.8),
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(msmith_extrisk_d0$r),
      xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
@@
 mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(1.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+legend(1.2,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
 dev.off()
@@
 
-qrng<-runif(10000,min=0,max=r_bf_pennycuick)
+qrng<-runif(10000,min=1,max=r_bf_pennycuick) # r must be > 1
 qr<-quantile(qrng,probs=c(.025,.975))
@@
 plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = c(2,7),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
 arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,

### 2019-02-20 ext_risk_copula.Rmd
rlist modified

@@
 
-qrng<-runif(10000,min=0,max=r_bf_ricker)
-qr<-quantile(qrng,probs=c(.025,.975))
+rng1<-runif(10000,min=0,max=r_bf_ricker)
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_ricker)
+q2<-abs(unname(qr1[2])-rc_ricker)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
 
-rlist<-unique(c(seq(from=range(qr)[1],to=rc_ricker,length.out = 5),c(seq(from=rc_ricker,to=range(qr)[2],length.out = 5))))
+rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Ricker model !!!",immediate.=T,call.=T)
+}
 
@@
 
-qrng<-runif(10000,min=1,max=r_bf_hassell) # r must be > 1
-qr<-quantile(qrng,probs=c(.025,.975))
+rng1<-runif(10000,min=1,max=r_bf_hassell) # r must be > 1
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_hassell)
+q2<-abs(unname(qr1[2])-rc_hassell)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
 
-rlist<-unique(c(seq(from=range(qr)[1],to=rc_hassell,length.out = 5),c(seq(from=rc_hassell,to=range(qr)[2],length.out = 5))))
+rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Hassell model !!!",immediate.=T,call.=T)
+}
 
@@
 
-qrng<-runif(10000,min=1,max=r_bf_msmith) #r must be > 1 for +ve eqm pt.
-qr<-quantile(qrng,probs=c(.025,.975))
+rng1<-runif(10000,min=1,max=r_bf_msmith) #r must be > 1 for +ve eqm pt.
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_msmith)
+q2<-abs(unname(qr1[2])-rc_msmith)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
 
-rlist<-unique(c(seq(from=range(qr)[1],to=rc_msmith,length.out = 5),c(seq(from=rc_msmith,to=range(qr)[2],length.out = 5))))
+rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Maynard-smith model !!!",immediate.=T,call.=T)
+}
 
@@
 
-qrng<-runif(10000,min=1,max=r_bf_pennycuick) # r must be > 1
-qr<-quantile(qrng,probs=c(.025,.975))
+rng1<-runif(10000,min=1,max=r_bf_pennycuick) #r must be > 1 for +ve eqm pt.
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_pennycuick)
+q2<-abs(unname(qr1[2])-rc_pennycuick)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
 
-rlist<-unique(c(seq(from=range(qr)[1],to=rc_pennycuick,length.out = 5),c(seq(from=rc_pennycuick,to=range(qr)[2],length.out = 5))))
+rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Pennycuick model !!!",immediate.=T,call.=T)
+}
 
@@
 
-qrng<-runif(10000,min=0,max=r_bf_verhulst)
-qr<-quantile(qrng,probs=c(.025,.975))
+rng1<-runif(10000,min=0,max=r_bf_verhulst)
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_verhulst)
+q2<-abs(unname(qr1[2])-rc_verhulst)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
 
-rlist<-unique(c(seq(from=range(qr)[1],to=rc_verhulst,length.out = 5),c(seq(from=rc_verhulst,to=range(qr)[2],length.out = 5))))
+rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Verhulst model !!!",immediate.=T,call.=T)
+}
 
@@
 
-qrng<-runif(10000,min=0,max=r_bf_abrewer)
-qr<-quantile(qrng,probs=c(.025,.975))
+rng1<-runif(10000,min=0,max=r_bf_abrewer)
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_abrewer)
+q2<-abs(unname(qr1[2])-rc_abrewer)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
 
-rlist<-unique(c(seq(from=range(qr)[1],to=rc_abrewer,length.out = 5),c(seq(from=rc_abrewer,to=range(qr)[2],length.out = 5))))
+rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Austin-Brewer model !!!",immediate.=T,call.=T)
+}
 

### 2019-02-21 ext_risk_cop_age_str.R
adding code for age structured 2var model

@@ -1,0 +1,149 @@
+library(copula)
+library(VineCopula)
+library(mvtnorm)
+
+# This is the simulator for age structured 2 variable pop model with fecundity and survival drawn from a specific copula family
+
+# Input :
+#       p0 : a vector of two numeric values : E0, A0 initial values for Egg and Adult
+#       ext_thrs : extinction threshold, generally = 1
+#       cop : a bivariate copula matrix
+#       par_dist : a vector containing 4 numbers for gamma and beta distribution parameters: gshape, gscale, bshape1, bshape2
+#       numsims : number of simulations
+#       numsteps : number of time steps
+#       ploton : logical to get optional plot
+#       resloc : location path to save optional plot
+
+# Output :
+#       A list of three : resE, resA : each is a numsims by numsteps+1 matrix for Egg and Adult stage
+#                         erA : a vector of length numsteps+1 : extinction risk for adult
+
+sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
+  
+  gshape<-par_dist[1]
+  gscale<-par_dist[2]
+  bshape1<-par_dist[3]
+  bshape2<-par_dist[4]
+  
+  #initialization for Egg and Adult pop variables
+  resE<-matrix(NA,nrow=numsims, ncol=numsteps+1)
+  resA<-matrix(NA,nrow=numsims, ncol=numsteps+1)
+  
+  resE[,1]<-p0[1]
+  resA[,1]<-p0[2]
+  
+  # make the transformation on cop
+  
+  # 1st column : fecundity of adult : gamma distributed
+  x <- qgamma(cop[,1],shape=gshape,scale=gscale)
+  
+  # 2nd column : survival of eggs : beta distributed
+  y <- qbeta(cop[,2],shape1=bshape1,shape2=bshape2) 
+  
+  transcop<-cbind(x,y)
+  
+  if(ploton==ploton){
+    pdf(paste(resloc,"fA_sE_gshape_",gshape,"_gscale_",gscale,"_bshape1_",bshape1,"_bshape2_",bshape2,".pdf",sep=""),height=5,width=10)
+    op<-par(mfrow=c(1,2))
[…]

### 2019-02-21 ext_risk_copula.Rmd
modified

@@
 #Ricker model :
-```{r sim_ricker, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_ricker, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-rlist<-seq(from=range(qr1)[iqmin],to=range(qr1)[iqmax], by = qs[iqmin]/4)
+fromr<-range(qr1)[iqmin]
+tor<-range(qr1)[iqmax]
+
+if(fromr>tor){
+  byr<--qs[iqmin]/4
+}else{
+  byr<-qs[iqmin]/4
+}
+
+rlist<-sort(seq(from=fromr,to=tor, by = byr))
 
@@
 
-```{r sim_hassell, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_hassell, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r sim_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_msmith, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r sim_pennycuick, echo=F,  results="hide", cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_pennycuick, echo=F, cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r sim_verhulst, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_verhulst, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r sim_abrewer, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_abrewer, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)

### 2019-02-21 age_structure.Rmd
age structured model results

---
title: "extinction risk with one patch age structured model"
author: "Shyamolina Ghosh"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
source("mtime.R")
```

# Age structure model :

```{r sim_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")

numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)

ktau<-0.6  # kendall's tau

# for clayton
copC<-claytonCopula(3)
parC<-iTau(copC,tau=ktau)
par<-parC
fam<-3

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

C_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(C_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------

# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same

fam<-13
par<-parC

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------

# Now for gumbel

fam<-4
copG<-gumbelCopula(2)
parG<-iTau(copG,tau = ktau)
par<-parG

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))

#---------------------------------------------------------------------------------

# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same

fam<-14
par<-parG

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------
```

```{r fig_extrisk_agestr_C_SC_G_SG, echo=F,results="hide"}

res_age_C<-readRDS("./Results/age_str_results/C_par_3_ktau_0.6/C_sim_age_str.RDS")
res_age_SC<-readRDS("./Results/age_str_results/SC_par_3_ktau_0.6/SC_sim_age_str.RDS")
res_age_G<-readRDS("./Results/age_str_results/G_par_2.5_ktau_0.6/G_sim_age_str.RDS")
res_age_SG<-readRDS("./Results/age_str_results/SG_par_2.5_ktau_0.6/SG_sim_age_str.RDS")

xC<-res_age_C$extdf
xSC<-res_age_SC$extdf
xG<-res_age_G$extdf
xSG<-res_age_SG$extdf

t<-c(0:(nrow(xC)-1))

pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_G_SG_ktau_",ktau,".pdf",sep=""),height=6,width=12)
op<-par(mfrow=c(1,2),mar=c(4,4,2,2),mgp=c(2.5,0.5,0))

# for C, SC
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())

x<-xC
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
lines(t,x$erA,col="black",type="l")

x<-xSC
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")

legend("topleft", c("C","SC"), lty=c(1,1),  
       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)

# for G, SG
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())

x<-xG
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
lines(t,x$erA,col="black",type="l")

x<-xSG
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")

legend("topleft", c("G","SG"), lty=c(1,1),  
       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)

par(op)
dev.off()
```

![Extinction risk from one patch age structured model, fecundity of adults and survival of eggs have different asymmetric copula atructure but with same kendall correlation = 0.6, numsims=10000.\label{fig_extrisk_agestr}](./Results/age_str_results/extrisk_age_str_C_SC_G_SG_ktau_0.6.pdf){ width=40%, height=40% }

### 2019-02-21 ext_risk_cop_age_str.R

@@ -16,7 +16,7 @@ library(mvtnorm)
 
 # Output :
 #       A list of three : resE, resA : each is a numsims by numsteps+1 matrix for Egg and Adult stage
-#                         erA : a vector of length numsteps+1 : extinction risk for adult
+#                         extdf : a dataframe (nrow= numsteps+1, ncol=3) : extinction risk for adult and its 95% CI
 
 sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
   
@@ -81,16 +81,18 @@ sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
   if(ploton==ploton){
   pdf(paste(resloc,BiCopName(family = fam),"_ext_riskA_vs_time.pdf",sep=""),height=5,width=5)
   op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-  plot(c(0:numsteps),erA,xlab="time",ylab="Extinction risk",cex.lab=1.5,cex=0.5,pch=16,col="black",ylim=c(0,1),panel.first = grid())
+  plot(c(0:numsteps),erA,xlab="time",ylab="Extinction risk",cex.lab=1.5,cex=0.3,pch=16,col="black",ylim=c(0,1),panel.first = grid())
   arrows(x0=c(1:numsteps),y0=CI0.025[2:(numsteps+1)],x1=c(1:numsteps),y1=CI0.975[2:(numsteps+1)],
-         angle=90,length=0.1,col="grey",code=3)
+         angle=90,length=0.01,col="grey",code=3)
   par(op)
   dev.off()
   }
   
+  extdf<-as.data.frame(cbind(CI0.025,erA,CI0.975))
+  
   return(list(resE=resE,
               resA=resA,
-              erA=erA))
+              extdf=extdf))
 }
 #----------------------------------------------------------------------------
 # function to calculate extinction risk
@@ -100,50 +102,13 @@ get_er<-function(mat){
 }
 
 #--------------------------------
-# Now calling the functions
-set.seed(seed=101)
-numsims<-10000
-numsteps<-25
-p0<-c(5,10)
-ext_thrs<-2
-fam<-14
-par<-4
-gshape<-2
-gscale<-1
-bshape1<-3
-bshape2<-1.5
-par_dist<-c(gshape,gscale,bshape1,bshape2)
[…]

### 2019-02-21 ExtremeTailDep.R
modified

@@ -5,13 +5,15 @@
 #n        Number of points to get
 #d        Dimension
 #rl       1 for right tail dependence, -1 for left tail dependence
+#mn       mean and 
+#sdev     standard deviation
 #
 #Output
 #A matrix of dimension n by d, marginals are standard normal.
 #
-retd<-function(n,d,rl)
+retd<-function(n,d,rl,mn,sdev)
 {
-  m<-matrix(rnorm(n*(d+1),mean=0,sd=1),n,d+1)
+  m<-matrix(rnorm(n*(d+1),mean=mn,sd=sdev),n,d+1)
   flag<-(m[,1]>0) #a coin toss
   m<-m[,2:(d+1)] #an n by d matrix of iid standard normals
   

### 2019-02-21 ext_risk_cop_age_str.R

@@ -79,7 +79,7 @@ sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
   CI0.975<-qbinom(p=0.975,size=numsims,prob=erA)/numsims
   
   if(ploton==ploton){
-  pdf(paste(resloc,BiCopName(family = fam),"_ext_riskA_vs_time.pdf",sep=""),height=5,width=5)
+  pdf(paste(resloc,"ext_riskA_vs_time.pdf",sep=""),height=5,width=5)
   op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
   plot(c(0:numsteps),erA,xlab="time",ylab="Extinction risk",cex.lab=1.5,cex=0.3,pch=16,col="black",ylim=c(0,1),panel.first = grid())
   arrows(x0=c(1:numsteps),y0=CI0.025[2:(numsteps+1)],x1=c(1:numsteps),y1=CI0.975[2:(numsteps+1)],

### 2019-02-21 mydispersal.R

@@ -195,7 +195,7 @@ get_avg_acf<-function(sims){
 
 plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,checkon,getacf){
  
-  ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
+  ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1,mn=0,sdev=1)# a righttail dep matrix(numpoints by numlocs,     
   #                                                      numpoints=numsteps*numsims)
   
   ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)

### 2019-02-22 ext_risk_cop_age_str.R
nonlinear age structure model added

@@ -9,6 +9,7 @@ library(mvtnorm)
 #       ext_thrs : extinction threshold, generally = 1
 #       cop : a bivariate copula matrix
 #       par_dist : a vector containing 4 numbers for gamma and beta distribution parameters: gshape, gscale, bshape1, bshape2
+#       omg : a number : if infinity then linear model, if finite nonlinear stage structured model
 #       numsims : number of simulations
 #       numsteps : number of time steps
 #       ploton : logical to get optional plot
@@ -18,7 +19,7 @@ library(mvtnorm)
 #       A list of three : resE, resA : each is a numsims by numsteps+1 matrix for Egg and Adult stage
 #                         extdf : a dataframe (nrow= numsteps+1, ncol=3) : extinction risk for adult and its 95% CI
 
-sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
+sim_age_str<-function(p0,ext_thrs,cop,par_dist,omg,numsims,numsteps,ploton,resloc){
   
   gshape<-par_dist[1]
   gscale<-par_dist[2]
@@ -58,12 +59,12 @@ sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
   
   #step the populations forward
   for (tct in 1:numsteps){
-   
+    
     L<-as.matrix(Ls[[tct]]) # this is the Leslie matrix (dim = numsims by numsims)
     fA<-L[,1]
     sE<-L[,2]
     
-    resE[,tct+1]<-fA * resA[,tct] 
+    resE[,tct+1]<-fA * resA[,tct] * exp(-(resA[,tct]/omg))
     resA[,tct+1]<-sE * resE[,tct] 
     
     #see if there is extinction on total Adult and Egg population
@@ -79,13 +80,13 @@ sim_age_str<-function(p0,ext_thrs,cop,par_dist,numsims,numsteps,ploton,resloc){
   CI0.975<-qbinom(p=0.975,size=numsims,prob=erA)/numsims
   
   if(ploton==ploton){
-  pdf(paste(resloc,"ext_riskA_vs_time.pdf",sep=""),height=5,width=5)
-  op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
[…]

### 2019-02-22 SuppMat.Rmd
renaming as suppmat

--- 2019-02-21 ext_risk_copula.Rmd
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewinton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ricker/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewinton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ricker/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
 

### 2019-02-22 MainText.Rmd
renaming for maintext

---
title: "extinction risk with one patch age structured model"
author: "Shyamolina Ghosh"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
source("mtime.R")
```

# Age structure model :

```{r sim_linear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")

numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf

ktau<-0.6  # kendall's tau

# for clayton
copC<-claytonCopula(3)
parC<-iTau(copC,tau=ktau)
par<-parC
fam<-3

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

C_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(C_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------

# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same

fam<-13
par<-parC

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------

# Now for gumbel

fam<-4
copG<-gumbelCopula(2)
parG<-iTau(copG,tau = ktau)
par<-parG

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))

#---------------------------------------------------------------------------------

# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same

fam<-14
par<-parG

cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------
```

```{r fig_extrisk_linear_agestr_C_SC_G_SG, echo=F,results="hide"}

res_agel_C<-readRDS("./Results/age_str_results/C_par_3_ktau_0.6_linear/C_sim_age_str.RDS")
res_agel_SC<-readRDS("./Results/age_str_results/SC_par_3_ktau_0.6_linear/SC_sim_age_str.RDS")
res_agel_G<-readRDS("./Results/age_str_results/G_par_2.5_ktau_0.6_linear/G_sim_age_str.RDS")
res_agel_SG<-readRDS("./Results/age_str_results/SG_par_2.5_ktau_0.6_linear/SG_sim_age_str.RDS")

xC<-res_agel_C$extdf
xSC<-res_agel_SC$extdf
xG<-res_agel_G$extdf
xSG<-res_agel_SG$extdf

t<-c(0:(nrow(xC)-1))

pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_G_SG_ktau_",ktau,"_linear.pdf",sep=""),height=6,width=12)
op<-par(mfrow=c(1,2),mar=c(4,4,2,2),mgp=c(2.5,0.5,0))

# for C, SC
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())

x<-xC
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
lines(t,x$erA,col="black",type="l")

x<-xSC
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")

legend("topleft", c("C","SC"), lty=c(1,1),  
       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)

# for G, SG
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())

x<-xG
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
lines(t,x$erA,col="black",type="l")

x<-xSG
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")

legend("topleft", c("G","SG"), lty=c(1,1),  
       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)

par(op)
dev.off()
```

![Extinction risk from one patch age structured model, fecundity of adults and survival of eggs have different asymmetric copula atructure but with same kendall correlation = 0.6, numsims=10000.\label{fig_extrisk_agestr}](./Results/age_str_results/extrisk_age_str_C_SC_G_SG_ktau_0.6_linear.pdf){ width=40%, height=40% }

```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")

numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf

# For extreme right tail dep.
mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
cop<-pnorm(mr) # to make normal to uniformly distributed

tempo<-paste("./Results/age_str_results/extreme_right_taildep_linear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

pdf(paste(resloc,"extreme_right_tail_cop.pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()

las_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(las_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))

# For extreme left tail dep.
ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
cop<-pnorm(ml) # to make normal to uniformly distributed

tempo<-paste("./Results/age_str_results/extreme_left_taildep_linear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

pdf(paste(resloc,"extreme_left_tail_cop.pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()

las_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(las_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
```

```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}

xL<-las_exL$extdf
xR<-las_exR$extdf

t<-c(0:(nrow(xL)-1))

pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=6,width=6)
op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))

plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())

x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
lines(t,x$erA,col="black",type="l")

x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")

legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)
par(op)
dev.off()
```

![Extinction risk from one patch linear age structured model (omg=Infinity), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep. copula structure, numsims=10000.\label{fig_extrisk_lagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf){ width=40%, height=40% }

```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")

numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-1000

# For extreme right tail dep.
mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
cop<-pnorm(mr) # to make normal to uniformly distributed

tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()

nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))

# For extreme left tail dep.
ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
cop<-pnorm(ml) # to make normal to uniformly distributed

tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")

pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()

nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
```

```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}

xL<-nlas_exL$extdf
xR<-nlas_exR$extdf

t<-c(0:(nrow(xL)-1))

pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=6,width=6)
op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))

plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())

x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
lines(t,x$erA,col="black",type="l")

x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")

legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)
par(op)
dev.off()
```

![Extinction risk from one patch non-linear age structured model (omg=1000), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep. copula structure, numsims=10000.\label{fig_extrisk_nlagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_1000.pdf){ width=40%, height=40% }

### 2019-02-22 MainText.Rmd
adding various files

@@
 ---
-title: "extinction risk with one patch age structured model"
+title: "Main text : Effect of copula on extinction risk"
 author: "Shyamolina Ghosh"
-output: pdf_document
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: 
+  pdf_document:
+    number_sections: no
+    keep_tex: yes
+    fig_caption: yes
+    includes:
+      in_header: head_maintext.sty
+
+tables: True
+link-citations: True
+urlcolor : blue
+indent : True
+
+csl: TheAmericanNaturalist.csl
+bibliography: Ref_extrisk.bib
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup_MainText, echo=F}
+knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
+options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
 seed<-101
-source("mtime.R")
 ```
-
-# Age structure model :
-
-```{r sim_linear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
-set.seed(seed)
-source("ext_risk_cop_age_str.R")
-
-numsims<-10000
-numsteps<-100
-p0<-c(5,10)
-ext_thrs<-2
-gshape<-2
-gscale<-1
-bshape1<-3
-bshape2<-2
-par_dist<-c(gshape,gscale,bshape1,bshape2)
-omg<-Inf
-
-ktau<-0.6  # kendall's tau
-
-# for clayton
-copC<-claytonCopula(3)
-parC<-iTau(copC,tau=ktau)
-par<-parC
-fam<-3
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-C_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(C_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-
-# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same
-
-fam<-13
-par<-parC
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-
-# Now for gumbel
-
-fam<-4
-copG<-gumbelCopula(2)
-parG<-iTau(copG,tau = ktau)
-par<-parG
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-
-#---------------------------------------------------------------------------------
-
-# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same
-
-fam<-14
-par<-parG
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-```
-
-```{r fig_extrisk_linear_agestr_C_SC_G_SG, echo=F,results="hide"}
-
-res_agel_C<-readRDS("./Results/age_str_results/C_par_3_ktau_0.6_linear/C_sim_age_str.RDS")
-res_agel_SC<-readRDS("./Results/age_str_results/SC_par_3_ktau_0.6_linear/SC_sim_age_str.RDS")
-res_agel_G<-readRDS("./Results/age_str_results/G_par_2.5_ktau_0.6_linear/G_sim_age_str.RDS")
-res_agel_SG<-readRDS("./Results/age_str_results/SG_par_2.5_ktau_0.6_linear/SG_sim_age_str.RDS")
-
-xC<-res_agel_C$extdf
-xSC<-res_agel_SC$extdf
-xG<-res_agel_G$extdf
-xSG<-res_agel_SG$extdf
-
-t<-c(0:(nrow(xC)-1))
-
-pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_G_SG_ktau_",ktau,"_linear.pdf",sep=""),height=6,width=12)
-op<-par(mfrow=c(1,2),mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-
-# for C, SC
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
-
-x<-xC
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
-lines(t,x$erA,col="black",type="l")
-
-x<-xSC
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
-lines(t,x$erA,col="red",type="l")
-
-legend("topleft", c("C","SC"), lty=c(1,1),  
-       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)
-
-# for G, SG
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
-
-x<-xG
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
-lines(t,x$erA,col="black",type="l")
-
-x<-xSG
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
-lines(t,x$erA,col="red",type="l")
-
-legend("topleft", c("G","SG"), lty=c(1,1),  
-       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)
-
-par(op)
-dev.off()
-```
-
-![Extinction risk from one patch age structured model, fecundity of adults and survival of eggs have different asymmetric copula atructure but with same kendall correlation = 0.6, numsims=10000.\label{fig_extrisk_agestr}](./Results/age_str_results/extrisk_age_str_C_SC_G_SG_ktau_0.6_linear.pdf){ width=40%, height=40% }
-
-```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
-set.seed(seed)
-source("ext_risk_cop_age_str.R")
-source("ExtremeTailDep.R")
-
-numsims<-10000
-numsteps<-100
-p0<-c(5,10)
-ext_thrs<-2
-gshape<-2
-gscale<-1
-bshape1<-3
-bshape2<-2
-par_dist<-c(gshape,gscale,bshape1,bshape2)
-omg<-Inf
-
-# For extreme right tail dep.
-mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
-cop<-pnorm(mr) # to make normal to uniformly distributed
-
-tempo<-paste("./Results/age_str_results/extreme_right_taildep_linear")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-pdf(paste(resloc,"extreme_right_tail_cop.pdf",sep=""),height = 5,width=5)
-plot(cop[,1],cop[,2])
-dev.off()
-
-las_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(las_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
-
-# For extreme left tail dep.
-ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
-cop<-pnorm(ml) # to make normal to uniformly distributed
-
-tempo<-paste("./Results/age_str_results/extreme_left_taildep_linear")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-pdf(paste(resloc,"extreme_left_tail_cop.pdf",sep=""),height = 5,width=5)
-plot(cop[,1],cop[,2])
-dev.off()
-
-las_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(las_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
-```
-
-```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}
-
-xL<-las_exL$extdf
-xR<-las_exR$extdf
-
-t<-c(0:(nrow(xL)-1))
-
-pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=6,width=6)
-op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
-
-x<-xL
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
-lines(t,x$erA,col="black",type="l")
-
-x<-xR
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
-lines(t,x$erA,col="red",type="l")
-
-legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
-       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)
-par(op)
-dev.off()
-```
-
-![Extinction risk from one patch linear age structured model (omg=Infinity), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep.
-copula structure, numsims=10000.\label{fig_extrisk_lagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf){ width=40%, height=40% }
-
-```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
-set.seed(seed)
-source("ext_risk_cop_age_str.R")
-source("ExtremeTailDep.R")
-
-numsims<-10000
-numsteps<-100
-p0<-c(5,10)
-ext_thrs<-2
-gshape<-2
-gscale<-1
-bshape1<-3
-bshape2<-2
-par_dist<-c(gshape,gscale,bshape1,bshape2)
-omg<-1000
-
-# For extreme right tail dep.
-mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
-cop<-pnorm(mr) # to make normal to uniformly distributed
-
-tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
-plot(cop[,1],cop[,2])
-dev.off()
-
-nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
-
-# For extreme left tail dep.
-ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
-cop<-pnorm(ml) # to make normal to uniformly distributed
-
-tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
-plot(cop[,1],cop[,2])
-dev.off()
-
-nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
-```
-
-```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}
-
-xL<-nlas_exL$extdf
-xR<-nlas_exR$extdf
-
-t<-c(0:(nrow(xL)-1))
-
-pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=6,width=6)
-op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
-
-x<-xL
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,0,0.3), border = NA)
-lines(t,x$erA,col="black",type="l")
-
-x<-xR
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
-lines(t,x$erA,col="red",type="l")
-
-legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
-       col=c('black', 'red'), horiz=F, bty='n', cex=1.2)
-par(op)
-dev.off()
-```
-
-![Extinction risk from one patch non-linear age structured model (omg=1000), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep.
-copula structure, numsims=10000.\label{fig_extrisk_nlagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_1000.pdf){ width=40%, height=40% }

### 2019-02-22 SuppMat.Rmd
adding various files

@@
 ---
-title: "ext_risk_copula"
+title: "Supporting Information : Effect of copula on extinction risk"
 author: "Shyamolina Ghosh"
-output: pdf_document
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
+
+output: 
+  pdf_document:
+    number_sections: yes
+    keep_tex: yes
+    fig_caption: yes
+    includes:
+      in_header: head_suppmat.sty
+
+tables: True
+link-citations: True
+urlcolor : blue
+indent : True
+
+csl: TheAmericanNaturalist.csl
+bibliography: Ref_extrisk.bib
 ---
+
+\pagenumbering{roman}
+\tableofcontents
+\listoffigures
+\listoftables
+
+\pagebreak
+\pagenumbering{arabic}
 
@@
 
+# Age structure model :
+
+```{r sim_linear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+set.seed(seed)
+source("ext_risk_cop_age_str.R")
+
+numsims<-10000
+numsteps<-100
+p0<-c(5,10)
+ext_thrs<-2
+gshape<-2
+gscale<-1
+bshape1<-3
+bshape2<-2
+par_dist<-c(gshape,gscale,bshape1,bshape2)
+omg<-Inf
+
+ktau<-0.6  # kendall's tau
+
+# for clayton
+copC<-claytonCopula(3)
+parC<-iTau(copC,tau=ktau)
+par<-parC
+fam<-3
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+C_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(C_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+#---------------------------------------------------------------------------------
+
+# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same
+
+fam<-13
+par<-parC
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+#---------------------------------------------------------------------------------
+
+# Now for gumbel
+
+fam<-4
+copG<-gumbelCopula(2)
+parG<-iTau(copG,tau = ktau)
+par<-parG
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+
+#---------------------------------------------------------------------------------
+
+# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same
+
+fam<-14
+par<-parG
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+#---------------------------------------------------------------------------------
+```
+
 #Ricker model :
+
 ```{r sim_ricker, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}

### 2019-02-22 makefile.Rmd
adding makefile

---
title: "makefile, compiliation notes"
author: "Daniel Reuman"
date: ""
output: pdf_document
---

<!--checkpoint package-->
```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
library(checkpoint)
if (!dir.exists("./.checkpoint/")){
  dir.create("./.checkpoint/")
}
checkpoint("2019-01-13",checkpointLocation = "./")
```

# Get tex files and pdf files for the supp mat, but the pdf files 
#are missing the external references

```{r tex_for_supp_mat}
library(rmarkdown)
rmarkdown::render(input="SuppMat.Rmd") 
```

# Get tex files and pdf files for the main text, but the pdf files 
#are missing the external references

```{r tex_for_main_text}
rmarkdown::render(input="MainText.Rmd") 
```

# Now recompile the latex

```{r recompile_latex}
#do it once
system2(command="pdflatex",args="SuppMat.tex",stdout=TRUE,stderr=TRUE)
system2(command="pdflatex",args="MainText.tex",stdout=TRUE,stderr=TRUE)

#do it again to make sure to get the external references right
system2(command="pdflatex",args="SuppMat.tex",stdout=TRUE,stderr=TRUE)
system2(command="pdflatex",args="MainText.tex",stdout=TRUE,stderr=TRUE)
```

### 2019-02-23 MainText.Rmd
checkpoint issue solved

@@
 ```{r setup_MainText, echo=F}
+library(rmarkdown)
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")

### 2019-02-25 mydispersal.R
verhulst scl=3 changed from 2.5 to ensure 20-80% extinction risk range

@@ -359,6 +359,8 @@ varying_d<-function(numsims,numsteps,numlocs,p0,params,ext_thrs=ext_thrs,scl,mod
   }
   
   if(plotteron==T){
+    
+    if(getacf==T){
     op<-par(mfrow=c(2,1), mar=c(5.2,4.2,1,1.2))
     plot(d_seq,risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
          xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
@@ -367,14 +369,24 @@ varying_d<-function(numsims,numsteps,numlocs,p0,params,ext_thrs=ext_thrs,scl,mod
            col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
     mtext(paste0("r = ", r))
     
-    plot(d_seq,avg_acf_left,xlim=c(0,1),ylim=c(-0.1,0.1),type="b",col="red",panel.first = grid(),pch=16,
+    plot(d_seq,pop_avg_acf_left,xlim=c(0,1),ylim=c(-0.1,0.1),type="b",col="red",panel.first = grid(),pch=16,
          xlab="d",ylab="avg. ACF",cex.lab=1.5,cex.axis=1.5)
-    lines(d_seq,avg_acf_right,type="b",col="blue",pch=16)
+    lines(d_seq,pop_avg_acf_right,type="b",col="blue",pch=16)
     legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(16,16), 
            col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
     #mtext(paste0("r = ", r))
     
     par(op)
+    }else{
+      op<-par(mar=c(5.2,4.2,1,1.2))
+      plot(d_seq,risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),
+           xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+      lines(d_seq,risk_right,type="b",col="blue")
+      legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+             col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+      mtext(paste0("r = ", r))
+      par(op)
+    }
   }
   
   return(data.frame(d_seq=d_seq,

### 2019-02-25 SuppMat.Rmd
verhulst scl=3 changed from 2.5 to ensure 20-80% extinction risk range
minor plot edit

@@
 model<-"verhulst"
-scl<-2.5
+scl<-3
 
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=2.5) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=2.5); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_2.5/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=3) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_3/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_3/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=3); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_3/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
 

### 2019-02-25 mydispersal.R
bug fixed in verhulst model

@@ -78,7 +78,7 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model,checkon){
       res[,,tct+1]<-lam_ricker*lam_sto #numsims by numlocs matrix
     }else if(model=="verhulst"){
       r<-params[1]
-      lam_verhulst<-1+(r*(1-res[,,tct]))*res[,,tct]
+      lam_verhulst<-(1+(r*(1-res[,,tct])))*res[,,tct]
       res[,,tct+1]<-lam_verhulst*lam_sto #numsims by numlocs matrix
     }else if(model=="hassell"){
       r<-params[1]

### 2019-02-25 SuppMat.Rmd
age str res: linear, nonlinear with extreme tail added

@@
 omg<-Inf
-
 ktau<-0.6  # kendall's tau
@@
 
+```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+source("ext_risk_cop_age_str.R")
+source("ExtremeTailDep.R")
+numsims<-10000
+numsteps<-100
+p0<-c(5,10)
+ext_thrs<-2
+gshape<-2
+gscale<-1
+bshape1<-3
+bshape2<-2
+par_dist<-c(gshape,gscale,bshape1,bshape2)
+omg<-Inf
+
+# For extreme right tail dep.
+mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
+cop<-pnorm(mr) # to make normal to uniformly distributed
+ tempo<-paste("./Results/age_str_results/extreme_right_taildep_linear")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+pdf(paste(resloc,"extreme_right_tail_cop.pdf",sep=""),height = 5,width=5)
+plot(cop[,1],cop[,2])
+dev.off()
+las_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(las_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
+
+# For extreme left tail dep.
+ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
+cop<-pnorm(ml) # to make normal to uniformly distributed
+ tempo<-paste("./Results/age_str_results/extreme_left_taildep_linear")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+pdf(paste(resloc,"extreme_left_tail_cop.pdf",sep=""),height = 5,width=5)
+plot(cop[,1],cop[,2])
+dev.off()
+ las_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(las_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
+```
+
+```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}
+xL<-las_exL$extdf
+xR<-las_exR$extdf
+t<-c(0:(nrow(xL)-1))
+pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=6,width=6)
+op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
+x<-xL
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
+lines(t,x$erA,col="red",type="l")
+x<-xR
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
+lines(t,x$erA,col="blue",type="l")
+legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.2)
+par(op)
+dev.off()
+```
+
+![Extinction risk from one patch linear age structured model (omg=Infinity), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep.
+copula structure, numsims=10000.\label{fig_extrisk_lagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf){ width=40%, height=40% }
+
+```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+source("ext_risk_cop_age_str.R")
+source("ExtremeTailDep.R")
+numsims<-10000
+numsteps<-100
+p0<-c(5,10)
+ext_thrs<-2
+gshape<-2
+gscale<-1
+bshape1<-3
+bshape2<-2
+par_dist<-c(gshape,gscale,bshape1,bshape2)
+omg<-100
+# For extreme right tail dep.
+mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
+cop<-pnorm(mr) # to make normal to uniformly distributed
+ tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
+plot(cop[,1],cop[,2])
+dev.off()
+nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
+ # For extreme left tail dep.
+ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
+cop<-pnorm(ml) # to make normal to uniformly distributed
+ tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
+plot(cop[,1],cop[,2])
+dev.off()
+nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
+```
+
+```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}
+xL<-nlas_exL$extdf
+xR<-nlas_exR$extdf
+t<-c(0:(nrow(xL)-1))
+pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=6,width=6)
+op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
+x<-xL
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
+lines(t,x$erA,col="red",type="l")
+x<-xR
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
+lines(t,x$erA,col="blue",type="l")
+legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.2)
+par(op)
+dev.off()
+```
+
+![Extinction risk from one patch non-linear age structured model (omg=100), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep.
+copula structure, numsims=10000.\label{fig_extrisk_nlagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf){ width=40%, height=40% }
+
 #Ricker model :
 
-```{r sim_ricker, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-tempo<-paste("./Results/",model,sep="")
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
 if (!dir.exists(tempo)){
@@
 
-```{r fig_ricker, echo=F, results="hide",cache=T, cache.extra=list(seed,res_ricker_local,res_ricker_global)}
+```{r fig_ricker_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_ricker_local,res_ricker_global)}
 pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
@@
 
-```{r fig_ricker_extrisk_d0, echo=F,results="hide"}
-set.seed(seed)
-
-ricker_extrisk_d0<-readRDS("./Results/ricker/ricker_extrisk_d0.RDS")
+```{r fig_ricker_extrisk_d0_scl_1, echo=F,results="hide"}
+set.seed(seed)
+
+ricker_extrisk_d0<-readRDS(paste(resloc,"ricker_extrisk_d0.RDS",sep=""))
 numsims<-10000
@@
 
-pdf("./Results/ricker/ricker_extrisk_d0.pdf",height = 6,width=6)
+pdf(paste(resloc,"ricker_extrisk_d0.pdf",sep=""),height=6,width=6)
 
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ricker/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ricker/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_d0}](./Results/ricker/ricker_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp}](./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker__risk_global_disp}](./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_d0}](./Results/ricker_scl_1/ricker_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-```{r sim_hassell, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 model<-"hassell"
-
-tempo<-paste("./Results/",model,sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
 
@@
 scl<-1
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
 rc_hassell<-((b/(b-1)))^b # transition from under to overcompensatory dynamics
@@
 
-```{r fig_hassell, echo=F, results="hide",cache=T, cache.extra=list(seed,res_hassell_local,res_hassell_global)}
-pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+```{r fig_hassell_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_hassell_local,res_hassell_global)}
+pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-```{r fig_hassell_extrisk_d0, echo=F,results="hide"}
-set.seed(seed)
-numsims<-10000
-hassell_extrisk_d0<-readRDS("./Results/hassell/hassell_extrisk_d0.RDS")
+```{r fig_hassell_extrisk_d0_scl_1, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+hassell_extrisk_d0<-readRDS(paste(resloc,"hassell_extrisk_d0.RDS",sep=""))
 
@@
 
-pdf("./Results/hassell/hassell_extrisk_d0.pdf",height = 6,width=6)
+pdf(paste(resloc,"hassell_extrisk_d0.pdf",sep=""),height=6,width=6)
 plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/hassell/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/hassell/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0}](./Results/hassell/hassell_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0}](./Results/hassell_scl_1/hassell_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-```{r sim_msmith, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r fig_msmith, echo=F, results="hide",cache=T, cache.extra=list(seed,res_msmith_local,res_msmith_global)}
-pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+```{r fig_msmith_scl_0.8, echo=F, results="hide",cache=T, cache.extra=list(seed,res_msmith_local,res_msmith_global)}
+pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-```{r fig_msmith_extrisk_d0, echo=F,results="hide"}
+```{r fig_msmith_extrisk_d0_scl_0.8, echo=F,results="hide"}
 set.seed(seed)
@@
 ```{r get_rc_pennycuick,echo=F,results="hide",cache=T,cache.extra=list(seed,mtime("1var_pop_model_bifurcation.R"))}
+set.seed(seed)
 source("./1var_pop_model_bifurcation.R")
@@
 
-```{r sim_pennycuick, echo=F, cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 model<-"pennycuick"
-
-tempo<-paste("./Results/",model,sep="")
+scl<-1
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
 if (!dir.exists(tempo)){
@@
 b<-b_pennycuick
-scl<-1
 
@@
 vxx<-var(xx)
-
 if(vxx>1e-10){
@@
 
-```{r fig_pennycuick, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+```{r fig_pennycuick_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-```{r fig_pennycuick_extrisk_d0, echo=F,results="hide"}
-set.seed(seed)
-numsims<-10000
-pennycuick_extrisk_d0<-readRDS("./Results/pennycuick/pennycuick_extrisk_d0.RDS")
+```{r fig_pennycuick_extrisk_d0_scl_1, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+pennycuick_extrisk_d0<-readRDS(paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
 
@@
 
-pdf("./Results/pennycuick/pennycuick_extrisk_d0.pdf",height = 6,width=6)
+pdf(paste(resloc,"pennycuick_extrisk_d0.pdf",sep=""),height=6,width=6)
 plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000\label{fig_pennycuick_risk_d0}](./Results/pennycuick/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
+
+```{r sim_pennycuick_scl_0.6, echo=F, cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+source("mydispersal.R")
+
+model<-"pennycuick"
+numsims<-10000
+numsteps<-25
+numlocs<-5
+a<-a_pennycuick
+b<-b_pennycuick
+scl<-0.6
+
+tempo<-paste("./Results/",model,"_scl_",scl,sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+
+rng1<-runif(10000,min=1,max=r_bf_pennycuick) #r must be > 1 for +ve eqm pt.
+qr1<-quantile(rng1,probs=c(.025,.975))
+q1<-abs(unname(qr1[1])-rc_pennycuick)
+q2<-abs(unname(qr1[2])-rc_pennycuick)
+qs<-c(q1,q2)
+iqmin<-which.min(qs)
+iqmax<-which.max(qs)
+
+fromr<-range(qr1)[iqmin]
+tor<-range(qr1)[iqmax]
+
+if(fromr>tor){
+  byr<--qs[iqmin]/4
+}else{
+  byr<-qs[iqmin]/4
+}
+
+rlist<-sort(seq(from=fromr,to=tor, by = byr))
+
+xx<-diff(rlist)
+vxx<-var(xx)
+
+if(vxx>1e-10){
+  warning("Not equal spacing in rlist : Pennycuick model !!!",immediate.=T,call.=T)
+}
+
+pennycuick_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) # a dataframe to store ext risk at dispersal d=0 for all r
+colnames(pennycuick_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+
+#for local dispersal in D matrix 
+res_pennycuick_local<-c()
+
+for(i in c(1:length(rlist))){
+    r<- rlist[i]
+    params<-c(r,a,b)
+    K_e<-(b/a)*(a+log(r-1))
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_pennycuick_local<-abind(res_pennycuick_local,temp_mat,along = 3)
+
+    pennycuick_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
+    pennycuick_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+}
+
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_pennycuick_local)<-list(NULL,cnm,anm)
+
+saveRDS(pennycuick_extrisk_d0,paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
+saveRDS(res_pennycuick_local,paste(resloc,"res_pennycuick_local.RDS",sep=""))
+
+# For global dispersal in D matrix 
+res_pennycuick_global<-c()
+for(r in rlist){
+    #cat("r=",r,"\n")
+    params<-c(r,a,b)
+    K_e<-(b/a)*(a+log(r-1))
+    ext_thrs<-K_e/10
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
+
+    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    res_pennycuick_global<-abind(res_pennycuick_global,temp_mat,along = 3)
+}
+anm<-paste("r_",rlist,sep="")
+cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+dimnames(res_pennycuick_global)<-list(NULL,cnm,anm)
+saveRDS(res_pennycuick_global,paste(resloc,"res_pennycuick_global.RDS",sep=""))
+```
+
+```{r fig_pennycuick_scl_0.6, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
+
+for(i in 1:dim(res_pennycuick_local)[3]){
+  s<-res_pennycuick_local[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)))
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
+op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
+
+for(i in 1:dim(res_pennycuick_global)[3]){
+  s<-res_pennycuick_global[,,i]
+  s<-as.data.frame(s)
+
+  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d,s$risk_right,type="b",col="blue")
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)))
+}
+par(op)
+dev.off()
+```
+
+```{r fig_pennycuick_extrisk_d0_scl_0.6, echo=F,results="hide"}
+set.seed(seed)
+numsims<-10000
+pennycuick_extrisk_d0<-readRDS(paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
+
+pennycuick_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
+pennycuick_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
+
+pennycuick_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
+pennycuick_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
+
+pdf(paste(resloc,"pennycuick_extrisk_d0.pdf",sep=""),height=6,width=6)
+plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_pennycuick,lty="dotted")
+
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025right,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
+legend(3,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick_scl_0.6/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick_scl_0.6/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=0.6\label{fig_pennycuick_risk_d0}](./Results/pennycuick_scl_0.6/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-```{r sim_verhulst, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_verhulst_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 model<-"verhulst"
-scl<-3
+scl<-1
 
@@
 
-```{r fig_verhulst, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
+```{r fig_verhulst_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
 pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
@@
   lines(s$d,s$risk_right,type="b",col="blue")
-  legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
          col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
@@
 
-```{r fig_verhulst_extrisk_d0, echo=F,results="hide"}
+```{r fig_verhulst_extrisk_d0_scl_1, echo=F,results="hide"}
 set.seed(seed)
@@
 mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(0.5,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-dev.off()
-```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=3) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_3/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=3) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_2.5}](./Results/verhulst_scl_3/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=3); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_2.5}](./Results/verhulst_scl_3/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
+legend(0.4,0.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_1}](./Results/verhulst_scl_1/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_1}](./Results/verhulst_scl_1/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_1}](./Results/verhulst_scl_1/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-```{r sim_abrewer, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_abrewer_scl_0.8, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r fig_abrewer, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
+```{r fig_abrewer_scl_0.8, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
 pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
@@
 
-```{r fig_abrewer_extrisk_d0, echo=F,results="hide"}
+```{r fig_abrewer_extrisk_d0_scl_0.8, echo=F,results="hide"}
 set.seed(seed)

### 2019-02-25 SuppMat.Rmd
bug fixed : pennycuick : min limit of r to get +ve eqm. pt.

@@
 
-rng1<-runif(10000,min=1,max=r_bf_pennycuick) #r must be > 1 for +ve eqm pt.
+rng1<-runif(10000,min=(1+exp(-a)),max=r_bf_pennycuick) # for +ve eqm pt.
 qr1<-quantile(rng1,probs=c(.025,.975))
@@
 }
-
 anm<-paste("r_",rlist,sep="")
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
-
-```{r sim_pennycuick_scl_0.6, echo=F, cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
-set.seed(seed)
-source("mydispersal.R")
-
-model<-"pennycuick"
-numsims<-10000
-numsteps<-25
-numlocs<-5
-a<-a_pennycuick
-b<-b_pennycuick
-scl<-0.6
-
-tempo<-paste("./Results/",model,"_scl_",scl,sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-
-rng1<-runif(10000,min=1,max=r_bf_pennycuick) #r must be > 1 for +ve eqm pt.
-qr1<-quantile(rng1,probs=c(.025,.975))
-q1<-abs(unname(qr1[1])-rc_pennycuick)
-q2<-abs(unname(qr1[2])-rc_pennycuick)
-qs<-c(q1,q2)
-iqmin<-which.min(qs)
-iqmax<-which.max(qs)
-
-fromr<-range(qr1)[iqmin]
-tor<-range(qr1)[iqmax]
-
-if(fromr>tor){
-  byr<--qs[iqmin]/4
-}else{
-  byr<-qs[iqmin]/4
-}
-
-rlist<-sort(seq(from=fromr,to=tor, by = byr))
-
-xx<-diff(rlist)
-vxx<-var(xx)
-
-if(vxx>1e-10){
-  warning("Not equal spacing in rlist : Pennycuick model !!!",immediate.=T,call.=T)
-}
-
-pennycuick_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) # a dataframe to store ext risk at dispersal d=0 for all r
-colnames(pennycuick_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
-
-#for local dispersal in D matrix 
-res_pennycuick_local<-c()
-
-for(i in c(1:length(rlist))){
-    r<- rlist[i]
-    params<-c(r,a,b)
-    K_e<-(b/a)*(a+log(r-1))
-    ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
-    res_pennycuick_local<-abind(res_pennycuick_local,temp_mat,along = 3)
-
-    pennycuick_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    pennycuick_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
-}
-
-anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
-dimnames(res_pennycuick_local)<-list(NULL,cnm,anm)
-
-saveRDS(pennycuick_extrisk_d0,paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
-saveRDS(res_pennycuick_local,paste(resloc,"res_pennycuick_local.RDS",sep=""))
-
-# For global dispersal in D matrix 
-res_pennycuick_global<-c()
-for(r in rlist){
-    #cat("r=",r,"\n")
-    params<-c(r,a,b)
-    K_e<-(b/a)*(a+log(r-1))
-    ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
-    res_pennycuick_global<-abind(res_pennycuick_global,temp_mat,along = 3)
-}
-anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
-dimnames(res_pennycuick_global)<-list(NULL,cnm,anm)
-saveRDS(res_pennycuick_global,paste(resloc,"res_pennycuick_global.RDS",sep=""))
-```
-
-```{r fig_pennycuick_scl_0.6, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
-
-for(i in 1:dim(res_pennycuick_local)[3]){
-  s<-res_pennycuick_local[,,i]
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", round(rlist[i],5)))
-}
-par(op)
-dev.off()
-
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
-
-for(i in 1:dim(res_pennycuick_global)[3]){
-  s<-res_pennycuick_global[,,i]
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", round(rlist[i],5)))
-}
-par(op)
-dev.off()
-```
-
-```{r fig_pennycuick_extrisk_d0_scl_0.6, echo=F,results="hide"}
-set.seed(seed)
-numsims<-10000
-pennycuick_extrisk_d0<-readRDS(paste(resloc,"pennycuick_extrisk_d0.RDS",sep=""))
-
-pennycuick_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
-pennycuick_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
-
-pennycuick_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
-pennycuick_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"pennycuick_extrisk_d0.pdf",sep=""),height=6,width=6)
-plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
-arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,
-       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_pennycuick,lty="dotted")
-
-lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025right,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(3,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-
-dev.off()
-```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick_scl_0.6/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick_scl_0.6/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=0.6\label{fig_pennycuick_risk_d0}](./Results/pennycuick_scl_0.6/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
 

### 2019-02-25 SuppMat.Rmd
modified

@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp}](./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker__risk_global_disp}](./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_d0}](./Results/ricker_scl_1/ricker_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_global_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_d0_scl_1}](./Results/ricker_scl_1/ricker_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0}](./Results/hassell_scl_1/hassell_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0_scl_1}](./Results/hassell_scl_1/hassell_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 ```{r fig_pennycuick_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=9,width=12)
-op<-par(mfrow=c(3,4),mar=c(5.2,4.2,1.2,1.5))
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
 
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0_scl_1}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
 

### 2019-02-25 SuppMat.Rmd
plotting d upto 0.5

@@
 for(i in 1:dim(res_ricker_local)[3]){
-  s<-res_ricker_local[,,i]
+  #s<-res_ricker_local[,,i]
+  s<-res_ricker_local[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_ricker_global)[3]){
-  s<-res_ricker_global[,,i]
+  #s<-res_ricker_global
+  s<-res_ricker_global[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_hassell_local)[3]){
-  s<-res_hassell_local[,,i]
+  s<-res_hassell_local[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_hassell_global)[3]){
-  s<-res_hassell_global[,,i]
+  s<-res_hassell_global[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_msmith_local)[3]){
-  s<-res_msmith_local[,,i]
+  s<-res_msmith_local[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_msmith_global)[3]){
-  s<-res_msmith_global[,,i]
+  s<-res_msmith_global[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_pennycuick_local)[3]){
-  s<-res_pennycuick_local[,,i]
+  s<-res_pennycuick_local[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_pennycuick_global)[3]){
-  s<-res_pennycuick_global[,,i]
+  s<-res_pennycuick_global[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_verhulst_local)[3]){
-  s<-res_verhulst_local[,,i]
+  s<-res_verhulst_local[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_verhulst_global)[3]){
-  s<-res_verhulst_global[,,i]
+  s<-res_verhulst_global[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_abrewer_local)[3]){
-  s<-res_abrewer_local[,,i]
+  s<-res_abrewer_local[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
@@
 for(i in 1:dim(res_abrewer_global)[3]){
-  s<-res_abrewer_global[,,i]
+  s<-res_abrewer_global[,,i][1:6,] # plot for d in range (0 to 0.5)
   s<-as.data.frame(s)
 
-  plot(s$d,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",
+  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),

### 2019-02-25 mydispersal.R

@@ -251,7 +251,7 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
     
     pop_sim1<-pops1[numsims,,]
     ylm<-max(c(abs(min(pop_sim1)),abs(max(pop_sim1))))
-    plot(0,0,xlim = c(1,numsteps+1),ylim = c(-ylm,ylm),type = "n",xlab="time",ylab="pops_with_noise_right_tail")
+    plot(0,0,xlim = c(1,numsteps+1),ylim = c(0,ylm),type = "n",xlab="time",ylab="pops_with_noise_right_tail")
     cl <- rainbow(numlocs)
     for(i in 1:numlocs){
       lines(c(1:c(numsteps+1)),pop_sim1[i,],col=cl[i],type="l")
@@ -259,7 +259,7 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
     
     pop_sim2<-pops2[numsims,,]
     ylm<-max(c(abs(min(pop_sim2)),abs(max(pop_sim2))))
-    plot(0,0,xlim = c(1,numsteps+1),ylim = c(-ylm,ylm),type = "n",xlab="time",ylab="pops_with_noise_left_tail")
+    plot(0,0,xlim = c(1,numsteps+1),ylim = c(0,ylm),type = "n",xlab="time",ylab="pops_with_noise_left_tail")
     cl <- rainbow(numlocs)
     for(i in 1:numlocs){
       lines(c(1:c(numsteps+1)),pop_sim2[i,],col=cl[i],type="l")

### 2019-02-26 SuppMat.Rmd
verhulst and abrewer with scl=0.6 added

@@
 
-```{r sim_verhulst_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 model<-"verhulst"
-scl<-1
+scl<-0.6
 
@@
 
-```{r fig_verhulst_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
+```{r fig_verhulst_scl_0.6, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
 pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
@@
 
-```{r fig_verhulst_extrisk_d0_scl_1, echo=F,results="hide"}
+```{r fig_verhulst_extrisk_d0_scl_0.6, echo=F,results="hide"}
 set.seed(seed)
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_1}](./Results/verhulst_scl_1/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_1}](./Results/verhulst_scl_1/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_1}](./Results/verhulst_scl_1/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}](./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}](./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=0.6); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_0.6}](./Results/verhulst_scl_0.6/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
 
@@
 
-```{r sim_abrewer_scl_0.8, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 model<-"austinbrewer"
-scl<-0.8
+scl<-0.6
 
@@
 
-```{r fig_abrewer_scl_0.8, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
+```{r fig_abrewer_scl_0.6, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
 pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
@@
 
-```{r fig_abrewer_extrisk_d0_scl_0.8, echo=F,results="hide"}
+```{r fig_abrewer_extrisk_d0_scl_0.6, echo=F,results="hide"}
 set.seed(seed)
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_abrewer_r_vary_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.8}](./Results/austinbrewer_scl_0.8/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.6); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf){ width=40%, height=40% }

### 2019-02-26 SuppMat.Rmd
age structure results plot upto 25 years

@@
 ```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}
-xL<-las_exL$extdf
-xR<-las_exR$extdf
+xL<-las_exL$extdf[1:26,]
+xR<-las_exR$extdf[1:26,]
 t<-c(0:(nrow(xL)-1))
@@
 ```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}
-xL<-nlas_exL$extdf
-xR<-nlas_exR$extdf
+xL<-nlas_exL$extdf[1:26,]
+xR<-nlas_exR$extdf[1:26,]
 t<-c(0:(nrow(xL)-1))

### 2019-02-27 MainText.Rmd
adding figures...continued

@@
 
+\listoffigures
+\listoftables
+
 ```{r setup_MainText, echo=F}
@@
 ```
+
+ <!--Generating maintext figures-->
+
+ <!--FIG : age structure model results : linear, nonlinear-->
+\begin{figure}[!h]
+\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
+\vspace{-0.9 cm}
+\begin{center}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
+\caption{Extinction risk from single patch age structured model $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$), fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
+\label{fig_extrisk_agestr}}
+\end{center}
+\end{figure}

### 2019-02-27 SuppMat.Rmd
adding figures...continued

--- 2019-02-26 SuppMat.Rmd
@@
 ```
-
-# Age structure model :
 
@@
 op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5,panel.first=grid())
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5)
 x<-xL
@@
 ```
-
-![Extinction risk from one patch linear age structured model (omg=Infinity), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep.
-copula structure, numsims=10000.\label{fig_extrisk_lagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf){ width=40%, height=40% }
 
@@
 
-![Extinction risk from one patch non-linear age structured model (omg=100), fecundity of adults and survival of eggs have either extreme left or extreme right tail dep.
-copula structure, numsims=10000.\label{fig_extrisk_nlagestr_extremetail}](./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf){ width=40%, height=40% }
-
-#Ricker model :
-
 ```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 ```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_global_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_d0_scl_1}](./Results/ricker_scl_1/ricker_extrisk_d0.pdf){ width=40%, height=40% }
-
-#Hassell Model :
 
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0_scl_1}](./Results/hassell_scl_1/hassell_extrisk_d0.pdf){ width=40%, height=40% }
-
-#Maynard Smith Model :
-
 ```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0_scl_0.8}](./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf){ width=40%, height=40% }
-
-# Pennycuick model :
-
 ```{r get_rc_pennycuick,echo=F,results="hide",cache=T,cache.extra=list(seed,mtime("1var_pop_model_bifurcation.R"))}
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0_scl_1}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
-
-# Verhulst model :
-
 ```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 ```
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}](./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : rc = 1, scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}](./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model with no dispersion : rc = 1, scl=0.6); numlocs=5, numsteps=25, numsims=10000\label{fig_verhulst_risk_d0_scl_0.6}](./Results/verhulst_scl_0.6/verhulst_extrisk_d0.pdf){ width=40%, height=40% }
-
-# Austin Brewer model :
 
@@
 
+<!--Ricker model-->
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_global_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_d0_scl_1}](./Results/ricker_scl_1/ricker_extrisk_d0.pdf){ width=40%, height=40% }
+
+<!--Hassell model-->
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0_scl_1}](./Results/hassell_scl_1/hassell_extrisk_d0.pdf){ width=40%, height=40% }
+
+<!--Maynard smith model-->
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0_scl_0.8}](./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf){ width=40%, height=40% }
+
+<!--Pennycuick model-->
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0_scl_1}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
+
+<!--Verhulst model-->
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}](./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}](./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0_scl_1}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
+
+<!--Austin brewer model-->
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
@@
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.6); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
+
+```{r create_maintext_fig3,echo=F,results="hide"}
+
+getplot<-function(x,rlist){
+lasta<-dim(x)[3]
+Af<-as.data.frame(x[,,1][1:6,])
+Al<-as.data.frame(x[,,lasta][1:6,])
+
+plot(Af$d,Af$risk_left,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+     xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+lines(Af$d,Af$risk_right,type="b",col="blue")
+
+mtext(paste0("r = ", round(rlist[1],5)))
+
+plot(Al$d,Al$risk_left,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+     xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+lines(Al$d,Al$risk_right,type="b",col="blue")
+mtext(paste0("r = ", round(rlist[lasta],5)))
+
+}
+
+pdf("./Results/allmodel_ext_risk_local_disp.pdf",height=12,width=9)
+
+op<-par(mfrow=c(6,3),mar=c(5.2,4.2,1.2,1.5))
+
+x<-res_ricker_local
+rlist<-ricker_extrisk_d0$r
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+legend(x=0.6,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
+mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(a = 0.5, b = 100, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_hassell_local
+rlist<-hassell_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(a = 1, b = 3, sd = 0.8)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_msmith_local
+rlist<-msmith_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(a = 0.1, b = 0.1, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_pennycuick_local
+rlist<-pennycuick_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_verhulst_local
+rlist<-verhulst_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_abrewer_local
+rlist<-abrewer_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+par(op)
+dev.off()
+```

### 2019-02-27 MainText.Rmd
adding figures...continued

@@
 \end{figure}
+
+<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_info} extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$ for local dispersion.
+Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime in deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
+\end{center}
+\end{figure}
+
+ <!--FIG : all nonlinear model results with d=0, extrisk vs. r-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0.pdf}
+\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0.pdf}\\
+\vspace{-0.6 cm}
+\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf}
+\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf}\\
+\vspace{-0.6 cm}
+\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0.pdf}
+\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf}\\
+\caption{Extinction risk varies with bifurcation parameter $(r)$ for all nonlinear density-dependence models without dispersal, numsims=10000. \label{fig_extrisk_d0}}
+\end{center}
+\end{figure}

### 2019-02-27 SuppMat.Rmd
adding figures...continued

@@
 ```{r fig_ricker_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_ricker_local,res_ricker_global)}
+
+pdf("./Results/common_legend.pdf",height=1,width=12)
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
+plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+legend("center", c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'),x.intersp = 0.2,
+         cex = 2.5, xpd = TRUE, horiz = T, inset = c(0,0), 
+         bty = "n")
+par(op)
+dev.off()
+
 pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
 #op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
        #panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
   lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", round(rlist[i],5)))
-
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
   #---------------------------------------------------
@@
 pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
        #panel.first = grid(),
-       xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
   lines(s$d,s$risk_right,type="b",col="blue")
-  legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-         col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
-  mtext(paste0("r = ", round(rlist[i],5)))
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
 }
@@
 
-pdf(paste(resloc,"ricker_extrisk_d0.pdf",sep=""),height=6,width=6)
-
+pdf(paste(resloc,"ricker_extrisk_d0.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
 plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",
-     ylim=c(0,1),xlim = range(ricker_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     ylim=c(0,1),xlim = range(ricker_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
 arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,
@@
 arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025right,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(0.2,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-
+legend("bottomright", c("Ricker",paste0("K = ",K),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+#legend("bottomright", c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=F, bty='n',cex=1.5)
+par(op)
 dev.off()
@@
 pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
-op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
-op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 
-pdf(paste(resloc,"hassell_extrisk_d0.pdf",sep=""),height=6,width=6)
+pdf(paste(resloc,"hassell_extrisk_d0.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
 plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
-     xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
 arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,
@@
 arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025right,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(2,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-
+legend("bottomright", c("Hassell",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
 dev.off()
@@
 pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
-op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
-op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5))
 
@@
 
-pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=6,width=6)
+pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
 plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(msmith_extrisk_d0$r),
-     xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
 arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,
@@
 arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(1.2,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+legend("bottomright", c("Maynard Smith",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
 dev.off()
@@
 pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
-op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(4,4),mar=c(5,4,1.2,1.5),mgp=c(2,0.5,0))
 
@@
 pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
-op<-par(mfrow=c(4,4),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 
-pdf(paste(resloc,"pennycuick_extrisk_d0.pdf",sep=""),height=6,width=6)
+pdf(paste(resloc,"pennycuick_extrisk_d0.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
 plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
 arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,
@@
 arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025right,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(3,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-
+legend("bottomright", c("Pennycuick",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
 dev.off()
@@
 pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 
-pdf(paste(resloc,"verhulst_extrisk_d0.pdf",sep=""),height=6,width=6)
+pdf(paste(resloc,"verhulst_extrisk_d0.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
 plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
 arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025left,x1=verhulst_extrisk_d0$r,
@@
        x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(0.4,0.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
+legend("bottomright", c("Verhulst",paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
 dev.off()
@@
 pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
-op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,1.5))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
 
@@
 abrewer_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
-pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=6,width=6)
+pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
 plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
-     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=1.5,pch=1)
+     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
 arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025left,x1=abrewer_extrisk_d0$r,
@@
        x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-mtext(paste0("d = 0 , numsims = ",numsims),side=3,line=0.2)
-legend(0.005,1.08, c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
-dev.off()
-```
-
-<!--Ricker model-->
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years ; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_global_disp_scl_1}](./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (ricker model with no dispersion : rc=1); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_d0_scl_1}](./Results/ricker_scl_1/ricker_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--Hassell model-->
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell_scl_1}](./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (hassell model with no dispersion : rc=2.73199, scl=1); numlocs=5, numsteps=25, numsims=10000\label{fig_hassell_risk_d0_scl_1}](./Results/hassell_scl_1/hassell_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--Maynard smith model-->
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 4/3, scl=0.8) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary_scl_0.8}](./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model with no dispersion : rc=4/3, scl=0.8); numlocs=5, numsteps=25, numsims=10000\label{fig_msmith_risk_d0_scl_0.8}](./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--Pennycuick model-->
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model : a = 0.1, b = 0.1, scl=1) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}](./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0_scl_1}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--Verhulst model-->
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}](./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Verhulst model : scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}](./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Pennycuick model with no dispersion : rc=4.3231657); numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_pennycuick_risk_d0_scl_1}](./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf){ width=40%, height=40% }
-
-<!--Austin brewer model-->
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.6) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model : K = 50, s = 0.2, rc = 0.02000091, scl=0.6) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Austin Brewer model with no dispersion : rc=0.02000091, scl=0.6); numlocs=5, numsteps=25, numsims=10000\label{fig_abrewer_risk_d0_scl_0.6}](./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf){ width=40%, height=40% }
-
-```{r create_maintext_fig3,echo=F,results="hide"}
+legend("bottomright", c("Austin-Brewer",paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+```
+
+<!--Ricker model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Ricker model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_global_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Hassell model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Hassell model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_hassell_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Hassell model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Hassell model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_hassell_risk_global_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Maynard smith model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Maynard Smith model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.8.
+\label{fig_msmith_risk_local_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Maynard smith model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Maynard Smith model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.8.
+\label{fig_msmith_risk_global_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Pennycuick model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Pennycuick model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1.
+\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Pennycuick model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Pennycuick model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1.
+\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Verhulst model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Verhulst model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
+\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Verhulst model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Verhulst model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
+\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Austin brewer model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Austin-Brewer model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
+\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Austin brewer model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
+\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Austin-Brewer model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
+\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+```{r create_maintext_fig_localdisp_summary,echo=F,results="hide"}
 
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-legend(x=0.6,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
        col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
@@
 mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 1, b = 3, sd = 0.8)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(paste("(a = 0.5, b = 4, sd = 0.8)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
 x<-res_msmith_local

### 2019-02-27 MainText.Rmd
several changes, Table added

@@
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_info} extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$ for local dispersion.
-Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime in deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary}, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$ for local dispersal.
+Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
 \end{center}
@@
 \end{figure}
+
+```{r read_tab_model_summary,echo=F, results='asis',message=F}
+library(kableExtra)
+library(dplyr)
+dt<-readRDS("./Results/tab_model_summary.RDS")
+knitr::kable(dt, escape=F,
+             format="latex",digits=4, align="l",
+             caption = "Summary of models.\\label{tab_model_summary}",
+             booktabs=T, linesep = "\\addlinespace"
+             )%>%column_spec(c(1,3,5,6),width="1.4cm")%>%kable_styling(latex_options = "basic",position = "center")
+             #column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
+```

### 2019-02-27 SuppMat.Rmd
several changes, Table added

@@
 t<-c(0:(nrow(xL)-1))
-pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=6,width=6)
-op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5)
+pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
 x<-xL
@@
 lines(t,x$erA,col="blue",type="l")
-legend("topleft", c("Left tail","Right tail"), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.2)
+legend("topleft", c("vital rates : LTdep.","vital rates : RTdep."), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
 par(op)
@@
 t<-c(0:(nrow(xL)-1))
-pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=6,width=6)
-op<-par(mar=c(4,4,2,2),mgp=c(2.5,0.5,0))
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=1.5)
+pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
 x<-xL
@@
     params<-c(r,K)
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_ricker_local<-abind(res_ricker_local,temp_mat,along = 3)
@@
     params<-c(r,K)
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_ricker_global<-abind(res_ricker_global,temp_mat,along = 3)
@@
   #s<-res_ricker_local[,,i]
-  s<-res_ricker_local[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_ricker_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,s$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  liness(s$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
   #s<-res_ricker_global
-  s<-res_ricker_global[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_ricker_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
     params<-c(r,a,b)
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_hassell_local<-abind(res_hassell_local,temp_mat,along = 3)
 
-    hassell_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    hassell_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    hassell_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
+    hassell_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
 }
@@
     params<-c(r,a,b)
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_hassell_global<-abind(res_hassell_global,temp_mat,along = 3)
@@
 for(i in 1:dim(res_hassell_local)[3]){
-  s<-res_hassell_local[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_hassell_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 for(i in 1:dim(res_hassell_global)[3]){
-  s<-res_hassell_global[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_hassell_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
     ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_msmith_local<-abind(res_msmith_local,temp_mat,along = 3)
 
-    msmith_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    msmith_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    msmith_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
+    msmith_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
 }
@@
     ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_msmith_global<-abind(res_msmith_global,temp_mat,along = 3)
@@
 for(i in 1:dim(res_msmith_local)[3]){
-  s<-res_msmith_local[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_msmith_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 for(i in 1:dim(res_msmith_global)[3]){
-  s<-res_msmith_global[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_msmith_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
     ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_pennycuick_local<-abind(res_pennycuick_local,temp_mat,along = 3)
 
-    pennycuick_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    pennycuick_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    pennycuick_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
+    pennycuick_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
 }
@@
     ext_thrs<-K_e/10
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_pennycuick_global<-abind(res_pennycuick_global,temp_mat,along = 3)
@@
 for(i in 1:dim(res_pennycuick_local)[3]){
-  s<-res_pennycuick_local[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_pennycuick_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 for(i in 1:dim(res_pennycuick_global)[3]){
-  s<-res_pennycuick_global[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_pennycuick_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_verhulst_local<-abind(res_verhulst_local,temp_mat,along = 3)
 
-    verhulst_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    verhulst_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    verhulst_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
+    verhulst_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
 }
@@
 
-    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-    temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+
+    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
     res_verhulst_global<-abind(res_verhulst_global,temp_mat,along = 3)
@@
 for(i in 1:dim(res_verhulst_local)[3]){
-  s<-res_verhulst_local[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_verhulst_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 for(i in 1:dim(res_verhulst_global)[3]){
-  s<-res_verhulst_global[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_verhulst_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
   r<- rlist[i]
-  K<-50
-  s<-0.2
   p0<-K
@@
 
-  s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-  temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+  ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+               scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+
+  temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
   res_abrewer_local<-abind(res_abrewer_local,temp_mat,along = 3)
 
-  abrewer_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-  abrewer_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+  abrewer_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
+  abrewer_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
 }
@@
 for(r in rlist){
-  #cat("r=",r,"\n")
-  K<-50
-  s<-0.2
   p0<-K
@@
 
-  s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc,checkon=F,getacf=F)
-
-  temp_mat<-cbind(s$d_seq,s$risk_left,s$risk_right,s$pop_avg_acf_left,s$pop_avg_acf_right)
+  ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
+               scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+
+  temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
   res_abrewer_global<-abind(res_abrewer_global,temp_mat,along = 3)
@@
 for(i in 1:dim(res_abrewer_local)[3]){
-  s<-res_abrewer_local[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_abrewer_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 for(i in 1:dim(res_abrewer_global)[3]){
-  s<-res_abrewer_global[,,i][1:6,] # plot for d in range (0 to 0.5)
-  s<-as.data.frame(s)
-
-  plot(s$d,s$risk_left,xlim=range(s$d),ylim=c(0,1),type="b",col="red",
+  ss<-res_abrewer_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(s$d,s$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
        x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Austin-Brewer",paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("bottomright", c("Austin-Brewer",paste0("s = ",s," , K = ",K),paste0(" sd =",scl)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
 mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(paste("(s = 0.2, K = 50, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
 x<-res_abrewer_local
@@
 ```
+
+```{r create_maintext_fig_globaldisp_summary,echo=F,results="hide"}  
+pdf("./Results/allmodel_ext_risk_global_disp.pdf",height=12,width=9)
+
+op<-par(mfrow=c(6,3),mar=c(5.2,4.2,1.2,1.5))
+
+x<-res_ricker_global
+rlist<-ricker_extrisk_d0$r
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
+mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(a = 0.5, b = 100, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_hassell_global
+rlist<-hassell_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(a = 0.5, b = 4, sd = 0.8)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_msmith_global
+rlist<-msmith_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(a = 0.1, b = 0.1, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_pennycuick_global
+rlist<-pennycuick_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_verhulst_global
+rlist<-verhulst_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("(s = 0.2, K = 50, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+x<-res_abrewer_global
+rlist<-abrewer_extrisk_d0$r
+getplot(x=x,rlist = rlist)
+
+par(op)
+dev.off()
+```
+
+```{r make_tab_model_summary,echo=F}
+library(tibble)
+
+dt1 <- tibble(Models = c("Ricker",
+           "Hassell",
+           "Maynard Smith",
+           "Pennycuick",
+           "Verhulst",
+           "Austin-Brewer")
+)
+
+dt2<-tibble(Equations = c("$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} P_t \\exp[r (1-P_t/K)$]",
+                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} rP_t/(1+aP_t)^b$",
+                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} rP_t/[1+(aP_t)^b]$",
+                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} rP_t/[1+\\exp \\{-a(1-P_t/b)\\}]$",
+                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} P_t [1+r(1-P_t)]$",
+                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} P_t [1+r(K-P_t)\\{1-\\exp(-sP_t)\\}]$"
+))
+
+dt3<-tibble("\\hspace{-0.1 cm} Parameters used" = c("K = 50, sd = 1",
+                            "a = 0.5, b = 100, sd = 1",
+                            "a = 0.5, b = 4, sd = 0.8",
+                           "a = 0.1, b= 0.1, sd = 1",
+                           "sd = 0.6",
+                           "s = 0.2, K = 50, sd = 0.6"
+))
+
+dt4<-tibble("$P_e$" = c("K",
+                        "$(r^{1/b}-1)/a$",
+                        "$[(r-1)^{1/b}]/a$",
+                        "$[b\\{a+ln(r-1)\\}]/a$",
+                        "1",
+                        "K"
+))
+
+dt5<-tibble("$r_c$" = c(rc_ricker,
+                        rc_hassell,
+                        rc_msmith,
+                        rc_pennycuick,
+                        rc_verhulst,
+                        rc_abrewer
+))
+
+dt6<-tibble("$r_{bf}$" = c(r_bf_ricker,
+                           r_bf_hassell,
+                           r_bf_msmith,
+                           r_bf_pennycuick,
+                           r_bf_verhulst,
+                           r_bf_abrewer
+))
+
+dt<-cbind(dt1,dt2,dt3,dt4,dt5,dt6)
+saveRDS(dt,"./Results/tab_model_summary.RDS")
+```

### 2019-02-27 mydispersal.R

@@ -53,7 +53,7 @@ Dispersal_mat<-function(numlocs,d,disp_everywhere){
 #--------Output------------------------------------------------------
 #A numsims by numlocs by numsteps+1 array of populations
 #----------------------------------------------------------------------
-popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model,checkon){
+popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
   numsims<-dim(ns)[1]
   numlocs<-dim(ns)[2]
   numsteps<-dim(ns)[3]
@@ -108,13 +108,6 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model,checkon){
       r<-params[1]
       K<-params[2]
       s<-params[3]
-      
-      if(checkon==T){
-        cat("tct=",tct,"\n")
-        saveRDS(res[,,tct],"./res_tct_abrewer.RDS")
-        saveRDS(lam_sto,"./lam_sto_abrewer.RDS")
-      }
-      
       lam_austin<-(1+(r*(K-res[,,tct])*(1-exp(-(s*res[,,tct])))))*res[,,tct]
       res[,,tct+1]<-lam_austin*lam_sto #numsims by numlocs matrix
     }else if(model=="varley"){
@@ -191,9 +184,9 @@ get_avg_acf<-function(sims){
 #       9) model : a character specifying model name
 #       10) ploton : logical(T or F) to get optional plot
 #       11) resloc : location to save plots
-#       checkon, getacf : these are tags
+#       getacf : these are tags
 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,checkon,getacf){
+plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,getacf){
  
   ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1,mn=0,sdev=1)# a righttail dep matrix(numpoints by numlocs,     
   #                                                      numpoints=numsteps*numsims)
@@ -202,11 +195,11 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
[…]

### 2019-02-27 SuppMat.Rmd
minor edit

@@
 
-    ricker_extrisk_d0$risk_left_d0[i]<-s$risk_left[1]
-    ricker_extrisk_d0$risk_right_d0[i]<-s$risk_right[1]
+    ricker_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
+    ricker_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
 }
@@
 
-  plot(ss$d,s$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),

### 2019-02-27 SuppMat.Rmd
minor edit

@@
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  liness(s$d,ss$risk_right,type="b",col="blue")
+  liness(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 

### 2019-02-27 SuppMat.Rmd
editing

--- 2019-02-27 SuppMat.Rmd
@@
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  liness(s$d,ss$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_right,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
   ## Plot the second plot and put axis scale on right
-  #plot(s$d, s$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
+  #plot(ss$d, ss$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
        #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
@@
  # abline(h=0,lty="dotted",col="grey")
- # lines(s$d, s$pop_avg_acf_right, col="skyblue")
+ # lines(ss$d, ss$pop_avg_acf_right, col="skyblue")
 
@@
 for(i in 1:dim(res_ricker_global)[3]){
-  #s<-res_ricker_global
+  #ss<-res_ricker_global
   ss<-res_ricker_global[,,i][1:6,] # plot for d in range (0 to 0.5)

### 2019-02-28 SuppMat.Rmd
editing

@@
   #     axes=FALSE, type="l", col="orange")
- # abline(h=0,lty="dotted",col="grey")
+ # abline(h=0,lty="dotted",col="darkgrey")
  # lines(ss$d, ss$pop_avg_acf_right, col="skyblue")
@@
        angle=90,length=0.1,col="red",code=3)
-abline(v=rc_ricker,lty="dotted") #rc=1
+abline(v=rc_ricker,lty="dotted",col="darkgrey") #rc=1
 
@@
        x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_hassell,lty="dotted")
+abline(v=rc_hassell,lty="dotted",col="darkgrey")
 
@@
        x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_msmith,lty="dotted")
+abline(v=rc_msmith,lty="dotted",col="darkgrey")
 
@@
        x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_pennycuick,lty="dotted")
+abline(v=rc_pennycuick,lty="dotted",col="darkgrey")
 
@@
        y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_verhulst,lty="dotted")
+abline(v=rc_verhulst,lty="dotted",col="darkgrey")
 
@@
        y1=abrewer_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_abrewer,lty="dotted")
+abline(v=rc_abrewer,lty="dotted",col="darkgrey")
 

### 2019-03-01 MainText.Rmd
summary of results

--- 2019-02-27 MainText.Rmd
@@
 
-\listoffigures
-\listoftables
-
 ```{r setup_MainText, echo=F}
@@
 
- <!--Generating maintext figures-->
+# Introduction
+
+objectives:
+
+1) Nature of tail dependence in environmental noise (or in vital rates) affects extinction risk of a population differently.
+       - Age structure stochastic matrix model : extrisk of left tail dep. > extrisk of right tail dep.
+         (This difference is small, could be explained by Tuljapurkar's approx. for linear model, see discussion)
+
+       - 6 Density-dependence models with extreme tail dep. in noise (local/ global dispersal)
+           - extrisk depends on tail dependnece as well as on model parameters (under-compensatory, over-compensatory)
+           - Rescue effect ??
+           - H0 : switching occurs near rc (without dispersal)
+
+# Methods
+
+# Results
+
+# Discussion
 
  <!--FIG : age structure model results : linear, nonlinear-->
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk from single patch age structured model $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$), fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
+\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models, fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
 \label{fig_extrisk_agestr}}
 \end{center}
-\end{figure}
+\end{figure*}
 
 <!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \begin{center}
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_info} extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$ for local dispersion.
-Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime in deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
+Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
 \end{center}
-\end{figure}
+\end{figure*}
 
  <!--FIG : all nonlinear model results with d=0, extrisk vs. r-->
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \begin{center}
@@
 \includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf}\\
-\caption{Extinction risk varies with bifurcation parameter $(r)$ for all nonlinear density-dependence models without dispersal, numsims=10000. \label{fig_extrisk_d0}}
+\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
+the metapopulations affected by extreme left tail dependent noise across
+$5$ multiple patches without dispersal ($d=0$), as intrinsic 
+growth rate ($r$) goes from under to over-compensatory regime. Opposite trends
+was observed with extreme right tail dependent noise.
+Error bars represent the standard errors associated with $10000$
+simulations for extinction risk computed at each $r$.
+\label{fig_extrisk_d0}}
 \end{center}
-\end{figure}
+\end{figure*}
+
+<!--Summary tables for 6 nonlinear models used-->
+```{r read_tab_model_summary,echo=F, results='asis',message=F}
+library(kableExtra)
+library(dplyr)
+dt<-readRDS("./Results/tab_model_summary.RDS")
+knitr::kable(dt, escape=F,
+             format="latex",digits=3, align="c",table.env='table*',
+             caption = "Summary of six density dependence models considered for each metapopulation patch with local or global dispersal; last three columns indicated equilibrium population value ($P_e$), critical growth rate ($r_c$) where deterministic analog of models showed transition from under- to over-compensatory dynamics and bifurcation parameter ($r_{bf}$) where $P_e$ became unstable from stable steady state, respectively.\\label{tab_model_summary}",
+             booktabs=T, linesep = "\\addlinespace"
+             )%>%column_spec(c(1,3,5,6),width="1.4cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.5 cm")
+```

### 2019-03-01 SuppMat.Rmd
summary of results

@@
 
-<!--Ricker model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_local_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Ricker model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Ricker model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_ricker_risk_global_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Hassell model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Hassell model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_hassell_risk_local_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Hassell model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Hassell model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1 \label{fig_hassell_risk_global_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Maynard smith model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Maynard Smith model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.8.
-\label{fig_msmith_risk_local_disp_scl_0.8}}
-\end{center}
-\end{figure}
-
-<!--Maynard smith model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Maynard Smith model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.8.
-\label{fig_msmith_risk_global_disp_scl_0.8}}
-\end{center}
-\end{figure}
-
-<!--Pennycuick model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Pennycuick model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1.
-\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Pennycuick model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Pennycuick model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=1.
-\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Verhulst model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Verhulst model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
-\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Verhulst model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Verhulst model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
-\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Austin brewer model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Austin-Brewer model with local dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
-\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Austin brewer model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
-\caption{Effect of non-Normal environmental noise copula on extinction risk after 25 years; Austin-Brewer model with global dispersion; numlocs=5, numsteps=25, numsims=10000, scl=0.6.
-\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
 ```{r create_maintext_fig_localdisp_summary,echo=F,results="hide"}
@@
 ```
+
+<!--Supp. FIG :  summary : all nonlinear model results for global dispersion : extrisk vs. d-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp.pdf}
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
+Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Ricker model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
+numsims=10000, sd=1. \label{fig_ricker_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Ricker model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
+numsims=10000, sd=1. \label{fig_ricker_risk_global_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Hassell model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
+numsims=10000, sd=1. \label{fig_hassell_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Hassell model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
+numsims=10000, sd=1. \label{fig_hassell_risk_global_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Maynard smith model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
+numsims=10000, sd=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Maynard smith model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
+numsims=10000, sd=0.8. \label{fig_msmith_risk_global_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Pennycuick model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
+numsims=10000, sd=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Pennycuick model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
+numsims=10000, sd=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Verhulst model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Verhulst model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Austin brewer model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Austin brewer model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}

### 2019-03-01 MainText.Rmd
minor change

@@
 
-       - 6 Density-dependence models with extreme tail dep. in noise (local/ global dispersal)
+       - 6 Density-dependent models with extreme tail dep. in noise (local/ global dispersal)
            - extrisk depends on tail dependnece as well as on model parameters (under-compensatory, over-compensatory)
@@
 $5$ multiple patches without dispersal ($d=0$), as intrinsic 
-growth rate ($r$) goes from under to over-compensatory regime. Opposite trends
+growth rate ($r$) goes from under to over-compensatory regime. Opposite trend
 was observed with extreme right tail dependent noise.
 Error bars represent the standard errors associated with $10000$
-simulations for extinction risk computed at each $r$.
+simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
+for each model where their deterministic analog switches from under- to over-compensatory dynamics.
 \label{fig_extrisk_d0}}

### 2019-03-04 SuppMat.Rmd
adding C-SC-G-SG linear age structured plots

@@
 #---------------------------------------------------------------------------------
+```
+
+```{r fig_agestr_C_SC_G_SG_linear, echo=F,results="hide"}
+xL<-C_age$extdf[1:26,]
+xR<-SC_age$extdf[1:26,]
+t<-c(0:(nrow(xL)-1))
+pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
+x<-xL
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
+lines(t,x$erA,col="red",type="l")
+x<-xR
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
+lines(t,x$erA,col="blue",type="l")
+legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
+par(op)
+dev.off()
+
+xL<-SG_age$extdf[1:26,]
+xR<-G_age$extdf[1:26,]
+t<-c(0:(nrow(xL)-1))
+pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
+x<-xL
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
+lines(t,x$erA,col="red",type="l")
+x<-xR
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
+lines(t,x$erA,col="blue",type="l")
+legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
+par(op)
+dev.off()
 ```
@@
 
+<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
+\begin{figure}[!h]
+\textbf{ \hspace{6.8 cm} (A) \hspace{4.4 cm} (B)} \\
+\vspace{-0.9 cm}
+\begin{center}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
+\caption{Extinction risk from single patch age structured linear ($\Omega = \infty$) models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5), numsims=10000.
+All copula has same Kendall correlation $\tau = 0.6$.
+\label{fig_extrisk_agestr_C_SC_G_SG_linear}}
+\end{center}
+\end{figure}
+
 <!--Supp. FIG :  summary : all nonlinear model results for global dispersion : extrisk vs. d-->
@@
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
 Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp}}

### 2019-03-05 SuppMat.Rmd
adding nonlinear age str results for C,SC,G,SG

@@
 pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
+x<-xL
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
+lines(t,x$erA,col="red",type="l")
+x<-xR
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
+lines(t,x$erA,col="blue",type="l")
+legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
+par(op)
+dev.off()
+```
+
+```{r sim_nonlinear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+set.seed(seed)
+source("ext_risk_cop_age_str.R")
+
+numsims<-10000
+numsteps<-100
+p0<-c(5,10)
+ext_thrs<-2
+gshape<-2
+gscale<-1
+bshape1<-3
+bshape2<-2
+par_dist<-c(gshape,gscale,bshape1,bshape2)
+omg<-100
+ktau<-0.6  # kendall's tau
+
+# for clayton
+copC<-claytonCopula(3)
+parC<-iTau(copC,tau=ktau)
+par<-parC
+fam<-3
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_nonlinear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+C_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(C_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+#---------------------------------------------------------------------------------
+
+# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same
+
+fam<-13
+par<-parC
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+#---------------------------------------------------------------------------------
+
+# Now for gumbel
+
+fam<-4
+copG<-gumbelCopula(2)
+parG<-iTau(copG,tau = ktau)
+par<-parG
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+
+#---------------------------------------------------------------------------------
+
+# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same
+
+fam<-14
+par<-parG
+
+cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
+
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
+            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
+saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
+#---------------------------------------------------------------------------------
+```
+
+```{r fig_agestr_C_SC_G_SG_nonlinear, echo=F,results="hide"}
+xL<-C_age$extdf[1:26,]
+xR<-SC_age$extdf[1:26,]
+t<-c(0:(nrow(xL)-1))
+pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
+x<-xL
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
+lines(t,x$erA,col="red",type="l")
+x<-xR
+polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
+lines(t,x$erA,col="blue",type="l")
+legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
+par(op)
+dev.off()
+
+xL<-SG_age$extdf[1:26,]
+xR<-G_age$extdf[1:26,]
+t<-c(0:(nrow(xL)-1))
+pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 \begin{figure}[!h]
-\textbf{ \hspace{6.8 cm} (A) \hspace{4.4 cm} (B)} \\
-\vspace{-0.9 cm}
 \begin{center}
+\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
+\vspace{-0.2 cm}
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
-\caption{Extinction risk from single patch age structured linear ($\Omega = \infty$) models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5), numsims=10000.
+\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
+\vspace{-0.2 cm}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
+\caption{Extinction risk from single patch age structured models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
+$(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$).
 All copula has same Kendall correlation $\tau = 0.6$.

### 2019-03-12 MainText.Rmd
adding C-SC-G-SG linear age structured plots
started writing methods

@@
 
+<!--stochastic matrix model : age structured-->
+To investigate how copula structure in noisy environment affect the extinction risk of
+populations in a single patch age-structured model,
+we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
+with two age classes: 
+one is egg ($E$) and the other is adult ($A$). We assumed
+the population was semelperous, i.e., adult ($A$) went through the
+reproductive episode only once before death. Eggs and adults in the 
+next generation depend upon the fecundity of adult ($f_a \geq 0$) 
+and survival of eggs ($0 \leq s_e \leq 1$) from present generation, respectively.
+The environmental stochasticity in the model was embedded in 
+the fluctuating vital rates ($f_a$ and $s_e$). We considered the fecundity ($f_a$)
+and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
+a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively,
+which were generated from a specific tail-dependent copula
+(for details see Appendix \ref{SM-age_str_vitalrates}).
+Eq. \ref{eq_vitalrates} represent the stochasticity in the vital rates.
+Demographic stochasticity was also taken into consideration when we included the 
+density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
+If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, the projection matrix was independent
+of populations from any age-class and the model was then called to be "linear", whereas
+its "nonlinear" version had finite $\Omega$.
+
+\begin{equation}
+\left({\begin{array}{cc} E \\ A \end{array}}\right)_{t+1} = 
+\left(\begin{array}{cc} 0 & f_a\\ s_e & 0 \end{array}\right)
+\left(\begin{array}{c} E \\ A \end{array}\right)_{t} \label{eq_age_str}
+\end{equation}
+
+\begin{equation}
+f_a = \Gamma (k, \theta)\exp(-A/\Omega), \hspace{0.2 cm} s_e = \mathcal{B}(\alpha,\beta) \label{eq_vitalrates}
+\end{equation}
+
+<!--calculation of ext risk with age-structured model-->
+
+<!--density-dependence metapopulation model-->
+
+<!--calculation of ext risk with density-dependence metapopulation model-->
+
 # Results
@@
  <!--FIG : age structure model results : linear, nonlinear-->
+
 \begin{figure*}[!h]

### 2019-03-13 MainText.Rmd
writing methods...continued, extrisk for 25 years

@@
 and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
-a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively,
-which were generated from a specific tail-dependent copula
-(for details see Appendix \ref{SM-age_str_vitalrates}).
-Eq. \ref{eq_vitalrates} represent the stochasticity in the vital rates.
-Demographic stochasticity was also taken into consideration when we included the 
-density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
+a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distribution 
+could be generated from iid noises showing specific tail-dependence (left or right tail)
+between fecundity and survival rates. Eq. \ref{eq_vitalrates} represent the stochasticity 
+in the vital rates. Demographic stochasticity was also taken into consideration when 
+we included the density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
 If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, the projection matrix was independent
@@
 <!--calculation of ext risk with age-structured model-->
+To calculate extinction risk after each time step, we simulated model (\ref{eq_age_str}) for $25$ years
+with $10000$ simulations. For each simulation, we started with initial population $(E,A)_{t=0} = (5,10)$
+and following Eq. \ref{eq_age_str} after each time step checked the total population $(E+A)_{t}$
+was below a certain extinction threshold (we set it as $2$) or not. If it was below that threshold then we 
+considered population went extinct for that simulation. We also calculated $2.5^{th}$ and $97.5^{th}$ quantiles
+of $10000$ simulations with the probability of computed extinction risk at each time step.
+
+We repeated the above analysis considering the vital rates coming from a 
+an asymmetric copula (Appendix \ref{SM-age_str_vitalrates}) with specific 
+left (Clayton, Survival Gumbel)
+or right (Survival Clayton, Gumbel) tail dependence
+or from an extreme tail-depentdent one, when vital rates were considered to be affected
+by extreme climatic events.
 

### 2019-03-13 SuppMat.Rmd
started writing methods
writing methods...continued, extrisk for 25 years

@@
 numsims<-10000
-numsteps<-100
+numsteps<-25
 p0<-c(5,10)
@@
 ```{r fig_agestr_C_SC_G_SG_linear, echo=F,results="hide"}
-xL<-C_age$extdf[1:26,]
-xR<-SC_age$extdf[1:26,]
+xL<-C_age$extdf
+xR<-SC_age$extdf
 t<-c(0:(nrow(xL)-1))
@@
 
-xL<-SG_age$extdf[1:26,]
-xR<-G_age$extdf[1:26,]
+xL<-SG_age$extdf
+xR<-G_age$extdf
 t<-c(0:(nrow(xL)-1))
@@
 ```{r fig_agestr_C_SC_G_SG_nonlinear, echo=F,results="hide"}
-xL<-C_age$extdf[1:26,]
-xR<-SC_age$extdf[1:26,]
+xL<-C_age$extdf
+xR<-SC_age$extdf
 t<-c(0:(nrow(xL)-1))
@@
 ```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}
-xL<-las_exL$extdf[1:26,]
-xR<-las_exR$extdf[1:26,]
+xL<-las_exL$extdf
+xR<-las_exR$extdf
 t<-c(0:(nrow(xL)-1))
@@
 ```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}
-xL<-nlas_exL$extdf[1:26,]
-xR<-nlas_exR$extdf[1:26,]
+xL<-nlas_exL$extdf
+xR<-nlas_exR$extdf
 t<-c(0:(nrow(xL)-1))
@@
 \end{figure}
+
+# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}
+
+A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
+(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
+We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
+(see `iTau` function from `copula` package). For fecundity of adults, we transformed the first column of matrix 
+$M$ as gamma ($\Gamma(k,\theta)$) distributed, $k = 2$ and $\theta = 1$ were shape and scale parameters, 
+respectively. For a given $\theta$, increasing 
+$k$ indicates distribution will shift towards higher fecundity which might 
+reduce the extinction risk if the other parameters
+remain unchanged. For survival rates of eggs, we took the second column of $M$ and applied a 
+beta ($\mathcal{B}(\alpha,\beta)$) transformation, $\alpha =3$ and
+$\beta = 2$ were the shape parameters which could control the distribution 
+of the survival rates of eggs. Entries within each clumn of matrix $M$, were thus independent of each others but 
+the two columns of $M$ had a specified Kendall's correlation ($\tau = 0.6$) and underlying tail-dependence. 
+For each $25$ timesteps from each of $10000$ simulations, the vital rates (fecundity and survival rates) 
+of the stochastic projrction matrix in model (\ref{MT-eq_age_str}) were drawn pairwise from each row of matrix $M$.
+
+For extreme tail-dependence algorithm,

### 2019-03-13 ext_risk_cop_age_str.R

@@ -46,10 +46,11 @@ sim_age_str<-function(p0,ext_thrs,cop,par_dist,omg,numsims,numsteps,ploton,reslo
   if(ploton==ploton){
     pdf(paste(resloc,"fA_sE_gshape_",gshape,"_gscale_",gscale,"_bshape1_",bshape1,"_bshape2_",bshape2,".pdf",sep=""),height=5,width=10)
     op<-par(mfrow=c(1,2))
-    hist(x,100,main="fecundity_adult : gamma dist.",col="grey",border=F,xlab="")
+    hist(x,100,main="",col="grey",border=F,xlab="Fecundity of adults")
+    #hist(x,100,main="fecundity of adult : gamma dist.",col="grey",border=F,xlab="")
     #mtext(paste0("gshape = ",gshape," , gscale = ", gscale),line=-1) 
     
-    hist(y,100,main="survival_egg : beta dist.",col="grey",border=F,xlab="")
+    hist(y,100,main="",col="grey",border=F,xlab="Survival of eggs")
     #mtext(paste0("bshape1 = ",bshape1," , bshape2 = ", bshape2),line=-1) 
     par(op)
     dev.off()

### 2019-03-13 MainText.Rmd
methods : writing age structure model completed

@@
 we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
-with two age classes: 
-one is egg ($E$) and the other is adult ($A$). We assumed
+with two age classes(egg $E$ and adult $A$). We assumed
 the population was semelperous, i.e., adult ($A$) went through the
@@
 between fecundity and survival rates. Eq. \ref{eq_vitalrates} represent the stochasticity 
-in the vital rates. Demographic stochasticity was also taken into consideration when 
-we included the density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
-If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, the projection matrix was independent
+in the vital rates. We included the density dependence effect in one of the vital rate,
+namely, fecundity ($f_a$). If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, 
+the projection matrix would be independent
 of populations from any age-class and the model was then called to be "linear", whereas
@@
 We repeated the above analysis considering the vital rates coming from a 
-an asymmetric copula (Appendix \ref{SM-age_str_vitalrates}) with specific 
+an asymmetric copula with specific 
 left (Clayton, Survival Gumbel)
 or right (Survival Clayton, Gumbel) tail dependence
-or from an extreme tail-depentdent one, when vital rates were considered to be affected
-by extreme climatic events.
+or from an extreme tail-depentdent one, 
+when vital rates were considered to be affected
+by extreme climatic events. A $N \times 2$ matrix, M, was generated
+using `BiCopSim` function
+from `VineCopula` package 
+for each of the left 
+(Clayton, Survival Gumbel) and right 
+(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
+where $N = numsteps(=25) \times numsims(=10000)$.
+We kept the same Kendall's correlation ($\tau = 0.6$) for 
+each bivariate copula from one parameter family which used its parameter accordingly
+(see `iTau` function from `copula` package). Entries within each clumn of
+matrix $M$, were thus 
+independent of each others but 
+the two columns of $M$ had a specified Kendall's 
+correlation ($\tau = 0.6$) and underlying tail-dependence. 
+Similarly, for extreme-tail dependence
+(either left or right tail) a matrix M with two column was generated as described in 
+Appendix \ref{SM-extr_tail_dep}.
+
+We made the first column of matrix 
+$M$ as gamma ($\Gamma(k,\theta)$) distributed. 
+$k = 2$ and $\theta = 1$ were shape and scale parameters, 
+respectively. This column represented the distribution of fecundity of the adult population, $f_a$.
+For a given $\theta$, increasing 
+$k$ indicates distribution will shift towards higher fecundity which might 
+reduce the extinction risk if the other parameters
+remain unchanged. We took the second column of $M$ and applied a 
+beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
+$\beta = 2$ were the shape parameters which could control the distribution 
+of the survival rates of eggs, $s_e$. 
+For each $25$ timesteps from each of $10000$ simulations,
+the vital rates (fecundity $f_a$ and survival rates $s_e$) 
+of the stochastic projrction matrix in model (\ref{eq_age_str})
+were drawn pairwise from each row of matrix $M$.
+[`sim_age_str`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R) 
+is the function we used to calculate extinction risk for age structured model.
 

### 2019-03-13 SuppMat.Rmd
methods : writing age structure model completed

@@
 
+<!--
 # Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}
-
 A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
@@
 We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
-(see `iTau` function from `copula` package). For fecundity of adults, we transformed the first column of matrix 
-$M$ as gamma ($\Gamma(k,\theta)$) distributed, $k = 2$ and $\theta = 1$ were shape and scale parameters, 
-respectively. For a given $\theta$, increasing 
-$k$ indicates distribution will shift towards higher fecundity which might 
-reduce the extinction risk if the other parameters
-remain unchanged. For survival rates of eggs, we took the second column of $M$ and applied a 
-beta ($\mathcal{B}(\alpha,\beta)$) transformation, $\alpha =3$ and
-$\beta = 2$ were the shape parameters which could control the distribution 
-of the survival rates of eggs. Entries within each clumn of matrix $M$, were thus independent of each others but 
-the two columns of $M$ had a specified Kendall's correlation ($\tau = 0.6$) and underlying tail-dependence. 
-For each $25$ timesteps from each of $10000$ simulations, the vital rates (fecundity and survival rates) 
-of the stochastic projrction matrix in model (\ref{MT-eq_age_str}) were drawn pairwise from each row of matrix $M$.
-
-For extreme tail-dependence algorithm,
+(see `iTau` function from `copula` package).-->
+
+# Generating matrix with extreme tail dependence  \label{extr_tail_dep}
+
+For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
+standard-normal random draws (in general), where $N = numsteps(=25) \times numsims(=10000)$.
+Then, each row of
+M was identified for a first set of modifications
+randomly and independently with a 50\% chance; for those rows identified, every
+element of the row was replaced with the absolute
+value of the first element of the row (including the first element itself).
+For other rows, every element in the row was replaced by the negative of 
+its own absolute value. This produces right-tail dependent
+matrix, normally distributed for each column. If left-tail 
+dependent noise is desired, take the negative of the entire matrix.
+Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
+generating extreme tail dependence matrix is flexible enough, so that 
+matrix M can be generated initially with independent random draws from a normal 
+distribution with a specified mean ($m$) and standard deviation $(sd)$.

### 2019-03-14 MainText.Rmd
wrting methods : density dep. metapop. model...continued

@@
 
-# Introduction
+# Introduction\label{intro}
 
-objectives:
-
-1) Nature of tail dependence in environmental noise (or in vital rates) affects extinction risk of a population differently.
+<!--
+Nature of tail dependence in environmental noise (or in vital rates) affects extinction risk of a population differently.
        - Age structure stochastic matrix model : extrisk of left tail dep. > extrisk of right tail dep.
@@
            - H0 : switching occurs near rc (without dispersal)
+-->
+
+We explored the potential of asymmetric tail-dependence in copula 
+for understanding the extinction risk of populations towards extreme climatic event in ecology
+by addressing the following specific questions: (Q1) Does nature of tail dependence in vital rates
+of a single-patch age-structured population model affect extinction risk of that population differently?
+We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
+(Q2) How does asymmetric tail-dependent noise across multiple patches influence extinction 
+risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
+from density-independent metapopulation model (higher extinction risk if noise across patches were 
+left-tail dependent than the one if noise came from right-tail dependent) holds true only in
+the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
+(Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
+near the critical parameter value where the deterministic model shows transition from 
+under-compensatory to over-compensatory regime.
 
@@
 <!--stochastic matrix model : age structured-->
-To investigate how copula structure in noisy environment affect the extinction risk of
+To investigate (Q1) from \nameref{intro} that how copula structure in noisy 
+environment affect the extinction risk of
 populations in a single patch age-structured model,
 we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
-with two age classes(egg $E$ and adult $A$). We assumed
+with two age classes (egg $E$ and adult $A$). We assumed
 the population was semelperous, i.e., adult ($A$) went through the
@@
 
-We repeated the above analysis considering the vital rates coming from a 
+To test our hypothesis from (Q1), we repeated the above analysis 
+considering the vital rates coming from a 
 an asymmetric copula with specific 
@@
 <!--density-dependence metapopulation model-->
+We explored (Q2) by examining how nature of the tail dependence of an
+environmental variable across space could influence the extinction risk of a 
+metapopulation model with density-dependence feedback. We considered six density-dependent 
+population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
+in noisy environment (see Table \ref{tab_model_summary}) 
+from a metapopulation context.
+Models were adapted from earlier work
+by @cohen1995unexpected, excluding only two : Varley model (this is not a continuous one) 
+and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
+All the models could be represented via 
+a generalized form : $\vec{P_{t+1}} = D \lambda_t \vec{P_t}$,
+where $\vec{P_t}$ was a vector of population densities, 
+the $i^{th}$ component $P_t^i$ representing
+the density in the $i^{th}$ habitat patch ($i=1,\ldots,n$) at time $t$. 
+The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
+of stochastic density dependence form $\lambda_i(t)$ (see second column 
+of Table \ref{tab_model_summary}). For simplicity, 
+we considered $n$ patches in each model were arranged, evenly spaced, in a line
+with the provision of "local" or "global" dispersal among patches. A n $\times$ n dispersal
+matrix, $D$ was constructed assuming that a fraction $d$ of 
+each population patch dispersed during any given time step, equally distributed to
+the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
+$d$ that dispersed
+from each patch was divided up equally among all other patches. We called $d$ as dispersal
+rate with $d$ being zero there was no dispersal, and threby no chance to rescue an extinct 
+patch through migration of nearby metapopulation patches.
 
-<!--calculation of ext risk with density-dependence metapopulation model-->
+The environmental noises from any two patch, say, 
+$(\epsilon_t^i, \epsilon_t^j)$ were normally 
+distributed within each patch with mean $m=0$ and standard deviation $sd$, 
+(see the specified $sd$ for each model in 3$^{rd}$ 
+column of Table \ref{tab_model_summary}), were 
+independent through time, and showed the same 
+spatial correlations between patches (to within sampling error) for every simulation. 
+But they were made to exhibit right- or left-tail dependence between patches in
+different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
+$\epsilon$, was made from these extreme tail-dependence noise 
+matrix generated via algorithm stated in 
+Appendix \ref{SM-extr_tail_dep} and used to calculate extinction risk 
+for each of the specified model in Table \ref{tab_model_summary} for $numsteps = 25$ years, $n =$ 5 patches
+and $numsims = 10000$ simulations.
+
+<!--say how you set ext. threshold = Pe/10-->
+
+<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
+
+<!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->
 
@@
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
-Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal
+among $n=5$ patches, extinction risk after $numsteps = 25$ years for $numsims = 10000$ simulations were calculated.
+Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
+Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
 \end{center}
@@
              format="latex",digits=3, align="c",table.env='table*',
-             caption = "Summary of six density dependence models considered for each metapopulation patch with local or global dispersal; last three columns indicated equilibrium population value ($P_e$), critical growth rate ($r_c$) where deterministic analog of models showed transition from under- to over-compensatory dynamics and bifurcation parameter ($r_{bf}$) where $P_e$ became unstable from stable steady state, respectively.\\label{tab_model_summary}",
+             caption = "Summary of six density dependence metapopulation models considered; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$) where deterministic analog of models showed transition from under- to over-compensatory dynamics and bifurcation parameter ($r_{bf}^{i}$) where $P_e^{i}$ became unstable from stable steady state, respectively.
+Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"
-             )%>%column_spec(c(1,3,5,6),width="1.4cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.5 cm")
+             )%>%column_spec(c(1,3,5,6),width="1.5 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
 ```
+
+\clearpage
+
+# Literature cited

### 2019-03-14 SuppMat.Rmd
wrting methods : density dep. metapop. model...continued

@@
 
-dt1 <- tibble(Models = c("Ricker",
+dt1 <- tibble("Models $(P_{t+1}^{i} = \\lambda_t^{i} P_t^{i})$"= c("Ricker",
            "Hassell",
@@
 
-dt2<-tibble(Equations = c("$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} P_t \\exp[r (1-P_t/K)$]",
-                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} rP_t/(1+aP_t)^b$",
-                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} rP_t/[1+(aP_t)^b]$",
-                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} rP_t/[1+\\exp \\{-a(1-P_t/b)\\}]$",
-                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} P_t [1+r(1-P_t)]$",
-                          "$P_{t+1} = \\exp(\\epsilon_t) \\hspace{0.1 cm} P_t [1+r(K-P_t)\\{1-\\exp(-sP_t)\\}]$"
+dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
+                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/(1+aP_t^{i})^b$",
+                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+(aP_t^{i})^b]$",
+                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_t^{i}/b)\\}]$",
+                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i})]$",
+                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(K-P_t^{i})\\{1-\\exp(-sP_t^{i})\\}]$"
 ))
@@
 
-dt4<-tibble("$P_e$" = c("K",
+dt4<-tibble("$P_e^{i}$" = c("K",
                         "$(r^{1/b}-1)/a$",
@@
 
-dt5<-tibble("$r_c$" = c(rc_ricker,
+dt5<-tibble("$r_c^{i}$" = c(rc_ricker,
                         rc_hassell,
@@
 
-dt6<-tibble("$r_{bf}$" = c(r_bf_ricker,
+dt6<-tibble("$r_{bf}^{i}$" = c(r_bf_ricker,
                            r_bf_hassell,
@@
 distribution with a specified mean ($m$) and standard deviation $(sd)$.
+
+# Stability analysis of deterministic metapopulation models  \label{model_stability}

### 2019-03-14 MainText.Rmd
modified plots

@@
 near the critical parameter value where the deterministic model shows transition from 
-under-compensatory to over-compensatory regime.
+under-compensatory to over-compensatory dynamics.
 
@@
 rate with $d$ being zero there was no dispersal, and threby no chance to rescue an extinct 
-patch through migration of nearby metapopulation patches.
+patch through migration from nearby metapopulation patches.
 
@@
 distributed within each patch with mean $m=0$ and standard deviation $sd$, 
-(see the specified $sd$ for each model in 3$^{rd}$ 
+(see the specified value $sd$ for each model in 3$^{rd}$ 
 column of Table \ref{tab_model_summary}), were 
@@
 <!--say how you set ext. threshold = Pe/10-->
+To estimate extinction risk, we started with 
+initial population $P_{t=0}^i = P_e^i$ for all patches ($i = 1,2,...,n$) where 
+$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
+for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
+We assumed population went extinct after each time step if it was below a certain threshold.
+This extinction threshold was set to 10\% of initial population $P_e$ for individual model.
 

### 2019-03-14 mydispersal.R

@@ -220,42 +220,41 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
     pdf(paste0(resloc,"r_",params[1],"_noise_lastsim_timeseries_all_locs.pdf",sep=""),height=5,width=10)
     op<-par(mfrow=c(1,2),mar=c(3.5,4.5,2,3.5),mgp=c(1.9,0.5,0))
     
-    ns_sim1<-ns1[numsims,,]
-    ylm<-max(c(abs(min(ns_sim1)),abs(max(ns_sim1))))
-    plot(0,0,xlim = c(1,numsteps),ylim = c(-ylm,ylm),type = "n",xlab="time",ylab="noise_right_tail")
-    cl <- rainbow(numlocs)
-    for(i in 1:numlocs){
-      lines(c(1:numsteps),ns_sim1[i,],col=cl[i],type="l")
-    }
-    
     ns_sim1<-ns2[numsims,,]
     ylm<-max(c(abs(min(ns_sim1)),abs(max(ns_sim1))))
-    plot(0,0,xlim = c(1,numsteps),ylim = c(-ylm,ylm),type = "n",xlab="time",ylab="noise_left_tail")
+    plot(0,0,xlim = c(0,numsteps),ylim = c(-ylm,ylm),type = "n",xlab="Time",ylab="Noise : LTdep.")
     cl <- rainbow(numlocs)
     for(i in 1:numlocs){
-      lines(c(1:numsteps),ns_sim1[i,],col=cl[i],type="l")
+      lines(c(0:numsteps),c(0,ns_sim1[i,]),col=cl[i],type="l") # at t=0 noise was also 0, for t>0 , there was finite noise in each time step
     }
     
+    ns_sim1<-ns1[numsims,,]
+    ylm<-max(c(abs(min(ns_sim1)),abs(max(ns_sim1))))
+    plot(0,0,xlim = c(0,numsteps),ylim = c(-ylm,ylm),type = "n",xlab="Time",ylab="Noise : RTdep.")
+    cl <- rainbow(numlocs)
+    for(i in 1:numlocs){
+      lines(c(0:numsteps),c(0,ns_sim1[i,]),col=cl[i],type="l") # at t=0 noise was also 0, for t>0 , there was finite noise in each time step
+    }
     par(op)
     dev.off()
     
[…]

### 2019-03-15 MainText.Rmd
writing methods...continued

@@
 ---
-title: "Main text : Effect of copula on extinction risk"
+title: "Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology"
 author: "Shyamolina Ghosh"
@@
 ```
+
+\noindent\makebox[\linewidth]{\rule{\columnwidth}{1pt}}
+
+\scriptsize
+
+\textbf{ABSTRACT:}  \small write abstract here
+
+\textbf{\textit{Keywords:}}
+
+\noindent\makebox[\linewidth]{\rule{\columnwidth}{1pt}}
+
+\normalsize
 
@@
 All the models could be represented via 
-a generalized form : $\vec{P_{t+1}} = D \lambda_t \vec{P_t}$,
+a generalized form : $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
 where $\vec{P_t}$ was a vector of population densities, 
 the $i^{th}$ component $P_t^i$ representing
-the density in the $i^{th}$ habitat patch ($i=1,\ldots,n$) at time $t$. 
+the density in the $i^{th}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
 The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
@@
 spatial correlations between patches (to within sampling error) for every simulation. 
-But they were made to exhibit right- or left-tail dependence between patches in
+But they were made to exhibit extreme right- or extreme left-tail dependence between patches in
 different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
@@
 
+The models stated in Table \ref{tab_model_summary} had one thing in 
+common that they showed undercompensatory
+dynamics (monotonic approach towards stable steady state $P_e$) when
+growth parameter $r$ remained below a critical 
+level, say, $r_c$. If $r$ was beyond $r_c$, model showed 
+overcompensatory dynamics with oscillatory approach 
+towards stable equilibrium $P_e$. But if $r$ became too high and went
+beyond $r_{bf}$, the parameter value responsible for Hopf-bifurcation,
+model equilibrium $P_e$ became unstable and model could show sustained
+oscillation and eventually led to chaos. Appendix \ref{SM-model_stability}
+summarizes the stability analysis of each model in their deterministic form.
+
 <!--say how you set ext. threshold = Pe/10-->
-To estimate extinction risk, we started with 
-initial population $P_{t=0}^i = P_e^i$ for all patches ($i = 1,2,...,n$) where 
+We started with $n$ patches of initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ where 
 $P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
@@
 We assumed population went extinct after each time step if it was below a certain threshold.
-This extinction threshold was set to 10\% of initial population $P_e$ for individual model.
+That extinction threshold was set to 10\% of initial population $P_e$ for individual model.
 
 <!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
+To test our hypothesis from (Q2) of \nameref{intro} for each metapopulation model,
+we estimated extinction risk for a wide range of parameter $r$ 
+which covered the undercompensatory and overcompensatory regime. We 
+chose equispaced values of $r$ from the 95\% quantiles 
+of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
+for which the model equilibrium $P_e$ had always a finite, positive initial 
+population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ of each 
+model where metapopulation could disperse locally with varying dispersal 
+rate $d= 0,0.1,0.2,..0.5$, we calculated extinction risk with left 
+tail dependent and right tail dependent environmental noise 
+in different simulations, respectively. We also repeated the analysis 
+for global dispersal for each metapopulation model. Script 
+[`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R) was used for this 
+analysis.
 
 <!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->
+Given hypothesis from (Q2) was valid which scrutinized that whereas 
+left tail dependence in noise made extinction risk 
+higher in undercompensatory regime ($r < r_c$), it was the 
+right tail dependent noise
+which appeared to evoke higher extinction risk for metapopulations
+in overcompensatory regime ($r > r_c$), we tested our third hypothesis for (Q3).
+We considered
+the switching to take place near $r_c$, not exactly at $r = r_c$ because the transition point $r_c$
+from under to overcompensatory dynamics was expected to differ in stochastic environment than its 
+deterministic analog depending on the underlying nonlinearity of the model.
+We generated plots with extinction risk along y-axis
+for both left tail and right tail dependent noises 
+against the varying range of $r$ when no dispersal ($d=0$) was considered.
+We excluded the possibility of dispersal in this case to avoid any rescue effect among patches. 
+For each model with extreme left and extreme right tail dependent noises, we calculated 
+2.5$^{th}$ and 97.5$^{th}$ quantiles
+of 10000 simulations with the probability of computed
+extinction risk at each $r$. We compared $r_c$ and the value of $r$ at which 
+95\% CI of extinction risk from both noises overlapped
+with each other. This could give us an idea for supporting our hypothesis.
 

### 2019-03-15 SuppMat.Rmd
writing methods...continued

@@
 ---
-title: "Supporting Information : Effect of copula on extinction risk"
+title: "Supporting Information : Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology"
 author: "Shyamolina Ghosh"

### 2019-03-16 MainText.Rmd
suppmat description of density dependence model : upto Hassell added

@@
 
-<!--
-Nature of tail dependence in environmental noise (or in vital rates) affects extinction risk of a population differently.
-       - Age structure stochastic matrix model : extrisk of left tail dep. > extrisk of right tail dep.
-         (This difference is small, could be explained by Tuljapurkar's approx. for linear model, see discussion)
-
-       - 6 Density-dependent models with extreme tail dep. in noise (local/ global dispersal)
-           - extrisk depends on tail dependnece as well as on model parameters (under-compensatory, over-compensatory)
-           - Rescue effect ??
-           - H0 : switching occurs near rc (without dispersal)
--->
-
 We explored the potential of asymmetric tail-dependence in copula 
@@
 risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
-from density-independent metapopulation model (higher extinction risk if noise across patches were 
+from density-independent metapopulation model [BIVAN paper results of Lewinton-Cohen model]
+(higher extinction risk if noise across patches were 
 left-tail dependent than the one if noise came from right-tail dependent) holds true only in

### 2019-03-16 SuppMat.Rmd
suppmat description of density dependence model : upto Hassell added

@@
 
-dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
+dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise $(\\epsilon_t^{i})$" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
                           "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/(1+aP_t^{i})^b$",
@@
 ```
+
+<!--
+# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}
+A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
+(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
+We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
+(see `iTau` function from `copula` package).-->
+
+# Generating matrix with extreme tail dependence  \label{extr_tail_dep}
+
+For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
+standard-normal random draws (in general), where $N = numsteps(=25) \times numsims(=10000)$.
+Then, each row of
+M was identified for a first set of modifications
+randomly and independently with a 50\% chance; for those rows identified, every
+element of the row was replaced with the absolute
+value of the first element of the row (including the first element itself).
+For other rows, every element in the row was replaced by the negative of 
+its own absolute value. This produces right-tail dependent
+matrix, normally distributed for each column. If left-tail 
+dependent noise is desired, take the negative of the entire matrix.
+Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
+generating extreme tail dependence matrix is flexible enough, so that 
+matrix M can be generated initially with independent random draws from a normal 
+distribution with a specified mean ($m$) and standard deviation $(sd)$.
+
+# Stability analysis of deterministic metapopulation models  \label{model_stability}
+
+First, we will briefly discuss about the stability analysis for a difference 
+equation which is frequently found in population ecology model. Let's say, $P_{t+1} = F(P_t)$
+represents the model dynamics of a single population $P$ which progress recursively via 
+density-dependence function $F$ with discrete time variable $t$. Fixed-point analysis can
+be used to get information about this type of nonlinear difference equation. The idea is to
+employ the concept that the system at fixed point (or equilibrium $P_e$) does not change with time 
+(i.e. $P_{t+1} = P_t$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
+returns to its original state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
+Mathematically, from standard linear stability analysis we can say an equilibrium is locally 
+asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P_t=P_e}$.
+If the approach towards its equilibrium is monotonic (when $0<{F}'<1$), the dynamics is said to be
+in under-compensatory regime. An over-compensatory dynamics means the approach towards its stable equlibrium
+is oscillatory ($-1<{F}'<0$). Following we will describe stability condition of the model dynamics considered for each 
+patch of metapopulation models in the maintext (see Table \ref{MT-tab_model_summary}).
+
+**Ricker model** [@ricker1954stock], originally introduced in the context of modelling
+stock and recruitment of salmon populations, is a classic 
+discrete population model 
+and frequently used in fisheries management problem. The model dynamics
+is represented as 
+$P_{t+1}=P_t \exp[r(1-P_t/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
+growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
+The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ and considered 
+as the extinction equilibrium 
+in population ecology context. Depending on parameter $r$, 
+the other equilibrium $P_e = K$ can be stable or unstable as ${F}' = (1-r)$. Following generalized
+concept, the model can show monotonic approach to its stable equilibrium $P_e = K$ if $0<r<1$ 
+or can approach $P_e = K$ in an oscillating fashion if $1<r<2$. When 
+$r>2$, system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
+as $r_c$, the critical point where system goes from under- to over-compensatory regime and 
+$r=2$ as $r_{bf}$ where bifurcation (transition from stable to unstable equilibrium) occurs.
+
+**Hassell model** [@hassell1975density], motivated for modelling insect populations,
+can be represented as $P_{t+1} = rP_t/(1+aP_t)^b$, where $r>0$ is the intrinsic growth rate 
+of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
+Beverton-Holt model [@beverton2012dynamics] used in fisheries. In general, if $b \sim 1$,
+then we say that there is contest and $F$ will saturate eventually at a finite
+level for large populations, which indicates that the population stabilizes by
+rejecting too many newborns. For scramble, $b$ is large and $F$ tends to
+zero for large populations which indicates that the resources are over-utilized
+leading to eventual extinction. Like Ricker, Hassel model too has two equilibrim:
+$P_e =0$ (trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
+For $P_e = P^\star$, model remains in under-compensatory regime if $1<r<[b/(b-1)]^b$ 
+and in over-compensatory regime if $[b/(b-1)]^b<r<[b/(b-2)]^b$ with $r_c = [b/(b-1)]^b$.
+Hopf bifurcation occurs at $r_{bf} = [b/(b-2)]^b$, so if $r>r_{bf}$, $P_e=P^\star$
+becomes unstable equilibrium.
+
+**Maynard Smith model**
+
+**Pennycuick model**
+
+**Verhulst model**
+
+**Austin-Brewer model**
+
+# References
+
+\clearpage
 
@@
 \end{figure}
-
-<!--
-# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}
-A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
-(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
-We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
-(see `iTau` function from `copula` package).-->
-
-# Generating matrix with extreme tail dependence  \label{extr_tail_dep}
-
-For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
-standard-normal random draws (in general), where $N = numsteps(=25) \times numsims(=10000)$.
-Then, each row of
-M was identified for a first set of modifications
-randomly and independently with a 50\% chance; for those rows identified, every
-element of the row was replaced with the absolute
-value of the first element of the row (including the first element itself).
-For other rows, every element in the row was replaced by the negative of 
-its own absolute value. This produces right-tail dependent
-matrix, normally distributed for each column. If left-tail 
-dependent noise is desired, take the negative of the entire matrix.
-Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
-generating extreme tail dependence matrix is flexible enough, so that 
-matrix M can be generated initially with independent random draws from a normal 
-distribution with a specified mean ($m$) and standard deviation $(sd)$.
-
-# Stability analysis of deterministic metapopulation models  \label{model_stability}

### 2019-03-16 SuppMat.Rmd
written upto Pennycuick, need to include K in verhulst model I think

@@
 First, we will briefly discuss about the stability analysis for a difference 
-equation which is frequently found in population ecology model. Let's say, $P_{t+1} = F(P_t)$
+equation which is frequently used in population ecology model. Let's say, $P_{t+1} = F(P_t,r)$
 represents the model dynamics of a single population $P$ which progress recursively via 
-density-dependence function $F$ with discrete time variable $t$. Fixed-point analysis can
+a nonlinear density-dependence function $F$ with discrete time variable $t$. $r$ is the model parameter 
+which plays major role in assessing the model's dynamical stability. Fixed-point analysis can
 be used to get information about this type of nonlinear difference equation. The idea is to
@@
 
-**Ricker model** [@ricker1954stock], originally introduced in the context of modelling
+**Ricker model** [@ricker1954stock; @may1976bifurcations], originally introduced in the context of modelling
 stock and recruitment of salmon populations, is a classic 
@@
 is represented as 
-$P_{t+1}=P_t \exp[r(1-P_t/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
+$P_{t+1}=F(P_t,r,K)=P_t \exp[r(1-P_t/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
 growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
-The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ and considered 
+The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ for a positive $r$ and considered 
 as the extinction equilibrium 
@@
 or can approach $P_e = K$ in an oscillating fashion if $1<r<2$. When 
-$r>2$, system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
+$r>2$ (i.e., ${F}'<-1$), system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
 as $r_c$, the critical point where system goes from under- to over-compensatory regime and 
@@
 
-**Hassell model** [@hassell1975density], motivated for modelling insect populations,
-can be represented as $P_{t+1} = rP_t/(1+aP_t)^b$, where $r>0$ is the intrinsic growth rate 
+**Hassell model** [@hassell1975density; @may1976bifurcations], motivated for modelling insect populations,
+can be represented as $P_{t+1} = F(P_t,r,a,b)=rP_t/(1+aP_t)^b$, where $r>0$ is the intrinsic growth rate 
 of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
@@
 
-**Maynard Smith model**
-
-**Pennycuick model**
+**Maynard Smith model** [@cohen1995unexpected; @maynard1978models; @may1976bifurcations] follows the dynamics 
+$P_{t+1}=F(P_t,r,a,b)=rP_t/[1+(aP_t)^b]$. $r$ is the growth rate, $a, b$ are the parameters with positive values.
+Equilibriums are: $P_e=0$(as expected extinction equilibrium) and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
+For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$. This $P^\star$ will be stable
+in under-compensatory regime for $1<r<r_c(=b/(b-1))$ and in over-compensatory regime for $r_c<r<b/(b-2)$.
+$P^\star$ will become unstable when $r>r_{bf}(=b/(b-2))$.
+
+**Pennycuick model** [@pennycuick1968computer; @may1976bifurcations] was 
+initially developed
+to follow change in age-structured total population
+level which assumed
+to have a constant sex 
+ratio and therefore change in female population was 
+expected through the change 
+in total population density. 
+A two population variant could also be possible
+in prey-predator perspective as indicated in [@pennycuick1968computer].
+In general, model dynamics is expressed as $P_{t+1}=F(P_t,r,a,b)=rP_t/[1+\exp \left\{-a(1-P_t/b)\right\}]$, where 
+$r$ is the intrinsic growth rate and $a>0,b>0$ are the model parameters. Model
+equilibriums are 
+$P_e=0$ (extinction equilibrium) and $P_e=P^\star=\frac{b}{a}[a+\ln{(r-1)}]$. For 
+a finite equilibrium population 
+$r$ must be greater than $r_{min}$, where 
+$r_{min} = 1+\exp(-a)$. The critical transition point $r_c$ (from under- to 
+over-compensatory dynamics) and Hopf-bifurcation point $r_{bf}$ can be obtained
+by numerical solution of ${F}'=0$ and ${F}'=-1$ at $P_e=P^{\star}$, respectively.
 

### 2019-03-16 SuppMat.Rmd
modified Verhulst with K=1

@@
 numlocs<-5
-p0<-1
+K<-1
+p0<-K
+ext_thrs<-p0/10
 
@@
     r<-rlist[i]
-    params<-r
-    ext_thrs<-p0/10
+    params<-c(r,K)
 
@@
 for(r in rlist){
-    params<-r
-    ext_thrs<-p0/10
+    params<-c(r,K)
 
@@
 ```{r fig_verhulst_scl_0.6, echo=F, results="hide",cache=T, cache.extra=list(seed,res_verhulst_local,res_verhulst_global)}
+
 pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
@@
 ```{r fig_verhulst_extrisk_d0_scl_0.6, echo=F,results="hide"}
+
 set.seed(seed)
@@
        x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Verhulst",paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("bottomright", c("Verhulst",paste0("K =1, sd =",scl)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
 mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(paste("(K = 1, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
 x<-res_verhulst_local
@@
 mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(paste("(K = 1, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
 x<-res_verhulst_global
@@
                           "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_t^{i}/b)\\}]$",
-                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i})]$",
+                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i}/K)]$",
                           "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(K-P_t^{i})\\{1-\\exp(-sP_t^{i})\\}]$"
@@
                            "a = 0.1, b= 0.1, sd = 1",
-                           "sd = 0.6",
+                           "K = 1, sd = 0.6",
                            "s = 0.2, K = 50, sd = 0.6"
@@
                         "$[b\\{a+ln(r-1)\\}]/a$",
-                        "1",
+                        "K",
                         "K"
@@
 
-<!--
-# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}
-A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
-(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
-We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
-(see `iTau` function from `copula` package).-->
+<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
+\begin{figure}[!h]
+\begin{center}
+\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
+\vspace{-0.2 cm}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
+\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
+\vspace{-0.2 cm}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
+\caption{Extinction risk from single patch age structured models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
+$(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$).
+All copula has same Kendall correlation $\tau = 0.6$.
+\label{fig_extrisk_agestr_C_SC_G_SG_linear}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG :  summary : all nonlinear model results for global dispersion : extrisk vs. d-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp.pdf}
+\caption{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
+Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Ricker model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
+numsims=10000, sd=1. \label{fig_ricker_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Ricker model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
+numsims=10000, sd=1. \label{fig_ricker_risk_global_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Hassell model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
+numsims=10000, sd=1. \label{fig_hassell_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Hassell model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
+numsims=10000, sd=1. \label{fig_hassell_risk_global_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Maynard smith model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
+numsims=10000, sd=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Maynard smith model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
+numsims=10000, sd=0.8. \label{fig_msmith_risk_global_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Pennycuick model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
+numsims=10000, sd=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Pennycuick model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
+numsims=10000, sd=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Verhulst model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Verhulst model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Austin brewer model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Austin brewer model, global dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
+numsims=10000, sd=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
 
@@
 \clearpage
-
-<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
-\begin{figure}[!h]
-\begin{center}
-\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
-\vspace{-0.2 cm}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
-\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
-\vspace{-0.2 cm}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk from single patch age structured models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
-$(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$).
-All copula has same Kendall correlation $\tau = 0.6$.
-\label{fig_extrisk_agestr_C_SC_G_SG_linear}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG :  summary : all nonlinear model results for global dispersion : extrisk vs. d-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
-Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Ricker model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-numsims=10000, sd=1. \label{fig_ricker_risk_local_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Ricker model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
-numsims=10000, sd=1. \label{fig_ricker_risk_global_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Hassell model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-numsims=10000, sd=1. \label{fig_hassell_risk_local_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Hassell model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
-numsims=10000, sd=1. \label{fig_hassell_risk_global_disp_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Maynard smith model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-numsims=10000, sd=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Maynard smith model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
-numsims=10000, sd=0.8. \label{fig_msmith_risk_global_disp_scl_0.8}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Pennycuick model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-numsims=10000, sd=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Pennycuick model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
-numsims=10000, sd=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Verhulst model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Verhulst model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Austin brewer model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Austin brewer model, global dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
-\end{center}
-\end{figure}

### 2019-03-16 1var_pop_models.R

@@ -86,24 +86,27 @@ Msmith(r=r,a=a,b=b,p0=K_e-0.1,lensim=200) #see changes as r=5,a=0.5 and vary b :
 #--------------------------------------------------------------------------------------------------
 
 # Exploring Verhulst model
-Verhulst<-function(r,p0,lensim){
+Verhulst<-function(r,K,p0,lensim){
   pop<-c(p0)
   time<-c(0)
   # po<-rep(K,r)
   # print(po)
   
   for(it in c(1:lensim)){
-    pt<-p0*(1+r*(1-p0))
+    pt<-p0*(1+r*(1-(p0/K))) 
     time<-c(time,it)
     pop<-c(pop,pt)
     p0<-pt
   }
   
-  plot(time,pop,type="l")
+  plot(time,pop,type="b")
 }
 
 #call the function
-Verhulst(r=1.8,p0=0.4,lensim=400) # eqm. point =1
+K<-2
+r<-1.5
+p0<-K-0.1
+Verhulst(r=r,K=K,p0=p0,lensim=100) # eqm. point =1
 
 #-------------------------------------------------------------
 # Exploring pennycuick model

### 2019-03-16 mydispersal.R

@@ -78,7 +78,8 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
       res[,,tct+1]<-lam_ricker*lam_sto #numsims by numlocs matrix
     }else if(model=="verhulst"){
       r<-params[1]
-      lam_verhulst<-(1+(r*(1-res[,,tct])))*res[,,tct]
+      K<-params[2]
+      lam_verhulst<-(1+(r*(1-(res[,,tct]/K))))*res[,,tct]
       res[,,tct+1]<-lam_verhulst*lam_sto #numsims by numlocs matrix
     }else if(model=="hassell"){
       r<-params[1]

### 2019-03-17 MainText.Rmd
methods written

@@
 for which the model equilibrium $P_e$ had always a finite, positive initial 
-population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ of each 
+population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ from each 
 model where metapopulation could disperse locally with varying dispersal 
@@
 # Literature cited
+
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{1pt}
+\noindent

### 2019-03-17 SuppMat.Rmd
methods written

@@
 
-**Verhulst model**
-
-**Austin-Brewer model**
+**Verhulst model** [@cohen1995unexpected; @may1976bifurcations] is a discrete time difference 
+representation of logistic growth of non-human biological population and can be written 
+as $P_{t+1} = F(P_t,r,K) =P_t[1+r(1-P_t/K)]$, wirh $r>0$ and $K>0$ being growth rate and carrying capacity, respectively. 
+Model equilibrium $P_e=0$ is always unstable whereas $P_e=K$ is stable below Hopf-bifurcation 
+point $r_{bf}(=2)$. Model at stable equilibrium ($P_e=K$) lies in the under-compensatory regime when $0<r<r_c$
+and in over-compensatory regime when $rc<r<r_{bf}$ with $r_c$ being equal to 1.
+
+**Austin-Brewer model** [@austin1971world; @cohen1995unexpected], unlike from the first
+five models we considered,
+is used to model growth of human population. It follows the
+discrete dynamics : $P_{t+1} = F(P_t,r,K,s)= P_t[1+r(K-P_t) \left\{ 1-\exp (-sP_t)\right\}]$, where
+$r, K, s$ are positive, representing growth rate, carrying capacity 
+and model parameter, respectively.
+Usual extinction equilibrium is $P_e=0$. The other
+finite, positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. The equilibrium $P^\star$
+is stable when $r<r_{bf}$ and unstable beyond the bifurcatio point $r_{bf} = \frac{2}{K[1-\exp (-sK)]}$.
+After perturbation, model can return back monotonically to its stable equilibrium if parameter $r$ lies in the 
+under-compensatory regime ($0<r<r_c$), where $r_c = \frac{1}{K[1-\exp (-sK)]}$. In the 
+over-compensatory regime ($r_c<r<r_{bf}$), model approaches to stable equilibrium $P^\star$
+in oscillatory fashion.
+
+\clearpage
 
 # References 
-
-\clearpage
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{1pt}
+\noindent

### 2019-03-19 MainText.Rmd
keywords added
plot changed notation sigma for standard deviation

@@
 
-\textbf{\textit{Keywords:}}
+\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, copula, density-dependence model
 
@@
 correlation ($\tau = 0.6$) and underlying tail-dependence. 
-Similarly, for extreme-tail dependence
-(either left or right tail) a matrix M with two column was generated as described in 
+Similarly, for extreme-tail dependent bivariant copula
+(either left or right tail) a matrix M was generated from standard iid as described in 
 Appendix \ref{SM-extr_tail_dep}.
@@
 
-The environmental noises from any two patch, say, 
+The environmental noises from any two out of $n$ patches, say, 
 $(\epsilon_t^i, \epsilon_t^j)$ were normally 
-distributed within each patch with mean $m=0$ and standard deviation $sd$, 
-(see the specified value $sd$ for each model in 3$^{rd}$ 
-column of Table \ref{tab_model_summary}), were 
+distributed within each patch with mean $m=0$ and standard deviation $\sigma$ 
+(see the specified value $\sigma$ for each model in 3$^{rd}$ 
+column of Table \ref{tab_model_summary}). Noises were 
 independent through time, and showed the same 
-spatial correlations between patches (to within sampling error) for every simulation. 
-But they were made to exhibit extreme right- or extreme left-tail dependence between patches in
+spatial correlations between patches (to within sampling error) for every simulation, 
+but they were made to exhibit extreme right- or extreme left-tail dependence between patches in
 different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
@@
 To test our hypothesis from (Q2) of \nameref{intro} for each metapopulation model,
-we estimated extinction risk for a wide range of parameter $r$ 
+we estimated extinction risk for a wide range of parameter $r$, 
 which covered the undercompensatory and overcompensatory regime. We 
@@
              format="latex",digits=3, align="c",table.env='table*',
-             caption = "Summary of six density dependence metapopulation models considered; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$) where deterministic analog of models showed transition from under- to over-compensatory dynamics and bifurcation parameter ($r_{bf}^{i}$) where $P_e^{i}$ became unstable from stable steady state, respectively.
+             caption = "Summary of six density dependence metapopulation models considered at each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
 Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"
-             )%>%column_spec(c(1,3,5,6),width="1.5 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
+             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
 ```

### 2019-03-19 SuppMat.Rmd
plot changed notation sigma for standard deviation

@@
 arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025right,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Ricker",paste0("K = ",K),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
-#legend("bottomright", c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=F, bty='n',cex=1.5)
+legend("topleft", legend=c("Ricker"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("K = ",K),side = 3, line=-16, adj=0.95, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", legend=c("Ricker",paste0("K = ",K)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
 arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025right,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Hassell",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("topleft", legend=c("Hassell"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Hassell",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
 arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Maynard Smith",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("topleft", legend=c("Maynard Smith"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
 par(op)
@@
 arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025right,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Pennycuick",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("topleft", legend=c("Pennycuick"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Pennycuick",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
        x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Verhulst",paste0("K =1, sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("topleft", legend=c("Verhulst"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("K = ",K),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Verhulst",paste0("K =1, sd =",scl)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
        x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
-legend("bottomright", c("Austin-Brewer",paste0("s = ",s," , K = ",K),paste0(" sd =",scl)), horiz=F, bty='n',cex=1.8)
+legend("topleft", legend=c("Austin-Brewer"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("s = ",s,", K = ",K),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Austin-Brewer",paste0("s = ",s," , K = ",K),paste0(" sd =",scl)), horiz=F, bty='n',cex=1.8)
 par(op)
@@
 mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+#mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 getplot(x=x,rlist = rlist)
@@
 mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 0.5, b = 100, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_hassell_local
@@
 mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 0.5, b = 4, sd = 0.8)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_msmith_local
@@
 mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 0.1, b = 0.1, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_pennycuick_local
@@
 mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(K = 1, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_verhulst_local
@@
 mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(s = 0.2, K = 50, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_abrewer_local
@@
 mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 getplot(x=x,rlist = rlist)
@@
 mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 0.5, b = 100, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_hassell_global
@@
 mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 0.5, b = 4, sd = 0.8)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_msmith_global
@@
 mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(a = 0.1, b = 0.1, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_pennycuick_global
@@
 mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(K = 1, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_verhulst_global
@@
 mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
-mtext(paste("(s = 0.2, K = 50, sd = 0.6)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
 x<-res_abrewer_global
@@
 
-dt1 <- tibble("Models $(P_{t+1}^{i} = \\lambda_t^{i} P_t^{i})$"= c("Ricker",
+dt1 <- tibble("Models"= c("Ricker",
            "Hassell",
@@
 
-dt3<-tibble("\\hspace{-0.1 cm} Parameters used" = c("K = 50, sd = 1",
-                            "a = 0.5, b = 100, sd = 1",
-                            "a = 0.5, b = 4, sd = 0.8",
-                           "a = 0.1, b= 0.1, sd = 1",
-                           "K = 1, sd = 0.6",
-                           "s = 0.2, K = 50, sd = 0.6"
+dt3<-tibble("\\hspace{-0.1 cm} Parameters used" = c("K = 50, $\\sigma$ = 1",
+                            "a = 0.5, b = 100, $\\sigma$ = 1",
+                            "a = 0.5, b = 4, $\\sigma$ = 0.8",
+                           "a = 0.1, b= 0.1, $\\sigma$ = 1",
+                           "K = 1, $\\sigma$ = 0.6",
+                           "s = 0.2, K = 50, $\\sigma$ = 0.6"
 ))
@@
 for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-numsims=10000, sd=1. \label{fig_ricker_risk_local_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_ricker_risk_local_disp_scl_1}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
-numsims=10000, sd=1. \label{fig_ricker_risk_global_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_ricker_risk_global_disp_scl_1}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-numsims=10000, sd=1. \label{fig_hassell_risk_local_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
-numsims=10000, sd=1. \label{fig_hassell_risk_global_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_hassell_risk_global_disp_scl_1}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-numsims=10000, sd=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
+numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
-numsims=10000, sd=0.8. \label{fig_msmith_risk_global_disp_scl_0.8}}
+numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_global_disp_scl_0.8}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-numsims=10000, sd=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
-numsims=10000, sd=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
-numsims=10000, sd=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
 \end{center}
@@
 matrix M can be generated initially with independent random draws from a normal 
-distribution with a specified mean ($m$) and standard deviation $(sd)$.
+distribution with a specified mean and standard deviation.
 

### 2019-03-20 MainText.Rmd
add guidelines for intro and results

@@
 
+<!--Background for extinction risk in ecological model-->
+
+<!--How extinction risk is related to extreme events : past studies-->
+
+<!--How copula could be relevent in estimating extreme event's influence on extinction risk : introduce Taildep. concept-->
+
+<!--Tail dependence of copula and recent result from BIVAN showing LTdep with higher extrisk than UTdep, but 
+not explored age structured or in metapopulation context with density dependence effect-->
+
+<!--our objective ?-->
 We explored the potential of asymmetric tail-dependence in copula 
@@
 
+<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula-->
+
+<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula-->
+
+<!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->
+
+<!--Q3 : metapopulation results : switching near rc for d=0-->
+
 # Discussion
@@
  <!--FIG : age structure model results : linear, nonlinear-->
-
 \begin{figure*}[!h]
@@
              format="latex",digits=3, align="c",table.env='table*',
-             caption = "Summary of six density dependence metapopulation models considered at each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
+             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
 Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",

### 2019-04-03 SuppMat.Rmd
add guidelines for intro and results
modified documentation

@@
 
-```{r get_rc_pennycuick,echo=F,results="hide",cache=T,cache.extra=list(seed,mtime("1var_pop_model_bifurcation.R"))}
+```{r get_rc_pennycuick,echo=F,results="hide",cache=T,cache.extra=list(seed,mtime("bifurcation_pennycuick.R"))}
 set.seed(seed)
-source("./1var_pop_model_bifurcation.R")
+source("./bifurcation_pennycuick.R")
 a_pennycuick<-0.1
@@
 
+<!--Supp. FIG : Hassell model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
+numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Maynard smith model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
+numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Pennycuick model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
+numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Verhulst model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
+<!--Supp. FIG : Austin brewer model, local dispersal-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
+\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+\end{center}
+\end{figure}
+
 <!--Supp. FIG : Ricker model, global dispersal-->
@@
 
-<!--Supp. FIG : Hassell model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1}}
-\end{center}
-\end{figure}
-
 <!--Supp. FIG : Hassell model, global dispersal-->
@@
 
-<!--Supp. FIG : Maynard smith model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
-\end{center}
-\end{figure}
-
 <!--Supp. FIG : Maynard smith model, global dispersal-->
@@
 
-<!--Supp. FIG : Pennycuick model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
-\end{center}
-\end{figure}
-
 <!--Supp. FIG : Pennycuick model, global dispersal-->
@@
 
-<!--Supp. FIG : Verhulst model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
 <!--Supp. FIG : Verhulst model, global dispersal-->
@@
 numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
-\end{center}
-\end{figure}
-
-<!--Supp. FIG : Austin brewer model, local dispersal-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
 \end{center}

### 2019-04-03 bifurcation_pennycuick.R

@@ -1,0 +1,14 @@
+# This function evaluates 1st order derrivative at equilibrium for pennycuick model
+# Args
+#     params : a vector containing model parameters
+#     model : a character with model name (default = "pennycuick")
+fn_d1_eqm<-function(params,model="pennycuick"){
+  fn<-expression((r*p)/(1+exp(-a*(1-(p/b)))))
+  fn_d1<-D(fn,"p")
+  r<-params[1]
+  a<-params[2]
+  b<-params[3]
+  p_e<-(b/a)*(a+log(r-1))
+  res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
+  return(res)
+}

### 2019-04-03 mydispersal.R

@@ -1,16 +1,7 @@
 #--------------------------------------------------------------------------
 source("ExtremeTailDep.R")
 #--------------------------------------------------------------------------
-#The simulator of the multi-location model.
-#
-#Args
-#p0       A length numlocs vector holding initial populations
-#ns       An numsims by numlocs by numsteps array of the epsilons, 
-#           where numsteps is the number of time steps you want
-#
-#Output
-#A numsims by numlocs by numsteps+1 array of populations
-#
+
 #-------------------------------------------------------------------------------
 # A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
 # Input :

### 2019-04-03 1var_pop_model_bifurcation.R
code not used

@@ -1,157 +1,0 @@
-# Bifurcation diagram for 1variable model
-#                    Ricker model : p(t+1)=p(t)*exp(r(1-(p(t)/K))) # rc=1
-#                    Hassell model : p(t+1)=r*p(t)/(1+(a*p(t)))^b  # rc=(b/(b-1))^b
-#                    Maynard-smith model : p(t+1)=(r*p(t))/(1+((a*p(t))^b)) # rc=b/(b-1)
-#                    Pennycuick model : p(t+1)=(r*p(t))/(1+exp(-a*(1-(p(t)/b)))) # rc calculate numerically
-
-
-fn_d1_eqm<-function(params,model){
-  if(model=="ricker"){
-    fn<-expression(p*exp(r*(1-(p/K))))
-    fn_d1<-D(fn,"p")
-    r<-params[1]
-    K<-params[2]
-    p_e<-K
-    res<-eval(fn_d1,list(r=r,K=K,p=p_e))
-  }
-  
-  if(model=="hassell"){
-    fn<-expression(r*p/(1+(a*p))^b)
-    fn_d1<-D(fn,"p")
-    r<-params[1]
-    a<-params[2]
-    b<-params[3]
-    p_e<-((r^(1/b))-1)/a
-    res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
-  }
-  
-  if(model=="msmith"){
-    fn<-expression((r*p)/(1+((a*p)^b)))
-    fn_d1<-D(fn,"p")
-    r<-params[1]
-    a<-params[2]
-    b<-params[3]
-    p_e<-((r-1)^(1/b))/a
-    res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
-  }
-  
-  if(model=="pennycuick"){
-    fn<-expression((r*p)/(1+exp(-a*(1-(p/b)))))
-    fn_d1<-D(fn,"p")
-    r<-params[1]
-    a<-params[2]
-    b<-params[3]
-    p_e<-(b/a)*(a+log(r-1))
-    res<-eval(fn_d1,list(r=r,a=a,b=b,p=p_e))
-  }
-  
-  if(model=="malthus"){
-    fn<-expression(r*p*(K-p+(L*log(p))))
-    fn_d1<-D(fn,"p")
-    r<-params[1]
-    K<-params[2]
-    L<-params[3]
-    p_e<-params[4]
-    res<-eval(fn_d1,list(r=r,K=K,L=L,p=p_e))
-  }
-  
[…]

### 2019-04-03 1var_pop_models.R

@@ -1,247 +1,0 @@
-# Exploring Ricker model
-Ricker<-function(r,K,p0,lensim){
-  pop<-c(p0)
-  time<-c(0)
-  # po<-rep(K,r)
-  # print(po)
-  
-  for(it in c(1:lensim)){
-    pt<-p0*exp(r*(1-(p0/K)))
-    time<-c(time,it)
-    pop<-c(pop,pt)
-    p0<-pt
-  }
-  
-  plot(time,pop,type="b")
-  
-}
-
-#call the function
-Ricker(r=1.1,K=50,p0=5,lensim=1000) # bifurcation point r=1
-#-------------------------------------------------------
-
-# Exploring Hassell model
-# help: https://jmahaffy.sdsu.edu/courses/s00a/math121/lectures/qual_discrete/qualdiscrete.html#Hassell
-#       https://www.jstor.org/stable/pdf/3863.pdf?refreqid=excelsior%3A348907abb1bca7f039109d94db5667b5
-#       http://lab.rockefeller.edu/cohenje/PDFs/231CohenUnexpectedDominanceHighFreqChaoticNonlinPpnModelsLet.pdf
-
-Hassell<-function(r,a,b,p0,lensim){
-  pop<-c(p0)
-  time<-c(0)
-  
-  for(it in c(1:lensim)){
-    pt<-(r*p0)/((1+(a*p0))^b)
-    time<-c(time,it)
-    pop<-c(pop,pt)
-    p0<-pt
-  }
-  
-  r1<-((b/(b-1)))^b
-  r2<-((b/(b-2)))^b
-  plot(time,pop,type="b")
-  return(pop)
-  
-  
-  #provided b>2 (to get finite r2), these are the stability conditions:
-  cat("For stable, monotonic approach towards eqm. upper limit of r should be",r1,"\n")
-
-  cat("For stable, oscillatory approach towards eqm. upper limit of r should be within ",r1,"to",r2,"\n")
-  
-}
-
-#call the function
-r<-2.72
-a<-0.5
-b<-100
-K_e<-((r^(1/b))-1)/a
-K_e
-x<-Hassell(r=r, a=a, b=b, p0=K_e-0.01, lensim=50) #b~1.5 : contest, b~100 : scramble
-diff(x)
[…]

### 2019-04-03 MainText.Rmd
minor edits

@@
 <!--Tail dependence of copula and recent result from BIVAN showing LTdep with higher extrisk than UTdep, but 
-not explored age structured or in metapopulation context with density dependence effect-->
+not explored in age structured model or in metapopulation context with density dependence effect-->
 
 <!--our objective ?-->
-We explored the potential of asymmetric tail-dependence in copula 
+We explored the potential use of asymmetric tail-dependence in copula 
 for understanding the extinction risk of populations towards extreme climatic event in ecology
-by addressing the following specific questions: (Q1) Does nature of tail dependence in vital rates
+by addressing the following questions: (Q1) Does nature of tail dependence in vital rates
 of a single-patch age-structured population model affect extinction risk of that population differently?
 We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
-(Q2) How does asymmetric tail-dependent noise across multiple patches influence extinction 
-risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
+(Q2) How does asymmetric tail-dependent noises across multiple patches influence extinction 
+risk of a nonlinear metapopulation model with density dependence? Our speculation is that the general finding 
 from density-independent metapopulation model [BIVAN paper results of Lewinton-Cohen model]
-(higher extinction risk if noise across patches were 
-left-tail dependent than the one if noise came from right-tail dependent) holds true only in
+(higher extinction risk if noises across patches were 
+left-tail dependent than the one if noises came with right-tail dependence) holds true only in
 the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
@@
 populations in a single patch age-structured model,
-we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
+we considered a basic stochastic age-structured model (Eq. \ref{eq_age_str}) 
 with two age classes (egg $E$ and adult $A$). We assumed
@@
 and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
-a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distribution 
+a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distributions 
 could be generated from iid noises showing specific tail-dependence (left or right tail)

### 2019-04-05 MainText.Rmd
minor edits

@@
 with $10000$ simulations. For each simulation, we started with initial population $(E,A)_{t=0} = (5,10)$
-and following Eq. \ref{eq_age_str} after each time step checked the total population $(E+A)_{t}$
+and following Eq. \ref{eq_age_str} we checked whether after each time step the total population $(E+A)_{t}$
 was below a certain extinction threshold (we set it as $2$) or not. If it was below that threshold then we 
@@
 
-<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula-->
+<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula : Fig. S1-->
 
-<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula-->
+<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula : Fig. 1 -->
 
@@
 
-<!--Q3 : metapopulation results : switching near rc for d=0-->
+<!--For local dispersal : all model summary : Fig. 2, individual model : Fig. S3-S8-->
+
+<!--For global dispersal : all model summary : Fig. S2, individual model : Fig. S9-S14-->
+
+<!--Q3 : metapopulation results : switching near rc for d=0, Fig. 3-->
 

### 2019-04-06 MainText.Rmd
started writing results

@@
 To test our hypothesis from (Q1), we repeated the above analysis 
-considering the vital rates coming from a 
-an asymmetric copula with specific 
+considering the vital rates coming from an asymmetric copula with specific 
 left (Clayton, Survival Gumbel)
@@
 <!--Q1 : age structured results : linear and nonlinear model with asymmetric copula : Fig. S1-->
+<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula : Fig. 1 -->
 
-<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula : Fig. 1 -->
+For single patch linear age-structured models, we found left-tail 
+dependence (Clayton, Suvival Gumbel copulas) in noisy vital 
+rates enhanced the extinction risk compared to the case when
+vital rates were right tail dependent (Survival clayton, Gumbel copulas).
+It was expected because a sudden "bad" year (e.g. with severe cold) can suppress 
+the fecundity of adult and survival/ maturation of eggs simultaneously 
+and thereby increasing the extinction risk of the population. For non-linear age-structured 
+models, we got the similar trend for extinction risks with different tail-dependence (left and right).
+With non-linear model the extinction risks appeared as higher than their corresponding 
+linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
+Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear} represents this scenario with 4 different 
+asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
+difference between extinction risks within a pair of different tail-dependent copulas
+(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear}(A,C) and
+G-SG in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear}(B,D)) and that difference was partly 
+dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
+vital rates coming from extreme-tail dependent (left and right in two different simulations)
+copulas and got the similar results but with much prominent difference between their extinction risks
+(Fig. \ref{fig_extrisk_agestr}). This result answered (Q1) and showed extinction risk of 
+a population could be considered as
+direct manifestation of tail-dependence measure between vital rates of different age-class
+within that population. <!--discussion later : add Tuljapurkar's approximation-->
 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models, fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
+\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models
+against time (shading represents 95\% confidence interval).
+Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
 \label{fig_extrisk_agestr}}

### 2019-04-06 SuppMat.Rmd
started writing results

@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk from single patch age structured models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
+\caption{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time; fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure.
+$(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
 $(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$).

### 2019-04-07 MainText.Rmd
started writing results for metapopulations

@@
 
-# Introduction\label{intro}
+# Introduction \label{intro}
 
@@
 <!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->
+We found density-dependence metapopulation models with $n=5$ patches in noisy 
+environment showed similar patterns for extinction risk as density-independent 
+metapopulation model showed in earlier studies [BIVAN paper]. For each density-dependent 
+model listed in Table \ref{tab_model_summary}, we observed higher (or lower) extinction risk for 
+the metapopulations when noises across multiple patches were left- (or right) tail dependent.
+But that observation remained valid when the model parameter, usually growth rate $(r)$, of each model was 
+sufficiently low so that under-compensatory dynamics (monotonous approach towards
+its stable equilibrium) for the deterministic analog of the model was ensured. When the model 
+was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
+value of $r$, we observed right (or left) tail dependent noise caused higher (or lower) extinction 
+risk for the metapopulations. All models were consistent with our prediction (see Q2 from \nameref{intro}) 
+across local and global dispersal (Figs. \ref{fig_extrisk_allmodel_localdisp} and 
+\ref{SM-fig_extrisk_allmodel_globaldisp}).
 

### 2019-04-18 MainText.Rmd
Results for Q2 written

@@
 <!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->
-We found density-dependence metapopulation models with $n=5$ patches in noisy 
+<!--For local dispersal : all model summary : Fig. 2, individual model : Fig. S3-S8-->
+<!--For global dispersal : all model summary : Fig. S2, individual model : Fig. S9-S14-->
+
+We found density-dependence metapopulation models (with $n=5$ patches) in noisy 
 environment showed similar patterns for extinction risk as density-independent 
@@
 value of $r$, we observed right (or left) tail dependent noise caused higher (or lower) extinction 
-risk for the metapopulations. All models were consistent with our prediction (see Q2 from \nameref{intro}) 
-across local and global dispersal (Figs. \ref{fig_extrisk_allmodel_localdisp} and 
-\ref{SM-fig_extrisk_allmodel_globaldisp}).
-
-<!--For local dispersal : all model summary : Fig. 2, individual model : Fig. S3-S8-->
-
-<!--For global dispersal : all model summary : Fig. S2, individual model : Fig. S9-S14-->
+risk for the metapopulations. Figs. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
+against local dispersal rate ($d$) for six density dependent metapopulation models 
+where growth rate $r$ was taken either from under-compensatory or from over-compensatory 
+regime. We also got similarly consisent results with 
+our prediction (see Q2 from \nameref{intro}) for global 
+dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
+Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6}) 
+shows how extinction risk changed against local dispersal rate for a whole range
+of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
+represents the same but when global dispersal took place between patches. 
+Both case showed the signature of 
+"rescue-effect" (i.e. by lowering extinction risk) when there was 
+a finite dispersal compared to the one with 
+$d=0$.
 
 <!--Q3 : metapopulation results : switching near rc for d=0, Fig. 3-->
+To examine (Q3), we evaluated extinction risk for each metapopulation model for a wide
 

### 2019-04-18 MainText.Rmd
results added as of now

@@
 
-# Methods
+# Methods \label{methods}
 
@@
 <!--Q3 : metapopulation results : switching near rc for d=0, Fig. 3-->
-To examine (Q3), we evaluated extinction risk for each metapopulation model for a wide
+Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) we hypothesized the switching 
+between the extinction risk generated by left-tail dependent and right-tail dependent noise 
+occurred near the critical parameter $r_c$ (see Q3). For simplicity we only consider 
+$d=0$ case, i.e., when there was no dispersal between patches (see \nameref{methods}). 
+Fig. \ref{fig_extrisk_d0} shows higher extinction risk with left-tail dependent noise 
+for $r<r_c$ and with right-tail dependent noise for $r>r_c$. $r_c$ was the vertical
+line in Fig. \ref{fig_extrisk_d0} indicating the value of $r$ for each model where 
+their deterministic analog switches from under- to over-compensatory dynamics. 
+Though the switching visually appered far from the theoretical $r_c$ value for Verhulst 
+and Austin-Brewer model, they were consistent with our hypothesis as they had comperatively 
+smaller range of $r$ than the other models.
 

### 2019-07-13 MainText.Rmd
started writing intro

@@
 ---
-title: "Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology"
-author: "Shyamolina Ghosh"
+title: "Extreme climatic events with asymmetric tail-dependence affect population-extinction risk"
+author: "Shyamolina Ghosh, Daniel Reuman"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+geometry: "left=1in,right=1in,top=1in,bottom=1in"
 output: 
@@
 
-\noindent\makebox[\linewidth]{\rule{\columnwidth}{1pt}}
+\noindent \emph{Affiliations:}
 
-\scriptsize
+\noindent Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
 
-\textbf{ABSTRACT:}  \small write abstract here
+\noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
+
+\newpage
+
+\noindent \textbf{ABSTRACT} 
+<!--200 words limit : J Am Nat-->
 
@@
 
-\noindent\makebox[\linewidth]{\rule{\columnwidth}{1pt}}
-
-\normalsize
+\newpage
 
@@
 
-<!--Background for extinction risk in ecological model-->
+<!--how common are the extreme events? 
+what are the implication of extreme events on life history? 
+any negative or positive effects? : past literatures-->
+For past few decades, there have been evidences that anthropogenic climate change 
+is causing more frequent extreme events [ref] (e.g. temperature or precipitation extremes,
+heat-waves, wild-fire) which show shifts in its mean, 
+in the variability of the distribution, as well as changes in its 
+symmetry or skewness. Such extreme climatic events (ECE) can pose severe
+impacts at the individual, population, community and ecosystem level (e.g. Parmesan,
+Root & Willig 2000; van de Pol et al. 2010b; Hoover,
+Knapp & Smith 2014). Therefore, understanding the long-term impacts of changes 
+in ECEs has been a longstanding interest in ecological research, specically in modelling 
+extreme-risks [ref:Risk Analysis, Vol. 32, No. 11, 2012].
 
-<!--How extinction risk is related to extreme events : past studies-->
+<!--How extinction risk is related to extreme events : past studies
+consequences of extreme events on population life history, community-->
 
@@
 risk of a nonlinear metapopulation model with density dependence? Our speculation is that the general finding 
-from density-independent metapopulation model [BIVAN paper results of Lewinton-Cohen model]
+from density-independent metapopulation model [@Ghosh_copula]
 (higher extinction risk if noises across patches were 
-left-tail dependent than the one if noises came with right-tail dependence) holds true only in
+left-tail dependent than the one if noises had right-tail dependence) holds true only in
 the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 

### 2019-07-13 SuppMat.Rmd
started writing intro

@@
 ---
-title: "Supporting Information : Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology"
-author: "Shyamolina Ghosh"
+title: "Supporting Information : Extreme climatic events with asymmetric tail-dependence affect population-extinction risk"
+author: "Shyamolina Ghosh, Daniel Reuman"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
+geometry: "left=1in,right=1in,top=1in,bottom=1in"
 

### 2019-07-15 MainText.Rmd
writing intro

@@
 
-<!--how common are the extreme events? 
-what are the implication of extreme events on life history? 
-any negative or positive effects? : past literatures-->
-For past few decades, there have been evidences that anthropogenic climate change 
-is causing more frequent extreme events [ref] (e.g. temperature or precipitation extremes,
-heat-waves, wild-fire) which show shifts in its mean, 
-in the variability of the distribution, as well as changes in its 
+<!--how common are the extreme events?-->
+For past few decades, there have been increasing evidences that anthropogenic climate change 
+is causing extreme events, e.g. temperature or precipitation extremes,
+heat-waves, wild-fire etc. [@ummenhofer2017extreme]. Frequent extreme events 
+can shift temperature or precipitation-mean, 
+change variability of the distribution, as well as change its 
 symmetry or skewness. Such extreme climatic events (ECE) can pose severe
-impacts at the individual, population, community and ecosystem level (e.g. Parmesan,
-Root & Willig 2000; van de Pol et al. 2010b; Hoover,
-Knapp & Smith 2014). Therefore, understanding the long-term impacts of changes 
-in ECEs has been a longstanding interest in ecological research, specically in modelling 
-extreme-risks [ref:Risk Analysis, Vol. 32, No. 11, 2012].
+impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
+@gutschick2003extreme; @jiguet2011community; 
+@hoover2014resistance]. Therefore, 
+understanding the long-term impacts of changes 
+in ECEs has been a longstanding interest in ecological research 
+[@katz2005statistics; @smith2011ecological; @diez2012will; @moritz2013future; @bailey2016tackling].
 
@@
 consequences of extreme events on population life history, community-->
+Earlier studies showed ECEs may affect species abundances (species re-ordering) through 
+life-history traits and this alteration in community structures could cascade 
+through higher hierarchical levels and lead to local extinction of 
+species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme]. For example, 
+a study based on annal survival data showed changes
+in prey communities induced by extreme flooding affected predators in 
+aquatic and adjacent riparian habitats [@chiu2013impact]. Similarly, 
+$43$ years of ring-recovery data was used to examine the demographic impact of ECEs 
+through temporal variability in juvenile, immature and adult survival
+of European shags on the Isle of May in eastern Scotland [@frederiksen2008demographic].
+A recent study [@mcdermott2017sensitivity] investigated the sensitivity 
+of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
+extreme precipitation, drought) over $37$ years (1976-2012) at the site level, across 
+each species' life-stages. Successive ECEs with short time intervals 
+could weaken the resistance of plant communities too [@dreesen2014successive]. 
+All these studies indicated that environmental stochasticity (through ECEs) 
+can lead to demographic stochasticity (through influence on reproductive or 
+mortality rate) of different stage class of a population making it vulnerable 
+to local extinction.
 

### 2019-07-15 MainText.Rmd
writing intro

@@
 impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
-@gutschick2003extreme; @jiguet2011community; 
+@gutschick2003extreme; @jiguet2011community; @mckechnie2009climate;
 @hoover2014resistance]. Therefore, 
@@
 can lead to demographic stochasticity (through influence on reproductive or 
-mortality rate) of different stage class of a population making it vulnerable 
+mortality rate) of different age class of a population making it vulnerable 
 to local extinction.
 
-<!--How copula could be relevent in estimating extreme event's influence on extinction risk : introduce Taildep. concept-->
+<!--How copula could be relevent in estimating extreme event's 
+influence on extinction risk : introduce Taildep. concept-->
+A recent study [@Ghosh_copula] showed the importance, causes 
+and consequences of copula structures in associations between ecologically
+important variables and demonstrated one example of density-independent metapopulation model where 
+copula was used to estimate extinction risk. Copula, a rank based measure, is much more informative 
+than linear correlation measure and a widely applied tool in dependence 
+modeling [@joe2014_dependence; @nelsen2006_copula]. We will hereby illaborate the idea why
+the potential use of copula in modelling extinction risk goes beyond the density-independent model 
+and even applicable for a simple age-structured population where its vital rates are influenced by ECEs. 
+<!--do I need to explain the details how to make a copula from raw data?-->
 
-<!--Tail dependence of copula and recent result from BIVAN showing LTdep with higher extrisk than UTdep, but 
-not explored in age structured model or in metapopulation context with density dependence effect-->
+<!--tail dep. of copula & extinction risk-->
+A major benefit of the copula approach is that it can effectively capture tail dependence. 
+Tail dependence, already exemplified in Fig..... (add Fig.), is associations between the more extreme 
+values of distributions. For two positively associated variables, left-tail 
+(respectively, right-tail) dependence is stronger association between
+values in the left or lower portions, i. e. smaller values 
+(respectively, the right or upper portions, i.e. higher values) 
+of the two distributions. For example, extreme cold winter during growing season
+might reduce the adult reproduction rate as well as larval survival (i.e. stronger 
+tail dependence between vital rates when temperature is too cold than normal), which 
+could lead to population crash and thereby inflate the extinction risk. Similarly, in context of 
+density-dependent metapopulation models, if extreme environmental fluctuations across sites 
+appear to be as synchronous (left tail dependent or right tail dependent between patches 
+with same correlation as in Fig.....) then it could wipe out some metapopulation patches 
+during unfavorable years unless they are rescued by the nearby populations. This scenario 
+could be more complex depending 
+on the dynamics of the model. Though tail-dependence 
+property of a copula has been used for risk-analysis in other fields like finance, 
+hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our knowledge
+copula have never been used in estimating the effect of ECEs through extinction risk of age structured 
+model or density-dependent metapopulations.
 
 <!--our objective ?-->
-We explored the potential use of asymmetric tail-dependence in copula 
+In this paper, we will explore the potential use of asymmetric tail-dependence in a copula 
 for understanding the extinction risk of populations towards extreme climatic event in ecology
-by addressing the following questions: (Q1) Does nature of tail dependence in vital rates
+by addressing the following questions: (Q1) Does nature of tail dependence 
+(left tail or right tail) in vital rates
 of a single-patch age-structured population model affect extinction risk of that population differently?
@@
 under-compensatory to over-compensatory dynamics.
+
+<!--Tuljapurkar estimation to be explained in Discussion-->
 
@@
 environment showed similar patterns for extinction risk as density-independent 
-metapopulation model showed in earlier studies [BIVAN paper]. For each density-dependent 
+metapopulation model showed in earlier studies [@Ghosh_copula]. For each density-dependent 
 model listed in Table \ref{tab_model_summary}, we observed higher (or lower) extinction risk for 

### 2019-07-15 MainText.Rmd
writing intro

@@
 important variables and demonstrated one example of density-independent metapopulation model where 
-copula was used to estimate extinction risk. Copula, a rank based measure, is much more informative 
+copula was used to estimate extinction risk. Copula, a rank based measure of association, is much more informative 
 than linear correlation measure and a widely applied tool in dependence 
-modeling [@joe2014_dependence; @nelsen2006_copula]. We will hereby illaborate the idea why
-the potential use of copula in modelling extinction risk goes beyond the density-independent model 
+modeling [@joe2014_dependence; @nelsen2006_copula]. We here introduce the concept of copula in a
+manner that follows [@Ghosh_copula]. Copulas separate the information content of a bivariate dataset,
+$(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
+distributions (which is not about the association between the variables) and the rest of 
+the information (which is solely about the association). The association information is revealed by 
+the copula plot ($v_n$ against $u_n$) of the raw variables ($y_n$ vs. $x_n$) in Fig...., where 
+$u_n$ is the rank of $x_n$ in the set ${x_1, x_2, \ldots , x_n }$, divided by $N + 1$; 
+and $v_n$ is the rank of $y_n$ in the set ${y_1, y_2, \ldots , y_n }$, also divided
+by $N + 1$. Here the rank of the smallest element of each sets is considered to be 1 and the rank of
+the largest element is $N$. We refer to the $u_n$ and $v_n$ as normalized ranks of the $x_n$ and $y_n$. 
+Ranking makes the marginal distributions uniform, isolating only the information on 
+association between the variables. Here, we will extend the application of copula 
+in estimating extinction risk as introduced by [@Ghosh_copula] on why the potential use of copula
+in modelling extinction risk goes beyond the density-independent model 
 and even applicable for a simple age-structured population where its vital rates are influenced by ECEs.
-<!--do I need to explain the details how to make a copula from raw data?-->
 

### 2019-07-16 MainText.Rmd
writing abstract

@@
 <!--200 words limit : J Am Nat-->
+
+With global climate change, extreme climatic events (ECEs) appear to be more common than usual. These ECEs 
+could pose severe impact on the resistance or resilience of a population or community level through
+life-history traits of the species. Therefore predicting the impact of ECEs 
+on extinction risk of a population is necessery. We here introduced the potential use of "tail-dependence" 
+property of a copula in modelling the extinction risk of an age-structured single patch model. We also 
+extended the generalized findings of density-independent metapopulation model from [@Ghosh_copula] 
+to density-dependence metapopulation models with this copula approach. For metapopulation models, 
+we showed extinction risk not only depended on the nature of tail-dependence of spatially correlated 
+noise but was also influenced by the inherent dynamics of the population-model.
 

### 2019-07-16 MainText.Rmd
edits, still suppmat fig caption short need to be added

@@
 
-\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, copula, density-dependence model
+\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, copula, density-dependent model
 
@@
 through higher hierarchical levels and lead to local extinction of 
-species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme]. For example, 
+species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme]. 
+<!--Should I mention idea of match-mismatch hypothesis?-->
+For example, 
 a study based on annal survival data showed changes
 in prey communities induced by extreme flooding affected predators in 
-aquatic and adjacent riparian habitats [@chiu2013impact]. Similarly, 
+aquatic and adjacent riparian habitats [@chiu2013impact] via trophic linkages. Similarly, 
 $43$ years of ring-recovery data was used to examine the demographic impact of ECEs 
@@
 extreme precipitation, drought) over $37$ years (1976-2012) at the site level, across 
-each species' life-stages. Successive ECEs with short time intervals 
+each species' life-stages. Experimental study showed successive ECEs with short time intervals 
 could weaken the resistance of plant communities too [@dreesen2014successive]. 
@@
 modeling [@joe2014_dependence; @nelsen2006_copula]. We here introduce the concept of copula in a
-manner that follows [@Ghosh_copula]. Copulas separate the information content of a bivariate dataset,
+manner that follows @Ghosh_copula. Copulas separate the information content of a bivariate dataset,
 $(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
@@
 <!--tail dep. of copula & extinction risk-->
-A major benefit of the copula approach is that it can effectively capture tail dependence. 
-Tail dependence, already exemplified in Fig..... (add Fig.), is associations between the more extreme 
+A major benefit of the copula approach is that it can effectively capture the tail dependence. 
+Tail dependence, already exemplified in Fig..... (add Fig.), is association between the more extreme 
 values of distributions. For two positively associated variables, left-tail 
 (respectively, right-tail) dependence is stronger association between
-values in the left or lower portions, i. e. smaller values 
-(respectively, the right or upper portions, i.e. higher values) 
-of the two distributions. For example, extreme cold winter during growing season
+values in the left or lower portions, i. e. for smaller values 
+(respectively, the right or upper portions, i.e. for higher values) 
+of the two distributions. Stronger dependence in either tail gives rise to asymmetric copula 
+structure for the joint distribution of two variables. 
+For example, extreme cold winter during growing season
 might reduce the adult reproduction rate as well as larval survival (i.e. stronger 
@@
 could be more complex depending 
-on the dynamics of the model. Though tail-dependence 
+on the dynamics of the model (see \nameref{results}). Though tail-dependence 
 property of a copula has been used for risk-analysis in other fields like finance, 
@@
 for both left tail and right tail dependent noises 
-against the varying range of $r$ when no dispersal ($d=0$) was considered.
+against the varying range of $r$ (along x-axis) when no dispersal ($d=0$) was considered.
 We excluded the possibility of dispersal in this case to avoid any rescue effect among patches. 
@@
 
-# Results
+# Results \label{results}
 
@@
 value of $r$, we observed right (or left) tail dependent noise caused higher (or lower) extinction 
-risk for the metapopulations. Figs. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
+risk for the metapopulations. Fig. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
 against local dispersal rate ($d$) for six density dependent metapopulation models 
@@
 \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
-represents the same but when global dispersal took place between patches. 
+represents the same but for the case when global dispersal took place between patches. 
 Both case showed the signature of 
@@
 Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) we hypothesized the switching 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) we hypothesized that the switching 
 between the extinction risk generated by left-tail dependent and right-tail dependent noise 

### 2019-07-16 SuppMat.Rmd
edits, still suppmat fig caption short need to be added
short caption added

@@
 \listoffigures
-\listoftables
 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time; fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure.
+\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C, SG) or right tail (SC, G) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time; fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure.
 $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
@@
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
+\caption[Extinction risk vs.
+dispersal rate for six density dependent metapopulation model with global dispersal.]{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
 Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp}}
@@
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
@@
 
-<!--Supp. FIG : Austin brewer model, local dispersal-->
+<!--Supp. FIG : Austin-Brewer model, local dispersal-->
 \begin{figure}[!h]
@@
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
-\caption{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 

### 2019-07-25 MainText.Rmd
edit intro

@@
 
-# Introduction \label{intro}
+# Introduction\label{Introduction}
 
@@
 and consequences of copula structures in associations between ecologically
-important variables and demonstrated one example of density-independent metapopulation model where 
-copula was used to estimate extinction risk. Copula, a rank based measure of association, is much more informative 
+important variables and demonstrated a basic example where 
+copula was used to model extinction risk of a density-independent 
+metapopulation. Present manuscript is an extension for the application of copula 
+in estimating the extinction risk as introduced by [@Ghosh_copula]. We will show 
+why the potential use of copula
+in modelling extinction risk goes beyond the density-independent model 
+and even applicable for a simple age-structured population where its vital rates are influenced by ECEs. 
+Following we will introduce the concept of copula in a manner that follows @Ghosh_copula.
+
+<!--brief intro for copula-->
+Copula, a rank based measure of association, is much more informative 
 than linear correlation measure and a widely applied tool in dependence 
-modeling [@joe2014_dependence; @nelsen2006_copula]. We here introduce the concept of copula in a
-manner that follows @Ghosh_copula. Copulas separate the information content of a bivariate dataset,
+modeling [@joe2014_dependence; @nelsen2006_copula]. Copulas separate the information content of a bivariate dataset,
 $(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
@@
 Ranking makes the marginal distributions uniform, isolating only the information on 
-association between the variables. Here, we will extend the application of copula 
-in estimating extinction risk as introduced by [@Ghosh_copula] on why the potential use of copula
-in modelling extinction risk goes beyond the density-independent model 
-and even applicable for a simple age-structured population where its vital rates are influenced by ECEs.
+association between the variables.
 
@@
 could be more complex depending 
-on the dynamics of the model (see \nameref{results}). Though tail-dependence 
+on the dynamics of the model (see \nameref{Results}). Though tail-dependence 
 property of a copula has been used for risk-analysis in other fields like finance, 
@@
 In this paper, we will explore the potential use of asymmetric tail-dependence in a copula 
-for understanding the extinction risk of populations towards extreme climatic event in ecology
+for understanding the extinction risk of populations towards extreme climatic event
 by addressing the following questions: (Q1) Does nature of tail dependence 
@@
 We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
-(Q2) How does asymmetric tail-dependent noises across multiple patches influence extinction 
-risk of a nonlinear metapopulation model with density dependence? Our speculation is that the general finding 
+(Q2) How does asymmetric noise copula (tail dependence across multiple patches) influence extinction 
+risk of a density-dependent metapopulation model? Our speculation is that the general finding 
 from density-independent metapopulation model [@Ghosh_copula]
 (higher extinction risk if noises across patches were 
-left-tail dependent than the one if noises had right-tail dependence) holds true only in
+left-tail dependent than the one if noises had right-tail dependence) would hold true only in
 the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
@@
 
-<!--Tuljapurkar estimation to be explained in Discussion-->
-
-# Methods \label{methods}
+# Methods\label{Methods}
 
 <!--stochastic matrix model : age structured-->
-To investigate (Q1) from \nameref{intro} that how copula structure in noisy 
+To investigate (Q1) from \nameref{Introduction} that how copula structure in noisy 
 environment affect the extinction risk of
@@
 <!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
-To test our hypothesis from (Q2) of \nameref{intro} for each metapopulation model,
+To test our hypothesis from (Q2) of \nameref{Introduction} for each metapopulation model,
 we estimated extinction risk for a wide range of parameter $r$, 
@@
 
-# Results \label{results}
+# Results\label{Results}
 
@@
 regime. We also got similarly consisent results with 
-our prediction (see Q2 from \nameref{intro}) for global 
+our prediction (see Q2 from \nameref{Introduction}) for global 
 dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
@@
 occurred near the critical parameter $r_c$ (see Q3). For simplicity we only consider 
-$d=0$ case, i.e., when there was no dispersal between patches (see \nameref{methods}). 
+$d=0$ case, i.e., when there was no dispersal between patches (see \nameref{Methods}). 
 Fig. \ref{fig_extrisk_d0} shows higher extinction risk with left-tail dependent noise 
@@
 # Discussion
+
+<!--summary results-->
+
+<!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
+
+<!--future study: more complex stage or age structured model, extinction risk how propagated through a food web-->
+
+<!--future study: worthy for PVA, IUCN red list-->
+
+<!--match-mismatch hypothesis leads to extinction? our phenology paper in prep.-->
 

### 2019-07-27 MainText.Rmd
started writing discussion

@@
 property of a copula in modelling the extinction risk of an age-structured single patch model. We also 
-extended the generalized findings of density-independent metapopulation model from [@Ghosh_copula] 
+extended the generalized findings of density-independent metapopulation model from @Ghosh_copula 
 to density-dependence metapopulation models with this copula approach. For metapopulation models, 
@@
 a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distributions 
-could be generated from iid noises showing specific tail-dependence (left or right tail)
-between fecundity and survival rates. Eq. \ref{eq_vitalrates} represent the stochasticity 
+were generated from iid noises showing specific tail-dependence (either in left- or in right-tail)
+between fecundity and survival rates. Tail dependence emerged if extreme bad years ("catastrophs") 
+would suppress both vital rates (hence, $f_a$ and $s_e$ would show a left tail dependent copula), whereas 
+extreme good years ("bonanzas") would increase both of them (i.e., $f_a$ and $s_e$ would show 
+a right tail dependent copula). Eq. \ref{eq_vitalrates} represent the stochasticity 
 in the vital rates. We included the density dependence effect in one of the vital rate,
@@
 by extreme climatic events. A $N \times 2$ matrix, M, was generated
-using `BiCopSim` function
-from `VineCopula` package 
-for each of the left 
-(Clayton, Survival Gumbel) and right 
+using `BiCopSim` function from `VineCopula` package 
+for each of the left (Clayton, Survival Gumbel) and right 
 (Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
-where $N = numsteps(=25) \times numsims(=10000)$.
+where $N$ is number of years ($25$) multiplied by number of simulations ($10000$) we used. 
 We kept the same Kendall's correlation ($\tau = 0.6$) for 
@@
 
-<!--summary results-->
+<!--summary results:-->
+Our results show the potential use of "asymmetric" copula in estimating the impact of 
+extreme ecological events on population-extinction risk. We show that through left and right tail 
+dependence in vital rates (respectively, spatially correlated environmental noises), years of extreme events 
+("catastrophs" and "bonanzas") affect extinction risk of a single-patch age-structured population
+(respectively, density dependent metapopulations) differently. Though left tail dependence is 
+expected to show higher extinction risk than the one for right tail dependence for age-structured 
+and density dependent metapopulation models, it is not always true 
+when metapopulation models with dispersal show over-compensatory dynamics. In the over-compensatory regime,
+right tail dependent noises result in higher extinction risk than the left tail dependent one. We show 
+that in absence of dispersal, the above-said switching in extinction risk measure occurs near the 
+critical parameter value where the deterministic model shows transition from 
+under-compensatory to over-compensatory regime.
 
@@
 
+<!--future study: worthy for PVA, IUCN red list
+1. Identify which life stage or demographic process as management target?
+2. multisite age str. model with dispersal, directional change in spatio-temporal variation for vital rates.
+3. Old study (Doak1994, Ferguson1995) showed including correlation btw vital rates for PVA makes a difference, 
+   what if we include TD through copula?
+4. What if TD with negative correlation?
+e.g. some events that will kill adults will allow more successful recruitment of seeds or other propagules.
+-->
+
 <!--future study: more complex stage or age structured model, extinction risk how propagated through a food web-->
-
-<!--future study: worthy for PVA, IUCN red list-->
 

### 2019-07-28 MainText.Rmd
writing discussion

@@
 change variability of the distribution, as well as change its 
-symmetry or skewness. Such extreme climatic events (ECE) can pose severe
+symmetry or skewness [@hansen2012perception]. Such extreme climatic events (ECE) can pose severe
 impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
@@
 we considered a basic stochastic age-structured model (Eq. \ref{eq_age_str}) 
-with two age classes (egg $E$ and adult $A$). We assumed
+with two age classes (egg $E$ and female adult $A$). We assumed
 the population was semelperous, i.e., adult ($A$) went through the
@@
 linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
-Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear} represents this scenario with 4 different 
+Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG} represents this scenario with 4 different 
 asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
 difference between extinction risks within a pair of different tail-dependent copulas
-(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear}(A,C) and
-G-SG in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear}(B,D)) and that difference was partly 
+(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}(A,C) and
+G-SG in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}(B,D)) and that difference was partly 
 dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
@@
 <!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
+Though we found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
+the right tail dependent ones, the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}) 
+unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr}). This finding could be explained 
+by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
+The approximation considered that the growth rate depends on the most dominant eigen value of the mean projection matrix 
+and the temporal variance of the log(population growth rate) caused by environmental stochasticity, which was based
+on the linear covariance measure between any two matrix elements, not on their tail-dependence properties. Therefore,
+according to Tuljapurkar's approximation, population growth rate in an iid environment 
+remained same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependent noises.
 
@@
 1. Identify which life stage or demographic process as management target?
-2. multisite age str. model with dispersal, directional change in spatio-temporal variation for vital rates.
-3. Old study (Doak1994, Ferguson1995) showed including correlation btw vital rates for PVA makes a difference, 
+2. Old study (Doak1994, Ferguson1995) showed including correlation btw vital rates for PVA makes a difference, 
    what if we include TD through copula?
-4. What if TD with negative correlation?
+3. multisite age str. model with dispersal, directional change in spatio-temporal variation for vital rates.
+-->
+Although we demonstrated that tail dependence between vital rates of an age structured model can
+influence extinction risk, substantial work remains to determine the importance of this effect in the 
+following scenarios. First, we demonstrated the effect of tail-dependence on extinction risk 
+with a very simplistic age structured (two age stages) model where we considered the total semelperous 
+population would go extinct if there would be only one female adult and one egg. But in reality population 
+viability analysis (PVA) deals with a little bit higher threshold with several age, stage or size class and a variety 
+of populations including semelperous and iteroperous [@morris2002quantitative]. Future studies should include 
+tail-dependence effect on extinction risk 
+for a multiple age-, size- and stage- structured model and estimate extinction risk 
+for each age, size and stage class separately 
+so that we can identify (via multisite-PVA) which life-stage or demographic process 
+to target for population management. Earlier study by @doak1994modeling showed PVA models including correlation 
+between different demographic rates indicated marked difference in predicting confidence interval for 
+future population size than the model which did not include correlation. But when 
+such correlation cannot be estimated directly, more generally where the
+forms of statistical dependence among the model's parameters are unknown, making incorrect
+assumptions about correlations and dependencies may lead to severe over-estimation 
+or under-estimation of extinction risks [@ferson1995correlations]. We believe modeling 
+dependencies with copula could help in this case. Second, we did not consider multiple sites of 
+age-structured metapopulation which could be linked via dispersal. So, possible future investigation 
+would be to explore the viability for the metapopulations if vital rates across sites 
+showed specific tail dependence with spatially correlated extreme climatic events.
+
+<!--What if TD with negative correlation?
 e.g. some events that will kill adults will allow more successful recruitment of seeds or other propagules.

### 2019-07-28 MainText.Rmd
discussion written for 1st draft

@@
 <!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
-Though we found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
-the right tail dependent ones, the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}) 
+We found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
+the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}) 
 unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr}). This finding could be explained 
@@
 2. Old study (Doak1994, Ferguson1995) showed including correlation btw vital rates for PVA makes a difference, 
-   what if we include TD through copula?
-3. multisite age str. model with dispersal, directional change in spatio-temporal variation for vital rates.
+   what if we include TD through copula? Multisite age str. model with dispersal, directional change
+   in spatio-temporal variation for vital rates.
+3. Tdep. with -ve correlation between vital rates?
+4. extinction risk how propagated through a food web?
 -->
@@
 age-structured metapopulation which could be linked via dispersal. So, possible future investigation 
-would be to explore the viability for the metapopulations if vital rates across sites 
-showed specific tail dependence with spatially correlated extreme climatic events.
-
-<!--What if TD with negative correlation?
-e.g. some events that will kill adults will allow more successful recruitment of seeds or other propagules.
--->
-
-<!--future study: more complex stage or age structured model, extinction risk how propagated through a food web-->
+would be to explore the viability for the metapopulations if vital rates (e.g. fecundity or 
+survival of different stage class) across sites showed specific tail dependence 
+with spatially correlated extreme climatic events. Third, we considered tail dependence 
+between vital rates when they were positively correlated, but what would be the extinction risk 
+and viability of a population if demographic rates were negatively correlated with asymmetric tail dependence. 
+This scenario could be considered for those extreme events that kill adults and allow more successful
+recruitment of seeds or other propagules. Fourth, following the aforementioned idea, it would be worthy 
+to model more complex prey-predator model where asymmetric tail dependent copula could be used to mimic
+interspecific-dependence. An earlier study [@carnicer2011widespread] showed extreme events
+(i.e., heat waves and droughts) increased mortality and defoliation in southern European forest and its 
+dependent herbivorous insects and fungi through trophic linkage. ECEs can even truncate 
+multiyear insect outbreak dynamics [@esper20061200]. We believe copula (may be a multidimensional one) has 
+a lot more potential to estimate risk of cascading extinction in model food webs.
 

### 2019-08-22 mydispersal.R
moderate tail dep. noise : clayton with ktau=0.6

@@ -1,7 +1,8 @@
 #--------------------------------------------------------------------------
 source("ExtremeTailDep.R")
 #--------------------------------------------------------------------------
-
+library(copula)
+library(VineCopula)
 #-------------------------------------------------------------------------------
 # A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
 # Input :
@@ -156,7 +157,7 @@ get_avg_acf<-function(sims){
   }
   # which(is.na(acfs))
   ans<-mean(acfs,na.rm=T) # This line to remove NaN which is produced due to pop = 0 throughout the time at any patch
-                          # you can see that in the row of matrix m (for a given i)
+  # you can see that in the row of matrix m (for a given i)
   return(ans)
 }
 # I checked : ACF of noise (right tail) = ACF of noise (left tail)
@@ -176,45 +177,80 @@ get_avg_acf<-function(sims){
 #       9) model : a character specifying model name
 #       10) ploton : logical(T or F) to get optional plot
 #       11) resloc : location to save plots
-#       getacf : these are tags
+#       12) getacf : these are tags
 
 plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,getacf){
- 
+  
+  # ========================== when noises are extreme tail dep. ========================================
   ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1,mn=0,sdev=1)# a righttail dep matrix(numpoints by numlocs,     
-  #                                                      numpoints=numsteps*numsims)
+  # numpoints=numsteps*numsims)
   
   ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)
   ns1<-aperm(ns1,c(2,3,1)) # convert to an array (numsims by numlocs by numsteps)
   
   ns1<-scl*ns1
[…]

### 2019-08-22 mydispersal.R
modified

@@ -2,7 +2,6 @@
 source("ExtremeTailDep.R")
 #--------------------------------------------------------------------------
 library(copula)
-library(VineCopula)
 #-------------------------------------------------------------------------------
 # A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
 # Input :

### 2019-08-22 SuppMat.Rmd
writing discussion
moderate tail dep. code added

@@
 #initiate a dataframe to store ext risk at dispersal d=0 for all r
-ricker_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) 
-colnames(ricker_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+ricker_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist)),
+                              NA*numeric(length(rlist)),NA*numeric(length(rlist))) 
+colnames(ricker_extrisk_d0)<-c("r","risk_xleft_d0","risk_xright_d0","risk_mleft_d0","risk_mright_d0")
 
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_ricker_local<-abind(res_ricker_local,temp_mat,along = 3)
 
-    ricker_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
-    ricker_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
+    ricker_extrisk_d0$risk_xleft_d0[i]<-ss$risk_xleft[1]
+    ricker_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
+
+    ricker_extrisk_d0$risk_mleft_d0[i]<-ss$risk_mleft[1]
+    ricker_extrisk_d0$risk_mright_d0[i]<-ss$risk_mright[1]
 }
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_ricker_local)<-list(NULL,cnm,anm)
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_ricker_global<-abind(res_ricker_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_ricker_global)<-list(NULL,cnm,anm)
@@
 
-pdf(paste(resloc,"ricker_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"ricker_ext_risk_local_disp_extremenoise.pdf",sep=""),height=9,width=9)
 #op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
@@
 
-  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(ss$d,ss$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
   mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
-  #---------------------------------------------------
-  ## Allow a second plot on the same graph
-  #par(new=TRUE)
-
-  #ylm2<-c(-0.2,0.6)
-  ## Plot the second plot and put axis scale on right
-  #plot(ss$d, ss$pop_avg_acf_left, pch=16,  xlab="", ylab="", ylim=ylm2,xlim=c(0,1),
-       #ylim=range(-0.1,0.1), xlim=c(0,1),cex.axis=1.5,bty="n",
-  #     axes=FALSE, type="l", col="orange")
- # abline(h=0,lty="dotted",col="darkgrey")
- # lines(ss$d, ss$pop_avg_acf_right, col="skyblue")
-
-  ## a little farther out (line=3.2) to make room for labels
- # mtext("Avg. ACF",side=4,line=3.2,col="green4") 
- # axis(4, ylim=ylm2, las=1,col="green4",col.axis="green4")
-  #---------------------------------------------------------
-}
-par(op)
-dev.off()
-
-pdf(paste(resloc,"ricker_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"ricker_ext_risk_global_disp_extremenoise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"ricker_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=9,width=9)
+
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_ricker_local)[3]){
+  #s<-res_ricker_local[,,i]
+  ss<-res_ricker_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"ricker_ext_risk_global_disp_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_ricker_global)[3]){
+  #ss<-res_ricker_global
+  ss<-res_ricker_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 numsims<-10000
-ricker_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_left_d0)/numsims
-ricker_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_left_d0)/numsims
-
-ricker_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_right_d0)/numsims
-ricker_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"ricker_extrisk_d0.pdf",sep=""),height=5,width=5)
+ricker_extrisk_d0$CI0.025xleft<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_xleft_d0)/numsims
+ricker_extrisk_d0$CI0.975xleft<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_xleft_d0)/numsims
+
+ricker_extrisk_d0$CI0.025xright<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_xright_d0)/numsims
+ricker_extrisk_d0$CI0.975xright<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_xright_d0)/numsims
+
+ricker_extrisk_d0$CI0.025mleft<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_mleft_d0)/numsims
+ricker_extrisk_d0$CI0.975mleft<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_mleft_d0)/numsims
+
+ricker_extrisk_d0$CI0.025mright<-qbinom(p=0.025,size=numsims,prob=ricker_extrisk_d0$risk_mright_d0)/numsims
+ricker_extrisk_d0$CI0.975mright<-qbinom(p=0.975,size=numsims,prob=ricker_extrisk_d0$risk_mright_d0)/numsims
+
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"ricker_extrisk_d0_extremenoise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_left_d0,col="red",type="b",
+plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_xleft_d0,col="red",type="b",
      ylim=c(0,1),xlim = range(ricker_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
-arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025left,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975left,
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025xleft,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975xleft,
        angle=90,length=0.1,col="red",code=3)
@@
 
-lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025right,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025xright,x1=ricker_extrisk_d0$r,
+       y1=ricker_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+legend("topleft", legend=c("Ricker"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("K = ",K),side = 3, line=-16, adj=0.95, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", legend=c("Ricker",paste0("K = ",K)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"ricker_extrisk_d0_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_mleft_d0,col="red",type="b",
+     ylim=c(0,1),xlim = range(ricker_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025mleft,x1=ricker_extrisk_d0$r,
+       y1=ricker_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_ricker,lty="dotted",col="darkgrey") #rc=1
+
+lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_mright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025mright,
+       x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="blue",code=3)
 legend("topleft", legend=c("Ricker"), horiz=F, bty='n',cex=1.8)  
@@
 
-hassell_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(hassell_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+hassell_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist)),
+                               NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store 
+                                                                       # ext risk at dispersal d=0 for all r
+colnames(hassell_extrisk_d0)<-c("r","risk_xleft_d0","risk_xright_d0","risk_mleft_d0","risk_mright_d0")
 
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_hassell_local<-abind(res_hassell_local,temp_mat,along = 3)
 
-    hassell_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
-    hassell_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
+    hassell_extrisk_d0$risk_xleft_d0[i]<-ss$risk_xleft[1]
+    hassell_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
 }
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_hassell_local)<-list(NULL,cnm,anm)
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_hassell_global<-abind(res_hassell_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_hassell_global)<-list(NULL,cnm,anm)
@@
 ```{r fig_hassell_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_hassell_local,res_hassell_global)}
-pdf(paste(resloc,"hassell_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
+
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"hassell_ext_risk_local_disp_extremenoise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-pdf(paste(resloc,"hassell_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
+pdf(paste(resloc,"hassell_ext_risk_global_disp_extremenoise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        #panel.first = grid(),
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(ss$d,ss$risk_right,type="b",col="blue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"hassell_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_hassell_local)[3]){
+  ss<-res_hassell_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"hassell_ext_risk_global_disp_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_hassell_global)[3]){
+  ss<-res_hassell_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 
-hassell_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_left_d0)/numsims
-hassell_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_left_d0)/numsims
-
-hassell_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_right_d0)/numsims
-hassell_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"hassell_extrisk_d0.pdf",sep=""),height=5,width=5)
+hassell_extrisk_d0$CI0.025xleft<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_xleft_d0)/numsims
+hassell_extrisk_d0$CI0.975xleft<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_xleft_d0)/numsims
+
+hassell_extrisk_d0$CI0.025xright<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_xright_d0)/numsims
+hassell_extrisk_d0$CI0.975xright<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_xright_d0)/numsims
+
+hassell_extrisk_d0$CI0.025mleft<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_mleft_d0)/numsims
+hassell_extrisk_d0$CI0.975mleft<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_mleft_d0)/numsims
+
+hassell_extrisk_d0$CI0.025mright<-qbinom(p=0.025,size=numsims,prob=hassell_extrisk_d0$risk_mright_d0)/numsims
+hassell_extrisk_d0$CI0.975mright<-qbinom(p=0.975,size=numsims,prob=hassell_extrisk_d0$risk_mright_d0)/numsims
+
+pdf(paste(resloc,"hassell_extrisk_d0_extremenoise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
      xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
-arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025left,
-       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025xleft,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_hassell,lty="dotted",col="darkgrey")
 
-lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025right,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025xright,x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+legend("topleft", legend=c("Hassell"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Hassell",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+
+pdf(paste(resloc,"hassell_extrisk_d0_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_mleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025mleft,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_hassell,lty="dotted",col="darkgrey")
+
+lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_mright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025mright,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="blue",code=3)
 legend("topleft", legend=c("Hassell"), horiz=F, bty='n',cex=1.8)  
@@
 
-msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(msmith_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+msmith_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist)),
+                              NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store 
+                                                                                     # ext risk at dispersal d=0 for all r
+colnames(msmith_extrisk_d0)<-c("r","risk_xleft_d0","risk_xright_d0","risk_mleft_d0","risk_mright_d0")
 
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_msmith_local<-abind(res_msmith_local,temp_mat,along = 3)
 
-    msmith_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
-    msmith_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
+    msmith_extrisk_d0$risk_xleft_d0[i]<-ss$risk_xleft[1]
+    msmith_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
 }
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_msmith_local)<-list(NULL,cnm,anm)
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_msmith_global<-abind(res_msmith_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_msmith_global)<-list(NULL,cnm,anm)
@@
 ```{r fig_msmith_scl_0.8, echo=F, results="hide",cache=T, cache.extra=list(seed,res_msmith_local,res_msmith_global)}
-pdf(paste(resloc,"msmith_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
+
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"msmith_ext_risk_local_disp_extremenoise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-pdf(paste(resloc,"msmith_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
+pdf(paste(resloc,"msmith_ext_risk_global_disp_extremenoise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5))
@@
 dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"msmith_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_msmith_local)[3]){
+  ss<-res_msmith_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"msmith_ext_risk_global_disp_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5))
+
+for(i in 1:dim(res_msmith_global)[3]){
+  ss<-res_msmith_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 numsims<-10000
-msmith_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
-msmith_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_left_d0)/numsims
-
-msmith_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
-msmith_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"msmith_extrisk_d0.pdf",sep=""),height=5,width=5)
+msmith_extrisk_d0$CI0.025xleft<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_xleft_d0)/numsims
+msmith_extrisk_d0$CI0.975xleft<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_xleft_d0)/numsims
+
+msmith_extrisk_d0$CI0.025xright<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_xright_d0)/numsims
+msmith_extrisk_d0$CI0.975xright<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_xright_d0)/numsims
+
+msmith_extrisk_d0$CI0.025mleft<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_mleft_d0)/numsims
+msmith_extrisk_d0$CI0.975mleft<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_mleft_d0)/numsims
+
+msmith_extrisk_d0$CI0.025mright<-qbinom(p=0.025,size=numsims,prob=msmith_extrisk_d0$risk_mright_d0)/numsims
+msmith_extrisk_d0$CI0.975mright<-qbinom(p=0.975,size=numsims,prob=msmith_extrisk_d0$risk_mright_d0)/numsims
+
+pdf(paste(resloc,"msmith_extrisk_d0_extremenoise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),xlim = range(msmith_extrisk_d0$r),
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(msmith_extrisk_d0$r),
      xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025left,
-       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025xleft,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_msmith,lty="dotted",col="darkgrey")
 
-lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
-arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025right,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_xright_d0,col="blue",type="b",pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025xright,x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+legend("topleft", legend=c("Maynard Smith"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+par(op)
+dev.off()
+
+pdf(paste(resloc,"msmith_extrisk_d0_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_mleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(msmith_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025mleft,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_msmith,lty="dotted",col="darkgrey")
+
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_mright_d0,col="blue",type="b",pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025mright,x1=msmith_extrisk_d0$r,
+       y1=msmith_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="blue",code=3)
 legend("topleft", legend=c("Maynard Smith"), horiz=F, bty='n',cex=1.8)  
@@
 
-pennycuick_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) # a dataframe to store ext risk at dispersal d=0 for all r
-colnames(pennycuick_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+pennycuick_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist)),
+                                  NA*numeric(length(rlist)),NA*numeric(length(rlist))) # a dataframe to store 
+                                                                            # ext risk at dispersal d=0 for all r
+colnames(pennycuick_extrisk_d0)<-c("r","risk_xleft_d0","risk_xright_d0","risk_mleft_d0","risk_mright_d0")
 
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_pennycuick_local<-abind(res_pennycuick_local,temp_mat,along = 3)
 
-    pennycuick_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
-    pennycuick_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
+    pennycuick_extrisk_d0$risk_xleft_d0[i]<-ss$risk_xleft[1]
+    pennycuick_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
 }
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_pennycuick_local)<-list(NULL,cnm,anm)
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_pennycuick_global<-abind(res_pennycuick_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_pennycuick_global)<-list(NULL,cnm,anm)
@@
 ```{r fig_pennycuick_scl_1, echo=F, results="hide",cache=T, cache.extra=list(seed,res_pennycuick_local,res_pennycuick_global)}
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp.pdf",sep=""),height=12,width=12)
+
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp_extremenoise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.2,1.5),mgp=c(2,0.5,0))
@@
 
-  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(ss$d,ss$risk_right,type="b",col="blue")
-  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
    mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
@@
 
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp.pdf",sep=""),height=12,width=12)
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp_extremenoise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(ss$d,ss$risk_right,type="b",col="blue")
-  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.2,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_pennycuick_local)[3]){
+  ss<-res_pennycuick_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+   mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_pennycuick_global)[3]){
+  ss<-res_pennycuick_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
   mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
@@
 
-pennycuick_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
-pennycuick_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_left_d0)/numsims
-
-pennycuick_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
-pennycuick_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"pennycuick_extrisk_d0.pdf",sep=""),height=5,width=5)
+pennycuick_extrisk_d0$CI0.025xleft<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_xleft_d0)/numsims
+pennycuick_extrisk_d0$CI0.975xleft<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_xleft_d0)/numsims
+
+pennycuick_extrisk_d0$CI0.025xright<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_xright_d0)/numsims
+pennycuick_extrisk_d0$CI0.975xright<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_xright_d0)/numsims
+
+pennycuick_extrisk_d0$CI0.025mleft<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_mleft_d0)/numsims
+pennycuick_extrisk_d0$CI0.975mleft<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_mleft_d0)/numsims
+
+pennycuick_extrisk_d0$CI0.025mright<-qbinom(p=0.025,size=numsims,prob=pennycuick_extrisk_d0$risk_mright_d0)/numsims
+pennycuick_extrisk_d0$CI0.975mright<-qbinom(p=0.975,size=numsims,prob=pennycuick_extrisk_d0$risk_mright_d0)/numsims
+
+pdf(paste(resloc,"pennycuick_extrisk_d0_extremenoise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),
      xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
-arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025left,
-       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025xleft,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_pennycuick,lty="dotted",col="darkgrey")
 
-lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_right_d0,col="blue",type="b",ylim=c(0,1),pch=1)
-arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025right,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025xright,x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+legend("topleft", legend=c("Pennycuick"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Pennycuick",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+
+pdf(paste(resloc,"pennycuick_extrisk_d0_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_mleft_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(pennycuick_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025mleft,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_pennycuick,lty="dotted",col="darkgrey")
+
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_mright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025mright,x1=pennycuick_extrisk_d0$r,
+       y1=pennycuick_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="blue",code=3)
 legend("topleft", legend=c("Pennycuick"), horiz=F, bty='n',cex=1.8)  
@@
 
-verhulst_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(verhulst_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+verhulst_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist)),
+                                NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store 
+                                                                   # ext risk at dispersal d=0 for all r
+colnames(verhulst_extrisk_d0)<-c("r","risk_xleft_d0","risk_xright_d0","risk_mleft_d0","risk_mright_d0")
 
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_verhulst_local<-abind(res_verhulst_local,temp_mat,along = 3)
 
-    verhulst_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
-    verhulst_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
+    verhulst_extrisk_d0$risk_xleft_d0[i]<-ss$risk_xleft[1]
+    verhulst_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
 }
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_verhulst_local)<-list(NULL,cnm,anm)
@@
 
-    temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+    temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
     res_verhulst_global<-abind(res_verhulst_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_verhulst_global)<-list(NULL,cnm,anm)
@@
 
-pdf(paste(resloc,"verhulst_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"verhulst_ext_risk_local_disp_extremenoise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(ss$d,ss$risk_right,type="b",col="blue")
-  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
   mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
@@
 
-pdf(paste(resloc,"verhulst_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+pdf(paste(resloc,"verhulst_ext_risk_global_disp_extremenoise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"verhulst_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_verhulst_local)[3]){
+  ss<-res_verhulst_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"verhulst_ext_risk_global_disp_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_verhulst_global)[3]){
+  ss<-res_verhulst_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 ```{r fig_verhulst_extrisk_d0_scl_0.6, echo=F,results="hide"}
-
 set.seed(seed)
@@
 
-verhulst_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
-verhulst_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_left_d0)/numsims
-
-verhulst_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
-verhulst_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_right_d0)/numsims
-
-pdf(paste(resloc,"verhulst_extrisk_d0.pdf",sep=""),height=5,width=5)
+verhulst_extrisk_d0$CI0.025xleft<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_xleft_d0)/numsims
+verhulst_extrisk_d0$CI0.975xleft<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_xleft_d0)/numsims
+
+verhulst_extrisk_d0$CI0.025xright<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_xright_d0)/numsims
+verhulst_extrisk_d0$CI0.975xright<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_xright_d0)/numsims
+
+verhulst_extrisk_d0$CI0.025mleft<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_mleft_d0)/numsims
+verhulst_extrisk_d0$CI0.975mleft<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_mleft_d0)/numsims
+
+verhulst_extrisk_d0$CI0.025mright<-qbinom(p=0.025,size=numsims,prob=verhulst_extrisk_d0$risk_mright_d0)/numsims
+verhulst_extrisk_d0$CI0.975mright<-qbinom(p=0.975,size=numsims,prob=verhulst_extrisk_d0$risk_mright_d0)/numsims
+
+pdf(paste(resloc,"verhulst_extrisk_d0_extremenoise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),
      xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
-arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025left,x1=verhulst_extrisk_d0$r,
-       y1=verhulst_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025xleft,x1=verhulst_extrisk_d0$r,
+       y1=verhulst_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_verhulst,lty="dotted",col="darkgrey")
 
-lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
-arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025right,
-       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_xright_d0,col="blue",type="b",pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025xright,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+legend("topleft", legend=c("Verhulst"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("K = ",K),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Verhulst",paste0("K =1, sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+
+pdf(paste(resloc,"verhulst_extrisk_d0_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_mleft_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(verhulst_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025mleft,x1=verhulst_extrisk_d0$r,
+       y1=verhulst_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_verhulst,lty="dotted",col="darkgrey")
+
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_mright_d0,col="blue",type="b",pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025mright,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="blue",code=3)
 legend("topleft", legend=c("Verhulst"), horiz=F, bty='n',cex=1.8)  
@@
 
-abrewer_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store ext risk at dispersal d=0 for all r
-colnames(abrewer_extrisk_d0)<-c("r","risk_left_d0","risk_right_d0")
+abrewer_extrisk_d0<-data.frame(rlist,NA*numeric(length(rlist)),NA*numeric(length(rlist)),
+                               NA*numeric(length(rlist)),NA*numeric(length(rlist))) #initiate a dataframe to store 
+                                                                         # ext risk at dispersal d=0 for all r
+colnames(abrewer_extrisk_d0)<-c("r","risk_xleft_d0","risk_xright_d0","risk_mleft_d0","risk_mright_d0")
 
@@
 
-  temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+  temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
   res_abrewer_local<-abind(res_abrewer_local,temp_mat,along = 3)
 
-  abrewer_extrisk_d0$risk_left_d0[i]<-ss$risk_left[1]
-  abrewer_extrisk_d0$risk_right_d0[i]<-ss$risk_right[1]
+  abrewer_extrisk_d0$risk_xleft_d0[i]<-ss$risk_xleft[1]
+  abrewer_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
 }
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_abrewer_local)<-list(NULL,cnm,anm)
@@
 
-  temp_mat<-cbind(ss$d_seq,ss$risk_left,ss$risk_right,ss$pop_avg_acf_left,ss$pop_avg_acf_right)
+  temp_mat<-cbind(ss$d_seq,ss$risk_xleft,ss$risk_xright,ss$risk_mleft,ss$risk_mright)
   res_abrewer_global<-abind(res_abrewer_global,temp_mat,along = 3)
@@
 anm<-paste("r_",rlist,sep="")
-cnm<-c("d","risk_left","risk_right","pop_avg_acf_left","pop_avg_acf_right")
+cnm<-c("d","risk_xleft","risk_xright","risk_mleft","risk_mright")
 dimnames(res_abrewer_global)<-list(NULL,cnm,anm)
@@
 ```{r fig_abrewer_scl_0.6, echo=F, results="hide",cache=T, cache.extra=list(seed,res_abrewer_local,res_abrewer_global)}
-pdf(paste(resloc,"abrewer_ext_risk_local_disp.pdf",sep=""),height=9,width=9)
+#-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
+pdf(paste(resloc,"abrewer_ext_risk_local_disp_extremenoise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-pdf(paste(resloc,"abrewer_ext_risk_global_disp.pdf",sep=""),height=9,width=9)
+pdf(paste(resloc,"abrewer_ext_risk_global_disp_extremenoise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-  plot(ss$d,ss$risk_left,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
-  lines(ss$d,ss$risk_right,type="b",col="blue")
-  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
-  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"abrewer_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_abrewer_local)[3]){
+  ss<-res_abrewer_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"abrewer_ext_risk_global_disp_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_abrewer_global)[3]){
+  ss<-res_abrewer_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mright,type="b",col="blue")
   mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
@@
 
-abrewer_extrisk_d0$CI0.025left<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
-abrewer_extrisk_d0$CI0.975left<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_left_d0)/numsims
-abrewer_extrisk_d0$CI0.025right<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
-abrewer_extrisk_d0$CI0.975right<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_right_d0)/numsims
-pdf(paste(resloc,"abrewer_extrisk_d0.pdf",sep=""),height=5,width=5)
+abrewer_extrisk_d0$CI0.025xleft<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_xleft_d0)/numsims
+abrewer_extrisk_d0$CI0.975xleft<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_xleft_d0)/numsims
+abrewer_extrisk_d0$CI0.025xright<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_xright_d0)/numsims
+abrewer_extrisk_d0$CI0.975xright<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_xright_d0)/numsims
+
+abrewer_extrisk_d0$CI0.025mleft<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_mleft_d0)/numsims
+abrewer_extrisk_d0$CI0.975mleft<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_mleft_d0)/numsims
+abrewer_extrisk_d0$CI0.025mright<-qbinom(p=0.025,size=numsims,prob=abrewer_extrisk_d0$risk_mright_d0)/numsims
+abrewer_extrisk_d0$CI0.975mright<-qbinom(p=0.975,size=numsims,prob=abrewer_extrisk_d0$risk_mright_d0)/numsims
+
+pdf(paste(resloc,"abrewer_extrisk_d0_extremenoise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_left_d0,col="red",type="b",ylim=c(0,1),
+plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),
      xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
-arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025left,x1=abrewer_extrisk_d0$r,
-       y1=abrewer_extrisk_d0$CI0.975left,angle=90,length=0.1,col="red",code=3)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025xleft,x1=abrewer_extrisk_d0$r,
+       y1=abrewer_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
 abline(v=rc_abrewer,lty="dotted",col="darkgrey")
 
-lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_right_d0,col="blue",type="b",pch=1)
-arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025right,
-       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975right,angle=90,length=0.1,col="blue",code=3)
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_xright_d0,col="blue",type="b",pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025xright,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
 legend("topleft", legend=c("Austin-Brewer"), horiz=F, bty='n',cex=1.8)  
@@
 dev.off()
+
+pdf(paste(resloc,"abrewer_extrisk_d0_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_mleft_d0,col="red",type="b",ylim=c(0,1),
+     xlim = range(abrewer_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025mleft,x1=abrewer_extrisk_d0$r,
+       y1=abrewer_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_abrewer,lty="dotted",col="darkgrey")
+
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_mright_d0,col="blue",type="b",pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025mright,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="blue",code=3)
+legend("topleft", legend=c("Austin-Brewer"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("s = ",s,", K = ",K),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+#legend("bottomright", c("Austin-Brewer",paste0("s = ",s," , K = ",K),paste0(" sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
 ```
@@
 
-getplot<-function(x,rlist){
+getplot<-function(x,rlist,noisetype){
 lasta<-dim(x)[3]
@@
 
-plot(Af$d,Af$risk_left,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
+if(noisetype=="extreme"){
+  risk_left<-Af$risk_xleft
+  risk_right<-Af$risk_xright
+}
+
+if(noisetype=="moderate"){
+  risk_left<-Af$risk_mleft
+  risk_right<-Af$risk_mright
+}
+
+plot(Af$d,risk_left,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
      xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-lines(Af$d,Af$risk_right,type="b",col="blue")
+lines(Af$d,risk_right,type="b",col="blue")
 
@@
 
-plot(Al$d,Al$risk_left,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
-       #panel.first = grid(),
+plot(Al$d,risk_left,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
      xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-lines(Al$d,Al$risk_right,type="b",col="blue")
+lines(Al$d,risk_right,type="b",col="blue")
 mtext(paste0("r = ", round(rlist[lasta],5)))
@@
 
-pdf("./Results/allmodel_ext_risk_local_disp.pdf",height=12,width=9)
+pdf("./Results/allmodel_ext_risk_local_disp_extremenoise.pdf",height=12,width=9)
 
@@
 mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist,noisetype = "extreme")
 
@@
 rlist<-hassell_extrisk_d0$r
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist, noisetype = "extreme")
 
@@
 rlist<-msmith_extrisk_d0$r
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist, noisetype = "extreme")
 
@@
 rlist<-pennycuick_extrisk_d0$r
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist, noisetype = "extreme")
 
@@
 rlist<-verhulst_extrisk_d0$r
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist, noisetype = "extreme")
 
@@
 rlist<-abrewer_extrisk_d0$r
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist,noisetype = "extreme")
 
@@
 ```{r create_maintext_fig_globaldisp_summary,echo=F,results="hide"}
-pdf("./Results/allmodel_ext_risk_global_disp.pdf",height=12,width=9)
+
+#------------- for extreme tail dep. noise ---------------------
+
+pdf("./Results/allmodel_ext_risk_global_disp_extremenoise.pdf",height=12,width=9)
 
@@
 mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist, noisetype = "extreme")
 
@@
 rlist<-abrewer_extrisk_d0$r
-getplot(x=x,rlist = rlist)
+getplot(x=x,rlist = rlist, noisetype = "extreme")
+
+par(op)
+dev.off()
+
+#--------------------for moderate tail dep. noise -----------------------
+
+pdf("./Results/allmodel_ext_risk_global_disp_moderatenoise.pdf",height=12,width=9)
+
+op<-par(mfrow=c(6,3),mar=c(5.2,4.2,1.2,1.5))
+
+x<-res_ricker_global
+rlist<-ricker_extrisk_d0$r
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
+mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_hassell_global
+rlist<-hassell_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_msmith_global
+rlist<-msmith_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_pennycuick_global
+rlist<-pennycuick_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_verhulst_global
+rlist<-verhulst_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_abrewer_global
+rlist<-abrewer_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
 
@@
 All copula has same Kendall correlation $\tau = 0.6$.
-\label{fig_extrisk_agestr_C_SC_G_SG_linear}}
+\label{fig_extrisk_agestr_C_SC_G_SG}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp.pdf}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_extremenoise.pdf}
 \caption[Extinction risk vs.
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
@@
 \includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_extremenoise.pdf}\\
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)

### 2019-08-23 SuppMat.Rmd
bug fixed and code modified

@@
     hassell_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
+
+    hassell_extrisk_d0$risk_mleft_d0[i]<-ss$risk_mleft[1]
+    hassell_extrisk_d0$risk_mright_d0[i]<-ss$risk_mright[1]
 }
@@
     msmith_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
+
+    msmith_extrisk_d0$risk_mleft_d0[i]<-ss$risk_mleft[1]
+    msmith_extrisk_d0$risk_mright_d0[i]<-ss$risk_mright[1]
 }
@@
     pennycuick_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
+
+    pennycuick_extrisk_d0$risk_mleft_d0[i]<-ss$risk_mleft[1]
+    pennycuick_extrisk_d0$risk_mright_d0[i]<-ss$risk_mright[1]
 }
@@
     verhulst_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
+
+    verhulst_extrisk_d0$risk_mleft_d0[i]<-ss$risk_mleft[1]
+    verhulst_extrisk_d0$risk_mright_d0[i]<-ss$risk_mright[1]
 }
@@
   abrewer_extrisk_d0$risk_xright_d0[i]<-ss$risk_xright[1]
+
+  abrewer_extrisk_d0$risk_mleft_d0[i]<-ss$risk_mleft[1]
+  abrewer_extrisk_d0$risk_mright_d0[i]<-ss$risk_mright[1]
 }
@@
 if(noisetype=="extreme"){
-  risk_left<-Af$risk_xleft
-  risk_right<-Af$risk_xright
+  risk_left1<-Af$risk_xleft
+  risk_right1<-Af$risk_xright
+
+  risk_left2<-Al$risk_xleft
+  risk_right2<-Al$risk_xright
 }
@@
 if(noisetype=="moderate"){
-  risk_left<-Af$risk_mleft
-  risk_right<-Af$risk_mright
-}
-
-plot(Af$d,risk_left,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
+  risk_left1<-Af$risk_mleft
+  risk_right1<-Af$risk_mright
+
+  risk_left2<-Al$risk_mleft
+  risk_right2<-Al$risk_mright
+}
+
+plot(Af$d,risk_left1,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
      xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-lines(Af$d,risk_right,type="b",col="blue")
+lines(Af$d,risk_right1,type="b",col="blue")
 
@@
 
-plot(Al$d,risk_left,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
+plot(Al$d,risk_left2,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
      xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
-lines(Al$d,risk_right,type="b",col="blue")
+lines(Al$d,risk_right2,type="b",col="blue")
 mtext(paste0("r = ", round(rlist[lasta],5)))
@@
 
+#------------- for extreme tail dep. noise : local dispersal ---------------------
 pdf("./Results/allmodel_ext_risk_local_disp_extremenoise.pdf",height=12,width=9)
@@
 getplot(x=x,rlist = rlist,noisetype = "extreme")
+
+par(op)
+dev.off()
+
+#------------- for moderate tail dep. noise : local dispersal ---------------------
+pdf("./Results/allmodel_ext_risk_local_disp_moderatenoise.pdf",height=12,width=9)
+
+op<-par(mfrow=c(6,3),mar=c(5.2,4.2,1.2,1.5))
+
+x<-res_ricker_local
+rlist<-ricker_extrisk_d0$r
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
+mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+#mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+getplot(x=x,rlist = rlist,noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_hassell_local
+rlist<-hassell_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_msmith_local
+rlist<-msmith_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_pennycuick_local
+rlist<-pennycuick_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_verhulst_local
+rlist<-verhulst_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_abrewer_local
+rlist<-abrewer_extrisk_d0$r
+getplot(x=x,rlist = rlist,noisetype = "moderate")
 

### 2019-08-26 MainText.Rmd
deleting G-SG results, only include C-SC as moderate tail dep.

@@
 ---
-title: "Extreme climatic events with asymmetric tail-dependence affect population-extinction risk"
+title: "Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
-Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG} represents this scenario with 4 different 
+Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with 4 different 
 asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
 difference between extinction risks within a pair of different tail-dependent copulas
-(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}(A,C) and
-G-SG in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}(B,D)) and that difference was partly 
+(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_mtd}(A,C) and
+G-SG in Fig. \ref{SM-fig_extrisk_agestr_mtd}(B,D)) and that difference was partly 
 dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
@@
 copulas and got the similar results but with much prominent difference between their extinction risks
-(Fig. \ref{fig_extrisk_agestr}). This result answered (Q1) and showed extinction risk of 
+(Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
 a population could be considered as
@@
 We found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
-the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}) 
-unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr}). This finding could be explained 
+the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_mtd}) 
+unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr_etd}). This finding could be explained 
 by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models
-against time (shading represents 95\% confidence interval).
+\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models against time (shading represents 95\% confidence interval).
 Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
-\label{fig_extrisk_agestr}}
+\label{fig_extrisk_agestr_etd}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_extremenoise.pdf}
 \caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal
@@
 \includegraphics[width=10 cm]{./Results/common_legend.pdf}\\
-\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0.pdf}
-\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0.pdf}\\
+\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extremenoise.pdf}
+\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extremenoise.pdf}\\
 \vspace{-0.6 cm}
-\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf}
-\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf}\\
+\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extremenoise.pdf}
+\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extremenoise.pdf}\\
 \vspace{-0.6 cm}
-\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0.pdf}
-\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf}\\
+\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extremenoise.pdf}
+\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extremenoise.pdf}\\
 \caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 

### 2019-08-26 SuppMat.Rmd
deleting G-SG results, only include C-SC as moderate tail dep.

@@
 ---
-title: "Supporting Information : Extreme climatic events with asymmetric tail-dependence affect population-extinction risk"
+title: "Supporting Information : Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 source("mtime.R")
-```
-
-```{r sim_linear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+# Note : for moderate tail dep. we used Kendall's tau = 0.6 to get C or SC
+```
+
+```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+# extinction risk with moderate tail dep. for single patch linear age structure model
+
 set.seed(seed)
@@
 saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-
-# Now for gumbel
-
-fam<-4
-copG<-gumbelCopula(2)
-parG<-iTau(copG,tau = ktau)
-par<-parG
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-
-#---------------------------------------------------------------------------------
-
-# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same
-
-fam<-14
-par<-parG
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-```
-
-```{r fig_agestr_C_SC_G_SG_linear, echo=F,results="hide"}
+```
+
+```{r fig_agestr_mtd_linear, echo=F,results="hide"}
 xL<-C_age$extdf
@@
 dev.off()
-
-xL<-SG_age$extdf
-xR<-G_age$extdf
-t<-c(0:(nrow(xL)-1))
-pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf",sep=""),height=5,width=5)
-op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
-x<-xL
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
-lines(t,x$erA,col="red",type="l")
-x<-xR
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
-lines(t,x$erA,col="blue",type="l")
-legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
-par(op)
-dev.off()
-```
-
-```{r sim_nonlinear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+```
+
+```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+# extinction risk with moderate tail dep. for single patch non-linear age structure model
+
 set.seed(seed)
@@
 saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-
-# Now for gumbel
-
-fam<-4
-copG<-gumbelCopula(2)
-parG<-iTau(copG,tau = ktau)
-par<-parG
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-
-#---------------------------------------------------------------------------------
-
-# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same
-
-fam<-14
-par<-parG
-
-cop<-BiCopSim(N=(numsteps*numsims), family = fam, par = par)
-
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-
-SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
-            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
-saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
-#---------------------------------------------------------------------------------
-```
-
-```{r fig_agestr_C_SC_G_SG_nonlinear, echo=F,results="hide"}
+```
+
+```{r fig_agestr_mtd_nonlinear, echo=F,results="hide"}
 xL<-C_age$extdf
@@
 dev.off()
-
-xL<-SG_age$extdf[1:26,]
-xR<-G_age$extdf[1:26,]
-t<-c(0:(nrow(xL)-1))
-pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
-op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
-plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
-x<-xL
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
-lines(t,x$erA,col="red",type="l")
-x<-xR
-polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
-lines(t,x$erA,col="blue",type="l")
-legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
-par(op)
-dev.off()
 ```
@@
 ```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
+# extinction risk with extreme tail dep. for single patch linear age structure model
+
 set.seed(seed)
@@
 ```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
+# extinction risk with extreme tail dep. for single patch non-linear age structure model
+
 set.seed(seed)
@@
 
-<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
+<!--Supp. FIG :  age structure model results for C, SC with linear model (\Omega = infinity)-->
 \begin{figure}[!h]
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
-\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
-\vspace{-0.2 cm}
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
-\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C, SG) or right tail (SC, G) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time; fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure.
-$(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000.
-$(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$).
-All copula has same Kendall correlation $\tau = 0.6$.
-\label{fig_extrisk_agestr_C_SC_G_SG}}
+\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C) or right tail (SC) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time for 10000 simulations.
+Fecundity of adults and survival of eggs have moderate tail dependence in either left or right tail in their corresponding copula structure.
+Moderate left tail dependence (or right tail dependence) was generated by a bivariate Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6, $(A)$ for linear model ($\Omega = \infty$) and for $(B)$ non-linear model ($\Omega = 100$).\label{fig_extrisk_agestr_mtd}}
 \end{center}

### 2019-08-27 MainText.Rmd
editing methods...continued

@@
 We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
-(Q2) How does asymmetric noise copula (tail dependence across multiple patches) influence extinction 
-risk of a density-dependent metapopulation model? Our speculation is that the general finding 
-from density-independent metapopulation model [@Ghosh_copula]
-(higher extinction risk if noises across patches were 
-left-tail dependent than the one if noises had right-tail dependence) would hold true only in
-the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
+(Q2) How does environmental noises with asymmetric tail dependence across multiple patches
+influence extinction risk of a density-dependent metapopulation model? Our speculation 
+is that the general finding from density-independent metapopulation model [@Ghosh_copula]
+(higher extinction risk if noises across patches were left-tail dependent than the one 
+if noises had right-tail dependence) would hold true only in the under-compensatory regime, but 
+should be reverse if the model shows over-compensatory dynamics. 
 (Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
@@
 <!--stochastic matrix model : age structured-->
-To investigate (Q1) from \nameref{Introduction} that how copula structure in noisy 
-environment affect the extinction risk of
+To investigate (Q1) from \nameref{Introduction} that how does asymmetric tail 
+dependent noise affect the extinction risk of
 populations in a single patch age-structured model,
-we considered a basic stochastic age-structured model (Eq. \ref{eq_age_str}) 
+we considered a stochastic age-structured model (Eq. \ref{eq_age_str}) 
 with two age classes (egg $E$ and female adult $A$). We assumed
@@
 reproductive episode only once before death. Eggs and adults in the 
-next generation depend upon the fecundity of adult ($f_a \geq 0$) 
-and survival of eggs ($0 \leq s_e \leq 1$) from present generation, respectively.
-The environmental stochasticity in the model was embedded in 
-the fluctuating vital rates ($f_a$ and $s_e$). We considered the fecundity ($f_a$)
-and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
-a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distributions 
-were generated from iid noises showing specific tail-dependence (either in left- or in right-tail)
-between fecundity and survival rates. Tail dependence emerged if extreme bad years ("catastrophs") 
-would suppress both vital rates (hence, $f_a$ and $s_e$ would show a left tail dependent copula), whereas 
-extreme good years ("bonanzas") would increase both of them (i.e., $f_a$ and $s_e$ would show 
-a right tail dependent copula). Eq. \ref{eq_vitalrates} represent the stochasticity 
-in the vital rates. We included the density dependence effect in one of the vital rate,
-namely, fecundity ($f_a$). If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, 
-the projection matrix would be independent
-of populations from any age-class and the model was then called to be "linear", whereas
-its "nonlinear" version had finite $\Omega$.
+next generation evolved depending on the fecundity of adult ($f_a \geq 0$) 
+and survival of eggs ($0 \leq s_e \leq 1$) from present generation.
+Fluctuating vital rates ($f_a$ and $s_e$) captured the environmental
+stochasticity (Eq. \ref{eq_vitalrates}). We made adult's fecundity 
+($f_a$) negative density-dependent with its own population. When $\Omega = \infty$ in 
+Eq. \ref{eq_vitalrates}, the projection matrix was independent of populations 
+(i.e., Eq. \ref{eq_age_str} was "linear" model then), 
+but when $\Omega$ was finite, Eq. \ref{eq_age_str} became "nonlinear".
+Fecundity ($f_a$) and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
+a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. We generated those distributions 
+from iid noises with a varied degree of asymmetric tail-dependence between 
+fecundity ($f_a$) and survival ($s_e$) rates. Degree of tail dependence had two broad 
+categories: moderate and extreme. Each category had specific direction in tail dependence: 
+either in left or in right tail of vital rates. Vital rates ($f_a$ and $s_e$) were
+left tail dependent when unusual bad years ("catastrophs") suppressed both vital rates synchronously, 
+and right tail dependent when unexpected good years ("bonanzas") boosted both of them.
 
@@
 
+We used bivariate Clayton copula to model moderate left-tail and Survival Clayton copula to model 
+moderate right tail dependence in the vital rates. Both copula had the same Kendall 
+correlation ($=0.6$) but opposite dependence in their respective tails. 
+We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
+right (Survival Clayton) tail-dependent bivariate copulas (see `BiCopSim` function from
+`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
+by number of simulations ($=10000$) we used. 
+We chose parameters for each copula family using 
+`iTau` function from `copula` package so that both copula had
+same Kendall's correlation ($\tau = 0.6$). Entries within each clumn of
+matrix $M$, then became independent of each others but 
+entries between columns had a specific tail dependence
+with positive Kendall's correlation ($\tau = 0.6$). 
+Similarly, we generated extreme left- (or right-) tail dependent matrix M 
+from standard iid (Appendix \ref{SM-extr_tail_dep}) and made the marginals uniform so that
+it could mimic an extreme tail dependent copula.
+
+We made the first column of matrix 
+$M$ as gamma ($\Gamma(k,\theta)$) distributed 
+with shape parameter $k = 2$ and scale parameter $\theta = 1$.
+This distribution represented fecundity of the adult population, $f_a \geq 0$.
+For a given $\theta$, increasing $k$ would shift the distribution towards
+higher fecundity and reduce the extinction risk if the other parameters
+remain unchanged. We took the second column of $M$ and applied a 
+beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
+$\beta = 2$ were the shape parameters that could control the distribution 
+of the survival rates of eggs ($s_e$) ranging in between 0 to 1. 
+For each $25$ timesteps from each of $10000$ simulations,
+we feed pairwise entries from each row of matrix $M$ as the vital rates 
+(fecundity $f_a$ and survival rates $s_e$) 
+of the stochastic projrction matrix in model (\ref{eq_age_str}).
+
 <!--calculation of ext risk with age-structured model-->
@@
 and following Eq. \ref{eq_age_str} we checked whether after each time step the total population $(E+A)_{t}$
-was below a certain extinction threshold (we set it as $2$) or not. If it was below that threshold then we 
+was below a certain extinction threshold (we set it as $2$) or not. If total population was below that threshold then we 
 considered population went extinct for that simulation. We also calculated $2.5^{th}$ and $97.5^{th}$ quantiles
-of $10000$ simulations with the probability of computed extinction risk at each time step.
-
-To test our hypothesis from (Q1), we repeated the above analysis 
-considering the vital rates coming from an asymmetric copula with specific 
-left (Clayton, Survival Gumbel)
-or right (Survival Clayton, Gumbel) tail dependence
-or from an extreme tail-depentdent one, 
-when vital rates were considered to be affected
-by extreme climatic events. A $N \times 2$ matrix, M, was generated
-using `BiCopSim` function from `VineCopula` package 
-for each of the left (Clayton, Survival Gumbel) and right 
-(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
-where $N$ is number of years ($25$) multiplied by number of simulations ($10000$) we used. 
-We kept the same Kendall's correlation ($\tau = 0.6$) for 
-each bivariate copula from one parameter family which used its parameter accordingly
-(see `iTau` function from `copula` package). Entries within each clumn of
-matrix $M$, were thus 
-independent of each others but 
-the two columns of $M$ had a specified Kendall's 
-correlation ($\tau = 0.6$) and underlying tail-dependence. 
-Similarly, for extreme-tail dependent bivariant copula
-(either left or right tail) a matrix M was generated from standard iid as described in 
-Appendix \ref{SM-extr_tail_dep}.
-
-We made the first column of matrix 
-$M$ as gamma ($\Gamma(k,\theta)$) distributed. 
-$k = 2$ and $\theta = 1$ were shape and scale parameters, 
-respectively. This column represented the distribution of fecundity of the adult population, $f_a$.
-For a given $\theta$, increasing 
-$k$ indicates distribution will shift towards higher fecundity which might 
-reduce the extinction risk if the other parameters
-remain unchanged. We took the second column of $M$ and applied a 
-beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
-$\beta = 2$ were the shape parameters which could control the distribution 
-of the survival rates of eggs, $s_e$. 
-For each $25$ timesteps from each of $10000$ simulations,
-the vital rates (fecundity $f_a$ and survival rates $s_e$) 
-of the stochastic projrction matrix in model (\ref{eq_age_str})
-were drawn pairwise from each row of matrix $M$.
-[`sim_age_str`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R) 
-is the function we used to calculate extinction risk for age structured model.
+of $10000$ simulations with the probability of computed extinction risk at each time step (for details see [`code`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R)).
 
@@
 and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
-All the models could be represented via 
-a generalized form : $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
+All models assumed a generalized form : $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
 where $\vec{P_t}$ was a vector of population densities, 
@@
 with the provision of "local" or "global" dispersal among patches. A n $\times$ n dispersal
-matrix, $D$ was constructed assuming that a fraction $d$ of 
+matrix, $D$ assumed that a fraction $d$ of 
 each population patch dispersed during any given time step, equally distributed to
 the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
-$d$ that dispersed
-from each patch was divided up equally among all other patches. We called $d$ as dispersal
+$d$ that dispersed from each patch was divided up equally among all other patches. We called $d$ as dispersal
 rate with $d$ being zero there was no dispersal, and threby no chance to rescue an extinct 
@@
 column of Table \ref{tab_model_summary}). Noises were 
-independent through time, and showed the same 
+independent through time with same 
 spatial correlations between patches (to within sampling error) for every simulation, 
-but they were made to exhibit extreme right- or extreme left-tail dependence between patches in
-different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
+but moderately (or extremely) left- and right-tail dependent between patches in
+different simulations.
+
+<!--Shya: ends here-->
+
+A three dimensional array, $n_s \times n \times n_t$, 
 $\epsilon$, was made from these extreme tail-dependence noise 

### 2019-08-28 MainText.Rmd
methods edited, started to edit intro

@@
 <!--how common are the extreme events?-->
-For past few decades, there have been increasing evidences that anthropogenic climate change 
-is causing extreme events, e.g. temperature or precipitation extremes,
+For past few decades, extreme climatic events (ECE) appeared more common than 
+usual due to anthropogenic climate change, e.g. temperature or precipitation extremes,
 heat-waves, wild-fire etc. [@ummenhofer2017extreme]. Frequent extreme events 
@@
 change variability of the distribution, as well as change its 
-symmetry or skewness [@hansen2012perception]. Such extreme climatic events (ECE) can pose severe
+symmetry or skewness [@hansen2012perception]. Extreme climatic events can pose severe
 impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
@@
 Earlier studies showed ECEs may affect species abundances (species re-ordering) through 
-life-history traits and this alteration in community structures could cascade 
-through higher hierarchical levels and lead to local extinction of 
+life-history traits, alter community structures through cascading effect on 
+higher hierarchical levels and lead to local extinction of 
 species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme]. 
 <!--Should I mention idea of match-mismatch hypothesis?-->
-For example, 
-a study based on annal survival data showed changes
-in prey communities induced by extreme flooding affected predators in 
-aquatic and adjacent riparian habitats [@chiu2013impact] via trophic linkages. Similarly, 
-$43$ years of ring-recovery data was used to examine the demographic impact of ECEs 
+For example, a study [@chiu2013impact] showed extreme flood affected 
+the annal survival of a prey community that in turn affected predators in 
+aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
+@frederiksen2008demographic used $43$ years of ring-recovery data 
+to examine the demographic impact of ECEs 
 through temporal variability in juvenile, immature and adult survival
-of European shags on the Isle of May in eastern Scotland [@frederiksen2008demographic].
-A recent study [@mcdermott2017sensitivity] investigated the sensitivity 
+of European shags on the Isle of May in eastern Scotland.
+A relatively recent study [@mcdermott2017sensitivity] investigated the sensitivity 
 of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
-extreme precipitation, drought) over $37$ years (1976-2012) at the site level, across 
+extreme precipitation, drought) over $37$ years at the site level, across 
 each species' life-stages. Experimental study showed successive ECEs with short time intervals 
@@
 
+<!--Earlier study on metapopulation extinction risk-->
+
 <!--How copula could be relevent in estimating extreme event's 
 influence on extinction risk : introduce Taildep. concept-->
-A recent study [@Ghosh_copula] showed the importance, causes 
-and consequences of copula structures in associations between ecologically
-important variables and demonstrated a basic example where 
-copula was used to model extinction risk of a density-independent 
-metapopulation. Present manuscript is an extension for the application of copula 
+A recent study [@Ghosh_copula] used the tail dependence feature 
+to estimate extinction risk of a density-independent 
+metapopulation model. Present manuscript is an extension for the application of copula 
 in estimating the extinction risk as introduced by [@Ghosh_copula]. We will show 
-why the potential use of copula
-in modelling extinction risk goes beyond the density-independent model 
+why the potential use of copula in modelling extinction risk goes beyond the density-independent model 
 and even applicable for a simple age-structured population where its vital rates are influenced by ECEs. 
@@
 considered population went extinct for that simulation. We also calculated $2.5^{th}$ and $97.5^{th}$ quantiles
-of $10000$ simulations with the probability of computed extinction risk at each time step (for details see [`code`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R)).
+of $10000$ simulations with the probability of computed extinction risk at each time step (see [`code`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R)).
 
@@
 the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
-$d$ that dispersed from each patch was divided up equally among all other patches. We called $d$ as dispersal
-rate with $d$ being zero there was no dispersal, and threby no chance to rescue an extinct 
+$d$ dispersed equally to other $(n-1)$ patches. We called $d$ as dispersal
+rate with $d$ being zero there was no dispersal, and hence no chance to rescue an extinct 
 patch through migration from nearby metapopulation patches.
 
-The environmental noises from any two out of $n$ patches, say, 
-$(\epsilon_t^i, \epsilon_t^j)$ were normally 
-distributed within each patch with mean $m=0$ and standard deviation $\sigma$ 
+We simulated our metapopulation models with noises $(\epsilon_t^i)$ that were 
+independent through time with same spatial correlations between patches (to within sampling error)
+for every simulation, but moderately (or extremely) left- and right-tail dependent between patches in
+different simulations. For moderately left-tail dependent noise we drew points from a $n$ dimensional 
+Clayton copula with same Kendall correlation ($= 0.6$) between any two out of $n$ patches and made 
+the marginals normal with zero mean and standard deviation $\sigma$ 
 (see the specified value $\sigma$ for each model in 3$^{rd}$ 
-column of Table \ref{tab_model_summary}). Noises were 
-independent through time with same 
-spatial correlations between patches (to within sampling error) for every simulation, 
-but moderately (or extremely) left- and right-tail dependent between patches in
-different simulations.
-
-<!--Shya: ends here-->
-
-A three dimensional array, $n_s \times n \times n_t$, 
-$\epsilon$, was made from these extreme tail-dependence noise 
-matrix generated via algorithm stated in 
-Appendix \ref{SM-extr_tail_dep} and used to calculate extinction risk 
-for each of the specified model in Table \ref{tab_model_summary} for $numsteps = 25$ years, $n =$ 5 patches
-and $numsims = 10000$ simulations.
-
-The models stated in Table \ref{tab_model_summary} had one thing in 
-common that they showed undercompensatory
-dynamics (monotonic approach towards stable steady state $P_e$) when
-growth parameter $r$ remained below a critical 
-level, say, $r_c$. If $r$ was beyond $r_c$, model showed 
-overcompensatory dynamics with oscillatory approach 
-towards stable equilibrium $P_e$. But if $r$ became too high and went
-beyond $r_{bf}$, the parameter value responsible for Hopf-bifurcation,
-model equilibrium $P_e$ became unstable and model could show sustained
-oscillation and eventually led to chaos. Appendix \ref{SM-model_stability}
-summarizes the stability analysis of each model in their deterministic form.
+column of Table \ref{tab_model_summary}). For moderately 
+right-tail dependent noise we flipped ($180^{\circ}$ rotation) the points drawn 
+from Clayton copula so that all the statistical properties remained 
+same except for the direction of tail-dependence. We generated
+extreme left- or right-tail dependent noises via algorithm stated in 
+Appendix \ref{SM-extr_tail_dep}. Our models (Table \ref{tab_model_summary}) 
+computed extinction risk for $n_t = 25$ years, $n = 5$ patches
+and $n_s = 10000$ simulations with both moderate and extreme tail dependence (left-, right-tail).
 
@@
 for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
-We assumed population went extinct after each time step if it was below a certain threshold.
-That extinction threshold was set to 10\% of initial population $P_e$ for individual model.
+We assumed population went extinct after each time step if it was below a 
+certain threshold (set to 10\% of initial population $P_e$ for individual model).
+
+The models stated in Table \ref{tab_model_summary} had one common 
+feature: they showed undercompensatory
+dynamics (monotonic approach towards stable steady state $P_e$) when
+growth parameter $r$ remained below a critical 
+level, say, $r_c$. Models showed 
+overcompensatory dynamics with oscillatory approach 
+towards stable equilibrium $P_e$ when $r$ was beyond $r_c$. 
+If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
+model equilibrium $P_e$ became unstable. Beyond Hopf-bifurcation point, model showed sustained
+oscillation and eventually led to chaos with increasing $r$. Appendix \ref{SM-model_stability}
+summarized the stability analysis of each model in their deterministic form.
 
@@
 To test our hypothesis from (Q2) of \nameref{Introduction} for each metapopulation model,
-we estimated extinction risk for a wide range of parameter $r$, 
-which covered the undercompensatory and overcompensatory regime. We 
+we estimated extinction risk for a wide range of parameter $r$ 
+that covered the undercompensatory and overcompensatory regime. We 
 chose equispaced values of $r$ from the 95\% quantiles 
 of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
-for which the model equilibrium $P_e$ had always a finite, positive initial 
-population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ from each 
-model where metapopulation could disperse locally with varying dispersal 
-rate $d= 0,0.1,0.2,..0.5$, we calculated extinction risk with left 
-tail dependent and right tail dependent environmental noise 
-in different simulations, respectively. We also repeated the analysis 
-for global dispersal for each metapopulation model. Script 
-[`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R) was used for this 
-analysis.
+to get a positive, finite equilibrium $P_e$ (for details, see Appendix 
+\ref{SM-model_stability}) to start with ($t=0$). For each chosen $r$,
+we calculated extinction risk after $25$ years with each metapopulation model 
+where population could disperse locally (or globally) with varying dispersal 
+rate $d= 0,0.1,0.2,..0.5$ and extreme (or moderate) left 
+tail- and right-tail dependent environmental noise 
+in different simulations (for code see [`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R)).
 
 <!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->
-Given hypothesis from (Q2) was valid which scrutinized that whereas 
-left tail dependence in noise made extinction risk 
-higher in undercompensatory regime ($r < r_c$), it was the 
-right tail dependent noise
-which appeared to evoke higher extinction risk for metapopulations
-in overcompensatory regime ($r > r_c$), we tested our third hypothesis for (Q3).
-We considered
-the switching to take place near $r_c$, not exactly at $r = r_c$ because the transition point $r_c$
-from under to overcompensatory dynamics was expected to differ in stochastic environment than its 
+Given hypothesis from (Q2) was valid - that left tail dependent noise made extinction risk 
+higher in undercompensatory regime ($r < r_c$) and
+right tail dependent noise made extinction risk higher
+in overcompensatory regime ($r > r_c$) - we tested our third hypothesis for (Q3).
+We expected the switching to take place near $r_c$, not exactly at $r = r_c$ because the transition point $r_c$
+from under to overcompensatory dynamics differed in stochastic environment than its 
 deterministic analog depending on the underlying nonlinearity of the model.
-We generated plots with extinction risk along y-axis
+We plotted extinction risk along y-axis
 for both left tail and right tail dependent noises 
-against the varying range of $r$ (along x-axis) when no dispersal ($d=0$) was considered.
+against the varying range of $r$ along x-axis when no dispersal ($d=0$) was considered.
 We excluded the possibility of dispersal in this case to avoid any rescue effect among patches. 
-For each model with extreme left and extreme right tail dependent noises, we calculated 
+For each model with extreme (or moderate) left- and right-tail dependent noises, we calculated 
 2.5$^{th}$ and 97.5$^{th}$ quantiles
@@
 \caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models against time (shading represents 95\% confidence interval).
-Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000.
+Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, $n_s=10000$.
 \label{fig_extrisk_agestr_etd}}
@@
 \caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal
-among $n=5$ patches, extinction risk after $numsteps = 25$ years for $numsims = 10000$ simulations were calculated.
+among $n=5$ patches, extinction risk after $n_t = 25$ years for $n_s = 10000$ simulations were calculated.
 Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.

### 2019-08-28 SuppMat.Rmd
editing methods...continued
methods edited, started to edit intro

@@
 For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
-standard-normal random draws (in general), where $N = numsteps(=25) \times numsims(=10000)$.
-Then, each row of
-M was identified for a first set of modifications
+standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->.
+Then, each row of M was identified for a first set of modifications
 randomly and independently with a 50\% chance; for those rows identified, every

### 2019-08-29 MainText.Rmd
intro edited

@@
 on extinction risk of a population is necessery. We here introduced the potential use of "tail-dependence" 
-property of a copula in modelling the extinction risk of an age-structured single patch model. We also 
+property of a copula in modeling the extinction risk of an age-structured single patch model. We also 
 extended the generalized findings of density-independent metapopulation model from @Ghosh_copula 
@@
 
-<!--Earlier study on metapopulation extinction risk-->
+<!--Earlier study on metapopulation extinction risk : Large-scale spatial synchrony may also be generated by
+other processes such as migration and predation, but regional
+stochasticity in the form of spatially correlated weather conditions is
+probably the dominant synchronizing mechanism (ref: Nature1998) but what about taildep. effect on extrisk?.-->
+In the metapopulation context, regional stochasticity 
+can simultaneously affected multiple habitat patches where spatially correlated weather conditions were 
+probably the dominant synchronizing mechanism for extinctions [@hanski1998metapopulation]. A 21 years-long
+study by @tack2015increasing on the Glanville fritillary butterfly demonstrated that frequency 
+of drought during early larval instars increased amplitude and span of spatial synchrony 
+over a metapopulation network of 4000 dry meadows and could affect the long-term population viability.
+Though earlier literature considered the effect of spatial and/ or temporal correlation in estimating extinction eisk of a metapopulation, but to our knowledge, using the concept of "tail dependence" (i.e.
+correlation would be stronger during extreme years) to model extinction risk remains unexplored.
 
@@
 influence on extinction risk : introduce Taildep. concept-->
-A recent study [@Ghosh_copula] used the tail dependence feature 
-to estimate extinction risk of a density-independent 
-metapopulation model. Present manuscript is an extension for the application of copula 
-in estimating the extinction risk as introduced by [@Ghosh_copula]. We will show 
-why the potential use of copula in modelling extinction risk goes beyond the density-independent model 
+We recently [@Ghosh_copula] used this tail dependence feature using "copula" - a statistical 
+tool broadly used in modeling dependence between variables - to estimate extinction risk of a density-independent 
+metapopulation model. Present manuscript is an extension of this concept 
+in estimating the extinction risk beyond the density-independent model 
 and even applicable for a simple age-structured population where its vital rates are influenced by ECEs. 
-Following we will introduce the concept of copula in a manner that follows @Ghosh_copula.
+Following we will reintroduce the concept of copula in a manner that follows @Ghosh_copula.
 
-<!--brief intro for copula-->
+<!--brief intro for copula: I am not sure whether it is needed in this paper or not?-->
 Copula, a rank based measure of association, is much more informative 
@@
 (respectively, the right or upper portions, i.e. for higher values) 
-of the two distributions. Stronger dependence in either tail gives rise to asymmetric copula 
-structure for the joint distribution of two variables. 
-For example, extreme cold winter during growing season
+of the two distributions. Stronger dependence in either tail causes asymmetric copula 
+structure for the joint distribution of two variables. For example, Clayton and Survival Clayton are 
+two copula families with asymmetric tail that are used to model exclusively 
+left and right tail dependence between variables, respectively.
+From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
 might reduce the adult reproduction rate as well as larval survival (i.e. stronger 
-tail dependence between vital rates when temperature is too cold than normal), which 
-could lead to population crash and thereby inflate the extinction risk. Similarly, in context of 
+tail dependence between vital rates when temperature is too cold than normal) and 
+thereby inflate the population viability. Similarly, in context of 
 density-dependent metapopulation models, if extreme environmental fluctuations across sites 
@@
 during unfavorable years unless they are rescued by the nearby populations. This scenario 
-could be more complex depending 
-on the dynamics of the model (see \nameref{Results}). Though tail-dependence 
+could be more complex and specific of the direction of dependence 
+depending on the dynamics of the model (see \nameref{Results}). Though tail-dependence 
 property of a copula has been used for risk-analysis in other fields like finance, 
-hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our knowledge
-copula have never been used in estimating the effect of ECEs through extinction risk of age structured 
+hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
+knowledge asymmetric tail dependence have never been considered 
+in estimating the effect of ECEs through extinction risk of age structured 
 model or density-dependent metapopulations.
@@
 <!--our objective ?-->
-In this paper, we will explore the potential use of asymmetric tail-dependence in a copula 
+In this paper, we will explore the potential use of asymmetric tail-dependence 
 for understanding the extinction risk of populations towards extreme climatic event
@@
 4. extinction risk how propagated through a food web?
+5. ref: see McCarthy2000: Most previous models of metapopulation viability have either ignored
+correlated extinction of patches, or have included uniform correlations, with the correlation between nearby patches being the same as the correlation between distant
+patches (Hanski 1994; Tilman et al. 1994; Lindenmayer et al. 1995; McCarthy
+1996). 
+6. see this: http://dx.doi.org/10.1098/rspb.2015.0173
 -->

### 2019-08-29 MainText.Rmd
started editing results

@@
 
-For single patch linear age-structured models, we found left-tail 
-dependence (Clayton, Suvival Gumbel copulas) in noisy vital 
-rates enhanced the extinction risk compared to the case when
-vital rates were right tail dependent (Survival clayton, Gumbel copulas).
-It was expected because a sudden "bad" year (e.g. with severe cold) can suppress 
+For single patch linear age-structured models, we found higher extinction risk for moderate left-tail 
+dependence (modeled by Clayton copula) in fluctuating vital 
+rates compared with the one when vital rates were right tail dependent (modeled by Survival Clayton copula).
+It was expected because an extreme "bad" year (e.g. with severe cold) can suppress 
 the fecundity of adult and survival/ maturation of eggs simultaneously 
-and thereby increasing the extinction risk of the population. For non-linear age-structured 
+and thereby can increase the extinction risk of the population. For non-linear age-structured 
 models, we got the similar trend for extinction risks with different tail-dependence (left and right).
-With non-linear model the extinction risks appeared as higher than their corresponding 
+Non-linear model showed higher extinction risk than its corresponding 
 linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
-Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with 4 different 
-asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
-difference between extinction risks within a pair of different tail-dependent copulas
-(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_mtd}(A,C) and
-G-SG in Fig. \ref{SM-fig_extrisk_agestr_mtd}(B,D)) and that difference was partly 
-dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
+Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with two copulas having different 
+asymmetric tails but same Kendall's correlation ($\tau = 0.6$). Extinction risks from 
+two different tail-dependent copulas were not strinkingly different (e.g. see pair of red and
+blue lines in Fig. \ref{SM-fig_extrisk_agestr_mtd}A or B) and we anticipated the difference being partly 
+dependent on the degree of tail dependence. Therefore we repeated the same analysis with 
 vital rates coming from extreme-tail dependent (left and right in two different simulations)
@@
 (Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
-a population could be considered as
-direct manifestation of tail-dependence measure between vital rates of different age-class
-within that population. <!--discussion later : add Tuljapurkar's approximation-->
+a population could be considered as direct manifestation of tail-dependence measure 
+between vital rates of different age-class
+of that population. <!--discussion later : add Tuljapurkar's approximation-->
 
@@
 
-We found density-dependence metapopulation models (with $n=5$ patches) in noisy 
+We found density-dependent metapopulation models in stochastic 
 environment showed similar patterns for extinction risk as density-independent 
-metapopulation model showed in earlier studies [@Ghosh_copula]. For each density-dependent 
-model listed in Table \ref{tab_model_summary}, we observed higher (or lower) extinction risk for 
-the metapopulations when noises across multiple patches were left- (or right) tail dependent.
-But that observation remained valid when the model parameter, usually growth rate $(r)$, of each model was 
+metapopulation model showed in recent study [@Ghosh_copula]. For each density-dependent 
+model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
+the metapopulations when noises across multiple patches were left-tail dependent than the one with right-tail dependence.
+But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - of each model was 
 sufficiently low so that under-compensatory dynamics (monotonous approach towards
-its stable equilibrium) for the deterministic analog of the model was ensured. When the model 
+its stable equilibrium) for the deterministic analog of the model was ensured.
+We observed right tail dependent noise caused higher extinction 
+risk for the metapopulations than left tail dependent noise when the model 
 was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
-value of $r$, we observed right (or left) tail dependent noise caused higher (or lower) extinction 
-risk for the metapopulations. Fig. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
+value of $r$. Fig. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
 against local dispersal rate ($d$) for six density dependent metapopulation models 

### 2019-08-29 SuppMat.Rmd
started editing results

@@
 
-pdf("./Results/common_legend.pdf",height=1,width=12)
+pdf("./Results/common_legend_extreme_noise.pdf",height=1,width=12)
 op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("center", c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
+legend("center", c("noise: extreme LTdep.","noise: extreme RTdep."), lty=c(1,1), pch=c(1,1), 
+         col=c('red', 'blue'),x.intersp = 0.2,
+         cex = 2.5, xpd = TRUE, horiz = T, inset = c(0,0), 
+         bty = "n")
+par(op)
+dev.off()
+
+pdf("./Results/common_legend_moderate_noise.pdf",height=1,width=12)
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
+plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+legend("center", c("noise: moderate LTdep.","noise: moderate RTdep."), lty=c(1,1), pch=c(1,1), 
          col=c('red', 'blue'),x.intersp = 0.2,
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
-       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
-mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+legend("topleft", c("noise: extreme LTdep.","noise: extreme RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0, x.intersp=0.2)
+mtext(paste("Ricker model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 #mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-mtext(paste("Hassell model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("Hassell model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-mtext(paste("Maynard Smith model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("Maynard Smith model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-mtext(paste("Pennycuick model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("Pennycuick model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-mtext(paste("Verhulst model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("Verhulst model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-mtext(paste("Austin-Brewer model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(paste("Austin-Brewer model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
-       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
-mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+legend("topleft", c("noise: moderate LTdep.","noise: moderate RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0, x.intersp=0.2)
+mtext(paste("Ricker model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 #mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
-       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
-mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+legend("topleft", c("noise: extreme LTdep.","noise: extreme RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0,x.intersp=0.2)
+mtext(paste("Ricker model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-legend(x=0.7,y=1.05, c("noise : LTdep.","noise : RTdep."), lty=c(1,1), pch=c(1,1), 
-       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0)
-mtext(paste("Ricker model : "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+legend("topleft", c("noise: moderate LTdep.","noise: moderate RTdep."), lty=c(1,1), pch=c(1,1), 
+       col=c('red', 'blue'), horiz=F, bty='o', cex=1.5, adj=0, x.intersp=0.2)
+mtext(paste("Ricker model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
 mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_extremenoise.pdf}\\
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_extremenoise.pdf}\\

### 2019-08-29 MainText.Rmd
fig,results editting continued

@@
 (Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
-a population could be considered as direct manifestation of tail-dependence measure 
+a population could be considered as direct manifestation of tail-dependence property 
 between vital rates of different age-class
-of that population. <!--discussion later : add Tuljapurkar's approximation-->
+of that population. <!--discussion later: add Tuljapurkar's approximation-->
 
@@
 was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
-value of $r$. Fig. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
-against local dispersal rate ($d$) for six density dependent metapopulation models 
+value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp_etd} and \ref{SM-fig_extrisk_allmodel_localdisp_mtd} show
+variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
 where growth rate $r$ was taken either from under-compensatory or from over-compensatory 
-regime. We also got similarly consisent results with 
+regime. We also got similar results with 
 our prediction (see Q2 from \nameref{Introduction}) for global 
-dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
-Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6}) 
+dispersal (Figs. \ref{SM-fig_extrisk_allmodel_globaldisp_etd} and \ref{SM-fig_extrisk_allmodel_globaldisp_mtd}). 
+Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_etd}) 
 shows how extinction risk changed against local dispersal rate for a whole range
-of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
+of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1_etd} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}) 
 represents the same but for the case when global dispersal took place between patches. 
@@
 <!--Q3 : metapopulation results : switching near rc for d=0, Fig. 3-->
-Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) we hypothesized that the switching 
+Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}) we hypothesized that the switching 
 between the extinction risk generated by left-tail dependent and right-tail dependent noise 
@@
 Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
-Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
+Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp_etd}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend.pdf}\\
+\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
 \includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extremenoise.pdf}
@@
 for each model where their deterministic analog switches from under- to over-compensatory dynamics.
-\label{fig_extrisk_d0}}
+\label{fig_extrisk_d0_etd}}
 \end{center}

### 2019-08-29 SuppMat.Rmd
fig,results editting continued

@@
 lines(t,x$erA,col="blue",type="l")
-legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
+legend(x=-2,y=1.05, c("vital rates: moderate LTdep.,C","vital rates: moderate RTdep.,SC"), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5, x.intersp = 0.2, seg.len = 1)
 par(op)
@@
 lines(t,x$erA,col="blue",type="l")
-legend("topleft", c("vital rates : LTdep.","vital rates : RTdep."), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
+legend(x=-2,y=1.05, c("vital rates: extreme LTdep.","vital rates: extreme RTdep."), lty=c(1,1),  
+       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5, x.intersp = 0.2)
 par(op)
@@
 
-<!--Supp. FIG :  age structure model results for C, SC with linear model (\Omega = infinity)-->
+<!--FIG:  age structure model results for moderate tail dep. (C, SC)-->
 \begin{figure}[!h]
@@
 
-<!--Supp. FIG :  summary : all nonlinear model results for global dispersion : extrisk vs. d-->
+<!--FIG:  summary: all nonlinear model results for local dispersion, moderate noise : extrisk vs. d-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_moderatenoise.pdf}
+\caption[Analogous figure as that of Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail dependence in spatially correlated noises.]{Analogous figure as that of Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail dependence in spatially correlated noises.\label{fig_extrisk_allmodel_localdisp_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG:  summary: all nonlinear model results for global dispersion, extreme noise : extrisk vs. d-->
 \begin{figure}[!h]
@@
 
-<!--Supp. FIG : Ricker model, local dispersal-->
+<!--FIG:  summary: all nonlinear model results for global dispersion, moderate noise : extrisk vs. d-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_moderatenoise.pdf}
+\caption[Analogous figure as that of Fig.
+\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail dependence in spatially correlated noises.]{Analogous figure as that of Fig.
+\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail dependence in spatially correlated noises.\label{fig_extrisk_allmodel_globaldisp_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Ricker model, local dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_ricker_risk_local_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_ricker_risk_local_disp_scl_1_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Hassell model, local dispersal-->
+<!--FIG : Hassell model, local dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Maynard smith model, local dispersal-->
+<!--FIG: Maynard smith model, local dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8}}
+numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Pennycuick model, local dispersal-->
+<!--FIG: Pennycuick model, local dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Verhulst model, local dispersal-->
+<!--FIG: Verhulst model, local dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Austin-Brewer model, local dispersal-->
+<!--FIG: Austin-Brewer model, local dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Ricker model, global dispersal-->
+======================
+<!--FIG: Ricker model, local dispersal, extreme noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of moderate tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
+numsims=10000, $\sigma$=1. \label{fig_ricker_risk_local_disp_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG : Hassell model, local dispersal, extreme noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_moderatenoise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
+numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Maynard smith model, local dispersal, extreme noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_moderatenoise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
+numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Pennycuick model, local dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_moderatenoise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
+numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Verhulst model, local dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_moderatenoise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Austin-Brewer model, local dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_moderatenoise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
+numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
+\end{center}
+\end{figure}
+=========================
+
+<!--FIG : Ricker model, global dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_ricker_risk_global_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_ricker_risk_global_disp_scl_1_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Hassell model, global dispersal-->
+<!--FIG: Hassell model, global dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_hassell_risk_global_disp_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_hassell_risk_global_disp_scl_1_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Maynard smith model, global dispersal-->
+<!--FIG: Maynard smith model, global dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
-numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_global_disp_scl_0.8}}
+numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_global_disp_scl_0.8_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Pennycuick model, global dispersal-->
+<!--FIG: Pennycuick model, global dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
+numsims=10000, $\sigma$=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Verhulst model, global dispersal-->
+<!--FIG: Verhulst model, global dispersal, extreme noise-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_etd}}
 \end{center}
@@
 
-<!--Supp. FIG : Austin brewer model, global dispersal-->
+<!--FIG: Austin brewer model, global dispersal-->
 \begin{figure}[!h]
@@
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
+numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}}
 \end{center}

### 2019-08-30 MainText.Rmd
results edited

@@
 
-<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula : Fig. S1-->
-<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula : Fig. 1 -->
+<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula-->
 
@@
 Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with two copulas having different 
-asymmetric tails but same Kendall's correlation ($\tau = 0.6$). Extinction risks from 
-two different tail-dependent copulas were not strinkingly different (e.g. see pair of red and
+asymmetric tails but same Kendall's correlation ($\tau = 0.6$). Extinction risks for
+two different tail-dependent vital rates were not strinkingly different (e.g. see pair of red and
 blue lines in Fig. \ref{SM-fig_extrisk_agestr_mtd}A or B) and we anticipated the difference being partly 
@@
 <!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->
-<!--For local dispersal : all model summary : Fig. 2, individual model : Fig. S3-S8-->
-<!--For global dispersal : all model summary : Fig. S2, individual model : Fig. S9-S14-->
 
@@
 model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
-the metapopulations when noises across multiple patches were left-tail dependent than the one with right-tail dependence.
-But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - of each model was 
+the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail dependence.
+But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - was 
 sufficiently low so that under-compensatory dynamics (monotonous approach towards
@@
 variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
-where growth rate $r$ was taken either from under-compensatory or from over-compensatory 
-regime. We also got similar results with 
-our prediction (see Q2 from \nameref{Introduction}) for global 
-dispersal (Figs. \ref{SM-fig_extrisk_allmodel_globaldisp_etd} and \ref{SM-fig_extrisk_allmodel_globaldisp_mtd}). 
-Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_etd}) 
-shows how extinction risk changed against local dispersal rate for a whole range
+(see Table \ref{tab_model_summary}) with growth rate $r$ considered either from under-compensatory or 
+from over-compensatory regime. We also got similar results with our prediction (see Q2 from \nameref{Introduction})
+for global dispersal (Figs. \ref{SM-fig_extrisk_allmodel_globaldisp_etd} and \ref{SM-fig_extrisk_allmodel_globaldisp_mtd}). 
+Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}) 
+demonstrate how extinction risk changed against local dispersal rate for a whole range
 of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1_etd} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}) 
-represents the same but for the case when global dispersal took place between patches. 
-Both case showed the signature of 
-"rescue-effect" (i.e. by lowering extinction risk) when there was 
-a finite dispersal compared to the one with 
-$d=0$.
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) 
+represent the same but for global dispersal among patches. 
+Both case showed the signature of "rescue-effect" (i.e. reduced extinction risk for 
+finite dispersal compared with $d=0$).
 
-<!--Q3 : metapopulation results : switching near rc for d=0, Fig. 3-->
+<!--Q3 : metapopulation results : switching near rc for d=0-->
 Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}) we hypothesized that the switching 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) we hypothesized that the switching 
 between the extinction risk generated by left-tail dependent and right-tail dependent noise 
-occurred near the critical parameter $r_c$ (see Q3). For simplicity we only consider 
+occurred near the critical parameter $r_c$ (see Q3). For simplicity we only considered 
 $d=0$ case, i.e., when there was no dispersal between patches (see \nameref{Methods}). 
-Fig. \ref{fig_extrisk_d0} shows higher extinction risk with left-tail dependent noise 
-for $r<r_c$ and with right-tail dependent noise for $r>r_c$. $r_c$ was the vertical
-line in Fig. \ref{fig_extrisk_d0} indicating the value of $r$ for each model where 
-their deterministic analog switches from under- to over-compensatory dynamics. 
-Though the switching visually appered far from the theoretical $r_c$ value for Verhulst 
-and Austin-Brewer model, they were consistent with our hypothesis as they had comperatively 
+Figs. \ref{fig_extrisk_d0_etd} and \ref{SM-fig_extrisk_d0_mtd} demonstrate higher extinction risk 
+with left-tail dependent noise for $r<r_c$ and with right-tail dependent noise 
+for $r>r_c$. $r_c$ was the vertical line in Figs. \ref{fig_extrisk_d0_etd} and \ref{SM-fig_extrisk_d0_mtd} 
+indicating the value of $r$ for each model where their deterministic analog switches 
+from under- to over-compensatory dynamics. Though the switching visually appered far away
+from the theoretical $r_c$ value for Verhulst and Austin-Brewer model, they were 
+consistent with our hypothesis as they had comperatively 
 smaller range of $r$ than the other models.
 
-# Discussion
+# Discussion\label{Discussion}
 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models against time (shading represents 95\% confidence interval).
-Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, $n_s=10000$.
+\caption{Extinction risk against time from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models for $10000$ simulations; shading represents 95\% confidence interval.
+Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence.
 \label{fig_extrisk_agestr_etd}}
@@
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_extremenoise.pdf}
-\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal
-among $n=5$ patches, extinction risk after $n_t = 25$ years for $n_s = 10000$ simulations were calculated.
+\caption{Extinction risks are computed after $25$ years for six nonlinear density-dependence models (Table \ref{tab_model_summary}) with local dispersal among $n=5$ patches and for $10000$ simulations.
 Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
@@
 
- <!--FIG : all nonlinear model results with d=0, extrisk vs. r-->
+ <!--FIG : all nonlinear model results with extreme noise, d=0, extrisk vs. r-->
 \begin{figure*}[!h]
@@
 $5$ multiple patches without dispersal ($d=0$), as intrinsic 
-growth rate ($r$) goes from under to over-compensatory regime. Opposite trend
-was observed with extreme right tail dependent noise.
+growth rate ($r$) goes from under to over-compensatory regime. Extinction risk for 
+extreme right tail dependent noise follows the opposite trend.
 Error bars represent the standard errors associated with $10000$
@@
              format="latex",digits=3, align="c",table.env='table*',
-             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
-Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
+             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
+Superscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"

### 2019-08-30 SuppMat.Rmd
results edited

@@
        angle=90,length=0.1,col="red",code=3)
-abline(v=rc_ricker,lty="dotted",col="darkgrey") #rc=1
+abline(v=rc_ricker,lty="dotted",col="black") #rc=1
 
@@
        y1=ricker_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_ricker,lty="dotted",col="darkgrey") #rc=1
+abline(v=rc_ricker,lty="dotted",col="black") #rc=1
 
@@
        x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_hassell,lty="dotted",col="darkgrey")
+abline(v=rc_hassell,lty="dotted",col="black")
 
@@
        x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_hassell,lty="dotted",col="darkgrey")
+abline(v=rc_hassell,lty="dotted",col="black")
 
@@
        x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_msmith,lty="dotted",col="darkgrey")
+abline(v=rc_msmith,lty="dotted",col="black")
 
@@
        x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_msmith,lty="dotted",col="darkgrey")
+abline(v=rc_msmith,lty="dotted",col="black")
 
@@
        x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_pennycuick,lty="dotted",col="darkgrey")
+abline(v=rc_pennycuick,lty="dotted",col="black")
 
@@
        x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_pennycuick,lty="dotted",col="darkgrey")
+abline(v=rc_pennycuick,lty="dotted",col="black")
 
@@
        y1=verhulst_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_verhulst,lty="dotted",col="darkgrey")
+abline(v=rc_verhulst,lty="dotted",col="black")
 
@@
        y1=verhulst_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_verhulst,lty="dotted",col="darkgrey")
+abline(v=rc_verhulst,lty="dotted",col="black")
 
@@
        y1=abrewer_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_abrewer,lty="dotted",col="darkgrey")
+abline(v=rc_abrewer,lty="dotted",col="black")
 
@@
        y1=abrewer_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="red",code=3)
-abline(v=rc_abrewer,lty="dotted",col="darkgrey")
+abline(v=rc_abrewer,lty="dotted",col="black")
 
@@
 dispersal rate for six density dependent metapopulation model with global dispersal.]{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
-Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp}}
+Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_ricker_risk_local_disp_scl_1_etd}}
+standard deviation $\sigma$=1 was used to generate extreme extreme tail dependent noises for 10000 simulations. \label{fig_ricker_risk_local_disp_scl_1_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1_etd}}
+standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_local_disp_scl_1_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8_etd}}
+standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations.. \label{fig_msmith_risk_local_disp_scl_0.8_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_etd}}
+standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_etd}}
+standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_etd}}
-\end{center}
-\end{figure}
-
-======================
-<!--FIG: Ricker model, local dispersal, extreme noise-->
+standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_etd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Ricker model, local dispersal, moderate noise-->
 \begin{figure}[!h]
@@
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of moderate tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_ricker_risk_local_disp_scl_1_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG : Hassell model, local dispersal, extreme noise-->
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Analogous figure as of Fig.
+\ref{fig_ricker_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_ricker_risk_local_disp_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG : Hassell model, local dispersal, moderate noise-->
 \begin{figure}[!h]
@@
 \includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_hassell_risk_local_disp_scl_1_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Maynard smith model, local dispersal, extreme noise-->
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Analogous figure as of Fig.
+\ref{fig_hassell_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_hassell_risk_local_disp_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Maynard smith model, local dispersal, moderate noise-->
 \begin{figure}[!h]
@@
 \includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with local dispersal.]{Analogous figure as of Fig.
+\ref{fig_msmith_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Analogous figure as of Fig.
+\ref{fig_pennycuick_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Analogous figure as of Fig.
+\ref{fig_verhulst_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
-\end{center}
-\end{figure}
-=========================
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Analogous figure as of Fig.
+\ref{fig_abrewer_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
+\end{center}
+\end{figure}
 
@@
 \caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_ricker_risk_global_disp_scl_1_etd}}
+for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations.
+\label{fig_ricker_risk_global_disp_scl_1_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_hassell_risk_global_disp_scl_1_etd}}
+standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_global_disp_scl_1_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
-numsims=10000, $\sigma$=0.8. \label{fig_msmith_risk_global_disp_scl_0.8_etd}}
+standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_msmith_risk_global_disp_scl_0.8_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
-numsims=10000, $\sigma$=1. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_etd}}
+standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_etd}}
+standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_etd}}
 \end{center}
@@
 for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
-numsims=10000, $\sigma$=0.6. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}}
-\end{center}
-\end{figure}
+standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Ricker model, global dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{fig_ricker_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_ricker_risk_global_disp_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG : Hassell model, global dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{fig_hassell_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_hassell_risk_global_disp_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Maynard smith model, global dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{fig_msmith_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_msmith_risk_global_disp_scl_0.8_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Pennycuick model, global dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{fig_pennycuick_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_risk_global_disp_pennycuick_r_vary_scl_1_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Verhulst model, global dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{fig_verhulst_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG: Austin-Brewer model, global dispersal, moderate noise-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_moderatenoise.pdf}\\
+\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{fig_abrewer_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}}
+\end{center}
+\end{figure}
+
+<!--FIG : all nonlinear model results with moderate noise, d=0, extrisk vs. r-->
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
+\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_moderatenoise.pdf}
+\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_moderatenoise.pdf}\\
+\vspace{-0.6 cm}
+\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_moderatenoise.pdf}
+\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_moderatenoise.pdf}\\
+\vspace{-0.6 cm}
+\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_moderatenoise.pdf}
+\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_moderatenoise.pdf}\\
+\caption[Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
+tail dependent noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
+tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated 
+by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between
+any two out of $n=5$ patches.\label{fig_extrisk_d0_mtd}}
+\end{center}
+\end{figure*}
 

### 2019-08-30 MainText.Rmd
discussion edited

@@
 <!--how common are the extreme events?-->
-For past few decades, extreme climatic events (ECE) appeared more common than 
+For past few decades, extreme climatic events (ECEs) appeared more common than 
 usual due to anthropogenic climate change, e.g. temperature or precipitation extremes,
@@
 Appendix \ref{SM-extr_tail_dep}. Our models (Table \ref{tab_model_summary}) 
-computed extinction risk for $n_t = 25$ years, $n = 5$ patches
-and $n_s = 10000$ simulations with both moderate and extreme tail dependence (left-, right-tail).
+computed extinction risk after $25$ years for $10000$ simulations at $n = 5$ patches
+with spatially correlated noises having either moderate or extreme tail 
+dependence in left and right-tail.
 
@@
 <!--summary results:-->
-Our results show the potential use of "asymmetric" copula in estimating the impact of 
-extreme ecological events on population-extinction risk. We show that through left and right tail 
-dependence in vital rates (respectively, spatially correlated environmental noises), years of extreme events 
-("catastrophs" and "bonanzas") affect extinction risk of a single-patch age-structured population
-(respectively, density dependent metapopulations) differently. Though left tail dependence is 
-expected to show higher extinction risk than the one for right tail dependence for age-structured 
-and density dependent metapopulation models, it is not always true 
-when metapopulation models with dispersal show over-compensatory dynamics. In the over-compensatory regime,
-right tail dependent noises result in higher extinction risk than the left tail dependent one. We show 
-that in absence of dispersal, the above-said switching in extinction risk measure occurs near the 
-critical parameter value where the deterministic model shows transition from 
-under-compensatory to over-compensatory regime.
+We showed the potential use of "asymmetric" tail dependence in modeling the impact of 
+extreme ecological events on population-extinction risk. Our results demonstrated left and right-tail 
+dependence in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
+in years of extreme events ("catastrophs" and "bonanzas"). Though we could expect left tail dependence 
+to show higher extinction risk than the one for right tail dependence with age-structured 
+and density dependent metapopulation models, it was not always true. In the over-compensatory regime,
+right tail dependent noise caused higher extinction risk than the left tail dependent noise. We showed 
+that in absence of dispersal the above-said switching took place near the critical parameter value where 
+the deterministic model showed transition from under-compensatory to over-compensatory regime.
 
@@
 by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
-The approximation considered that the growth rate depends on the most dominant eigen value of the mean projection matrix 
-and the temporal variance of the log(population growth rate) caused by environmental stochasticity, which was based
+The approximation considered growth rate being dependent on the most dominant eigen value of the mean projection matrix 
+and the temporal variance of the log(population growth rate) caused by environmental stochasticity - which was based
 on the linear covariance measure between any two matrix elements, not on their tail-dependence properties. Therefore,
-according to Tuljapurkar's approximation, population growth rate in an iid environment 
-remained same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependent noises.
+following @tuljapurkar1990lecture, population growth rate in an iid environment could be approximated
+as same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependence.
 
-<!--future study: worthy for PVA, IUCN red list
+<!--future study: worthy for PVA, IUCN red list:
 1. Identify which life stage or demographic process as management target?
@@
 4. extinction risk how propagated through a food web?
-5. ref: see McCarthy2000: Most previous models of metapopulation viability have either ignored
-correlated extinction of patches, or have included uniform correlations, with the correlation between nearby patches being the same as the correlation between distant
+5. Not sure : see McCarthy2000: Most previous models of metapopulation viability have either ignored
+correlated extinction of patches, or have included uniform correlations, with the correlation between 
+nearby patches being the same as the correlation between distant
 patches (Hanski 1994; Tilman et al. 1994; Lindenmayer et al. 1995; McCarthy
 1996). 
-6. see this: http://dx.doi.org/10.1098/rspb.2015.0173
 -->
@@
 following scenarios. First, we demonstrated the effect of tail-dependence on extinction risk 
-with a very simplistic age structured (two age stages) model where we considered the total semelperous 
+with a basic age structured (two age stages) model where we considered the total semelperous 
 population would go extinct if there would be only one female adult and one egg. But in reality population 
@@
 so that we can identify (via multisite-PVA) which life-stage or demographic process 
-to target for population management. Earlier study by @doak1994modeling showed PVA models including correlation 
+to target for population management. Earlier study by @doak1994modeling showed PVA models with correlation 
 between different demographic rates indicated marked difference in predicting confidence interval for 
-future population size than the model which did not include correlation. But when 
+future population size compared with the model which did not include correlation. But when 
 such correlation cannot be estimated directly, more generally where the
@@
 with spatially correlated extreme climatic events. Third, we considered tail dependence 
-between vital rates when they were positively correlated, but what would be the extinction risk 
+between positively correlated vital rates, but what would be the extinction risk 
 and viability of a population if demographic rates were negatively correlated with asymmetric tail dependence. 
-This scenario could be considered for those extreme events that kill adults and allow more successful
+We could consider this scenario to model extreme events that kill adults and allow more successful
 recruitment of seeds or other propagules. Fourth, following the aforementioned idea, it would be worthy 
@@
 (i.e., heat waves and droughts) increased mortality and defoliation in southern European forest and its 
-dependent herbivorous insects and fungi through trophic linkage. ECEs can even truncate 
-multiyear insect outbreak dynamics [@esper20061200]. We believe copula (may be a multidimensional one) has 
-a lot more potential to estimate risk of cascading extinction in model food webs.
-
-<!--match-mismatch hypothesis leads to extinction? our phenology paper in prep.-->
+dependent herbivorous insects and fungi through trophic linkage. <!--ECEs can even truncate 
+multiyear insect outbreak dynamics [@esper20061200].--> We believe copula (may be a multidimensional one) has 
+a lot more potential to estimate risk of cascading extinction in model food webs. <!--should I mention about 
+match-mismatch hypothesis leads to extinction? our phenology paper in prep.--> Fifth, we considered 
+our metapopulation patches were equispaced on a linear chain with equal global dispersal 
+rate between patches. But in realistic network models, as inter-patch distances 
+can be different, the dispersal rate should depend on distance: dispersal is possible at higher 
+rates between nearby patches than the patches which are far away. It is worthy to study 
+metapopulation viability with non-uniform spatial correlation (i.e. asymmetric tail dependence) in a 
+network with different clusters having various dispersal rates.
 

### 2019-08-30 SuppMat.Rmd
discussion edited

@@
 For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
-standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->.
+standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->
 Then, each row of M was identified for a first set of modifications

### 2019-08-30 MainText.Rmd
abstract edited

@@
 
-With global climate change, extreme climatic events (ECEs) appear to be more common than usual. These ECEs 
-could pose severe impact on the resistance or resilience of a population or community level through
-life-history traits of the species. Therefore predicting the impact of ECEs 
-on extinction risk of a population is necessery. We here introduced the potential use of "tail-dependence" 
-property of a copula in modeling the extinction risk of an age-structured single patch model. We also 
-extended the generalized findings of density-independent metapopulation model from @Ghosh_copula 
-to density-dependence metapopulation models with this copula approach. For metapopulation models, 
-we showed extinction risk not only depended on the nature of tail-dependence of spatially correlated 
-noise but was also influenced by the inherent dynamics of the population-model.
+With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
+could pose severe impact on the resistance or resilience of a population and/ or community level through
+life-history traits of the species. Therefore predicting the impact of extreme events 
+on population extinction risk is necessery. We here introduce the potential use of "tail-dependence" 
+property of two correlated ecological variables in modeling the extinction risk of 
+an age-structured single patch model. We also extend this idea 
+to density-dependent metapopulation models. For metapopulation models, 
+we show extinction risk not only depends on the nature of tail-dependence of 
+spatially correlated noise but is also influenced by the inherent dynamics 
+of the population-model. To illaborate, we demonstrate left tail dependent noise across spatially correlated patches 
+cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
+But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail-dependence instead of simple correlation in population viability analysis.
 
-\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, copula, density-dependent model
+\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, density-dependent model
 

### 2019-10-07 MainText.Rmd
some comments starting the process of gettin this paper ready for Ecosphere

@@
 
+<!--***DAN: 
+1) Shyamolina,  was reminded during my editing of BIVAN that "tail dependence" has 
+a specific mathematical definition (now given in the Background section of BIVAN), and that
+is not what we mean here when we say "tail dependence". So we cannot go on calling it "tail
+dependence" in this manuscript - that would be mike me calling my house "the white house" - 
+there already is one of those, and it is much better known than my house, so we cannot
+attempt to hijack the name. So can you change terminology throughout, please? Let's refer to
+it using terminology such as "associations in the tails" or 
+"tail associations". I would like you to provide a 
+definition at an appropriate place early in the manuscript that we will call a pair
+of positively associated variables "left-tail associated" if their association in their 
+left tails is stronger than their association in their right tails (in a sense to be defined
+mathematically in Methods, and then in methods we will define things based on cor_l-cor_u).
+We call a pair of positively associated variable "right-tail associated" if their association
+in their right tails is stronger than their association in their left tails.
+
+2) You need to start getting this paper ready for submission - if you want to send it to 
+the Ecosphere special issue (what do you think of that?) then we have to get moving. Please 
+let me know soon, after reading about Ecospere on their website. If you think this is a 
+good idea, let me know and I will send a summary of the paper to Jon Walter to make sure he is still willing to have us submit the paper to him, and to make sure the paper is close enough
+to the theme of the special issue that he is going for, so that he will be willing to see it
+submitted there.
+
+3) Please
+add the figure you have mentioned in the text but that does not exist. Please run a spell
+check. Please go and check every sentence throughout the manuscript for grammatical 
+errors. I think you will find some 
+and will be able to correct them if you focus on that - there are many, and by correcting them
+you will not only make the paper better you will get better at writing grammatically correct 
+sentences in English.
+
+4) Assuming you want to go for Ecosphere and Jon is OK with us submitting there, please read the "guidelines for authors" section of the Ecosphere website 
+and format the paper for Ecosphere.
+
+5) Please make all the improvements that you can see for yourself
+are needed! I suspect if you read the paper carefully now after having worked on something
+else for a while you will see many things you can improve!
+-->
+
 # Introduction\label{Introduction}
@@
 consequences of extreme events on population life history, community-->
-Earlier studies showed ECEs may affect species abundances (species re-ordering) through 
+Earlier studies showed ECEs may affect species abundances (species re-ordering) 
+<!--***DAN: Shyamolina, I do not know what the parenthesis refers to -->
+<!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
+in the appropriate literature, and that is great!-->
+through 
 life-history traits, alter community structures through cascading effect on 
@@
 to local extinction. 
+<!--***DAN: Shyamolina, demogra[hic stochasticity is a specific thing. I suspect you may
+just want to say instead something like "All these studies indicated that environmental stochasticity (through ECEs) can make a population vulnerable to extirpation."] -->
 
@@
 correlation would be stronger during extreme years) to model extinction risk remains unexplored.
+<!--***DAN: Shyamolina, you cannot intro tail dependence (please change to calling it
+tail association) in the last sentence of a paragraph. The reader does not know what that is, 
+and a whole paragraph is needed to intro it.-->
+
+<!--***DAN: Shyamolina, you need to make two major changes in how tail associations
+are introdcued. 
+  First, do it prior to saying anything about copulas. We have been wanting
+to disassociate this paper, to some extent, from BIVAN. Well, probably the best way
+to do that is by introducing tail dependence first, before copulas. You can 
+describe it intuitively as association in the tails, and you can use pictures to describe it.
+You can introduce the partial Spearman correlation (briefly in the Intro, more detail in Methods). 
+You will need an explnation of why we 
+use Spearman instead of Pearson (talk to me if you do not know, but you may want to read the 
+new Background section of BIVAN again first). One does not need to know anything about copulas
+to understand tail associations.
+  Second, the introduction to copulas that is given is outmoded. See the BAckground section
+to BIVAN for improvements. But you do not want to say anything like that much about copulas 
+here, you just want to briefly link to the concept, cite some of the Intro work that
+has been done (includeing BIVAN and some of the paper BIVAN cites for Intro work),
+mention that tail associations are one aspect of copula structure that we have found important,
+and introduce any specific things you will need, like the Clayton copula.
+
+Right now this is quite a short paper. You have space to explain things and show pedagogical
+figures.
+-->
 
@@
 <!--brief intro for copula: I am not sure whether it is needed in this paper or not?-->
+<!--***DAN: Shyamolina, to answer your above Q, well, you use Clayton and survival 
+Clayton in your simulations, so you probably have to explain copulas enough to explain those.-->
 Copula, a rank based measure of association, is much more informative 
@@
 structure for the joint distribution of two variables. For example, Clayton and Survival Clayton are 
-two copula families with asymmetric tail that are used to model exclusively 
+two copula families 
+<!--***DAN: Shyamolina, we cannot just mention these, no one will know what they are.
+As I said I think we should intro tail associations without copulas, and then subsequently link 
+to copulas afterwards.-->
+with asymmetric tail that are used to model exclusively 
 left and right tail dependence between variables, respectively.

### 2019-10-07 makefile.Rmd
checkpoint issue solved
some comments starting the process of gettin this paper ready for Ecosphere

@@
 ---
-title: "makefile, compiliation notes"
+title: "extinction risk makefile, compiliation notes"
 author: "Daniel Reuman"
@@
 ```{r tex_for_supp_mat}
-library(rmarkdown)
 rmarkdown::render(input="SuppMat.Rmd") 

### 2019-10-07 MainText.Rmd
edits

@@
 For example, a study [@chiu2013impact] showed extreme flood affected 
-the annal survival of a prey community that in turn affected predators in 
+the annual survival of a prey community that in turn affected predators in 
 aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
@@
 with positive Kendall's correlation ($\tau = 0.6$).
-Similarly, we generated extreme left- (or right-) tail dependent matrix M 
+
+On a similar note, we generated extreme left- (or right-) tail dependent matrix (M)
 from standard iid (Appendix \ref{SM-extr_tail_dep}) and made the marginals uniform so that
-it could mimic an extreme tail dependent copula.
-
-We made the first column of matrix 
+it could mimic an extreme tail dependent copula. We made the first column of matrix 
 $M$ as gamma ($\Gamma(k,\theta)$) distributed 
@@
 with spatially correlated noises having either moderate or extreme tail 
-dependence in left and right-tail.
-
-<!--say how you set ext. threshold = Pe/10-->
-We started with $n$ patches of initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ where 
+dependence in left and right-tail. <!--say how you set ext. threshold = Pe/10-->
+We started with initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ for each patch, where 
 $P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
@@
 growth parameter $r$ remained below a critical 
-level, say, $r_c$. Models showed 
-overcompensatory dynamics with oscillatory approach 
+level (say, $r_c$) and overcompensatory dynamics with oscillatory approach 
 towards stable equilibrium $P_e$ when $r$ was beyond $r_c$. 

### 2019-10-10 SuppMat.Rmd
code edited to get common kendall's tau

@@
 
-```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+```{r get_ktau_common, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+library(copula)
+source("./ExtremeTailDep.R")
+
+np<-100000
+
+#for(i in 1:5){ # for loop to ensure the consistency of Kendall's tau with specified np
+  xtc_r<-retd(n=np,d=2,rl=1,mn=0,sdev=1) # copula with extreme right tail
+
+  start_time <- Sys.time()
+  ktau_common<- cor(xtc_r[,1],xtc_r[,2],method="kendall") #get the kendall's tau
+  end_time <- Sys.time()
+
+#  cat("points = ",np, ", ktau_common = ", ktau_common, ", run time = ", (end_time -start_time),"\n")
+#}
+
+# see the change in 3rd decimal place
+#points =  1e+05 , ktau_common =  0.7512526 , run time =  1.837576 
+#points =  1e+05 , ktau_common =  0.7507594 , run time =  1.864947 
+#points =  1e+05 , ktau_common =  0.7497681 , run time =  1.862879 
+#points =  1e+05 , ktau_common =  0.7487744 , run time =  1.854317 
+#points =  1e+05 , ktau_common =  0.7516491 , run time =  1.840355
+
+#--------- Plot for intro + methods ------------------------------------
+
+# flip xtc_r to get xtc_l
+#xtc_l<-1-xtc_r
+
+# make marginals uniform 
+xtc_r<-pnorm(xtc_r)
+xtc_l<-1-xtc_r
+
+# for clayton : left tail dep.
+copC<-claytonCopula(3)
+parC<-iTau(copC,tau=ktau_common)
+
+# first generate Clayton copula 
+cc<-claytonCopula(par=parC,dim=2)
+mtc_l<-rCopula(np,cc)
+
+# flip Clayton copula 
+mtc_r<-1-mtc_l
+
+pdf("./Results/fig_4_copula.pdf", height=3, width=12)
+op<-par(mfrow=c(1,4),mar=c(4,4,2,2),mgp=c(2,1,0))
+
+# extreme left tail
+plot(xtc_l[,1],xtc_l[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
+legend(x="topleft",legend="A",bty="n",cex=2)
+
+# moderate left tail
+plot(mtc_l[,1],mtc_l[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
+legend(x="topleft",legend="B",bty="n",cex=2)
+
+# moderate right tail
+plot(mtc_r[,1],mtc_r[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
+legend(x="topleft",legend="C",bty="n",cex=2)
+
+# extreme right tail
+plot(xtc_r[,1],xtc_r[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
+legend(x="topleft",legend="D",bty="n",cex=2)
+
+par(op)
+dev.off()
+```
+
+```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("ext_risk_cop_age_str.R"))}
 # extinction risk with moderate tail dep. for single patch linear age structure model
@@
 omg<-Inf
-ktau<-0.6  # kendall's tau
+ktau<-ktau_common  # kendall's tau
 
@@
 
-```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
+```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("ext_risk_cop_age_str.R"))}
 # extinction risk with moderate tail dep. for single patch non-linear age structure model
@@
 omg<-100
-ktau<-0.6  # kendall's tau
+ktau<-ktau_common  # kendall's tau
 
@@
 
-```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-1
+ktau<-ktau_common
 ext_thrs<-K/10
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 
-```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-1
+ktau<-ktau_common
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 
-```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-0.8
+ktau<-ktau_common
 
@@
 
-```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,ktau_common,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-0.6
+ktau<-ktau_common
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 
-```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
   ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+               scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
   ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+               scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 

### 2019-10-10 mydispersal.R

@@ -173,12 +173,13 @@ get_avg_acf<-function(sims){
 #       6) params : a vector containing model parameters
 #       7) ext_thrs : extinction threshold below which populations go extinct
 #       8) scl : a scaling factor which is multiplied with noise generated
-#       9) model : a character specifying model name
-#       10) ploton : logical(T or F) to get optional plot
-#       11) resloc : location to save plots
-#       12) getacf : these are tags
+#       9) ktau : given kendall's tau to generate moderate tail dep. copula
+#       10) model : a character specifying model name
+#       11) ploton : logical(T or F) to get optional plot
+#       12) resloc : location to save plots
+#       13) getacf : these are tags
 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,model,ploton,resloc,getacf){
+plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,ktau,model,ploton,resloc,getacf){
   
   # ========================== when noises are extreme tail dep. ========================================
   ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1,mn=0,sdev=1)# a righttail dep matrix(numpoints by numlocs,     
@@ -209,7 +210,7 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
   }
   
   # ========================== when noises have moderate tail dep. ========================================
-  ktau<-0.6  # kendall's tau
+  #ktau<-0.6  # kendall's tau
   
   # for clayton : left tail dep.
   copC<-claytonCopula(3)
@@ -368,17 +369,18 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,mod
 #       3) numlocs : an integer : number of locations
 #       4) p0 : initial population to start with
 #       5) params : a vector with model parameters
-#       6) scl : a scaling factor which is multiplied with noise generated
-#       7) ext_thrs extinction threshold below which populations go extinct
-#       8) model : a character specifying model name
-#       9) disp_everywhere :logical:
+#       6) ext_thrs extinction threshold below which populations go extinct
+#       7) scl : a scaling factor which is multiplied with noise generated
[…]

### 2019-10-11 MainText.Rmd
started editing intro...continued

@@
 ---
-title: "Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
+title: "Asymmetric tail-association in extreme climatic events affects population-extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
-options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
+options(scipen = 1, digits = 3) #This option round all numbers appeared in the inline r code upto 5th digit
 seed<-101
@@
 
+# Alternative titles
+1) Tail associations in environmental variables and extinction risk
+2) The influence of correlations in the extremes on extinction risk
+
 \newpage
@@
 life-history traits of the species. Therefore predicting the impact of extreme events 
-on population extinction risk is necessery. We here introduce the potential use of "tail-dependence" 
+on population extinction risk is necessery. We here introduce the potential use of "tail-association" 
 property of two correlated ecological variables in modeling the extinction risk of 
@@
 to density-dependent metapopulation models. For metapopulation models, 
-we show extinction risk not only depends on the nature of tail-dependence of 
+we show extinction risk not only depends on the nature of tail-association of 
 spatially correlated noise but is also influenced by the inherent dynamics 
@@
 cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
-But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail-dependence instead of simple correlation in population viability analysis.
+But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail-association instead of simple correlation in population viability analysis.
 
-\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, density-dependent model
+\textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model
 
@@
 
+```{r read_res, echo=F}
+ktau_common<-readRDS("./Results/ktau_common.RDS")
+```
+
 # Introduction\label{Introduction}
@@
 @gutschick2003extreme; @jiguet2011community; @mckechnie2009climate;
-@hoover2014resistance]. Therefore, 
+@felton2017integrating]. Therefore, 
 understanding the long-term impacts of changes 
@@
 Earlier studies showed ECEs may affect species abundances (species re-ordering) 
-<!--***DAN: Shyamolina, I do not know what the parenthesis refers to -->
+<!--***DAN: Shyamolina, I do not know what the parenthesis refers to
+
+***Shyamolina: species reordering is used to access the change in community 
+in response to extreme events or global environmental changes, it relates to change 
+in relative abundances of dominant species (mainly). For example, as different species have different 
+sensitivity levels to extreme condition (means the extreme threshold is not 
+constant for all the species in a community), some years with extreme drought favors say c3 grass 
+but in the next year c4 grass could be dominant. Such alteration in community composition has "impact on 
+ecosystem function, leading to an extreme ecological response". For detailed overview you may see 
+@hoover2014resistance, Jones et al. Ecol. lett. 2017 (doi: 10.1111/ele.12864), 
+Smith et al., Ecology 2009 (doi:10.1890/08-1815.1).
+Actually LAN K. KNAPP and MELINDA D. SMITH did a lot of work on this to access changes in 
+plant community response to extreme events 
+(this review article: Felton et al. 2017, Phil. Trans. R. Soc. B 372: 20160142 is a nice summary).
+-->
 <!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
@@
 higher hierarchical levels and lead to local extinction of 
-species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme]. 
+species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
 <!--Should I mention idea of match-mismatch hypothesis?-->
@@
 over a metapopulation network of 4000 dry meadows and could affect the long-term population viability.
-Though earlier literature considered the effect of spatial and/ or temporal correlation in estimating extinction eisk of a metapopulation, but to our knowledge, using the concept of "tail dependence" (i.e.
-correlation would be stronger during extreme years) to model extinction risk remains unexplored.
-<!--***DAN: Shyamolina, you cannot intro tail dependence (please change to calling it
-tail association) in the last sentence of a paragraph. The reader does not know what that is, 
-and a whole paragraph is needed to intro it.-->
+Therefore, understanding the synchronized fluctuation over a metapopulation-network
+remains important to the conservationists because well connected patches, if synchronous, are more 
+prone to simultaneous extinction during extreme bad years [@earn2000coherence].
+
+<!--concept for correlation in the extreme tail: tail-association-->
+Though earlier literature considered the influence of spatial and/ or temporal 
+correlations in environmental variables on extinction risk of a metapopulation, to 
+our knowledge the separate influence on extinction risk of 'tail associations' between
+environmental time series in different locations has hardly been considered.
+
+” But this is not the topic sentence.
+The topic sentence comes second and is something like “Tail associations are…” or “We say two positively associated variables are more right- than left-tail associated if…”.
+Then the para continues, its point being to explain what asymmetric tail associations are.
+You can use a pedagogical figure.
+One thing you probably want to include is the logic (see BIVAN) that if extreme events are getting stronger and more extreme/intense events are also more spatially widespread, then it seems reasonable to expect that spatial patterns of tail associations in environmental variables will also be changing.
+You may also want to mention (one sentence only) that we use the statistical concept of the copula to control tail dependence in our simulations – more will be said about copulas in the Methods.
 
@@
 to understand tail associations.
+
   Second, the introduction to copulas that is given is outmoded. See the BAckground section
@@
 influence on extinction risk : introduce Taildep. concept-->
-We recently [@Ghosh_copula] used this tail dependence feature using "copula" - a statistical 
+We recently [@Ghosh_copula] used this tail association feature using "copula" - a statistical 
 tool broadly used in modeling dependence between variables - to estimate extinction risk of a density-independent 
@@
 <!--tail dep. of copula & extinction risk-->
-A major benefit of the copula approach is that it can effectively capture the tail dependence. 
-Tail dependence, already exemplified in Fig..... (add Fig.), is association between the more extreme 
+A major benefit of the copula approach is that it can effectively capture the tail association. 
+Tail association, already exemplified in Fig..... (add Fig.), is association between the more extreme 
 values of distributions. For two positively associated variables, left-tail 
-(respectively, right-tail) dependence is stronger association between
+(respectively, right-tail) association is stronger association between
 values in the left or lower portions, i. e. for smaller values 
 (respectively, the right or upper portions, i.e. for higher values) 
-of the two distributions. Stronger dependence in either tail causes asymmetric copula 
+of the two distributions. Stronger association in either tail causes asymmetric copula 
 structure for the joint distribution of two variables. For example, Clayton and Survival Clayton are 
@@
 with asymmetric tail that are used to model exclusively 
-left and right tail dependence between variables, respectively.
+left and right tail association between variables, respectively.
 From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
 might reduce the adult reproduction rate as well as larval survival (i.e. stronger 
-tail dependence between vital rates when temperature is too cold than normal) and 
+tail association between vital rates when temperature is too cold than normal) and 
 thereby inflate the population viability. Similarly, in context of 
@@
 during unfavorable years unless they are rescued by the nearby populations. This scenario 
-could be more complex and specific of the direction of dependence 
-depending on the dynamics of the model (see \nameref{Results}). Though tail-dependence 
+could be more complex and specific of the direction of association 
+depending on the dynamics of the model (see \nameref{Results}). Though tail-association 
 property of a copula has been used for risk-analysis in other fields like finance, 
 hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
-knowledge asymmetric tail dependence have never been considered 
+knowledge asymmetric tail association have never been considered 
 in estimating the effect of ECEs through extinction risk of age structured 
@@
 <!--our objective ?-->
-In this paper, we will explore the potential use of asymmetric tail-dependence 
+In this paper, we will explore the potential use of asymmetric tail-association 
 for understanding the extinction risk of populations towards extreme climatic event
-by addressing the following questions: (Q1) Does nature of tail dependence 
+by addressing the following questions: (Q1) Does nature of tail association 
 (left tail or right tail) in vital rates
@@
 We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
-(Q2) How does environmental noises with asymmetric tail dependence across multiple patches
+(Q2) How does environmental noises with asymmetric tail association across multiple patches
 influence extinction risk of a density-dependent metapopulation model? Our speculation 
@@
 (higher extinction risk if noises across patches were left-tail dependent than the one 
-if noises had right-tail dependence) would hold true only in the under-compensatory regime, but 
+if noises had right-tail association) would hold true only in the under-compensatory regime, but 
 should be reverse if the model shows over-compensatory dynamics. 
@@
 under-compensatory to over-compensatory dynamics.
+
+<!--FIG : pedagogical fig showing 4 types of tail associations: 
+extreme left, moderate left, moderate right, extreme right-->
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
+\caption{Two random variables can have same Kendall 
+correlation ($\tau=$ `r ktau_common`), but different degree of associations
+in their extreme values: (A) extreme left tail, (B) moderate left tail, (C) moderate right tail, and
+(D) extreme right tail association.\label{fig_intro_cop}}
+\end{center}
+\end{figure*}
 
@@
 a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. We generated those distributions 
-from iid noises with a varied degree of asymmetric tail-dependence between 
-fecundity ($f_a$) and survival ($s_e$) rates. Degree of tail dependence had two broad 
-categories: moderate and extreme. Each category had specific direction in tail dependence: 
+from iid noises with a varied degree of asymmetric tail-association between 
+fecundity ($f_a$) and survival ($s_e$) rates. Degree of tail association had two broad 
+categories: moderate and extreme. Each category had specific direction in tail association: 
 either in left or in right tail of vital rates. Vital rates ($f_a$ and $s_e$) were
@@
 We used bivariate Clayton copula to model moderate left-tail and Survival Clayton copula to model 
-moderate right tail dependence in the vital rates. Both copula had the same Kendall 
-correlation ($=0.6$) but opposite dependence in their respective tails. 
+moderate right tail association in the vital rates. Both copula had the same Kendall 
+correlation ($=0.6$) but opposite association in their respective tails. 
 We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
@@
 matrix $M$, then became independent of each others but 
-entries between columns had a specific tail dependence
+entries between columns had a specific tail association
 with positive Kendall's correlation ($\tau = 0.6$).
@@
 <!--density-dependence metapopulation model-->
-We explored (Q2) by examining how nature of the tail dependence of an
+We explored (Q2) by examining how nature of the tail association of an
 environmental variable across space could influence the extinction risk of a 
@@
 from Clayton copula so that all the statistical properties remained 
-same except for the direction of tail-dependence. We generated
+same except for the direction of tail-association. We generated
 extreme left- or right-tail dependent noises via algorithm stated in 
@@
 with spatially correlated noises having either moderate or extreme tail 
-dependence in left and right-tail. <!--say how you set ext. threshold = Pe/10-->
+association in left and right-tail. <!--say how you set ext. threshold = Pe/10-->
 We started with initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ for each patch, where 
@@
 For single patch linear age-structured models, we found higher extinction risk for moderate left-tail 
-dependence (modeled by Clayton copula) in fluctuating vital 
+association (modeled by Clayton copula) in fluctuating vital 
 rates compared with the one when vital rates were right tail dependent (modeled by Survival Clayton copula).
@@
 and thereby can increase the extinction risk of the population. For non-linear age-structured 
-models, we got the similar trend for extinction risks with different tail-dependence (left and right).
+models, we got the similar trend for extinction risks with different tail-association (left and right).
 Non-linear model showed higher extinction risk than its corresponding 
@@
 blue lines in Fig. \ref{SM-fig_extrisk_agestr_mtd}A or B) and we anticipated the difference being partly 
-dependent on the degree of tail dependence. Therefore we repeated the same analysis with 
+dependent on the degree of tail association. Therefore we repeated the same analysis with 
 vital rates coming from extreme-tail dependent (left and right in two different simulations)
@@
 (Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
-a population could be considered as direct manifestation of tail-dependence property 
+a population could be considered as direct manifestation of tail-association property 
 between vital rates of different age-class
@@
 model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
-the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail dependence.
+the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail association.
 But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - was 
@@
 <!--summary results:-->
-We showed the potential use of "asymmetric" tail dependence in modeling the impact of 
+We showed the potential use of "asymmetric" tail association in modeling the impact of 
 extreme ecological events on population-extinction risk. Our results demonstrated left and right-tail 
-dependence in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
-in years of extreme events ("catastrophs" and "bonanzas"). Though we could expect left tail dependence 
-to show higher extinction risk than the one for right tail dependence with age-structured 
+association in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
+in years of extreme events ("catastrophs" and "bonanzas"). Though we could expect left tail association 
+to show higher extinction risk than the one for right tail association with age-structured 
 and density dependent metapopulation models, it was not always true. In the over-compensatory regime,
@@
 <!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
-We found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
+We found left tail association in vital rates (induced by ECEs) caused higher extinction risk than 
 the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_mtd}) 
-unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr_etd}). This finding could be explained 
+unless the tail association was at its extreme (Fig. \ref{fig_extrisk_agestr_etd}). This finding could be explained 
 by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
@@
 and the temporal variance of the log(population growth rate) caused by environmental stochasticity - which was based
-on the linear covariance measure between any two matrix elements, not on their tail-dependence properties. Therefore,
+on the linear covariance measure between any two matrix elements, not on their tail-association properties. Therefore,
 following @tuljapurkar1990lecture, population growth rate in an iid environment could be approximated
-as same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependence.
+as same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail association.
 
@@
 -->
-Although we demonstrated that tail dependence between vital rates of an age structured model can
+Although we demonstrated that tail association between vital rates of an age structured model can
 influence extinction risk, substantial work remains to determine the importance of this effect in the 
-following scenarios. First, we demonstrated the effect of tail-dependence on extinction risk 
+following scenarios. First, we demonstrated the effect of tail-association on extinction risk 
 with a basic age structured (two age stages) model where we considered the total semelperous 
@@
 of populations including semelperous and iteroperous [@morris2002quantitative]. Future studies should include 
-tail-dependence effect on extinction risk 
+tail-association effect on extinction risk 
 for a multiple age-, size- and stage- structured model and estimate extinction risk 
@@
 would be to explore the viability for the metapopulations if vital rates (e.g. fecundity or 
-survival of different stage class) across sites showed specific tail dependence 
-with spatially correlated extreme climatic events. Third, we considered tail dependence 
+survival of different stage class) across sites showed specific tail association 
+with spatially correlated extreme climatic events. Third, we considered tail association 
 between positively correlated vital rates, but what would be the extinction risk 
-and viability of a population if demographic rates were negatively correlated with asymmetric tail dependence. 
+and viability of a population if demographic rates were negatively correlated with asymmetric tail association. 
 We could consider this scenario to model extreme events that kill adults and allow more successful
 recruitment of seeds or other propagules. Fourth, following the aforementioned idea, it would be worthy 
-to model more complex prey-predator model where asymmetric tail dependent copula could be used to mimic
+to model more complex prey-predator model where asymmetric tail association could feature the
 interspecific-dependence. An earlier study [@carnicer2011widespread] showed extreme events
@@
 rates between nearby patches than the patches which are far away. It is worthy to study 
-metapopulation viability with non-uniform spatial correlation (i.e. asymmetric tail dependence) in a 
+metapopulation viability with non-uniform spatial correlation (i.e. asymmetric tail association) in a 
 network with different clusters having various dispersal rates.
@@
 \caption{Extinction risk against time from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models for $10000$ simulations; shading represents 95\% confidence interval.
-Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence.
+Fecundity of adults and survival of eggs have either extreme left or extreme right tail association.
 \label{fig_extrisk_agestr_etd}}

### 2019-10-11 SuppMat.Rmd
started editing intro...continued

@@
 ---
-title: "Supporting Information : Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
+title: "Supporting Information : Asymmetric tail-association in extreme climatic events affects population-extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 
-#--------- Plot for intro + methods ------------------------------------
-
 # flip xtc_r to get xtc_l
@@
 mtc_r<-1-mtc_l
-
-pdf("./Results/fig_4_copula.pdf", height=3, width=12)
-op<-par(mfrow=c(1,4),mar=c(4,4,2,2),mgp=c(2,1,0))
+```
+
+```{r plot_intro_fig, echo=F, results="hide"}
+
+#--------- Plot for intro + methods ------------------------------------
+
+# for plotting purpose I just showed only first 500 points
+pdf("./Results/introfig_copula.pdf", height=3, width=12)
+op<-par(mfrow=c(1,4),mar=c(6,6,2,2),mgp=c(3,1,0))
 
 # extreme left tail
-plot(xtc_l[,1],xtc_l[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
-legend(x="topleft",legend="A",bty="n",cex=2)
+plot(xtc_l[1:500,1],xtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
+legend(x="topleft",legend="A",bty="n",cex=2.5)
 
 # moderate left tail
-plot(mtc_l[,1],mtc_l[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
-legend(x="topleft",legend="B",bty="n",cex=2)
+plot(mtc_l[1:500,1],mtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
+legend(x="topleft",legend="B",bty="n",cex=2.5)
 
 # moderate right tail
-plot(mtc_r[,1],mtc_r[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
-legend(x="topleft",legend="C",bty="n",cex=2)
+plot(mtc_r[1:500,1],mtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
+legend(x="topleft",legend="C",bty="n",cex=2.5)
 
 # extreme right tail
-plot(xtc_r[,1],xtc_r[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
-legend(x="topleft",legend="D",bty="n",cex=2)
-
-par(op)
-dev.off()
+plot(xtc_r[1:500,1],xtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
+legend(x="topleft",legend="D",bty="n",cex=2.5)
+
+par(op)
+dev.off()
+
+saveRDS(ktau_common,"./Results/ktau_common.RDS")
 ```
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
-\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C) or right tail (SC) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time for 10000 simulations.
-Fecundity of adults and survival of eggs have moderate tail dependence in either left or right tail in their corresponding copula structure.
-Moderate left tail dependence (or right tail dependence) was generated by a bivariate Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6, $(A)$ for linear model ($\Omega = \infty$) and for $(B)$ non-linear model ($\Omega = 100$).\label{fig_extrisk_agestr_mtd}}
+\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C) or right tail (SC) association.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time for 10000 simulations.
+Fecundity of adults and survival of eggs have moderate tail association in either left or right tail in their corresponding copula structure.
+Moderate left tail association (or right tail association) was generated by a bivariate Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6, $(A)$ for linear model ($\Omega = \infty$) and for $(B)$ non-linear model ($\Omega = 100$).\label{fig_extrisk_agestr_mtd}}
 \end{center}
@@
 \caption[Analogous figure as that of Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail dependence in spatially correlated noises.]{Analogous figure as that of Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail dependence in spatially correlated noises.\label{fig_extrisk_allmodel_localdisp_mtd}}
+\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail association in spatially correlated noises.]{Analogous figure as that of Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail association in spatially correlated noises.\label{fig_extrisk_allmodel_localdisp_mtd}}
 \end{center}
@@
 \caption[Analogous figure as that of Fig.
-\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail dependence in spatially correlated noises.]{Analogous figure as that of Fig.
-\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail dependence in spatially correlated noises.\label{fig_extrisk_allmodel_globaldisp_mtd}}
+\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail association in spatially correlated noises.]{Analogous figure as that of Fig.
+\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail association in spatially correlated noises.\label{fig_extrisk_allmodel_globaldisp_mtd}}
 \end{center}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_ricker_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_ricker_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_ricker_risk_local_disp_scl_1_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_hassell_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_hassell_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_hassell_risk_local_disp_scl_1_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_msmith_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_msmith_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_pennycuick_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_pennycuick_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_verhulst_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_verhulst_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_abrewer_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_abrewer_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_ricker_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_ricker_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_ricker_risk_global_disp_scl_1_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_hassell_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_hassell_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_hassell_risk_global_disp_scl_1_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_msmith_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_msmith_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_msmith_risk_global_disp_scl_0.8_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_pennycuick_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_pennycuick_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_verhulst_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_verhulst_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_mtd}}
@@
 \caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_abrewer_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
+\ref{fig_abrewer_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
 \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}}
@@
 tail dependent noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated 
+tail dependent noises; Moderate left tail association (or right tail association) was generated 
 by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between
@@
 
-# Generating matrix with extreme tail dependence  \label{extr_tail_dep}
-
-For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
+# Generating matrix with extreme tail association  \label{extr_tail_dep}
+
+For extreme tail-association algorithm, a $N \times n$ matrix, M was generated of independent 
 standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->
@@
 Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
-generating extreme tail dependence matrix is flexible enough, so that 
+generating extreme tail association matrix is flexible enough, so that 
 matrix M can be generated initially with independent random draws from a normal 

### 2019-10-11 MainText.Rmd
extreme and moderate results combined in single plot

@@
 could weaken the resistance of plant communities too [@dreesen2014successive]. 
-All these studies indicated that environmental stochasticity (through ECEs) 
+All these studies indicated that environmental stochasticity (through ECEs) can 
+make a population vulnerable to extirpation.
+<!--All these studies indicated that environmental stochasticity (through ECEs) 
 can lead to demographic stochasticity (through influence on reproductive or 
 mortality rate) of different age class of a population making it vulnerable 
-to local extinction. 
+to local extinction.--> 
 <!--***DAN: Shyamolina, demogra[hic stochasticity is a specific thing. I suspect you may
@@
 correlations in environmental variables on extinction risk of a metapopulation, to 
-our knowledge the separate influence on extinction risk of 'tail associations' between
-environmental time series in different locations has hardly been considered.
+our knowledge the separate influence on extinction risk of "tail associations" between
+environmental time series in different locations has hardly been considered. "Tail associations" are
+the association between two random variables in their extreme values. If two positively correlated 
+variables are more associated in their smaller values than larger, they are said to have stronger
+"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). Similarly, if two positively correlated 
+variables are more associated in their larger values than smaller, they are said to have stronger
+"right-tail" or "upper-tail" association (Figure \ref{fig_intro_cop}C, D). Figure \ref{fig_intro_cop} shows 
+two random variables can have a wide range of assciation from extreme to moderate degree in their extreme tails, but
+with a fixed Kendall's $\tau$.
 

### 2019-10-11 SuppMat.Rmd
extreme and moderate results combined in single plot

@@
 
+pdf("./Results/common_legend_extreme_and_moderate_noise.pdf",height=1,width=12)
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
+plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+legend("center", c("extreme LT","moderate LT", "moderate RT","extreme RT"), lty=c(1,1), pch=c(1,1,1,1), 
+         col=c('red', 'orange','skyblue','blue'),x.intersp = 0.1,
+         cex = 2.2, xpd = TRUE, horiz = T, inset = c(0,0), 
+         bty = "n")
+par(op)
+dev.off()
+
 #-------------------------------------- plot for extreme tail dep. noise ---------------------------------------
@@
   lines(ss$d,ss$risk_mright,type="b",col="blue")
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+#----------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"ricker_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=9,width=9)
+#op<-par(mfrow=c(3,3),mar=c(5.2,4.2,1.2,6.2))
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_ricker_local)[3]){
+  #s<-res_ricker_local[,,i]
+  ss<-res_ricker_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
+  #       col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"ricker_ext_risk_global_disp_extreme_and_moderate_noise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_ricker_global)[3]){
+  #ss<-res_ricker_global
+  ss<-res_ricker_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
   #legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), 
@@
 #legend("bottomright", legend=c("Ricker",paste0("K = ",K)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+
+#-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"ricker_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+
+plot(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_xleft_d0,col="red",type="b",
+     ylim=c(0,1),xlim = range(ricker_extrisk_d0$r),xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025xleft,x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975xleft,
+       angle=90,length=0.1,col="red",code=3)
+abline(v=rc_ricker,lty="dotted",col="black") #rc=1
+
+lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_mleft_d0,col="orange",type="b",ylim=c(0,1),pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025mleft,
+       x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="orange",code=3)
+
+lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_mright_d0,col="skyblue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025mright,
+       x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="skyblue",code=3)
+
+lines(ricker_extrisk_d0$r,ricker_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=ricker_extrisk_d0$r,y0=ricker_extrisk_d0$CI0.025xright,
+       x1=ricker_extrisk_d0$r,y1=ricker_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+
+legend("topleft", legend=c("Ricker"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("K = ",K),side = 3, line=-16, adj=0.95, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+
 par(op)
@@
 dev.off()
+
+#----------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_hassell_local)[3]){
+  ss<-res_hassell_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"hassell_ext_risk_global_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_hassell_global)[3]){
+  ss<-res_hassell_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       #panel.first = grid(),
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 #legend("bottomright", c("Hassell",paste0("a = ",a," ,b = ",b),paste0("sd =",scl)), horiz=F, bty='n',cex=1.8)
+par(op)
+dev.off()
+
+#-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"hassell_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+
+plot(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(hassell_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025xleft,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_hassell,lty="dotted",col="black")
+
+lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_mleft_d0,col="orange",type="b",ylim=c(0,1),pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025mleft,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="orange",code=3)
+
+lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_mright_d0,col="skyblue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025mright,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="skyblue",code=3)
+
+lines(hassell_extrisk_d0$r,hassell_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=hassell_extrisk_d0$r,y0=hassell_extrisk_d0$CI0.025xright,
+       x1=hassell_extrisk_d0$r,y1=hassell_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+
+legend("topleft", legend=c("Hassell"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+
 par(op)
@@
 dev.off()
+
+#---------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"msmith_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_msmith_local)[3]){
+  ss<-res_msmith_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"msmith_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5))
+
+for(i in 1:dim(res_msmith_global)[3]){
+  ss<-res_msmith_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+par(op)
+dev.off()
+
+#-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"msmith_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+
+plot(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(msmith_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025xleft,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_msmith,lty="dotted",col="black")
+
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_mleft_d0,col="orange",type="b",ylim=c(0,1),pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025mleft,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="orange",code=3)
+
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_mright_d0,col="skyblue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025mright,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="skyblue",code=3)
+
+lines(msmith_extrisk_d0$r,msmith_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=msmith_extrisk_d0$r,y0=msmith_extrisk_d0$CI0.025xright,
+       x1=msmith_extrisk_d0$r,y1=msmith_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+
+legend("topleft", legend=c("Maynard Smith"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+
 par(op)
@@
 dev.off()
+
+#--------------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.2,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_pennycuick_local)[3]){
+  ss<-res_pennycuick_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_pennycuick_global)[3]){
+  ss<-res_pennycuick_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 dev.off()
+
+#-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"pennycuick_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+
+plot(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(pennycuick_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025xleft,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_pennycuick,lty="dotted",col="black")
+
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_mleft_d0,col="orange",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025mleft,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="orange",code=3)
+
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_mright_d0,col="skyblue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025mright,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="skyblue",code=3)
+
+lines(pennycuick_extrisk_d0$r,pennycuick_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=pennycuick_extrisk_d0$r,y0=pennycuick_extrisk_d0$CI0.025xright,
+       x1=pennycuick_extrisk_d0$r,y1=pennycuick_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+
+legend("topleft", legend=c("Pennycuick"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("a = ",a,", b = ",b),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+
+par(op)
+dev.off()
 ```
@@
 dev.off()
+
+#------------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"verhulst_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_verhulst_local)[3]){
+  ss<-res_verhulst_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"verhulst_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_verhulst_global)[3]){
+  ss<-res_verhulst_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 dev.off()
+
+#-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"verhulst_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+
+plot(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(verhulst_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025xleft,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_verhulst,lty="dotted",col="black")
+
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_mleft_d0,col="orange",type="b",ylim=c(0,1),pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025mleft,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="orange",code=3)
+
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_mright_d0,col="skyblue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025mright,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="skyblue",code=3)
+
+lines(verhulst_extrisk_d0$r,verhulst_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=verhulst_extrisk_d0$r,y0=verhulst_extrisk_d0$CI0.025xright,
+       x1=verhulst_extrisk_d0$r,y1=verhulst_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+
+legend("topleft", legend=c("Verhulst"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("K = ",K),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+
+par(op)
+dev.off()
 ```
@@
 dev.off()
+
+#--------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"abrewer_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_abrewer_local)[3]){
+  ss<-res_abrewer_local[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
+
+pdf(paste(resloc,"abrewer_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
+
+for(i in 1:dim(res_abrewer_global)[3]){
+  ss<-res_abrewer_global[,,i][1:6,] # plot for d in range (0 to 0.5)
+  ss<-as.data.frame(ss)
+
+  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+       xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
+  lines(ss$d,ss$risk_mleft,type="b",col="orange")
+  lines(ss$d,ss$risk_mright,type="b",col="skyblue")
+  lines(ss$d,ss$risk_xright,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[i],5)),cex=1.2)
+}
+par(op)
+dev.off()
 ```
@@
 dev.off()
+
+#-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+pdf(paste(resloc,"abrewer_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
+
+plot(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_xleft_d0,col="red",type="b",ylim=c(0,1),xlim = range(abrewer_extrisk_d0$r),
+     xlab="r",ylab="Extinction risk",cex.lab=2,cex.axis=1.5,pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025xleft,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975xleft,angle=90,length=0.1,col="red",code=3)
+abline(v=rc_abrewer,lty="dotted",col="black")
+
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_mleft_d0,col="orange",type="b",ylim=c(0,1),pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025mleft,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975mleft,angle=90,length=0.1,col="orange",code=3)
+
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_mright_d0,col="skyblue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025mright,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975mright,angle=90,length=0.1,col="skyblue",code=3)
+
+lines(abrewer_extrisk_d0$r,abrewer_extrisk_d0$risk_xright_d0,col="blue",type="b",ylim=c(0,1),pch=1)
+arrows(x0=abrewer_extrisk_d0$r,y0=abrewer_extrisk_d0$CI0.025xright,
+       x1=abrewer_extrisk_d0$r,y1=abrewer_extrisk_d0$CI0.975xright,angle=90,length=0.1,col="blue",code=3)
+
+legend("topleft", legend=c("Austin-Brewer"), horiz=F, bty='n',cex=1.8)  
+mtext(paste0("s = ",s,", K = ",K),side = 3, line=-16, adj=0.9, cex=1.8, col="black")
+legend("bottomright",legend=bquote(sigma==.(scl)), horiz=F,bty='n',cex=1.8)
+
+par(op)
+dev.off()
 ```
@@
 
-if(noisetype=="extreme"){
+if(noisetype=="both"){
+   plot(Af$d,Af$risk_xleft,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
+     xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+   lines(Af$d,Af$risk_mleft,type="b",col="orange")
+   lines(Af$d,Af$risk_mright,type="b",col="skyblue")
+   lines(Af$d,Af$risk_xright,type="b",col="blue")
+
+  mtext(paste0("r = ", round(rlist[1],5)))
+
+  plot(Al$d,Al$risk_xleft,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
+     xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+   lines(Al$d,Al$risk_mleft,type="b",col="orange")
+   lines(Al$d,Al$risk_mright,type="b",col="skyblue")
+   lines(Al$d,Al$risk_xright,type="b",col="blue")
+
+  mtext(paste0("r = ", round(rlist[lasta],5)))
+
+ }else if(noisetype=="extreme"){
   risk_left1<-Af$risk_xleft
@@
   risk_right2<-Al$risk_xright
-}
-
-if(noisetype=="moderate"){
-  risk_left1<-Af$risk_mleft
-  risk_right1<-Af$risk_mright
-
-  risk_left2<-Al$risk_mleft
-  risk_right2<-Al$risk_mright
-}
 
@@
 
-}
-
+ }else if(noisetype=="moderate"){
+  risk_left1<-Af$risk_mleft
+  risk_right1<-Af$risk_mright
+
+  risk_left2<-Al$risk_mleft
+  risk_right2<-Al$risk_mright
+
+  plot(Af$d,risk_left1,xlim=range(Af$d),ylim=c(0,1),type="b",col="red",
+     xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(Af$d,risk_right1,type="b",col="blue")
+
+  mtext(paste0("r = ", round(rlist[1],5)))
+
+  plot(Al$d,risk_left2,xlim=range(Al$d),ylim=c(0,1),type="b",col="red",
+     xlab="d",ylab="Extinction risk",cex.lab=1.5,cex.axis=1.5)
+  lines(Al$d,risk_right2,type="b",col="blue")
+  mtext(paste0("r = ", round(rlist[lasta],5)))
+
+ }else{
+  cat("specify noisetype argument!")
+  }
+}
 #------------- for extreme tail dep. noise : local dispersal ---------------------
@@
 dev.off()
+
+#------------- for extreme and moderate tail dep. noise : local dispersal summary fig. ---------------------
+pdf("./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf",height=12,width=9)
+
+op<-par(mfrow=c(6,3),mar=c(5.2,4.2,1.2,1.5))
+
+x<-res_ricker_local
+rlist<-ricker_extrisk_d0$r
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+legend("topleft", c("extreme LT","moderate LT","moderate RT","extreme RT"), lty=c(1,1,1,1), pch=c(1,1,1.1), 
+       col=c('red', 'orange','skyblue','blue'), horiz=F, bty='o', cex=1.5, adj=0, x.intersp=0.2)
+mtext(paste("Ricker model: "), side = 3, line=-9, adj=0.3, cex=1, col="black")
+#mtext(paste("(K = 50, sd = 1)"), side = 3, line=-8.2, adj=0.6, cex=1, col="black")
+mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-11, adj=0.6, cex=1, col="black")
+getplot(x=x,rlist = rlist,noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Hassell model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_hassell_local
+rlist<-hassell_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Maynard Smith model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_msmith_local
+rlist<-msmith_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Pennycuick model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_pennycuick_local
+rlist<-pennycuick_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Verhulst model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_verhulst_local
+rlist<-verhulst_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Austin-Brewer model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_abrewer_local
+rlist<-abrewer_extrisk_d0$r
+getplot(x=x,rlist = rlist,noisetype = "both")
+
+par(op)
+dev.off()
 ```
@@
 getplot(x=x,rlist = rlist, noisetype = "moderate")
+
+par(op)
+dev.off()
+
+#--------------------for extreme and moderate tail dep. noise : global dispersal summary fig-----------------------
+
+pdf("./Results/allmodel_ext_risk_global_disp_extreme_and_moderatenoise.pdf",height=12,width=9)
+
+op<-par(mfrow=c(6,3),mar=c(5.2,4.2,1.2,1.5))
+
+x<-res_ricker_global
+rlist<-ricker_extrisk_d0$r
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+legend("topleft", c("extreme LT","moderate LT","moderate RT","extreme RT"), lty=c(1,1,1,1), pch=c(1,1,1.1), 
+       col=c('red', 'orange','skyblue','blue'), horiz=F, bty='o', cex=1.5, adj=0, x.intersp=0.2)
+mtext(paste("Ricker model: "), side = 3, line=-9, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( K = 50, ", sigma,"=1 )" )),side = 3, line=-11, adj=0.6, cex=1, col="black")
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Hassell model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 100, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_hassell_global
+rlist<-hassell_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Maynard Smith model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.5, b = 4, ", sigma,"=0.8 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_msmith_global
+rlist<-msmith_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Pennycuick model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( a = 0.1, b = 0.1, ", sigma,"=1 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_pennycuick_global
+rlist<-pennycuick_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Verhulst model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( K = 1, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_verhulst_global
+rlist<-verhulst_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
+
+plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
+mtext(paste("Austin-Brewer model: "), side = 3, line=-7, adj=0.3, cex=1, col="black")
+mtext(expression( paste("( s = 0.2, K = 50, ", sigma,"=0.6 )" )),side = 3, line=-9, adj=0.6, cex=1, col="black")
+x<-res_abrewer_global
+rlist<-abrewer_extrisk_d0$r
+getplot(x=x,rlist = rlist, noisetype = "both")
 

### 2019-10-12 MainText.Rmd
rearranging figs...continued

@@
 was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
-value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp_etd} and \ref{SM-fig_extrisk_allmodel_localdisp_mtd} show
+value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp} shows
 variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
@@
 from over-compensatory regime. We also got similar results with our prediction (see Q2 from \nameref{Introduction})
-for global dispersal (Figs. \ref{SM-fig_extrisk_allmodel_globaldisp_etd} and \ref{SM-fig_extrisk_allmodel_globaldisp_mtd}). 
+for global dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
 Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}) 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
-\caption{Extinction risk against time from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models for $10000$ simulations; shading represents 95\% confidence interval.
-Fecundity of adults and survival of eggs have either extreme left or extreme right tail association.
-\label{fig_extrisk_agestr_etd}}
+\end{center}
+\textbf{ \hspace{5.4 cm} (C) \hspace{4 cm} (D)} \\
+\vspace{-0.9 cm}
+\begin{center}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
+\caption{Extinction risk against time from single patch age structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(C, D)$ models for $10000$ simulations; shading represents 95\% confidence interval.
+$(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have extreme left tail (LT) or extreme right tail (RT) association, $(C, D)$ show the analogous scenario but with moderate tail association between vital rates.
+\label{fig_extrisk_agestr}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_extremenoise.pdf}
+\includegraphics[width=14.5 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
 \caption{Extinction risks are computed after $25$ years for six nonlinear density-dependence models (Table \ref{tab_model_summary}) with local dispersal among $n=5$ patches and for $10000$ simulations.
-Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
-Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp_etd}}
+Spatially correlated
+noises have varying degree of tail associations (extreme left tail, moderate left tail, moderate right tail and extreme right tail) as introduced in Fig.
+\ref{fig_intro_cop}.
+Last two columns represent variation of extinction risk against dispersal rate ($d$), calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
 \end{center}
@@
 
- <!--FIG : all nonlinear model results with extreme noise, d=0, extrisk vs. r-->
+<!--FIG : all nonlinear model results, d=0, extrisk vs. r-->
 \begin{figure*}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extremenoise.pdf}
-\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extremenoise.pdf}\\
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=6 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extreme_and_moderate_noise.pdf}
+\includegraphics[width=6 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extreme_and_moderate_noise.pdf}\\
 \vspace{-0.6 cm}
-\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extremenoise.pdf}
-\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extremenoise.pdf}\\
+\includegraphics[width=6 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extreme_and_moderate_noise.pdf}
+\includegraphics[width=6 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extreme_and_moderate_noise.pdf}\\
 \vspace{-0.6 cm}
-\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extremenoise.pdf}
-\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extremenoise.pdf}\\
+\includegraphics[width=6 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extreme_and_moderate_noise.pdf}
+\includegraphics[width=6 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extreme_and_moderate_noise.pdf}\\
 \caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 

### 2019-10-12 SuppMat.Rmd
rearranging figs...continued

@@
 lines(t,x$erA,col="blue",type="l")
-legend(x=-2,y=1.05, c("vital rates: moderate LTdep.,C","vital rates: moderate RTdep.,SC"), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5, x.intersp = 0.2, seg.len = 1)
+legend(x=-2,y=1.05, c("moderate LT","moderate RT", expression(paste(Omega,"= ", infinity))), lty=c(1,1, NA),  
+       col=c('red', 'blue', NA), horiz=F, bty='n', cex=1.5, x.intersp = 0.2
+       #, seg.len = 1
+       )
 par(op)
@@
 lines(t,x$erA,col="blue",type="l")
-legend(x=-2,y=1.05, c("vital rates: moderate LTdep.,C","vital rates: moderate RTdep.,SC"), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5, x.intersp = 0.2, seg.len = 1)
+legend(x=-2,y=1.05, c("moderate LT","moderate RT", as.expression(bquote(Omega==.(omg)))), lty=c(1,1,NA),  
+       col=c('red', 'blue',NA), horiz=F, bty='n', cex=1.5, x.intersp = 0.2
+       #, seg.len = 1
+       )
 par(op)
@@
 lines(t,x$erA,col="blue",type="l")
-legend(x=-2,y=1.05, c("vital rates: extreme LTdep.","vital rates: extreme RTdep."), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5, x.intersp = 0.2)
+legend(x=-2,y=1.05, c("extreme LT","extreme RT", expression(paste(Omega,"= ", infinity))), lty=c(1,1, NA),  
+       col=c('red', 'blue', NA), horiz=F, bty='n', cex=1.5, x.intersp = 0.2
+       #, seg.len = 1
+       )
 par(op)
@@
 lines(t,x$erA,col="blue",type="l")
-legend(x=-2,y=1.05, c("vital rates: extreme LTdep.","vital rates: extreme RTdep."), lty=c(1,1),  
-       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5, x.intersp = 0.2)
+legend(x=-2,y=1.05, c("extreme LT","extreme RT", as.expression(bquote(Omega==.(omg)))), lty=c(1,1,NA),  
+       col=c('red', 'blue',NA), horiz=F, bty='n', cex=1.5, x.intersp = 0.2
+       #, seg.len = 1
+       )
 par(op)
@@
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("center", c("extreme LT","moderate LT", "moderate RT","extreme RT"), lty=c(1,1), pch=c(1,1,1,1), 
+legend("center", c("extreme LT,","moderate LT,", "moderate RT,","extreme RT"), lty=c(1,1), pch=c(1,1,1,1), 
          col=c('red', 'orange','skyblue','blue'),x.intersp = 0.1,
@@
 
-#-------------------------------------- plot for moderate tail dep. noise ---------------------------------------
+#-------------------------------------- plot for moderate tail dep. noise --------------------------------------
 pdf(paste(resloc,"ricker_ext_risk_local_disp_moderatenoise.pdf",sep=""),height=9,width=9)
@@
 #-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"ricker_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+pdf(paste(resloc,"ricker_extrisk_d0_extreme_and_moderate_noise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 
-#----------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
+#----------------------- plot for extreme and moderate tail dep. noise --------------------------------------
 pdf(paste(resloc,"hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
@@
 #-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"hassell_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+pdf(paste(resloc,"hassell_extrisk_d0_extreme_and_moderate_noise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 
-#---------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"msmith_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+#---------------------- plot for extreme and moderate tail dep. noise --------------------------------------
+pdf(paste(resloc,"msmith_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-  plot(ss$d,ss$risk_mleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
+  plot(ss$d,ss$risk_xleft,xlim=range(ss$d),ylim=c(0,1),type="b",col="red",
        xlab="d",ylab="Extinction risk",cex.lab=2,cex.axis=1.5)
@@
 
-pdf(paste(resloc,"msmith_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+pdf(paste(resloc,"msmith_ext_risk_global_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5))
@@
 #-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"msmith_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+pdf(paste(resloc,"msmith_extrisk_d0_extreme_and_moderate_noise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 
-#--------------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"pennycuick_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+#--------------------------- plot for extreme and moderate tail dep. noise --------------------------------------
+pdf(paste(resloc,"pennycuick_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.2,1.5),mgp=c(2,0.5,0))
@@
 
-pdf(paste(resloc,"pennycuick_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=12,width=12)
+pdf(paste(resloc,"pennycuick_ext_risk_global_disp_extreme_and_moderate_noise.pdf",sep=""),height=12,width=12)
 op<-par(mfrow=c(4,4),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 #-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"pennycuick_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+pdf(paste(resloc,"pennycuick_extrisk_d0_extreme_and_moderate_noise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 
-#------------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"verhulst_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+#------------------------- plot for extreme and moderate tail dep. noise --------------------------------------
+pdf(paste(resloc,"verhulst_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-pdf(paste(resloc,"verhulst_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+pdf(paste(resloc,"verhulst_ext_risk_global_disp_extreme_and_moderate_noise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 #-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"verhulst_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+pdf(paste(resloc,"verhulst_extrisk_d0_extreme_and_moderate_noise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 
-#--------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"abrewer_ext_risk_local_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+#--------------------- plot for extreme and moderate tail dep. noise --------------------------------------
+pdf(paste(resloc,"abrewer_ext_risk_local_disp_extreme_and_moderate_noise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 
-pdf(paste(resloc,"abrewer_ext_risk_global_disp_extreme_and_moderatenoise.pdf",sep=""),height=9,width=9)
+pdf(paste(resloc,"abrewer_ext_risk_global_disp_extreme_and_moderate_noise.pdf",sep=""),height=9,width=9)
 op<-par(mfrow=c(3,3),mar=c(5,4,1.5,1.5),mgp=c(2,0.5,0))
@@
 #-------------------- plot for extreme and moderate tail dep. noise ---------------------------------------
-pdf(paste(resloc,"abrewer_extrisk_d0_extreme_and_moderatenoise.pdf",sep=""),height=5,width=5)
+pdf(paste(resloc,"abrewer_extrisk_d0_extreme_and_moderate_noise.pdf",sep=""),height=5,width=5)
 op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
@@
 
-#------------- for extreme and moderate tail dep. noise : local dispersal summary fig. ---------------------
+#------------- for extreme and moderate tail dep. noise : local dispersal summary fig. -------------------
 pdf("./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf",height=12,width=9)
@@
 plot(1, 0, bty = "n", xaxt = "n", yaxt = "n",xlab="",ylab="",type="n")
-legend("topleft", c("extreme LT","moderate LT","moderate RT","extreme RT"), lty=c(1,1,1,1), pch=c(1,1,1.1), 
-       col=c('red', 'orange','skyblue','blue'), horiz=F, bty='o', cex=1.5, adj=0, x.intersp=0.2)
+legend(x=0.6,y=1.3, c("extreme LT","moderate LT","moderate RT","extreme RT"), lty=c(1,1,1,1), pch=c(1,1,1.1), 
+       col=c('red', 'orange','skyblue','blue'), horiz=F, bty='n', cex=1.6, adj=0, x.intersp=0.2)
 mtext(paste("Ricker model: "), side = 3, line=-9, adj=0.3, cex=1, col="black")
@@
 
-pdf("./Results/allmodel_ext_risk_global_disp_extreme_and_moderatenoise.pdf",height=12,width=9)
+pdf("./Results/allmodel_ext_risk_global_disp_extreme_and_moderate_noise.pdf",height=12,width=9)
 
@@
 
-<!--FIG:  age structure model results for moderate tail dep. (C, SC)-->
+<!--FIG:  age structure model results for moderate tail dep. (C, SC)
 \begin{figure}[!h]
@@
 \end{center}
-\end{figure}
-
-<!--FIG:  summary: all nonlinear model results for local dispersion, moderate noise : extrisk vs. d-->
+\end{figure}-->
+
+<!--FIG:  summary: all nonlinear model results for global dispersion: extrisk vs. d-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_moderatenoise.pdf}
-\caption[Analogous figure as that of Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail association in spatially correlated noises.]{Analogous figure as that of Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail association in spatially correlated noises.\label{fig_extrisk_allmodel_localdisp_mtd}}
+\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_extreme_and_moderate_noise.pdf}
+\caption[Extinction risk vs.
+dispersal rate for six density dependent metapopulation model with global dispersal.]{Analogous figure as of Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text but for global dispersal between metapopulation patches.\label{fig_extrisk_allmodel_globaldisp}}
 \end{center}
@@
 
-<!--FIG:  summary: all nonlinear model results for global dispersion, extreme noise : extrisk vs. d-->
+<!--FIG: Ricker model, local dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_extremenoise.pdf}
-\caption[Extinction risk vs.
-dispersal rate for six density dependent metapopulation model with global dispersal.]{For each of six nonlinear density-dependence models presented in Table \ref{MT-tab_model_summary} with global dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$.
-Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_globaldisp_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
+standard deviation $\sigma$=1 was used to generate extreme extreme tail dependent noises for 10000 simulations. \label{fig_ricker_risk_local_disp_scl_1}}
 \end{center}
@@
 
-<!--FIG:  summary: all nonlinear model results for global dispersion, moderate noise : extrisk vs. d-->
+<!--FIG : Hassell model, local dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_moderatenoise.pdf}
-\caption[Analogous figure as that of Fig.
-\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail association in spatially correlated noises.]{Analogous figure as that of Fig.
-\ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail association in spatially correlated noises.\label{fig_extrisk_allmodel_globaldisp_mtd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
+standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_local_disp_scl_1}}
 \end{center}
@@
 
-<!--FIG: Ricker model, local dispersal, extreme noise-->
+<!--FIG: Maynard smith model, local dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme extreme tail dependent noises for 10000 simulations. \label{fig_ricker_risk_local_disp_scl_1_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
+standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations.. \label{fig_msmith_risk_local_disp_scl_0.8}}
 \end{center}
@@
 
-<!--FIG : Hassell model, local dispersal, extreme noise-->
+<!--FIG: Pennycuick model, local dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_local_disp_scl_1_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
+standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
 \end{center}
@@
 
-<!--FIG: Maynard smith model, local dispersal, extreme noise-->
+<!--FIG: Verhulst model, local dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations.. \label{fig_msmith_risk_local_disp_scl_0.8_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
+standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
 \end{center}
@@
 
-<!--FIG: Pennycuick model, local dispersal, extreme noise-->
+<!--FIG: Austin-Brewer model, local dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
+standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
 \end{center}
@@
 
-<!--FIG: Verhulst model, local dispersal, extreme noise-->
+<!--FIG : Ricker model, global dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Analogous figure as of Fig.
+\ref{fig_ricker_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
+\ref{fig_ricker_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.
+\label{fig_ricker_risk_global_disp_scl_1}}
 \end{center}
@@
 
-<!--FIG: Austin-Brewer model, local dispersal, extreme noise-->
+<!--FIG: Hassell model, global dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
-standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Analogous figure as of Fig.
+\ref{fig_hassell_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
+\ref{fig_hassell_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.
+\label{fig_hassell_risk_global_disp_scl_1}}
 \end{center}
@@
 
-<!--FIG: Ricker model, local dispersal, moderate noise-->
+<!--FIG: Maynard smith model, global dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_ricker_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_ricker_risk_local_disp_scl_1_mtd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Analogous figure as of Fig.
+\ref{fig_msmith_risk_local_disp_scl_0.8} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
+\ref{fig_msmith_risk_local_disp_scl_0.8} but for global dispersal between metapopulation patches.
+\label{fig_msmith_risk_global_disp_scl_0.8}}
 \end{center}
@@
 
-<!--FIG : Hassell model, local dispersal, moderate noise-->
+<!--FIG: Pennycuick model, global dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_hassell_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_hassell_risk_local_disp_scl_1_mtd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Analogous figure as of Fig.
+\ref{fig_risk_local_disp_pennycuick_r_vary_scl_1} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
+\ref{fig_risk_local_disp_pennycuick_r_vary_scl_1} but for global dispersal between metapopulation patches.
+\label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
 \end{center}
@@
 
-<!--FIG: Maynard smith model, local dispersal, moderate noise-->
+<!--FIG: Verhulst model, global dispersal-->
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_msmith_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Pennycuick model, local dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_pennycuick_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Verhulst model, local dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_verhulst_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Austin-Brewer model, local dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Analogous figure as of Fig.
-\ref{fig_abrewer_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG : Ricker model, global dispersal, extreme noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with global dispersal; standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations.
-\label{fig_ricker_risk_global_disp_scl_1_etd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Hassell model, global dispersal, extreme noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with global dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_global_disp_scl_1_etd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Maynard smith model, global dispersal, extreme noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with global dispersal; 
-standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_msmith_risk_global_disp_scl_0.8_etd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Pennycuick model, global dispersal, extreme noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with global dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_etd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Verhulst model, global dispersal, extreme noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with global dispersal; 
-standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Analogous figure as of Fig.
+\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
+\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} but for global dispersal between metapopulation patches.
+\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_extremenoise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with global dispersal; 
-standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_etd}}
+\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
+\includegraphics[width=15 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
+\caption[Analogous figure as of Fig.
+\ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
+\ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} but for global dispersal between metapopulation patches.
+\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}
 \end{center}
 \end{figure}
-
-<!--FIG: Ricker model, global dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_ricker_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_ricker_risk_global_disp_scl_1_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG : Hassell model, global dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_hassell_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_hassell_risk_global_disp_scl_1_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Maynard smith model, global dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_msmith_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_msmith_risk_global_disp_scl_0.8_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Pennycuick model, global dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_pennycuick_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_risk_global_disp_pennycuick_r_vary_scl_1_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Verhulst model, global dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_verhulst_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG: Austin-Brewer model, global dispersal, moderate noise-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_moderatenoise.pdf}\\
-\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{fig_abrewer_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches.
-\label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}}
-\end{center}
-\end{figure}
-
-<!--FIG : all nonlinear model results with moderate noise, d=0, extrisk vs. r-->
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
-\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_moderatenoise.pdf}
-\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_moderatenoise.pdf}\\
-\vspace{-0.6 cm}
-\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_moderatenoise.pdf}
-\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_moderatenoise.pdf}\\
-\vspace{-0.6 cm}
-\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_moderatenoise.pdf}
-\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_moderatenoise.pdf}\\
-\caption[Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
-tail dependent noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
-tail dependent noises; Moderate left tail association (or right tail association) was generated 
-by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between
-any two out of $n=5$ patches.\label{fig_extrisk_d0_mtd}}
-\end{center}
-\end{figure*}
 

### 2019-10-12 MainText.Rmd
editing intro...continued

@@
 in the appropriate literature, and that is great!-->
-through 
-life-history traits, alter community structures through cascading effect on 
+through life-history traits, alter community structures through cascading effect on 
 higher hierarchical levels and lead to local extinction of 
@@
 to local extinction.-->
-<!--***DAN: Shyamolina, demogra[hic stochasticity is a specific thing. I suspect you may
-just want to say instead something like "All these studies indicated that environmental stochasticity (through ECEs) can make a population vulnerable to extirpation."] -->
 
-<!--Earlier study on metapopulation extinction risk : Large-scale spatial synchrony may also be generated by
-other processes such as migration and predation, but regional
+<!--Earlier study on metapopulation extinction risk : Large-scale spatial synchrony 
+may also be generated by other processes such as migration and predation, but regional
 stochasticity in the form of spatially correlated weather conditions is
@@
 variables are more associated in their smaller values than larger, they are said to have stronger
-"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). Similarly, if two positively correlated 
-variables are more associated in their larger values than smaller, they are said to have stronger
-"right-tail" or "upper-tail" association (Figure \ref{fig_intro_cop}C, D). Figure \ref{fig_intro_cop} shows 
-two random variables can have a wide range of assciation from extreme to moderate degree in their extreme tails, but
-with a fixed Kendall's $\tau$.
+"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). Similarly, if two positively correlated variables are more associated in their larger values than smaller, they are 
+said to have stronger "right-tail" or "upper-tail" association (Figure \ref{fig_intro_cop}C, D). 
+Figure \ref{fig_intro_cop} shows two random variables can have a wide 
+range of assciation from extreme to moderate degree in one of their 
+extreme tails - often refers to having 'asymmetric tail' - but with a 
+fixed Kendall's $\tau$. One can control the degree of "tail-asymmetry" - moderate 
+to extreme - using "copula" (for details see @Ghosh_copula and \nameref{Methods}), a statistical 
+tool generally used to model dependencies between correlated variables. 
+This 'asymmetric tail' is commonly found in many 
+bivariate or multivariate dependence structures for environmental, species-trait, 
+phenological, population, community, and ecosystem functioning datasets 
+and can help to interpret the causal mechanisms as well as 
+to draw consequential inferences for such patterns [@Ghosh_copula]. For example, it 
+is well known that spatial dependence of environmental
+variables is an important driver of spatial dependence in fluctuating 
+population time series (called 'spatial synchrony'). Therefore, one may expect 
+the synchronous populations over a larger network would suppress for extreme bad 
+years of limiting resources than moderately 'good' years. Any two such populations, that are 
+affected badly in a similar manner due to common food shortage, are expected to be more
+"left-tail" associated (Figure \ref{fig_intro_cop}A, B) rather than "right-tail" (Figure \ref{fig_intro_cop}C, D) as the environmental fluctuation 
+time-series have more "left-tail" association for those extreme bad years
+(i.e., 'Moran-effect' like mechanism). Similarly, during 
+favorable years if pest metapopulations 
+grow abundantly compared to other average years, population time series look like 
+\ref{fig_intro_cop}C or D with stronger "right-tail" association, turning local 
+outbreaks into regional epidemics. In a 
+nutshell, "tail-asymmetry" could have profound effect in understanding 
+risk for population viability or ecosystem function
+considering their relation with extreme climatic events.
 
-” But this is not the topic sentence.
-The topic sentence comes second and is something like “Tail associations are…” or “We say two positively associated variables are more right- than left-tail associated if…”.
-Then the para continues, its point being to explain what asymmetric tail associations are.
-You can use a pedagogical figure.
-One thing you probably want to include is the logic (see BIVAN) that if extreme events are getting stronger and more extreme/intense events are also more spatially widespread, then it seems reasonable to expect that spatial patterns of tail associations in environmental variables will also be changing.
-You may also want to mention (one sentence only) that we use the statistical concept of the copula to control tail dependence in our simulations – more will be said about copulas in the Methods.
+<!--How copula could be relevent in estimating extreme event's 
+influence on extinction risk : introduce Taildep. concept-->
+A recent study [@Ghosh_copula] used the above mentioned feature of "tail-asymmetry" 
+to estimate extinction risk for a density-independent metapopulation model. The study 
+showed "left-tail" associated density-independent metapopulations 
+to be more prone to extinctions than "right-tail" associated one as we may expect species 
+are more vulnerable to "catastrophs" (extreme bad years) than to "bonanzas" (extreme good years).
+Present manuscript is an extension of that concept in estimating the extinction risk 
+beyond the density-independent model. Population growth rate, when depending 
+on its own density, serves as a more realistic model to regulate unbounded growth, hence 
+models with "density-dependence" are often studied in population ecology to 
+predict population viability over long-term period [@hassell1975density; @fowler1987review; 
+@courchamp1999inverse]. Here, we will explore the general finding from @Ghosh_copula in the context of
+the density-dependent metapopulation patches which experience same spatially correlated noise but having a variety of tail asymmetry (from extreme to moderate, Fig.
+\ref{fig_intro_cop}).
+It is expected that extreme left-tail associated noise might wipe out some metapopulation patches
+during unfavorable years unless they are rescued by the nearby populations. This scenario 
+could be more complex and specific of the direction of association 
+depending on the dynamics of the model (see \nameref{Results}).
+
+<!--and even applicable for a simple age-structured population where its vital rates are influenced by ECEs.
+Following we will reintroduce the concept of copula in a manner that follows @Ghosh_copula.-->
+
+<!--***DAN: One thing you probably want to include is the logic (see BIVAN) that if extreme events are getting stronger and more extreme/intense events are also more spatially widespread, then it seems reasonable to expect that spatial patterns of tail associations in environmental variables will also be changing.
+You may also want to mention (one sentence only) that we use the statistical concept of the copula to control tail dependence in our simulations – more will be said about copulas in the Methods.-->
 
@@
 
-<!--How copula could be relevent in estimating extreme event's 
-influence on extinction risk : introduce Taildep. concept-->
-We recently [@Ghosh_copula] used this tail association feature using "copula" - a statistical 
-tool broadly used in modeling dependence between variables - to estimate extinction risk of a density-independent 
-metapopulation model. Present manuscript is an extension of this concept 
-in estimating the extinction risk beyond the density-independent model 
-and even applicable for a simple age-structured population where its vital rates are influenced by ECEs. 
-Following we will reintroduce the concept of copula in a manner that follows @Ghosh_copula.
-
-<!--brief intro for copula: I am not sure whether it is needed in this paper or not?-->
-<!--***DAN: Shyamolina, to answer your above Q, well, you use Clayton and survival 
-Clayton in your simulations, so you probably have to explain copulas enough to explain those.-->
-Copula, a rank based measure of association, is much more informative 
-than linear correlation measure and a widely applied tool in dependence 
-modeling [@joe2014_dependence; @nelsen2006_copula]. Copulas separate the information content of a bivariate dataset,
-$(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
-distributions (which is not about the association between the variables) and the rest of 
-the information (which is solely about the association). The association information is revealed by 
-the copula plot ($v_n$ against $u_n$) of the raw variables ($y_n$ vs. $x_n$) in Fig...., where 
-$u_n$ is the rank of $x_n$ in the set ${x_1, x_2, \ldots , x_n }$, divided by $N + 1$; 
-and $v_n$ is the rank of $y_n$ in the set ${y_1, y_2, \ldots , y_n }$, also divided
-by $N + 1$. Here the rank of the smallest element of each sets is considered to be 1 and the rank of
-the largest element is $N$. We refer to the $u_n$ and $v_n$ as normalized ranks of the $x_n$ and $y_n$. 
-Ranking makes the marginal distributions uniform, isolating only the information on 
-association between the variables.
-
-<!--tail dep. of copula & extinction risk-->
-A major benefit of the copula approach is that it can effectively capture the tail association. 
-Tail association, already exemplified in Fig..... (add Fig.), is association between the more extreme 
-values of distributions. For two positively associated variables, left-tail 
-(respectively, right-tail) association is stronger association between
-values in the left or lower portions, i. e. for smaller values 
-(respectively, the right or upper portions, i.e. for higher values) 
-of the two distributions. Stronger association in either tail causes asymmetric copula 
-structure for the joint distribution of two variables. For example, Clayton and Survival Clayton are 
-two copula families 
-<!--***DAN: Shyamolina, we cannot just mention these, no one will know what they are.
-As I said I think we should intro tail associations without copulas, and then subsequently link 
-to copulas afterwards.-->
-with asymmetric tail that are used to model exclusively 
-left and right tail association between variables, respectively.
 From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
@@
 tail association between vital rates when temperature is too cold than normal) and 
-thereby inflate the population viability. Similarly, in context of 
+thereby inflate the population viability.
+
+Similarly, in context of 
 density-dependent metapopulation models, if extreme environmental fluctuations across sites 
@@
 could be more complex and specific of the direction of association 
-depending on the dynamics of the model (see \nameref{Results}). Though tail-association 
+depending on the dynamics of the model (see \nameref{Results}).
+
+Though tail-association 
 property of a copula has been used for risk-analysis in other fields like finance, 

### 2019-10-14 MainText.Rmd
intro edited as of now, last para need to modify..Questions for DCR

@@
 fixed Kendall's $\tau$. One can control the degree of "tail-asymmetry" - moderate 
-to extreme - using "copula" (for details see @Ghosh_copula and \nameref{Methods}), a statistical 
-tool generally used to model dependencies between correlated variables. 
+to extreme - using "copula", a statistical 
+tool generally used to model dependencies between correlated variables (for details see @Ghosh_copula and \nameref{Methods}). 
 This 'asymmetric tail' is commonly found in many 
@@
 
-<!--How copula could be relevent in estimating extreme event's 
-influence on extinction risk : introduce Taildep. concept-->
 A recent study [@Ghosh_copula] used the above mentioned feature of "tail-asymmetry" 
@@
 @courchamp1999inverse]. Here, we will explore the general finding from @Ghosh_copula in the context of
-the density-dependent metapopulation patches which experience same spatially correlated noise but having a variety of tail asymmetry (from extreme to moderate, Fig.
-\ref{fig_intro_cop}).
-It is expected that extreme left-tail associated noise might wipe out some metapopulation patches
+the density-dependent metapopulation patches which experience environmental noises with same spatial correlation but with a variety of tail 
+asymmetry (from extreme to moderate, Fig. \ref{fig_intro_cop}). It is expected 
+that extreme left-tail associated noise might wipe out some metapopulation patches 
 during unfavorable years unless they are rescued by the nearby populations. This scenario 
@@
 
-<!--and even applicable for a simple age-structured population where its vital rates are influenced by ECEs.
-Following we will reintroduce the concept of copula in a manner that follows @Ghosh_copula.-->
-
-<!--***DAN: One thing you probably want to include is the logic (see BIVAN) that if extreme events are getting stronger and more extreme/intense events are also more spatially widespread, then it seems reasonable to expect that spatial patterns of tail associations in environmental variables will also be changing.
-You may also want to mention (one sentence only) that we use the statistical concept of the copula to control tail dependence in our simulations – more will be said about copulas in the Methods.-->
+Tail-asymmetry in estimating extinction risk can also be applied to an age structured model. From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
+might reduce the adult reproduction rate (fecundity) as well as larval survival.
+This scenario would resemble Fig.\ref{fig_intro_cop}(A, B) with stronger tail association between vital rates of two stages at their smaller values than larger ones and thereby inflate the population viability compared to a scenario as of Fig.\ref{fig_intro_cop}(C, D).
+<!--***Shyamolina: Should I tell this?
+It is reasonable to assume cold winter-temperature as the limiting factor 
+for each age-structured population - thereby also showing "left-tail" asymmetry between 
+the temperature and each age-population time series.(Ref: driver-synchrony relationship as shown in our aphid-plankton paper)--> Though tail-association 
+property of a copula has been used for risk-analysis in other fields like finance, 
+hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
+knowledge asymmetric tail association have never been considered 
+in estimating the effect of ECEs with age structured 
+model or density-dependent metapopulations.
 
@@
 -->
-
-From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
-might reduce the adult reproduction rate as well as larval survival (i.e. stronger 
-tail association between vital rates when temperature is too cold than normal) and 
-thereby inflate the population viability.
-
-Similarly, in context of 
-density-dependent metapopulation models, if extreme environmental fluctuations across sites 
-appear to be as synchronous (left tail dependent or right tail dependent between patches 
-with same correlation as in Fig.....) then it could wipe out some metapopulation patches 
-during unfavorable years unless they are rescued by the nearby populations. This scenario 
-could be more complex and specific of the direction of association 
-depending on the dynamics of the model (see \nameref{Results}).
-
-Though tail-association 
-property of a copula has been used for risk-analysis in other fields like finance, 
-hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
-knowledge asymmetric tail association have never been considered 
-in estimating the effect of ECEs through extinction risk of age structured 
-model or density-dependent metapopulations.
 

### 2019-10-14 MainText.Rmd
methods editing continued

@@
 ---
-title: "Asymmetric tail-association in extreme climatic events affects population-extinction risk"
+title: "Asymmetric tail association in extreme climatic events affects population-extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 life-history traits of the species. Therefore predicting the impact of extreme events 
-on population extinction risk is necessery. We here introduce the potential use of "tail-association" 
+on population extinction risk is necessery. We here introduce the potential use of "tail association" 
 property of two correlated ecological variables in modeling the extinction risk of 
@@
 to density-dependent metapopulation models. For metapopulation models, 
-we show extinction risk not only depends on the nature of tail-association of 
+we show extinction risk not only depends on the nature of tail association of 
 spatially correlated noise but is also influenced by the inherent dynamics 
@@
 cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
-But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail-association instead of simple correlation in population viability analysis.
+But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail association instead of simple correlation in population viability analysis.
 
@@
 
-<!--concept for correlation in the extreme tail: tail-association-->
+<!--concept for correlation in the extreme tail: tail association-->
 Though earlier literature considered the influence of spatial and/ or temporal 
@@
 extreme tails - often refers to having 'asymmetric tail' - but with a 
-fixed Kendall's $\tau$. One can control the degree of "tail-asymmetry" - moderate 
+fixed Kendall's $\tau$. One can control the degree of "tail asymmetry" - moderate 
 to extreme - using "copula", a statistical 
@@
 outbreaks into regional epidemics. In a 
-nutshell, "tail-asymmetry" could have profound effect in understanding 
+nutshell, "tail asymmetry" could have profound effect in understanding 
 risk for population viability or ecosystem function
@@
 
-A recent study [@Ghosh_copula] used the above mentioned feature of "tail-asymmetry" 
+A recent study [@Ghosh_copula] used the above mentioned feature of "tail asymmetry" 
 to estimate extinction risk for a density-independent metapopulation model. The study 
@@
 to be more prone to extinctions than "right-tail" associated one as we may expect species 
-are more vulnerable to "catastrophs" (extreme bad years) than to "bonanzas" (extreme good years).
+are more vulnerable to "catastrophes" (extreme bad years) than to "bonanzas" (extreme good years).
 Present manuscript is an extension of that concept in estimating the extinction risk 
@@
 
-Tail-asymmetry in estimating extinction risk can also be applied to an age structured model. From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
+Tail asymmetry in estimating extinction risk can also be applied to an age structured model. From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
 might reduce the adult reproduction rate (fecundity) as well as larval survival.
@@
 for each age-structured population - thereby also showing "left-tail" asymmetry between 
-the temperature and each age-population time series.(Ref: driver-synchrony relationship as shown in our aphid-plankton paper)--> Though tail-association 
+the temperature and each age-population time series.(Ref: driver-synchrony relationship as shown in our aphid-plankton paper)--> Though tail association 
 property of a copula has been used for risk-analysis in other fields like finance, 
@@
 <!--our objective ?-->
-In this paper, we will explore the potential use of asymmetric tail-association 
+In this paper, we will explore the potential use of asymmetric tail association 
 for understanding the extinction risk of populations towards extreme climatic event
@@
 We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
-(Q2) How does environmental noises with asymmetric tail association across multiple patches
+(Q2) How do asymmetric tail associations in environmental noises across multiple patches influence extinction risk of a density dependent metapopulation model, and how do the effects 
+interact with the nature of density dependence? Will they remain consistent with the 
+general finding from density-independent metapopulation model [@Ghosh_copula]
+(higher extinction risk if noises across patches were left-tail associated than the one 
+if noises had right-tail association) irrespective of the nature of density dependence?
+
+<!--
+How does environmental noises with asymmetric tail association across multiple patches
 influence extinction risk of a density-dependent metapopulation model? Our speculation 
@@
 near the critical parameter value where the deterministic model shows transition from 
-under-compensatory to over-compensatory dynamics.
+under-compensatory to over-compensatory dynamics.-->
 
@@
 <!--stochastic matrix model : age structured-->
-To investigate (Q1) from \nameref{Introduction} that how does asymmetric tail 
-dependent noise affect the extinction risk of
+To investigate (Q1) from \nameref{Introduction} that how do vital rates with asymmetric tail 
+association affect the extinction risk of
 populations in a single patch age-structured model,
-we considered a stochastic age-structured model (Eq. \ref{eq_age_str}) 
-with two age classes (egg $E$ and female adult $A$). We assumed
-the population was semelperous, i.e., adult ($A$) went through the
+we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
+with two age classes (egg $E$ and female adult $A$).
+
+\begin{equation}
+\left({\begin{array}{cc} E \\ A \end{array}}\right)_{t+1} = 
+\left(\begin{array}{cc} 0 & f_a\\ s_e & 0 \end{array}\right)
+\left(\begin{array}{c} E \\ A \end{array}\right)_{t} \label{eq_age_str}
+\end{equation}
+
+\begin{equation}
+f_a \sim \Gamma (k, \theta)\exp(-A/\Omega), \hspace{0.2 cm} s_e \sim \mathcal{B}(\alpha,\beta) \label{eq_vitalrates}
+\end{equation}
+
+We assumed the population was semelperous, i.e., adult ($A$) went through the
 reproductive episode only once before death. Eggs and adults in the 
@@
 (i.e., Eq. \ref{eq_age_str} was "linear" model then), 
-but when $\Omega$ was finite, Eq. \ref{eq_age_str} became "nonlinear".
+but when $\Omega$ was finite, Eq. \ref{eq_age_str} became "nonlinear" with density dependence.
 Fecundity ($f_a$) and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
 a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. We generated those distributions 
-from iid noises with a varied degree of asymmetric tail-association between 
+from iid noises with a varied degree of asymmetric tail association between 
 fecundity ($f_a$) and survival ($s_e$) rates. Degree of tail association had two broad 
 categories: moderate and extreme. Each category had specific direction in tail association: 
-either in left or in right tail of vital rates. Vital rates ($f_a$ and $s_e$) were
-left tail dependent when unusual bad years ("catastrophs") suppressed both vital rates synchronously, 
+either in left or in right tail of vital rates (as depicted in Fig. \ref{fig_intro_cop}). Vital 
+rates ($f_a$ and $s_e$) were left tail dependent when unusual bad years ("catastrophes") 
+suppressed both vital rates synchronously, 
 and right tail dependent when unexpected good years ("bonanzas") boosted both of them.
 
-\begin{equation}
-\left({\begin{array}{cc} E \\ A \end{array}}\right)_{t+1} = 
-\left(\begin{array}{cc} 0 & f_a\\ s_e & 0 \end{array}\right)
-\left(\begin{array}{c} E \\ A \end{array}\right)_{t} \label{eq_age_str}
-\end{equation}
-
-\begin{equation}
-f_a = \Gamma (k, \theta)\exp(-A/\Omega), \hspace{0.2 cm} s_e = \mathcal{B}(\alpha,\beta) \label{eq_vitalrates}
-\end{equation}
-
-We used bivariate Clayton copula to model moderate left-tail and Survival Clayton copula to model 
-moderate right tail association in the vital rates. Both copula had the same Kendall 
-correlation ($=0.6$) but opposite association in their respective tails. 
+Moderate tail associatation was modeled by using the 
+concept of "copula", a statistical tool to model 
+dependencies. Appendix \ref{SM-intro_copula} summarized
+a brief introduction of "copulas" and their relation to tail asscociations, symmetries 
+with examples of available copula families 
+from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
+copula to model moderate left-tail and Survival Clayton copula to model 
+moderate right-tail association in the vital rates. Both copula had the same Kendall 
+correlation ($\tau =$ `r ktau_common`) but opposite association in their respective tails. 
 We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
-right (Survival Clayton) tail-dependent bivariate copulas (see `BiCopSim` function from
+right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
 `VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
@@
 `iTau` function from `copula` package so that both copula had
-same Kendall's correlation ($\tau = 0.6$). Entries within each clumn of
+same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each clumn of
 matrix $M$, then became independent of each others but 
 entries between columns had a specific tail association
-with positive Kendall's correlation ($\tau = 0.6$).
+with positive Kendall's correlation.
 
@@
 For each $25$ timesteps from each of $10000$ simulations,
-we feed pairwise entries from each row of matrix $M$ as the vital rates 
-(fecundity $f_a$ and survival rates $s_e$) 
-of the stochastic projrction matrix in model (\ref{eq_age_str}).
+we incorporated pairwise entries from each row of matrix $M$ in the expression for 
+vital rates (fecundity $f_a$ and survival rates $s_e$, Eq. \ref{eq_vitalrates}) of 
+the stochastic projrction matrix in model (\ref{eq_age_str}).
 
@@
 environmental variable across space could influence the extinction risk of a 
-metapopulation model with density-dependence feedback. We considered six density-dependent 
+metapopulation model with density-dependent feedback. We considered six density-dependent 
 population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
@@
 and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
-All models assumed a generalized form : $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
+All models assumed a generalized form: $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
 where $\vec{P_t}$ was a vector of population densities, 
@@
 We simulated our metapopulation models with noises $(\epsilon_t^i)$ that were 
-independent through time with same spatial correlations between patches (to within sampling error)
-for every simulation, but moderately (or extremely) left- and right-tail dependent between patches in
-different simulations. For moderately left-tail dependent noise we drew points from a $n$ dimensional 
-Clayton copula with same Kendall correlation ($= 0.6$) between any two out of $n$ patches and made 
+independent through time with same spatial correlations 
+between patches (to within sampling error)
+for every simulation, but were moderately (or extremely) left- and 
+right-tail associated between patches in
+different simulations. For moderately left-tail associated 
+noises we drew points from a $n$ dimensional 
+Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
+any two out of $n$ patches and made 
 the marginals normal with zero mean and standard deviation $\sigma$ 
@@
 column of Table \ref{tab_model_summary}). For moderately 
-right-tail dependent noise we flipped ($180^{\circ}$ rotation) the points drawn 
+right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
 from Clayton copula so that all the statistical properties remained 
-same except for the direction of tail-association. We generated
-extreme left- or right-tail dependent noises via algorithm stated in 
+same except for the direction of tail association. We generated
+extreme left- or right-tail associated noises via algorithm stated in 
 Appendix \ref{SM-extr_tail_dep}. Our models (Table \ref{tab_model_summary}) 
@@
 to get a positive, finite equilibrium $P_e$ (for details, see Appendix 
-\ref{SM-model_stability}) to start with ($t=0$). For each chosen $r$,
+\ref{SM-model_stability}) to start with (at $t=0$). For each chosen $r$,
 we calculated extinction risk after $25$ years with each metapopulation model 
 where population could disperse locally (or globally) with varying dispersal 
-rate $d= 0,0.1,0.2,..0.5$ and extreme (or moderate) left 
-tail- and right-tail dependent environmental noise 
+rate $d= 0,0.1,0.2,..0.5$ and moderate to extreme left 
+tail- and right-tail associated environmental noises with same Kendall's $\tau$ (= `r ktau_common`)
 in different simulations (for code see [`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R)).
+
+<!--start from here-->
 
@@
 and thereby can increase the extinction risk of the population. For non-linear age-structured 
-models, we got the similar trend for extinction risks with different tail-association (left and right).
+models, we got the similar trend for extinction risks with different tail association (left and right).
 Non-linear model showed higher extinction risk than its corresponding 
@@
 (Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
-a population could be considered as direct manifestation of tail-association property 
+a population could be considered as direct manifestation of tail association property 
 between vital rates of different age-class
@@
 association in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
-in years of extreme events ("catastrophs" and "bonanzas"). Though we could expect left tail association 
+in years of extreme events ("catastrophes" and "bonanzas"). Though we could expect left tail association 
 to show higher extinction risk than the one for right tail association with age-structured 
@@
 and the temporal variance of the log(population growth rate) caused by environmental stochasticity - which was based
-on the linear covariance measure between any two matrix elements, not on their tail-association properties. Therefore,
+on the linear covariance measure between any two matrix elements, not on their tail association properties. Therefore,
 following @tuljapurkar1990lecture, population growth rate in an iid environment could be approximated
@@
 influence extinction risk, substantial work remains to determine the importance of this effect in the 
-following scenarios. First, we demonstrated the effect of tail-association on extinction risk 
+following scenarios. First, we demonstrated the effect of tail association on extinction risk 
 with a basic age structured (two age stages) model where we considered the total semelperous 
@@
 of populations including semelperous and iteroperous [@morris2002quantitative]. Future studies should include 
-tail-association effect on extinction risk 
+tail association effect on extinction risk 
 for a multiple age-, size- and stage- structured model and estimate extinction risk 

### 2019-10-14 SuppMat.Rmd
methods editing continued

@@
 ---
-title: "Supporting Information : Asymmetric tail-association in extreme climatic events affects population-extinction risk"
+title: "Supporting Information : Asymmetric tail association in extreme climatic events affects population-extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 
-# Generating matrix with extreme tail association  \label{extr_tail_dep}
-
-For extreme tail-association algorithm, a $N \times n$ matrix, M was generated of independent 
+# A brief introduction about "copulas"  \label{intro_copula}
+
+# Algorithm to generate extreme tail association  \label{extr_tail_dep}
+
+For extreme tail association algorithm, a $N \times n$ matrix, M was generated of independent 
 standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->

### 2019-10-15 SuppMat.Rmd
edited maintext

@@
 
+<!--needs editing, I just dumped old text here-->
+Copula, a rank based measure of association, is much more informative 
+than linear correlation measure and a widely applied tool in dependence 
+modeling [@joe2014_dependence; @nelsen2006_copula]. Copulas separate 
+the information content of a bivariate dataset,
+$(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
+distributions (which is not about the association between the variables) and the rest of 
+the information (which is solely about the association). The association information is revealed by 
+the copula plot ($v_n$ against $u_n$) of the raw variables ($y_n$ vs. $x_n$) in Fig...., where 
+$u_n$ is the rank of $x_n$ in the set ${x_1, x_2, \ldots , x_n }$, divided by $N + 1$; 
+and $v_n$ is the rank of $y_n$ in the set ${y_1, y_2, \ldots , y_n }$, also divided
+by $N + 1$. Here the rank of the smallest element of each sets is considered to be 1 and the rank of
+the largest element is $N$. We refer to the $u_n$ and $v_n$ as normalized ranks of the $x_n$ and $y_n$. 
+Ranking makes the marginal distributions uniform, isolating only the information on 
+association between the variables.
+
 # Algorithm to generate extreme tail association  \label{extr_tail_dep}

### 2019-10-15 SuppMat.Rmd
removed a hyphen in title

--- 2019-10-14 SuppMat.Rmd
@@
 
-# A brief introduction about "copulas"  \label{intro_copula}
-
-# Algorithm to generate extreme tail association  \label{extr_tail_dep}
-
-For extreme tail association algorithm, a $N \times n$ matrix, M was generated of independent 
+# Generating matrix with extreme tail association  \label{extr_tail_dep}
+
+For extreme tail-association algorithm, a $N \times n$ matrix, M was generated of independent 
 standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->

### 2019-10-15 MainText.Rmd

--- 2019-10-14 MainText.Rmd
@@
 <!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
-To test our hypothesis from (Q2) of \nameref{Introduction} for each metapopulation model,
+To answer (Q2) of \nameref{Introduction} for each metapopulation model,
 we estimated extinction risk for a wide range of parameter $r$ 
@@
 
-<!--start from here-->
-
-<!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->
-Given hypothesis from (Q2) was valid - that left tail dependent noise made extinction risk 
-higher in undercompensatory regime ($r < r_c$) and
-right tail dependent noise made extinction risk higher
-in overcompensatory regime ($r > r_c$) - we tested our third hypothesis for (Q3).
-We expected the switching to take place near $r_c$, not exactly at $r = r_c$ because the transition point $r_c$
-from under to overcompensatory dynamics differed in stochastic environment than its 
-deterministic analog depending on the underlying nonlinearity of the model.
-We plotted extinction risk along y-axis
-for both left tail and right tail dependent noises 
+We plotted extinction risks along y-axis
+for both left tail and right tail associated noises 
 against the varying range of $r$ along x-axis when no dispersal ($d=0$) was considered.
-We excluded the possibility of dispersal in this case to avoid any rescue effect among patches. 
-For each model with extreme (or moderate) left- and right-tail dependent noises, we calculated 
+We excluded the possibility of dispersal to avoid any rescue effect among patches. 
+For each model with extreme (or moderate) left- and right-tail associated noises, we calculated 
 2.5$^{th}$ and 97.5$^{th}$ quantiles
 of 10000 simulations with the probability of computed
-extinction risk at each $r$. We compared $r_c$ and the value of $r$ at which 
-95\% CI of extinction risk from both noises overlapped
-with each other. This could give us an idea for supporting our hypothesis.
+extinction risks at each $r$. We compared $r_c$ and the value of $r$ at which 
+95\% CI of extinction risks from both noises overlapped
+with each other. This could give us an idea of how extinction risks affected by the model parameter 
+$r$ that controlled the underlying dynamics of the model, i.e., undercompensatory 
+or overcompensatory.
 
@@
 linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
-Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with two copulas having different 
-asymmetric tails but same Kendall's correlation ($\tau = 0.6$). Extinction risks for
-two different tail-dependent vital rates were not strinkingly different (e.g. see pair of red and
-blue lines in Fig. \ref{SM-fig_extrisk_agestr_mtd}A or B) and we anticipated the difference being partly 
+Fig.
+\ref{fig_extrisk_agestr} represents this scenario when vital rates have moderate to extreme tail association in either tail but with same Kendall's correlation ($\tau =$ `r ktau_common`).
+Extinction risks for two different tail-dependent vital rates were not strinkingly different (e.g.
+see pair of red and
+blue lines in Fig. \ref{fig_extrisk_agestr}A or B) and we anticipated the difference being partly 
 dependent on the degree of tail association. Therefore we repeated the same analysis with 
@@
 copulas and got the similar results but with much prominent difference between their extinction risks
-(Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
+(Fig. \ref{fig_extrisk_agestr}C or D). This result answered (Q1) and showed extinction risk of 
 a population could be considered as direct manifestation of tail association property 
@@
 environment showed similar patterns for extinction risk as density-independent 
-metapopulation model showed in recent study [@Ghosh_copula]. For each density-dependent 
+metapopulation model did in recent study [@Ghosh_copula]. For each density-dependent 
 model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
-the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail association.
-But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - was 
+the metapopulations when spatially correlated noises were left-tail associated 
+than the one with right-tail association. But that observation remained 
+valid when the model parameter - here considered as growth rate $(r)$ - was 
 sufficiently low so that under-compensatory dynamics (monotonous approach towards
 its stable equilibrium) for the deterministic analog of the model was ensured.
-We observed right tail dependent noise caused higher extinction 
-risk for the metapopulations than left tail dependent noise when the model 
+We observed right tail associated noise caused higher extinction 
+risk for the metapopulations than left tail associated noise when the model 
 was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
-value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp} shows
-variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
-(see Table \ref{tab_model_summary}) with growth rate $r$ considered either from under-compensatory or 
-from over-compensatory regime. We also got similar results with our prediction (see Q2 from \nameref{Introduction})
+value of $r$. Fig. \ref{fig_extrisk_allmodel_localdisp} shows
+variation of extinction risk against local dispersal rate ($d$) for 
+six density dependent metapopulation models (see Table \ref{tab_model_summary}) with growth 
+rate $r$ considered either from under-compensatory or 
+from over-compensatory regime. This finding addresses (Q2) from \nameref{Introduction}. We 
+also got similar results 
 for global dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
-Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}) 
+Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
+\ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6}) 
 demonstrate how extinction risk changed against local dispersal rate for a whole range
-of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1_etd} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) 
+of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
 represent the same but for global dispersal among patches. 
@@
 <!--Q3 : metapopulation results : switching near rc for d=0-->
-Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) we hypothesized that the switching 
-between the extinction risk generated by left-tail dependent and right-tail dependent noise 
-occurred near the critical parameter $r_c$ (see Q3). For simplicity we only considered 
-$d=0$ case, i.e., when there was no dispersal between patches (see \nameref{Methods}). 
-Figs. \ref{fig_extrisk_d0_etd} and \ref{SM-fig_extrisk_d0_mtd} demonstrate higher extinction risk 
-with left-tail dependent noise for $r<r_c$ and with right-tail dependent noise 
-for $r>r_c$. $r_c$ was the vertical line in Figs. \ref{fig_extrisk_d0_etd} and \ref{SM-fig_extrisk_d0_mtd} 
+Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) one can reasonably hypothesize
+that the switching between the extinction risk 
+generated by left-tail associated and right-tail associated noise 
+occurred near the critical parameter $r_c$. For simplicity, we only considered 
+$d=0$ case, i.e., when there was no dispersal between patches. 
+Fig. \ref{fig_extrisk_d0} demonstrates higher extinction risk 
+with left-tail associated noise for $r<r_c$ and with right-tail associated noise 
+for $r>r_c$. $r_c$ was the vertical line in Fig. \ref{fig_extrisk_d0} 
 indicating the value of $r$ for each model where their deterministic analog switches 
@@
 We found left tail association in vital rates (induced by ECEs) caused higher extinction risk than 
-the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_mtd}) 
-unless the tail association was at its extreme (Fig. \ref{fig_extrisk_agestr_etd}). This finding could be explained 
-by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
-The approximation considered growth rate being dependent on the most dominant eigen value of the mean projection matrix 
-and the temporal variance of the log(population growth rate) caused by environmental stochasticity - which was based
-on the linear covariance measure between any two matrix elements, not on their tail association properties. Therefore,
-following @tuljapurkar1990lecture, population growth rate in an iid environment could be approximated
-as same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail association.
+the right tail dependent ones, but the difference was not remarkable 
+unless the tail association was at its 
+extreme (Fig. \ref{fig_extrisk_agestr}). This finding could be explained 
+by the approximation @tuljapurkar1990lecture made for the 
+stochastic log(growth rate) in an iid environment.
+The approximation considered growth rate being 
+dependent on the most dominant eigen value of the mean projection matrix 
+and the temporal variance of the log(population growth rate) caused 
+by environmental stochasticity - which was based
+on the linear covariance measure between any 
+two matrix elements, not on their tail association properties. Therefore,
+following @tuljapurkar1990lecture, population growth 
+rate in an iid environment could be approximated
+as same for our model (Eq. \ref{eq_age_str}) with 
+left-tail or right-tail association.
+
+<!--Intution/ possible explanation for expected results/ hypo?-->
 
@@
 \begin{center}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}\\
 \end{center}
@@
 \begin{center}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
-\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
-\caption{Extinction risk against time from single patch age structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(C, D)$ models for $10000$ simulations; shading represents 95\% confidence interval.
-$(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have extreme left tail (LT) or extreme right tail (RT) association, $(C, D)$ show the analogous scenario but with moderate tail association between vital rates.
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
+\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}
+\caption{Extinction risk against time from single patch age structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(B, D)$ models for $10000$ simulations; shading represents 95\% confidence interval.
+$(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have moderate left tail (LT) or moderate right tail (RT) association, $(C, D)$ show the analogous scenario but with extreme tail association between vital rates.
 \label{fig_extrisk_agestr}}
@@
 for each model where their deterministic analog switches from under- to over-compensatory dynamics.
-\label{fig_extrisk_d0_etd}}
+\label{fig_extrisk_d0}}
 \end{center}

### 2019-10-23 MainText.Rmd
edited the Intro

@@
 
-# Alternative titles
+<!--
+Alternative titles
 1) Tail associations in environmental variables and extinction risk
 2) The influence of correlations in the extremes on extinction risk
+DAN FAVORITE SO FAR 3) Tail associations in environmental variables and their impact on extinction risk
+-->
 
@@
 
+<!--***DAN: Dan has not edited this--> 
 With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
@@
 <!--***DAN: 
-1) Shyamolina,  was reminded during my editing of BIVAN that "tail dependence" has 
-a specific mathematical definition (now given in the Background section of BIVAN), and that
-is not what we mean here when we say "tail dependence". So we cannot go on calling it "tail
-dependence" in this manuscript - that would be mike me calling my house "the white house" - 
-there already is one of those, and it is much better known than my house, so we cannot
-attempt to hijack the name. So can you change terminology throughout, please? Let's refer to
-it using terminology such as "associations in the tails" or 
-"tail associations". I would like you to provide a 
-definition at an appropriate place early in the manuscript that we will call a pair
-of positively associated variables "left-tail associated" if their association in their 
-left tails is stronger than their association in their right tails (in a sense to be defined
-mathematically in Methods, and then in methods we will define things based on cor_l-cor_u).
-We call a pair of positively associated variable "right-tail associated" if their association
-in their right tails is stronger than their association in their left tails.
-
-2) You need to start getting this paper ready for submission - if you want to send it to 
-the Ecosphere special issue (what do you think of that?) then we have to get moving. Please 
-let me know soon, after reading about Ecospere on their website. If you think this is a 
-good idea, let me know and I will send a summary of the paper to Jon Walter to make sure he is still willing to have us submit the paper to him, and to make sure the paper is close enough
-to the theme of the special issue that he is going for, so that he will be willing to see it
-submitted there.
-
 3) Please
@@
 sentences in English. 
-
-4) Assuming you want to go for Ecosphere and Jon is OK with us submitting there, please read the "guidelines for authors" section of the Ecosphere website 
-and format the paper for Ecosphere.
-
-5) Please make all the improvements that you can see for yourself
-are needed! I suspect if you read the paper carefully now after having worked on something
-else for a while you will see many things you can improve!
 -->
@@
 
-<!--how common are the extreme events?-->
-For past few decades, extreme climatic events (ECEs) appeared more common than 
-usual due to anthropogenic climate change, e.g. temperature or precipitation extremes,
-heat-waves, wild-fire etc. [@ummenhofer2017extreme]. Frequent extreme events 
-can shift temperature or precipitation-mean, 
-change variability of the distribution, as well as change its 
-symmetry or skewness [@hansen2012perception]. Extreme climatic events can pose severe
-impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
+<!--ECEs have a lot of impacts, it is important to understand their impacts-->
+Over the past few decades, extreme climatic events (ECEs) such as
+temperature or precipitation extremes, heat-waves, and extreme wildfires
+have become more common or more extreme due to anthropogenic climate change 
+[@ummenhofer2017extreme]. ECEs are well known to produce severe
+impacts at the population, community and ecosystem levels [@bragazza2008climatic; 
 @gutschick2003extreme; @jiguet2011community; @mckechnie2009climate;
-@felton2017integrating]. Therefore, 
-understanding the long-term impacts of changes 
-in ECEs has been a longstanding interest in ecological research 
-[@katz2005statistics; @smith2011ecological; @diez2012will; @moritz2013future; @bailey2016tackling].
-
-<!--How extinction risk is related to extreme events : past studies
-consequences of extreme events on population life history, community-->
-Earlier studies showed ECEs may affect species abundances (species re-ordering) 
-<!--***DAN: Shyamolina, I do not know what the parenthesis refers to
-
-***Shyamolina: species reordering is used to access the change in community 
-in response to extreme events or global environmental changes, it relates to change 
-in relative abundances of dominant species (mainly). For example, as different species have different 
-sensitivity levels to extreme condition (means the extreme threshold is not 
-constant for all the species in a community), some years with extreme drought favors say c3 grass 
-but in the next year c4 grass could be dominant. Such alteration in community composition has "impact on 
-ecosystem function, leading to an extreme ecological response". For detailed overview you may see 
-@hoover2014resistance, Jones et al. Ecol. lett. 2017 (doi: 10.1111/ele.12864), 
-Smith et al., Ecology 2009 (doi:10.1890/08-1815.1).
-Actually LAN K. KNAPP and MELINDA D. SMITH did a lot of work on this to access changes in 
-plant community response to extreme events 
-(this review article: Felton et al. 2017, Phil. Trans. R. Soc. B 372: 20160142 is a nice summary).
--->
-<!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
-in the appropriate literature, and that is great!-->
-through life-history traits, alter community structures through cascading effect on 
-higher hierarchical levels and lead to local extinction of 
-species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
-<!--Should I mention idea of match-mismatch hypothesis?-->
-For example, a study [@chiu2013impact] showed extreme flood affected 
-the annual survival of a prey community that in turn affected predators in 
-aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
-@frederiksen2008demographic used $43$ years of ring-recovery data 
-to examine the demographic impact of ECEs 
-through temporal variability in juvenile, immature and adult survival
-of European shags on the Isle of May in eastern Scotland.
-A relatively recent study [@mcdermott2017sensitivity] investigated the sensitivity 
-of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
-extreme precipitation, drought) over $37$ years at the site level, across 
-each species' life-stages. Experimental study showed successive ECEs with short time intervals 
-could weaken the resistance of plant communities too [@dreesen2014successive]. 
-All these studies indicated that environmental stochasticity (through ECEs) can 
-make a population vulnerable to extirpation.
-<!--All these studies indicated that environmental stochasticity (through ECEs) 
-can lead to demographic stochasticity (through influence on reproductive or 
-mortality rate) of different age class of a population making it vulnerable 
-to local extinction.-->
+@felton2017integrating], and thus the goal of improving our understanding of
+the variety of ecological impacts of ECEs is an important applied goal for 
+modern ecology [@katz2005statistics; @smith2011ecological; @diez2012will; @moritz2013future; @bailey2016tackling]. Frequent extreme events 
+can help alter the mean temperature or precipitation of an area, 
+can be associated with changes through time in the overall variance of 
+climate signals, and can directly contribute to changes in the symmetry or
+skewness of distributions of climate signals through time 
+[@hansen2012perception]. All such changes have the potential to influence 
+populations, including their risk of extirpation from an area 
+[@GarciaCarerras2013; @Vasseuretal2014].
+For instance, @frederiksen2008demographic used $43$ years of ring-recovery data 
+to examine the demographic impact of ECEs on European shags 
+via their effects on temporal variability in 
+juvenile, immature and adult survival rates.
+@mcdermott2017sensitivity investigated the sensitivity 
+of $41$ UK butterfly species to four different types of ECEs. 
+At the community level, research has showed that ECEs can reorder
+the dominance heirarchy of species in an area, sometimes leading to 
+selective local extictions, and can also produce effects that cascade 
+across trophic levels [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
+For example, @chiu2013impact showed that extreme floods affected 
+the annual survival of a prey community, in turn influencing predators in 
+aquatic as well as adjacent riparian habitats. 
+@dreesen2014successive, in an experimental study, showed that 
+successive ECEs with short time intervals between them
+can weaken the resistance of plant communities.
+<!--***DAN: Shyamolina, what, precisely, does it mean to "weaken the resistance of a
+plant community"? Resistance to what? Please clarify.-->
 
@@
 probably the dominant synchronizing mechanism (ref: Nature1998) but what about taildep. effect on extrisk?.-->
-In the metapopulation context, regional stochasticity 
-can simultaneously affected multiple habitat patches where spatially correlated weather conditions were 
-probably the dominant synchronizing mechanism for extinctions [@hanski1998metapopulation]. A 21 years-long
-study by @tack2015increasing on the Glanville fritillary butterfly demonstrated that frequency 
-of drought during early larval instars increased amplitude and span of spatial synchrony 
-over a metapopulation network of 4000 dry meadows and could affect the long-term population viability.
-Therefore, understanding the synchronized fluctuation over a metapopulation-network
-remains important to the conservationists because well connected patches, if synchronous, are more 
-prone to simultaneous extinction during extreme bad years [@earn2000coherence].
+
+<!--ECEs have impacts in the spatial context as well-->
+ECEs also have impacts in ecological scenarios for which spatial dynamics are
+important.
+In the metapopulation context, spatial synchrony of population dynamics, which can
+be caused by spatial synchrony of climatic fluctuations, is linked to metapopulation
+extinction risk [@hanski1998metapopulation; @earn2000coherence].
+If ECEs are also spatially more extensive than moderate events,
+as seems likely for some types of events, extreme events may accentuate
+population synchrony and thereby increase extinction risk.
+For instance, a 21-year-long
+study of @tack2015increasing on the Glanville fritillary butterfly demonstrated that frequency 
+of drought during early larval instars increased the strength and extent 
+of spatial synchrony 
+over a metapopulation network of 4000 dry meadows, and could thereby influence
+the long-term viability of the species in the area of study.
 
 <!--concept for correlation in the extreme tail: tail association-->
-Though earlier literature considered the influence of spatial and/ or temporal 
-correlations in environmental variables on extinction risk of a metapopulation, to 
-our knowledge the separate influence on extinction risk of "tail associations" between
-environmental time series in different locations has hardly been considered. "Tail associations" are
-the association between two random variables in their extreme values. If two positively correlated 
-variables are more associated in their smaller values than larger, they are said to have stronger
-"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). Similarly, if two positively correlated variables are more associated in their larger values than smaller, they are 
-said to have stronger "right-tail" or "upper-tail" association (Figure \ref{fig_intro_cop}C, D). 
-Figure \ref{fig_intro_cop} shows two random variables can have a wide 
-range of assciation from extreme to moderate degree in one of their 
-extreme tails - often refers to having 'asymmetric tail' - but with a 
-fixed Kendall's $\tau$. One can control the degree of "tail asymmetry" - moderate 
-to extreme - using "copula", a statistical 
-tool generally used to model dependencies between correlated variables (for details see @Ghosh_copula and \nameref{Methods}). 
-This 'asymmetric tail' is commonly found in many 
-bivariate or multivariate dependence structures for environmental, species-trait, 
-phenological, population, community, and ecosystem functioning datasets 
-and can help to interpret the causal mechanisms as well as 
-to draw consequential inferences for such patterns [@Ghosh_copula]. For example, it 
-is well known that spatial dependence of environmental
-variables is an important driver of spatial dependence in fluctuating 
-population time series (called 'spatial synchrony'). Therefore, one may expect 
-the synchronous populations over a larger network would suppress for extreme bad 
-years of limiting resources than moderately 'good' years. Any two such populations, that are 
-affected badly in a similar manner due to common food shortage, are expected to be more
-"left-tail" associated (Figure \ref{fig_intro_cop}A, B) rather than "right-tail" (Figure \ref{fig_intro_cop}C, D) as the environmental fluctuation 
-time-series have more "left-tail" association for those extreme bad years
-(i.e., 'Moran-effect' like mechanism). Similarly, during 
-favorable years if pest metapopulations 
-grow abundantly compared to other average years, population time series look like 
-\ref{fig_intro_cop}C or D with stronger "right-tail" association, turning local 
-outbreaks into regional epidemics. In a 
-nutshell, "tail asymmetry" could have profound effect in understanding 
-risk for population viability or ecosystem function
-considering their relation with extreme climatic events.
+Though earlier literature considered the influence of spatial synchrony 
+in environmental variables on metapopulation extinction risk, to 
+our knowledge the distinct influence of "asymmetric tail associations" between
+environmental time series in different locations has hardly been considered. 
+"Tail associations" are
+associations between two random quantities in their extreme values. 
+For instance, if two positively associated 
+variables are more associated in their smaller values than in their larger values,
+they are said to have stronger
+"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). 
+If two positively associated variables are more associated in their larger 
+values than in their smaller values, they are 
+said to have stronger "right-tail" or "upper-tail" association 
+(Figure \ref{fig_intro_cop}C, D). 
+Figure \ref{fig_intro_cop} shows that two variables can have a wide 
+range of patterns of tail association, from extreme left- to extreme right-tail
+association, with the same overall correlation. So tail association
+is a distinct concept from correlation itself, and may have distinct ecological
+ramifications [@Ghosh_copula]. Asymmetric tail associations have been documented 
+between environmental variables, and between the same environmental variable 
+measured in different places [@Serinaldi2008; @li2013; @She2018; @Goswami2018]. 
+Asymmetric tail associations have also been documented between numerous ecological
+variables and between environmental and ecological variables [@Ghosh_copula].
 
-A recent study [@Ghosh_copula] used the above mentioned feature of "tail asymmetry" 
-to estimate extinction risk for a density-independent metapopulation model. The study 
-showed "left-tail" associated density-independent metapopulations 
-to be more prone to extinctions than "right-tail" associated one as we may expect species 
-are more vulnerable to "catastrophes" (extreme bad years) than to "bonanzas" (extreme good years).
-Present manuscript is an extension of that concept in estimating the extinction risk 
-beyond the density-independent model. Population growth rate, when depending 
-on its own density, serves as a more realistic model to regulate unbounded growth, hence 
-models with "density-dependence" are often studied in population ecology to 
-predict population viability over long-term period [@hassell1975density; @fowler1987review; 
-@courchamp1999inverse]. Here, we will explore the general finding from @Ghosh_copula in the context of
-the density-dependent metapopulation patches which experience environmental noises with same spatial correlation but with a variety of tail 
-asymmetry (from extreme to moderate, Fig. \ref{fig_intro_cop}). It is expected 
-that extreme left-tail associated noise might wipe out some metapopulation patches 
-during unfavorable years unless they are rescued by the nearby populations. This scenario 
-could be more complex and specific of the direction of association 
-depending on the dynamics of the model (see \nameref{Results}).
+<!--ECEs probably produce tail associations-->
+There are reasons to believe that ECEs may produce asymmetric tail 
+associations between climatic variables
+measured in different locations. For instance, if a single ECE is associated
+with spatially widespread, large values 
+of some climatic variable, but small values of the
+variable are associated with local weather rather than ECEs, the result could
+be stronger right-tail association between measurements of the climatic variable
+made in different locations, through time.
+For example, rain storms which deposit exceptionally large amounts of rain
+may do so over large areas; but months with less rain may experience
+scattered showers only, which may be much less correlated across locales.
 
-Tail asymmetry in estimating extinction risk can also be applied to an age structured model. From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
-might reduce the adult reproduction rate (fecundity) as well as larval survival.
-This scenario would resemble Fig.\ref{fig_intro_cop}(A, B) with stronger tail association between vital rates of two stages at their smaller values than larger ones and thereby inflate the population viability compared to a scenario as of Fig.\ref{fig_intro_cop}(C, D).
-<!--***Shyamolina: Should I tell this?
-It is reasonable to assume cold winter-temperature as the limiting factor 
-for each age-structured population - thereby also showing "left-tail" asymmetry between 
-the temperature and each age-population time series.(Ref: driver-synchrony relationship as shown in our aphid-plankton paper)--> Though tail association 
-property of a copula has been used for risk-analysis in other fields like finance, 
-hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
-knowledge asymmetric tail association have never been considered 
-in estimating the effect of ECEs with age structured 
-model or density-dependent metapopulations.
+<!--***DAN: May be scope for a discussion para linking what is known about
+how ECEs are changing to what effect that may have on how tail assiciations
+in environmental variables might be changing, and what influence that may have
+on extinction risk. If little is known, then this might be a good topic for 
+future work.-->
 
-<!--***DAN: Shyamolina, you need to make two major changes in how tail associations
-are introdcued. 
-  First, do it prior to saying anything about copulas. We have been wanting
-to disassociate this paper, to some extent, from BIVAN. Well, probably the best way
-to do that is by introducing tail dependence first, before copulas. You can 
-describe it intuitively as association in the tails, and you can use pictures to describe it.
-You can introduce the partial Spearman correlation (briefly in the Intro, more detail in Methods). 
-You will need an explnation of why we 
-use Spearman instead of Pearson (talk to me if you do not know, but you may want to read the 
-new Background section of BIVAN again first). One does not need to know anything about copulas
-to understand tail associations.
+There are also reasons to hypothesize that asymmetry of tail association in 
+environmental variables may influence extinction risks for metapopulations.
+For instance, assuming for the sake of a simple example that low values of an environmental
+variable are "bad" for populations of a given species and high values are "good",
+greater left-tail associations between measurements of the variable in different
+habitat patches may cause higher metapopulation extinction risk because then
+very bad years for component populations would occur at the same time in 
+many patches, reducing the potential for rescue effects. Greater left-tail associations
+would probably not have the same effect, even if overall correlations between locations
+were the same, because it that case it would be very good years for the component 
+populations which would occur at the same time in many patches [@Ghosh_copulas].  
+@Ghosh_copula substantiated this intuition by showing that it held true for a 
+modelling setup using a spatial version of a very simple non-density dependent
+population growth model (the Lewontin-Cohen model). In essence, if environmental
+"catastrophes" (i.e., extremely bad years for a population) are widely 
+spatially synchronized it should create much greater metapopulation extinction 
+risk than if "bonanzas" (i.e., extremely good years) are widely synchronized, even
+if overall (non-tail-specific) levels of environmental synchrony are the same
+in both these cases. The present study extends tests of this idea to metapopulation
+models incorporating density dependence.
 
-  Second, the introduction to copulas that is given is outmoded. See the BAckground section
-to BIVAN for improvements. But you do not want to say anything like that much about copulas 
-here, you just want to briefly link to the concept, cite some of the Intro work that
-has been done (includeing BIVAN and some of the paper BIVAN cites for Intro work),
-mention that tail associations are one aspect of copula structure that we have found important,
-and introduce any specific things you will need, like the Clayton copula.
+Extinction risk is also very commonly estimated for age-structured models 
+such as stochastic matrix models [@MorrisDoak2002 ; @Caswell2006] and 
+integral projection models
+[@Merowetal2014]. For such models, extinction risk depends, in part, on the long-term
+population growth rate, a population parameter which in turn 
+depends on the means, variances and covariances of population vital rates such 
+as the survival and fecundity rates of age classes. Extinction risks may,
+*a priori*, also depend on possible asymmetries of tail association between
+population vital rates. For instance, it seems reasonable to imagine
+a case in which both adult fecundity and juvenile survival rates 
+depend on an environmental variable such as spring rainfall in similar ways, but 
+only when that variable takes very low values. When rainfall takes
+typical or high values, fecundity and survival may be limited instead 
+by different factors. This could produce left-tail dependence between these 
+vital rates, possibly resulting in different extinction risk for the model
+compared to a case in which vital rates were symmetrically associated in 
+their tails but otherwise statistically similar.
 
-Right now this is quite a short paper. You have space to explain things and show pedagogical
-figures.
--->
-
-<!--our objective ?-->
-In this paper, we will explore the potential use of asymmetric tail association 
-for understanding the extinction risk of populations towards extreme climatic event
-by addressing the following questions: (Q1) Does nature of tail association 
-(left tail or right tail) in vital rates
-of a single-patch age-structured population model affect extinction risk of that population differently?
-We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
-(Q2) How do asymmetric tail associations in environmental noises across multiple patches influence extinction risk of a density dependent metapopulation model, and how do the effects 
-interact with the nature of density dependence? Will they remain consistent with the 
-general finding from density-independent metapopulation model [@Ghosh_copula]
-(higher extinction risk if noises across patches were left-tail associated than the one 
-if noises had right-tail association) irrespective of the nature of density dependence?
+<!--objectives, last para of intro-->
+In this paper, we will explore the potential for asymmtric tail associations to 
+influence extinction risk in populations and metapopulations. Specifically,
+we will address the following sets of questions. Q1) Can the nature of tail associations
+between vital rates in an age-structured population model influence
+extinction risk for the model, and to what extent? Q2) Does the nature
+of tail associations between measurements of an environmental variable in different
+locations influence extinction risk of a density dependendent metapopulation model
+that depends on the environmental variable? How does the nature of the density
+dependence mediate this influence? Will density dependence reverse or reproduce
+the earlier result [@Ghosh_copula], using non-density-dependent models, that 
+left-tail dependence in environmental variables accentuates and right-tail dependence 
+mitigates extinction risks?
+For Q1, we use a simple 2-stage model. We hypothesize that left-tail dependence
+between juvenile survival and adult fecundity rates increases extinction risk
+relative to right-tail dependence, even when overall correlation is kept fixed. 
+For Q2, we consider spatial extensions of several simple and commonly
+used population models. Though tail associations are used
+to assess risks in fields such as finance and hydrology 
+[@alexander2008market; @li2013; @chen2011flood; @Goswami2018], to our
+knowledge asymmetric tail associations have rarely been assessed for their
+importance for extinction risks of populations and species (but see @Ghosh_copulas). 
+Our questions are an important step for understanding
+the potential influence of ECEs on extinction risks if ECEs are associated with
+tail dependence in environmental variables.
 
@@
 \includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
-\caption{Two random variables can have same Kendall 
-correlation ($\tau=$ `r ktau_common`), but different degree of associations
-in their extreme values: (A) extreme left tail, (B) moderate left tail, (C) moderate right tail, and
-(D) extreme right tail association.\label{fig_intro_cop}}
+\caption{Two stochastic quantities can have same Kendall 
+correlation, but different degrees of tail association: 
+(A) extreme left-tail association, (B) moderate left-tail association, (C) 
+moderate right-tail association, and (D) extreme right-tail association.
+In all cases Kendall's $\tau$ was the same ($\tau \approx$ `r ktau_common`),
+up to sample variation (the sample correlation for each dataset is displayed
+on its panel). The extreme left-tail (respectively, right-tail) dependent 
+case is perfectly correlated below (respectively, above) the threshold of 0.
+See Methods for information on how bivariate noise was generated with
+asymmetric tail associations.
+\label{fig_intro_cop}}
 \end{center}
 \end{figure*}
+<!--***DAN: Shyamolina, can you please put the actual Kendall tau that pertained
+for each dataset on each corresponding
+panel? Use, say, 3-4 digits accuracy for display. These values should be very similar, 
+but are expected to differ due to sampling variation. I would like to make that
+clear.-->
 

### 2019-10-24 MainText.Rmd
response added

@@
 <!--***DAN: Dan has not edited this--> 
-With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
-could pose severe impact on the resistance or resilience of a population and/ or community level through
-life-history traits of the species. Therefore predicting the impact of extreme events 
-on population extinction risk is necessery. We here introduce the potential use of "tail association" 
-property of two correlated ecological variables in modeling the extinction risk of 
-an age-structured single patch model. We also extend this idea 
-to density-dependent metapopulation models. For metapopulation models, 
-we show extinction risk not only depends on the nature of tail association of 
-spatially correlated noise but is also influenced by the inherent dynamics 
-of the population-model. To illaborate, we demonstrate left tail dependent noise across spatially correlated patches 
-cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
-But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail association instead of simple correlation in population viability analysis.
+With global climate change, extreme climatic events (ECEs) become more frequent.
+These ECEs could pose a severe impact on the resistance or resilience of a population and/ or community level through life-history traits of the species.
+Therefore predicting the impact of extreme events on population extinction risk is necessary.
+We here introduce the potential use of "tail association" property of two correlated ecological variables in modeling the extinction risk of an age-structured single patch model.
+We also extend this idea to density-dependent metapopulation models.
+For metapopulation models, we show extinction risk not only depends on the nature of the tail association of spatially correlated noise but is also influenced by the inherent dynamics of the population-model.
+To elaborate, we demonstrate left tail dependent noise across spatially correlated patches cause higher extinction risk than the right tail dependent noise when the model show undercompensatory dynamics.
+But in the overcompensatory regime, the trend is the opposite.
+Our study paves the path for future research considering this tail association instead of a simple correlation in population viability analysis.
 
@@
 sentences in English.
+
+***Shya: 
+Figure added, spell and grammar check done for the entire Maintext, but I think I should do it again just before submission.
 -->
@@
 can be associated with changes through time in the overall variance of 
-climate signals, and can directly contribute to changes in the symmetry or
+climate signals, and can directly contribute to changes in the symmetry or the 
 skewness of distributions of climate signals through time 
@@
 of $41$ UK butterfly species to four different types of ECEs. 
-At the community level, research has showed that ECEs can reorder
-the dominance heirarchy of species in an area, sometimes leading to 
-selective local extictions, and can also produce effects that cascade 
+At the community level, research has shown that ECEs can reorder
+the dominance hierarchy of species in an area, sometimes leading to 
+selective local extinctions, and can also produce effects that cascade 
 across trophic levels [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
 For example, @chiu2013impact showed that extreme floods affected 
-the annual survival of a prey community, in turn influencing predators in 
+the annual survival of a prey community, in turn, influencing predators in 
 aquatic as well as adjacent riparian habitats. 
@@
 <!--***DAN: Shyamolina, what, precisely, does it mean to "weaken the resistance of a
-plant community"? Resistance to what? Please clarify.-->
+plant community"? Resistance to what? Please clarify.
+
+***Shya: From @dreesen2014successive, authors concluded 
+"Our results show that when experimental temperate herbaceous assemblages were exposed to an extreme climatic
+event, their resistance to a subsequent event, in terms of leaf
+survival, is reduced only when the recovery period is relatively short (14 days for the assemblages studied here) and
+not when it is longer."-->
 
@@
 For instance, if two positively associated 
-variables are more associated in their smaller values than in their larger values,
+variables are more associated with their smaller values than in their larger values,
 they are said to have stronger
@@
 range of patterns of tail association, from extreme left- to extreme right-tail
-association, with the same overall correlation. So tail association
-is a distinct concept from correlation itself, and may have distinct ecological
+association, with the same overall correlation. So the tail association
+is a distinct concept from correlation itself and may have distinct ecological
 ramifications [@Ghosh_copula]. Asymmetric tail associations have been documented 
-between environmental variables, and between the same environmental variable 
+between environmental variables and between the same environmental variable 
 measured in different places [@Serinaldi2008; @li2013; @She2018; @Goswami2018]. 
@@
 variable are associated with local weather rather than ECEs, the result could
-be stronger right-tail association between measurements of the climatic variable
+be a stronger right-tail association between measurements of the climatic variable
 made in different locations, through time.
-For example, rain storms which deposit exceptionally large amounts of rain
-may do so over large areas; but months with less rain may experience
+For example, rainstorms which deposit exceptionally large amounts of rain
+may do so over large areas, but months with less rain may experience
 scattered showers only, which may be much less correlated across locales.
@@
 were the same, because it that case it would be very good years for the component 
-populations which would occur at the same time in many patches [@Ghosh_copulas].  
+populations that would occur at the same time in many patches [@Ghosh_copula]. 
 @Ghosh_copula substantiated this intuition by showing that it held true for a 
-modelling setup using a spatial version of a very simple non-density dependent
+modeling setup using a spatial version of a very simple non-density dependent
 population growth model (the Lewontin-Cohen model). In essence, if environmental
@@
 Extinction risk is also very commonly estimated for age-structured models 
-such as stochastic matrix models [@MorrisDoak2002 ; @Caswell2006] and 
+such as stochastic matrix models [@MorrisDoak2002; @Caswell2006] and 
 integral projection models
@@
 <!--objectives, last para of intro-->
-In this paper, we will explore the potential for asymmtric tail associations to 
+In this paper, we will explore the potential for asymmetric tail associations to 
 influence extinction risk in populations and metapopulations. Specifically,
@@
 of tail associations between measurements of an environmental variable in different
-locations influence extinction risk of a density dependendent metapopulation model
+locations influence the extinction risk of a density-dependent metapopulation model
 that depends on the environmental variable? How does the nature of the density
@@
 between juvenile survival and adult fecundity rates increases extinction risk
-relative to right-tail dependence, even when overall correlation is kept fixed. 
+relative to right-tail dependence, even when the overall correlation is kept fixed. 
 For Q2, we consider spatial extensions of several simple and commonly
@@
 knowledge asymmetric tail associations have rarely been assessed for their
-importance for extinction risks of populations and species (but see @Ghosh_copulas). 
+importance for extinction risks of populations and species (but see @Ghosh_copula). 
 Our questions are an important step for understanding
@@
 moderate right-tail association, and (D) extreme right-tail association.
-In all cases Kendall's $\tau$ was the same ($\tau \approx$ `r ktau_common`),
+In all cases, Kendall's $\tau$ was the same ($\tau \approx$ `r ktau_common`),
 up to sample variation (the sample correlation for each dataset is displayed
@@
 but are expected to differ due to sampling variation. I would like to make that
-clear.-->
+clear.
+***Shya: correlation added (rounded upto 4 digit) in each plot.-->
 
@@
 
-<!--stochastic matrix model : age structured-->
-To investigate (Q1) from \nameref{Introduction} that how do vital rates with asymmetric tail 
+<!--stochastic matrix model : age-structured-->
+To investigate (Q1) from \nameref{Introduction} that how do vital rates with an asymmetric tail 
 association affect the extinction risk of
@@
 
-We assumed the population was semelperous, i.e., adult ($A$) went through the
+We assumed the population was semelparous, i.e., adult ($A$) went through the
 reproductive episode only once before death. Eggs and adults in the 
@@
 
-Moderate tail associatation was modeled by using the 
+The moderate tail association was modeled by using the 
 concept of "copula", a statistical tool to model 
 dependencies. Appendix \ref{SM-intro_copula} summarized
-a brief introduction of "copulas" and their relation to tail asscociations, symmetries 
+a brief introduction of "copulas" and their relation to tail associations, symmetries 
 with examples of available copula families 
@@
 copula to model moderate left-tail and Survival Clayton copula to model 
-moderate right-tail association in the vital rates. Both copula had the same Kendall 
+moderate right-tail association in the vital rates. Both copulae had the same Kendall 
 correlation ($\tau =$ `r ktau_common`) but opposite association in their respective tails. 
@@
 `iTau` function from `copula` package so that both copula had
-same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each clumn of
-matrix $M$, then became independent of each others but 
+same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each column of
+matrix $M$, then became independent of each other but 
 entries between columns had a specific tail association
@@
 <!--density-dependence metapopulation model-->
-We explored (Q2) by examining how nature of the tail association of an
+We explored (Q2) by examining how the nature of the tail association of an
 environmental variable across space could influence the extinction risk of a 
@@
 Models were adapted from earlier work
-by @cohen1995unexpected, excluding only two : Varley model (this is not a continuous one) 
-and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
+by @cohen1995unexpected, excluding only two: Varley model (this is not a continuous one) 
+and the Malthus-Condorcet-Mill model (terms containing logarithm of the extinct population does not make sense). 
 All models assumed a generalized form: $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
@@
 we considered $n$ patches in each model were arranged, evenly spaced, in a line
-with the provision of "local" or "global" dispersal among patches. A n $\times$ n dispersal
+with the provision of "local" or "global" dispersal among patches. An n $\times$ n dispersal
 matrix, $D$ assumed that a fraction $d$ of 
@@
 for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
-We assumed population went extinct after each time step if it was below a 
-certain threshold (set to 10\% of initial population $P_e$ for individual model).
+We assumed the population went extinct after each time step if it was below a 
+certain threshold (set to 10\% of the initial population $P_e$ for individual model).
 
@@
 If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
-model equilibrium $P_e$ became unstable. Beyond Hopf-bifurcation point, model showed sustained
+model equilibrium $P_e$ became unstable. Beyond the Hopf-bifurcation point, the model showed sustained
 oscillation and eventually led to chaos with increasing $r$. Appendix \ref{SM-model_stability}
@@
 
-We plotted extinction risks along y-axis
+We plotted extinction risks along the y-axis
 for both left tail and right tail associated noises 
@@
 
-<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula-->
+<!--Q1 : age-structured results : linear and nonlinear model with asymmetric copula-->
 
@@
 and thereby can increase the extinction risk of the population. For non-linear age-structured 
-models, we got the similar trend for extinction risks with different tail association (left and right).
-Non-linear model showed higher extinction risk than its corresponding 
+models, we got a similar trend for extinction risks with the different tail associations (left and right).
+The non-linear model showed higher extinction risk than its corresponding 
 linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
@@
 vital rates coming from extreme-tail dependent (left and right in two different simulations)
-copulas and got the similar results but with much prominent difference between their extinction risks
+copulas and got similar results but with much prominent difference between their extinction risks
 (Fig. \ref{fig_extrisk_agestr}C or D). This result answered (Q1) and showed extinction risk of 
-a population could be considered as direct manifestation of tail association property 
+a population could be considered as a direct manifestation of tail association property 
 between vital rates of different age-class
@@
 environment showed similar patterns for extinction risk as density-independent 
-metapopulation model did in recent study [@Ghosh_copula]. For each density-dependent 
+metapopulation model did in a recent study [@Ghosh_copula]. For each density-dependent 
 model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
 the metapopulations when spatially correlated noises were left-tail associated 
-than the one with right-tail association. But that observation remained 
+than the one with the right-tail association. But that observation remained 
 valid when the model parameter - here considered as growth rate $(r)$ - was 
@@
 represent the same but for global dispersal among patches. 
-Both case showed the signature of "rescue-effect" (i.e. reduced extinction risk for 
+Both case showed the signature of "rescue-effect" (i.e. reduced the extinction risk for 
 finite dispersal compared with $d=0$).
@@
 indicating the value of $r$ for each model where their deterministic analog switches 
-from under- to over-compensatory dynamics. Though the switching visually appered far away
+from under- to over-compensatory dynamics. Though the switching visually appeared far away
 from the theoretical $r_c$ value for Verhulst and Austin-Brewer model, they were 
-consistent with our hypothesis as they had comperatively 
-smaller range of $r$ than the other models.
+consistent with our hypothesis as they had comparatively a smaller range of $r$ than the other models.
 
@@
 <!--summary results:-->
-We showed the potential use of "asymmetric" tail association in modeling the impact of 
+We showed the potential use of the "asymmetric" tail association in modeling the impact of 
 extreme ecological events on population-extinction risk. Our results demonstrated left and right-tail 
 association in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
-in years of extreme events ("catastrophes" and "bonanzas"). Though we could expect left tail association 
+in years of extreme events ("catastrophes" and "bonanzas"). Though we could expect the left tail association 
 to show higher extinction risk than the one for right tail association with age-structured 
-and density dependent metapopulation models, it was not always true. In the over-compensatory regime,
+and density-dependent metapopulation models, it was not always true. In the over-compensatory regime,
 right tail dependent noise caused higher extinction risk than the left tail dependent noise. We showed 
-that in absence of dispersal the above-said switching took place near the critical parameter value where 
-the deterministic model showed transition from under-compensatory to over-compensatory regime.
+that in the absence of dispersal the above-said switching took place near the critical parameter value where 
+the deterministic model showed the transition from under-compensatory to the over-compensatory regime.
 
 <!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
-We found left tail association in vital rates (induced by ECEs) caused higher extinction risk than 
+We found a left tail association in vital rates (induced by ECEs) caused higher extinction risk than 
 the right tail dependent ones, but the difference was not remarkable 
@@
 The approximation considered growth rate being 
-dependent on the most dominant eigen value of the mean projection matrix 
+dependent on the most dominant eigenvalue of the mean projection matrix 
 and the temporal variance of the log(population growth rate) caused 
@@
 two matrix elements, not on their tail association properties. Therefore,
-following @tuljapurkar1990lecture, population growth 
+following @tuljapurkar1990lecture, the population growth 
 rate in an iid environment could be approximated
@@
 -->
-Although we demonstrated that tail association between vital rates of an age structured model can
+Although we demonstrated that tail association between vital rates of an age-structured model can
 influence extinction risk, substantial work remains to determine the importance of this effect in the 
-following scenarios. First, we demonstrated the effect of tail association on extinction risk 
-with a basic age structured (two age stages) model where we considered the total semelperous 
+following scenarios. First, we demonstrated the effect of the tail association on extinction risk 
+with a basic age-structured (two age stages) model where we considered the total semelparous 
 population would go extinct if there would be only one female adult and one egg. But in reality population 
 viability analysis (PVA) deals with a little bit higher threshold with several age, stage or size class and a variety 
-of populations including semelperous and iteroperous [@morris2002quantitative]. Future studies should include 
+of populations including semelparous and iteroparous [@morris2002quantitative]. Future studies should include 
 tail association effect on extinction risk 
-for a multiple age-, size- and stage- structured model and estimate extinction risk 
+for a multiple age-, size- and stage-structured model and estimate extinction risk 
 for each age, size and stage class separately 
 so that we can identify (via multisite-PVA) which life-stage or demographic process 
-to target for population management. Earlier study by @doak1994modeling showed PVA models with correlation 
-between different demographic rates indicated marked difference in predicting confidence interval for 
+to target for population management. An earlier study by @doak1994modeling showed PVA models with correlated 
+demographic rates indicated a marked difference in predicting confidence interval for 
 future population size compared with the model which did not include correlation. But when 
@@
 dependencies with copula could help in this case. Second, we did not consider multiple sites of 
-age-structured metapopulation which could be linked via dispersal. So, possible future investigation 
+age-structured metapopulation which could be linked via dispersal. So, the possible future investigation 
 would be to explore the viability for the metapopulations if vital rates (e.g. fecundity or 
 survival of different stage class) across sites showed specific tail association 
-with spatially correlated extreme climatic events. Third, we considered tail association 
+with spatially correlated extreme climatic events. Third, we considered the tail association 
 between positively correlated vital rates, but what would be the extinction risk 
-and viability of a population if demographic rates were negatively correlated with asymmetric tail association. 
+and viability of a population if demographic rates were negatively correlated with the asymmetric tail association. 
 We could consider this scenario to model extreme events that kill adults and allow more successful
@@
 interspecific-dependence. An earlier study [@carnicer2011widespread] showed extreme events
-(i.e., heat waves and droughts) increased mortality and defoliation in southern European forest and its 
+(i.e., heatwaves and droughts) increased mortality and defoliation in the southern European forest and its 
 dependent herbivorous insects and fungi through trophic linkage. <!--ECEs can even truncate 
-multiyear insect outbreak dynamics [@esper20061200].--> We believe copula (may be a multidimensional one) has 
-a lot more potential to estimate risk of cascading extinction in model food webs. <!--should I mention about 
+multiyear insect outbreak dynamics [@esper20061200].--> We believe copula (maybe a multidimensional one) has 
+a lot more potential to estimate the risk of cascading extinction in model food webs. <!--should I mention about 
 match-mismatch hypothesis leads to extinction? our phenology paper in prep.--> Fifth, we considered 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}
-\caption{Extinction risk against time from single patch age structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(B, D)$ models for $10000$ simulations; shading represents 95\% confidence interval.
-$(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have moderate left tail (LT) or moderate right tail (RT) association, $(C, D)$ show the analogous scenario but with extreme tail association between vital rates.
+\caption{Extinction risk against time from single patch age-structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(B, D)$ models for $10000$ simulations; shading represents 95\% confidence interval.
+$(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have a moderate left tail (LT) or a moderate right tail (RT) association, $(C, D)$ show the analogous scenario but with the extreme tail association between vital rates.
 \label{fig_extrisk_agestr}}

### 2019-10-24 SuppMat.Rmd
response added

--- 2019-10-15 SuppMat.Rmd
@@
 
+saveRDS(ktau_common,"./Results/ktau_common.RDS")
+
 # flip xtc_r to get xtc_l
@@
 mtc_r<-1-mtc_l
+
+# calculate tau with sampling variation
+kcor_el<-cor(xtc_l,xtc_l,method = "kendall")
+kcor_el<-kcor_el[1,2]
+
+kcor_er<-cor(xtc_r,xtc_r,method = "kendall")
+kcor_er<-kcor_er[1,2]
+
+kcor_ml<-cor(mtc_l,mtc_l,method = "kendall")
+kcor_ml<-kcor_ml[1,2]
+
+kcor_mr<-cor(mtc_r,mtc_r,method = "kendall")
+kcor_mr<-kcor_mr[1,2]
 ```
@@
 # for plotting purpose I just showed only first 500 points
+
 pdf("./Results/introfig_copula.pdf", height=3, width=12)
@@
 plot(xtc_l[1:500,1],xtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
-legend(x="topleft",legend="A",bty="n",cex=2.5)
+#legend(x="topleft",legend="A",bty="n",cex=2.5)
+text(x=0.1,y=0.9,"A",cex=2.5)
+text(x=0.8,y=0.1,round(kcor_el,4),cex=2)
 
@@
 plot(mtc_l[1:500,1],mtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
-legend(x="topleft",legend="B",bty="n",cex=2.5)
+text(x=0.1,y=0.9,"B",cex=2.5)
+text(x=0.8,y=0.1,round(kcor_ml,4),cex=2)
 
@@
 plot(mtc_r[1:500,1],mtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
-legend(x="topleft",legend="C",bty="n",cex=2.5)
+text(x=0.1,y=0.9,"C",cex=2.5)
+text(x=0.8,y=0.1,round(kcor_mr,4),cex=2)
 
@@
 plot(xtc_r[1:500,1],xtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
-legend(x="topleft",legend="D",bty="n",cex=2.5)
-
-par(op)
-dev.off()
-
-saveRDS(ktau_common,"./Results/ktau_common.RDS")
+text(x=0.1,y=0.9,"D",cex=2.5)
+text(x=0.8,y=0.1,round(kcor_er,4),cex=2)
+
+par(op)
+dev.off()
 ```

### 2019-10-25 MainText.Rmd
deletions of a couple comments

--- 2019-10-23 MainText.Rmd
@@
 \newpage
-
-<!--***DAN: 
-3) Please
-add the figure you have mentioned in the text but that does not exist. Please run a spell
-check. Please go and check every sentence throughout the manuscript for grammatical 
-errors. I think you will find some 
-and will be able to correct them if you focus on that - there are many, and by correcting them
-you will not only make the paper better you will get better at writing grammatically correct 
-sentences in English. 
--->
 

### 2019-10-25 MainText.Rmd

--- 2019-10-24 MainText.Rmd
@@
 \newpage
-
-<!--***DAN: 
-3) Please
-add the figure you have mentioned in the text but that does not exist. Please run a spell
-check. Please go and check every sentence throughout the manuscript for grammatical 
-errors. I think you will find some 
-and will be able to correct them if you focus on that - there are many, and by correcting them
-you will not only make the paper better you will get better at writing grammatically correct 
-sentences in English.
-
-***Shya: 
-Figure added, spell and grammar check done for the entire Maintext, but I think I should do it again just before submission.
--->
 

### 2019-10-25 MainText.Rmd
working on methods, also some other points made

@@
 ---
-title: "Asymmetric tail association in extreme climatic events affects population-extinction risk"
+title: "Tail associations and their impact on extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 2) The influence of correlations in the extremes on extinction risk
-DAN FAVORITE SO FAR 3) Tail associations in environmental variables and their impact on extinction risk
+3) Tail associations in environmental variables and their impact on extinction risk
+4) Asymmetric tail association in extreme climatic events affects population-extinction risk
 -->
@@
 For example, @chiu2013impact showed that extreme floods affected 
-the annual survival of a prey community, in turn, influencing predators in 
+the annual survival of a prey community, in turn influencing predators in 
 aquatic as well as adjacent riparian habitats. 
 @dreesen2014successive, in an experimental study, showed that 
-successive ECEs with short time intervals between them
-can weaken the resistance of plant communities.
-<!--***DAN: Shyamolina, what, precisely, does it mean to "weaken the resistance of a
-plant community"? Resistance to what? Please clarify.
-
-***Shya: From @dreesen2014successive, authors concluded 
-"Our results show that when experimental temperate herbaceous assemblages were exposed to an extreme climatic
-event, their resistance to a subsequent event, in terms of leaf
-survival, is reduced only when the recovery period is relatively short (14 days for the assemblages studied here) and
-not when it is longer."-->
+ECEs can reduce the ability of plant communities to withstand subsequent ECEs
+that occur after a short time interval.
 
@@
 For instance, if two positively associated 
-variables are more associated with their smaller values than in their larger values,
+variables are more closely associated in their smaller values than in their larger values,
 they are said to have stronger
@@
 range of patterns of tail association, from extreme left- to extreme right-tail
-association, with the same overall correlation. So the tail association
-is a distinct concept from correlation itself and may have distinct ecological
+association, with the same overall correlation. So tail association
+is a distinct concept from correlation itself, and may have distinct ecological
 ramifications [@Ghosh_copula]. Asymmetric tail associations have been documented 
-between environmental variables and between the same environmental variable 
+between environmental variables, and between the same environmental variable 
 measured in different places [@Serinaldi2008; @li2013; @She2018; @Goswami2018]. 
@@
 variable are associated with local weather rather than ECEs, the result could
-be a stronger right-tail association between measurements of the climatic variable
+be stronger right-tail association between measurements of the climatic variable
 made in different locations, through time.
 For example, rainstorms which deposit exceptionally large amounts of rain
-may do so over large areas, but months with less rain may experience
+may do so over large areas; but months with less rain may experience
 scattered showers only, which may be much less correlated across locales.
@@
 would probably not have the same effect, even if overall correlations between locations
-were the same, because it that case it would be very good years for the component 
+were the same, because in that case it would be very good years for the component 
 populations that would occur at the same time in many patches [@Ghosh_copula]. 
@@
 mitigates extinction risks?
-For Q1, we use a simple 2-stage model. We hypothesize that left-tail dependence
+For Q1, we simulate a simple 2-stage model. We hypothesize that left-tail dependence
 between juvenile survival and adult fecundity rates increases extinction risk
 relative to right-tail dependence, even when the overall correlation is kept fixed. 
-For Q2, we consider spatial extensions of several simple and commonly
-used population models. Though tail associations are used
+For Q2, we simulate spatial extensions of several simple and commonly
+used population models. We intentionally use a simulation approach and simple models
+instead of attampting mathematical analysis and/or using realistic models
+because our intention here is to provide an initial exploration only of 
+the potential mechanisms outlined above which suggest that
+asymmetries of tail association may make a difference for extinction risks. 
+<!--***DAN: In Discussion need to talk about what analytic results are needed
+to take the next step of determining when tail depencence may make enough of
+a difference to think about. I think our results, together with Tulja's 
+approximation, suggest that tail associations probably don't make much difference
+for models without a spatial component. But our results illsutaret that it can
+make a big difference when there is a spatial component, apparently because of 
+the influence of tail associations on rescue effects. Could maybe even outline
+analytic approaches for non-dd models? Not sure what would be done with dd
+models. The point is you need to take the next step in the Discussion of 
+talking about how our results, which admitedly are just for specific models,
+could be turned into analytic results or general results.-->
+Though tail associations are used
 to assess risks in fields such as finance and hydrology 
@@
 moderate right-tail association, and (D) extreme right-tail association.
-In all cases, Kendall's $\tau$ was the same ($\tau \approx$ `r ktau_common`),
+In all cases, Kendall's $\tau$ was the same, $\tau = 0.75$,
 up to sample variation (the sample correlation for each dataset is displayed
@@
 \end{figure*}
-<!--***DAN: Shyamolina, can you please put the actual Kendall tau that pertained
-for each dataset on each corresponding
-panel? Use, say, 3-4 digits accuracy for display. These values should be very similar, 
-but are expected to differ due to sampling variation. I would like to make that
-clear.
-***Shya: correlation added (rounded upto 4 digit) in each plot.-->
+<!--***DAN: Shyamolina, it occured to be there are simple and good analytic reasons
+why ktau_common is exactly 0.75 (up to sample variation). 
+I bet you will see them too if you look up the
+definition of kendall tau and think about the extreme tail dependent cases. 
+Currently, around line 50 of the sup mat Rmd, you calculate ktau_common
+by simulating 100000 points from the extreme-tail-denepdnece noise and then
+calculating the sample kendall. Instead, can you please replace those lines 
+with the simple line "ktau_common<-0.75"? I would do this myself but I
+don't want to risk messing up your code in some way I don't understand. You should not
+need the seed anymore to ensure you get the same ktau_common on each run
+since you won't be using the sample correlartion anymore to estimate ktau_common,
+you'll just be using the analytic result that the population correlation
+of data from the extreme-tail-dependent algorithm (in the limit of infinite
+sample sizes) is exactly 0.75. Please
+let me know if you do not understand something here.-->
+<!--***DAN: Shyamolina, currently the sample kendalls are the same on A and D, and
+also on B and C, and I know why. This looks weird, and makes the reader
+wonder why this is the case. Rather than plotting the reversed version of the
+data from A on D, and the reversed version of the data from B on C, can you plase
+generate separate, independent data for A and D, and also for B and C? That
+will have the effect of making the sample kendalls on the 4 panels slightly
+different, which will mean readers think anything funny is going on.-->
 
@@
 <!--stochastic matrix model : age-structured-->
-To investigate (Q1) from \nameref{Introduction} that how do vital rates with an asymmetric tail 
-association affect the extinction risk of
-populations in a single patch age-structured model,
-we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
-with two age classes (egg $E$ and female adult $A$).
+To investigate (Q1) from the \nameref{Introduction},
+we considered a simple stochastic age-structured model
+with two age classes,  
+\begin{equation}
+\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
+\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
+\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
+\end{equation}
+\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
+and adults, respectively, at time $t$. We chose a very simple model
+because our goal was to carry out an initial evaluation, only, of 
+the realism of a potential mechanism. The stochastic vital rates $s_e(t)$ and $f_a(t)$
+were selected, independently across times, as follows. 
+The egg survival rate $s_e(t)$ was drawn from a beta distribution with parameters
+$\alpha=3$ and $\beta=2$. 
+Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
+was drawn from a gamma distribution with shape and scale parameters $k=2$
+and $\theta=1$, respectively, 
+and $\Omega$ is a density dependence parameter. Taking $\Omega=\infty$ made
+$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
+(it was then a standard stochastic matrix model). Finite, positive values of 
+$\Omega$ made the model nonlinear, with density dependence in the adult
+fecundity rate.
 
-\begin{equation}
-\left({\begin{array}{cc} E \\ A \end{array}}\right)_{t+1} = 
-\left(\begin{array}{cc} 0 & f_a\\ s_e & 0 \end{array}\right)
-\left(\begin{array}{c} E \\ A \end{array}\right)_{t} \label{eq_age_str}
-\end{equation}
-
-\begin{equation}
-f_a \sim \Gamma (k, \theta)\exp(-A/\Omega), \hspace{0.2 cm} s_e \sim \mathcal{B}(\alpha,\beta) \label{eq_vitalrates}
-\end{equation}
-
-We assumed the population was semelparous, i.e., adult ($A$) went through the
-reproductive episode only once before death. Eggs and adults in the 
-next generation evolved depending on the fecundity of adult ($f_a \geq 0$) 
-and survival of eggs ($0 \leq s_e \leq 1$) from present generation.
-Fluctuating vital rates ($f_a$ and $s_e$) captured the environmental
-stochasticity (Eq. \ref{eq_vitalrates}). We made adult's fecundity 
-($f_a$) negative density-dependent with its own population. When $\Omega = \infty$ in 
-Eq. \ref{eq_vitalrates}, the projection matrix was independent of populations 
-(i.e., Eq. \ref{eq_age_str} was "linear" model then), 
-but when $\Omega$ was finite, Eq. \ref{eq_age_str} became "nonlinear" with density dependence.
-Fecundity ($f_a$) and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
-a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. We generated those distributions 
-from iid noises with a varied degree of asymmetric tail association between 
-fecundity ($f_a$) and survival ($s_e$) rates. Degree of tail association had two broad 
-categories: moderate and extreme. Each category had specific direction in tail association: 
-either in left or in right tail of vital rates (as depicted in Fig. \ref{fig_intro_cop}). Vital 
-rates ($f_a$ and $s_e$) were left tail dependent when unusual bad years ("catastrophes") 
-suppressed both vital rates synchronously, 
-and right tail dependent when unexpected good years ("bonanzas") boosted both of them.
-
-The moderate tail association was modeled by using the 
-concept of "copula", a statistical tool to model 
-dependencies. Appendix \ref{SM-intro_copula} summarized
-a brief introduction of "copulas" and their relation to tail associations, symmetries 
-with examples of available copula families 
-from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
-copula to model moderate left-tail and Survival Clayton copula to model 
-moderate right-tail association in the vital rates. Both copulae had the same Kendall 
-correlation ($\tau =$ `r ktau_common`) but opposite association in their respective tails. 
-We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
-right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
-`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
-by number of simulations ($=10000$) we used. 
-We chose parameters for each copula family using 
-`iTau` function from `copula` package so that both copula had
-same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each column of
-matrix $M$, then became independent of each other but 
-entries between columns had a specific tail association
-with positive Kendall's correlation.
-
-On a similar note, we generated extreme left- (or right-) tail dependent matrix (M)
-from standard iid (Appendix \ref{SM-extr_tail_dep}) and made the marginals uniform so that
-it could mimic an extreme tail dependent copula. We made the first column of matrix 
-$M$ as gamma ($\Gamma(k,\theta)$) distributed 
-with shape parameter $k = 2$ and scale parameter $\theta = 1$.
-This distribution represented fecundity of the adult population, $f_a \geq 0$.
-For a given $\theta$, increasing $k$ would shift the distribution towards
-higher fecundity and reduce the extinction risk if the other parameters
-remain unchanged. We took the second column of $M$ and applied a 
-beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
-$\beta = 2$ were the shape parameters that could control the distribution 
-of the survival rates of eggs ($s_e$) ranging in between 0 to 1. 
-For each $25$ timesteps from each of $10000$ simulations,
-we incorporated pairwise entries from each row of matrix $M$ in the expression for 
-vital rates (fecundity $f_a$ and survival rates $s_e$, Eq. \ref{eq_vitalrates}) of 
-the stochastic projrction matrix in model (\ref{eq_age_str}).
+The stochastic quantities $s_e(t)$ and $\tilde{f}_a(t)$ were generated 
+to have the same overall association (Kendall's $\tau$ equal to $0.75$,
+up to sampling variation), but different degrees and types of 
+asymmetry of tail association in different simulations, as follows.
+For a simulation of length $T$, we used methods that can 
+generate $T$ independent pairs $(u(t),v(t))$ for $t=1,\ldots,T$ with
+$\tau=0.75$ (up to sample variation), but with either extreme left-,
+moderate left-, moderate right-, or extreme right-tail association 
+(e.g., Fig. \ref{fig_intro_cop}). The methods we used generated
+$(u(t),v(t))$ such that $u(t)$ was uniformly distributed on $(0,1)$,
+and likewise for $v(t)$. We then applied a monotonic transformation
+to the $u(t)$ so that the resulting numbers were distributed according to
+a beta distribution with parameters
+$\alpha=3$ and $\beta=2$ 
+and used those values for the $s_e(t)$. We also applied a (different)
+monotonic transformation to the $v(t)$ so that the resulting numbers were
+distributed according to a gamma distribution with parameters $k=2$
+and $\theta=1$ and used those values for the $\tilde{f}_a(t)$. Because
+the $u(t)$ and $v(t)$ were uniformly distributed, the inverse 
+cummulative distribution functions of the desired beta and
+gamma distributions were the transformations. Because these
+were monotonic transformations, the Kendall correlation of 
+$s_e(t)$ and $\tilde{f}_a(t)$ was the same as that of 
+$u(t)$ and $v(t)$. The quantities $(u(t),v(t))$ for $t=1,\ldots,T$ were
+generated as described in Appendix \ref{SM-makenoise}.
 
 <!--calculation of ext risk with age-structured model-->
-To calculate extinction risk after each time step, we simulated model (\ref{eq_age_str}) for $25$ years
+To calculate extinction risk for a given $\Omega$ and choice of tail dependence, 
+we simulated the model for $25$ time steps, $10000$ times,
+each simulation starting from $E(t)=5$ and $A(t)=10$. The population for a simulation
+was considered to have gone extinct if the total population $E(t)+A(t)$ ever 
+went below an extinction threshold (we used a threshold of $2$) within the 
+$25$-time-step time horizon. Extinction risk was the fraction of $10000$
+simulations that went extinct.
+
+<!--***DAN: Shyamolina, the below text makes me wonder if I am misunderstanding something, please
+explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
+of that and what does it have to do with extinction riks?
+
+after each time step, we simulated model (\ref{eq_age_str}) for $25$ years
 with $10000$ simulations. For each simulation, we started with initial population $(E,A)_{t=0} = (5,10)$
@@
 of $10000$ simulations with the probability of computed extinction risk at each time step (see [`code`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R)).
+-->
 

### 2019-10-25 SuppMat.Rmd
working on methods, also some other points made

@@
 ---
-title: "Supporting Information : Asymmetric tail association in extreme climatic events affects population-extinction risk"
+title: "Supporting information: Tail associations and their impact on extinction risk"
 author: "Shyamolina Ghosh, Daniel Reuman"
@@
 
-# A brief introduction about "copulas"  \label{intro_copula}
+# Generating noise with asymmetric tail associations  \label{makenoise}
 
 <!--needs editing, I just dumped old text here-->
+
 Copula, a rank based measure of association, is much more informative 
@@
 
-# Algorithm to generate extreme tail association  \label{extr_tail_dep}
+<!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
 
@@
 distribution with a specified mean and standard deviation.
+
+<!--some stuff that used to be in the main text-->
+
+The moderate tail association was modeled by using the 
+concept of "copula", a statistical tool to model 
+dependencies. Appendix \ref{SM-intro_copula} summarized
+a brief introduction of "copulas" and their relation to tail associations, symmetries 
+with examples of available copula families 
+from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
+copula to model moderate left-tail and Survival Clayton copula to model 
+moderate right-tail association in the vital rates. Both copulae had the same Kendall 
+correlation ($\tau =$ `r ktau_common`) but opposite association in their respective tails. 
+We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
+right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
+`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
+by number of simulations ($=10000$) we used. 
+We chose parameters for each copula family using 
+`iTau` function from `copula` package so that both copula had
+same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each column of
+matrix $M$, then became independent of each other but 
+entries between columns had a specific tail association
+with positive Kendall's correlation.
+
+<!--some more stuff from the main text-->
+
+On a similar note, we generated extreme left- (or right-) tail dependent matrix (M)
+from standard iid (Appendix \ref{SM-extr_tail_dep}) and made the marginals uniform so that
+it could mimic an extreme tail dependent copula. We made the first column of matrix 
+$M$ as gamma ($\Gamma(k,\theta)$) distributed 
+with shape parameter $k = 2$ and scale parameter $\theta = 1$.
+This distribution represented fecundity of the adult population, $f_a \geq 0$.
+For a given $\theta$, increasing $k$ would shift the distribution towards
+higher fecundity and reduce the extinction risk if the other parameters
+remain unchanged. We took the second column of $M$ and applied a 
+beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
+$\beta = 2$ were the shape parameters that could control the distribution 
+of the survival rates of eggs ($s_e$) ranging in between 0 to 1. 
+For each $25$ timesteps from each of $10000$ simulations,
+we incorporated pairwise entries from each row of matrix $M$ in the expression for 
+vital rates (fecundity $f_a$ and survival rates $s_e$, Eq. \ref{eq_vitalrates}) of 
+the stochastic projrction matrix in model (\ref{eq_age_str}).
 

### 2019-10-26 MainText.Rmd
elaboration of a comment
moer progress on methods

@@
 <!--stochastic matrix model : age-structured-->
-To investigate (Q1) from the \nameref{Introduction},
+To investigate Q1 from the \nameref{Introduction},
 we considered a simple stochastic age-structured model
@@
 explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
-of that and what does it have to do with extinction riks?
+of that and what does it have to do with extinction riks? I would have thought extinction risk
+was just the fraction of 10000 sims that went extinct.
 
@@
 <!--density-dependence metapopulation model-->
-We explored (Q2) by examining how the nature of the tail association of an
-environmental variable across space could influence the extinction risk of a 
-metapopulation model with density-dependent feedback. We considered six density-dependent 
-population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
-in noisy environment (see Table \ref{tab_model_summary}) 
-from a metapopulation context.
-Models were adapted from earlier work
-by @cohen1995unexpected, excluding only two: Varley model (this is not a continuous one) 
-and the Malthus-Condorcet-Mill model (terms containing logarithm of the extinct population does not make sense). 
-All models assumed a generalized form: $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
-where $\vec{P_t}$ was a vector of population densities, 
-the $i^{th}$ component $P_t^i$ representing
-the density in the $i^{th}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
-The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
-of stochastic density dependence form $\lambda_i(t)$ (see second column 
-of Table \ref{tab_model_summary}). For simplicity, 
-we considered $n$ patches in each model were arranged, evenly spaced, in a line
-with the provision of "local" or "global" dispersal among patches. An n $\times$ n dispersal
-matrix, $D$ assumed that a fraction $d$ of 
-each population patch dispersed during any given time step, equally distributed to
-the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
-$d$ dispersed equally to other $(n-1)$ patches. We called $d$ as dispersal
-rate with $d$ being zero there was no dispersal, and hence no chance to rescue an extinct 
-patch through migration from nearby metapopulation patches.
+To investigate Q2 from the \nameref{Introduction}, we considered spatialized, stochastic versions
+of six density dependent population models, the Ricker, Hassell, Maynard-Smith, Pennycuick, Verhulst, 
+and Austin-Brewer models as elaborated by @cohen1995unexpected (Table \ref{tab_model_summary}).
+All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
+where $\vec{P}(t)$ is a vector of population densities, 
+the $i^{\text{th}}$ component $P_i(t)$ representing
+the population density in the $i^{\text{th}}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
+The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{\text{th}}$ diagonal entry 
+$\lambda_i(t)$ depnding on the model used (see Table \ref{tab_model_summary}). The matrix $D$
+was an $n \times n$ dispersal matrix. For simplicity, 
+we assumed the $n$ habitat patches were arranged, evenly spaced, in a line.
+We considered both "local" and "global" dispersal among patches in different simulations. 
+For local dispersal, a fraction $d$ of 
+each population dispersed during each time step, equally distributed to
+the two or one nearest neighbors. For global dispersal the same fraction $d$ dispersed
+equally to the other $(n-1)$ patches. When $d$ was $0$, dispersal did not occur and 
+rescue effects were not possible.
+Two models that @cohen1995unexpected used were excluded from our analyses because they were not
+readily adapted to our simulation setup: the Varley model,
+which is not continuous;
+<!--***DAN: Shyamolina, I cannot remember this one and cannot see immediately why non-continuity 
+is a reason to exclude these models? I remember we had a good reason to exclude it, I am just 
+wanting a bit more explanation here, please. Can you add?-->
+and the Malthus-Condorcet-Mill model, which employs a logarithm of the population and therefore
+cannot be applied in the context of extinction.
+<!--***DAN: go back and look at that model when you get the paper and remind yourself-->
+
+<!--***DAN: got to here 2019 10 25-->
 

### 2019-10-27 MainText.Rmd
comment to Shyamolina on the Varley model on line 391

@@
 To investigate Q2 from the \nameref{Introduction}, we considered spatialized, stochastic versions
-of six density dependent population models, the Ricker, Hassell, Maynard-Smith, Pennycuick, Verhulst, 
-and Austin-Brewer models as elaborated by @cohen1995unexpected (Table \ref{tab_model_summary}).
+of six density dependent population models, the Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, 
+and Austin-Brewer models (@cohen1995unexpected; Table \ref{tab_model_summary}).
 All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
@@
 which is not continuous;
-<!--***DAN: Shyamolina, I cannot remember this one and cannot see immediately why non-continuity 
-is a reason to exclude these models? I remember we had a good reason to exclude it, I am just 
-wanting a bit more explanation here, please. Can you add?-->
+<!--***DAN: Shyamolina, looking back at the Cohen paper, using the value C=1 as they used, the
+Varley model *is* continuous. It is not differentiable, but it's not clear to me that 
+non-differentiability means we cannot use the model. Let's discuss.-->
 and the Malthus-Condorcet-Mill model, which employs a logarithm of the population and therefore

### 2019-10-27 MainText.Rmd
need to revisit why we used 6 of cohen's models and not the other two

@@
 Varley model *is* continuous. It is not differentiable, but it's not clear to me that 
-non-differentiability means we cannot use the model. Let's discuss.-->
-and the Malthus-Condorcet-Mill model, which employs a logarithm of the population and therefore
-cannot be applied in the context of extinction.
-<!--***DAN: go back and look at that model when you get the paper and remind yourself-->
+non-differentiability means we cannot use the model. Perhaps it does mean that, but I don't see 
+it. Let's discuss. I'm not saying we have to go back and use it, but we may have to justify
+differently why we used 6 of Cohen's models and not the other two.-->
+and the Malthus-Condorcet-Mill model, which employs a logarithm of the population and is therefore
+inconvenient to analyze in the context of extinction.
+<!--***DAN: Unless I am mistaken, the limit of the right hand side of the M-C-M model as
+described by Cohen, as P approaches 0 from the right, is 0. That means the model has
+a removable singularity at 0, i.e., we could just define the model so P_t+1=0 when P_t
+is 0, and we would get a perfectly valid continuous model. This means that using logs is 
+not really a valid reason to not use the models. Let's discuss. I'm not saying we have to go 
+back and use it, but we may have to justify differently why we used 6 of Cohen's models and not 
+the other two.-->
 

### 2019-10-28 MainText.Rmd
working on para intro-ing partial spearman

@@
 said to have stronger "right-tail" or "upper-tail" association 
-(Figure \ref{fig_intro_cop}C, D). 
+(Figure \ref{fig_intro_cop}C, D). A precise definition of a measure of tail 
+association is given in Methods.
 Figure \ref{fig_intro_cop} shows that two variables can have a wide 
@@
 very bad years for component populations would occur at the same time in 
-many patches, reducing the potential for rescue effects. Greater left-tail associations
+many patches, reducing the potential for rescue effects. Greater right-tail associations
 would probably not have the same effect, even if overall correlations between locations
@@
 @Ghosh_copula substantiated this intuition by showing that it held true for a 
-modeling setup using a spatial version of a very simple non-density dependent
+modeling setup using a spatial version of a very simple non-density-dependent
 population growth model (the Lewontin-Cohen model). In essence, if environmental
@@
 used population models. We intentionally use a simulation approach and simple models
-instead of attampting mathematical analysis and/or using realistic models
+instead of attempting mathematical analysis and/or using models that are more realistic, 
 because our intention here is to provide an initial exploration only of 
@@
 asymmetries of tail association may make a difference for extinction risks. 
+Some material in the Discussion outlines possible next steps.
 <!--***DAN: In Discussion need to talk about what analytic results are needed
@@
 talking about how our results, which admitedly are just for specific models,
-could be turned into analytic results or general results.-->
+could be turned into analytic results or general results. Interestingly, 
+Tulja's approximation may also apply to the spatial case, at least when 
+all locations can be reached from all other locations via dispersal. But the 
+approximation seems poor in that case. Need to think about it more, and maybe
+need more research.-->
 Though tail associations are used
@@
 will have the effect of making the sample kendalls on the 4 panels slightly
-different, which will mean readers think anything funny is going on.-->
+different, which will mean readers wont think anything funny is going on.-->
 
@@
 
-<!--***DAN: Shyamolina, the below text makes me wonder if I am misunderstanding something, please
+<!--***DAN: Insert para here talking about in what sense the variables
+$s_e(t)$ and $f_a(t)$ are asymmetrically associated, through time, in their 
+tails. This is where you intro the partial Spearman and may also make references
+to Genest and others indicating that, when you are interested in dependence,
+you want a rank-based notion of association - it's the only thing that
+separates the association from the marginals. I will insert some text now,
+but it won't fit perfectly until after we go through and replace Kendall with
+Spearman throughout.-->
+The variables $u(t)$ and $v(t)$ are visibly asymmetrically tail associated (Fig. \ref{fig_intro_cop}),
+but we have instead used the transformed variables $s_e(t)$ and $\tilde{f}_a(t)$
+in our model. In what sense are $s_e(t)$ and $\tilde{f}_a(t)$ asymmetrically
+tail associated? Given two boundary quantities $l_b$ and $u_b$ such
+that $0 \leq l_b < u_b \leq 1$, and letting $x(t)$ be the rank of 
+$u(t)$ in the set $\{u(t) | t=1,\ldots,T\}$ and letting 
+$y(t)$ be the rank of 
+$v(t)$ in the set $\{v(t) | i=1,\ldots,T\}$ (smallest elements
+are defined to have rank 1),
+consider the portions of the 
+lines $x+y = 2l_b$ and $x+y=2u_b$ which meet with the unit square
+(Fig. \ref{PartialSpearmanFig}). 
+<!--***DAN: need to make this figure-->
+We can then define the *partial Spearman correlation*,
+\begin{equation}
+\cor_{l_b,u_b}(u,v) = \frac{\sum (x(t)-\mean(x))(y(t)-\mean(y))}{(n-1)\sqrt{\var(x) \var(y)}},\label{PartialSpearmanEq}
+\end{equation}
+\noindent where sample means and variances in this equation
+are computed using all the data but the sum is computed only
+over the $t$ such that $x(t)+y(t) > 2l_b$ and $x(t)+y(t)<2u_b$. 
+The partial Spearman correlation is the component of the Spearman
+correlation of $u$ and $v$ which can be attributed to the points
+for which the ranks, $x$ and $y$, lie between the boundary lines
+we defined (Fig. \ref{PartialSpearmanFig}) in
+the unit square. The quantity $\cor_{0,0.5}(u,v)-\cor_{0.5,1}(u,v)$ 
+is a way to measure asymmetry of tail associations between
+$u$ and $v$. And this quantity is the same as 
+$\cor_{0,0.5}(s_e,\tilde{f}_a)-\cor_{0.5,1}(s_e,\tilde{f}_a)$, 
+because the partial Spearman correlation, like the Spearman
+correlation itself, is based on ranks. Ranks are the same
+for $u(t)$ and for $s_e(t)$ because these quantities are 
+related by a monotonic transformation (and likewise for 
+$v(t)$ and $\tilde{f}_a(t)$). Thus, in the sense captured
+by the statistic $\cor_{0,0.5}-\cor_{0.5,1}$, $s_e(t)$
+and $\tilde{f}_a(t)$ have the same asymmetry of tail
+association as the values $u(t)$ and $v(t)$ from which they
+were constructed.
+
+<!--***DAN: Shyamolina, the below commented text makes me wonder if I am misunderstanding something, please
 explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
@@
 
-<!--***DAN: got to here 2019 10 25-->
+We simulated our metapopulation models using environmental noises $\epsilon_i(t)$ (Table \ref{tab_model_summary}), 
+independent through time, that were generated to have the same overall associations 
+between all patches and in all simulations (Kendall's $\tau$ again $0.75$, up to sampling variation), 
+but with different degrees and types of asymmetry of tail associations 
+between patches in different simulations, as follows. For a simulation of length $T$,
+we used methods that can generate $T$ independent $n$-tuples $(u_1(t),\ldots,u_n(t))$
+for $t=1,\ldots,T$ with $\tau=0.75$ between $u_i(t)$ and $u_j(t)$ for all $i \neq j$
+(up to sampling variation), but with either extreme left-, moderate left-, moderate right-,
+or extreme right-tail association between all pairs of locations (as in Fig. \ref{fig_intro_cop}). 
+Pairs $u_i(t)$ and $u_j(t)$ looked like Fig. \ref{fig_intro_cop} and were structured, statistically,
+in the same way as the stochastic quantities $u(t)$ and $v(t)$ used for the non-spatial 
+simulations described above. For any $i$, $u_i(t)$ was again uniformly distributed on 
+the interval $(0,1)$. To obtain the values $\epsilon_i(t)$, the $u_i(t)$ were 
+transformed so that the resulting numbers were
+normally distributed with mean $0$ and standard deviation $\sigma$ (Table \ref{tab_model_summary}), 
+for all $i$. This was 
+accomplished with the inverse cummulative distribution function of the normal distribution
+with mean $0$ and standard deviation $\sigma$. Because this was a monotonic transformation,
+the Kendall correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
+as that between $u_i(t)$ and $u_j(t)$. The quantities $(u_1(t),\ldots,u_n(t))$
+for $t=1,\ldots,T$ were generated as described in Appendix \ref{SM-makenoise}.
 
-We simulated our metapopulation models with noises $(\epsilon_t^i)$ that were 
-independent through time with same spatial correlations 
-between patches (to within sampling error)
-for every simulation, but were moderately (or extremely) left- and 
-right-tail associated between patches in
-different simulations. For moderately left-tail associated 
-noises we drew points from a $n$ dimensional 
-Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
-any two out of $n$ patches and made 
-the marginals normal with zero mean and standard deviation $\sigma$ 
-(see the specified value $\sigma$ for each model in 3$^{rd}$ 
-column of Table \ref{tab_model_summary}). For moderately 
-right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
-from Clayton copula so that all the statistical properties remained 
-same except for the direction of tail association. We generated
-extreme left- or right-tail associated noises via algorithm stated in 
-Appendix \ref{SM-extr_tail_dep}. Our models (Table \ref{tab_model_summary}) 
+<!--***DAN: Dan is starting to wonder if referees will wonder something like "what 
+happens to tail associations when you do these transformatons." We do not wonder 
+this because we have in mind a defn for tail associations that only depends on ranks,
+but we have not presented that defn. We have instead only given an intuitive description
+of tail association. This may be a problem. Referees will see uf generating data with tail associatons
+and uniform marginals, but then transforming it, and putting the transformed values into 
+our models. And we are confortable with the idea that that transformation does not
+alter the tail associations, but they are not comfortable with that.
+We may need to define tail associations
+using partial Spearman or something, to get this idea across. The problem with that is, 
+we use Kendall throughout
+the rest of the ms, and so if we define partial Spearman that will create an inconsistency
+(using Kendall in some places, Spearman in others) that may raise eyebrows in a short paper 
+like this. Not sure right now what to do anout this, we need to think about it. Can we add a new
+para at the beginnign of methods giving some kind of formal introduction to tail associations
+that makes clear it is something we consider to be for ranks? Without using partial Spearman?
+There is no partial Kendall, since Kendall is based on concordant and discordant pairs.
+An alterntive is to replace Kendall wth Spearman throughout the paper, then go ahead
+and intro partial Spearmn as a way of making it clear tail associaton is for ranks. 
+Is this possible? It would definitely be a pain, so therefore want to avoid unless
+necessary. One thing this would require would be calculating the populaton (not sample) 
+Spearman of the extreme tail associated data. At least we could estimate it very precisely
+using a huge sample (would not have the same problems doing that as we did for Kendall),
+and may even be able to calculate it analytically. Let's just use Spearman and calculate it very
+precisely using a large sample. Well, maybe we could calculate it analytically. 
+-->
+
+To calculate extinction risk for a given model and a given choice of tail association,
+we simulated the model for $25$ time steps, $10000$ times, using $n=5$ habitat patches.
+The initial condition $P_e$ was used in each habitat patch ($P_e$ is the equilibrium value of the 
+one-patch, deterministic version of the model, see Table \ref{tab_model_summary}).
+After each time step, if $P_i(t)$ was less than $P_e/10$ then it was set to $0$.
+The whole metapopulation was considered to have gone extinct if all patches were $0$
+a time step $25$.
+<!--***DAN: Shyamolina, please check the above two sentences, is this what you did?-->
+Extinction risk was the fraction of 10000 metapopulations that went extinct.
+
+<!--***DAN: Dan plans to delete this text as it has been superseded by the above. Keeping it
+around until I am sure the above is right:
+Our models (Table \ref{tab_model_summary}) 
 computed extinction risk after $25$ years for $10000$ simulations at $n = 5$ patches
 with spatially correlated noises having either moderate or extreme tail 
-association in left and right-tail. <!--say how you set ext. threshold = Pe/10-->
+association in left and right-tail. 
 We started with initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ for each patch, where 
@@
 certain threshold (set to 10\% of the initial population $P_e$ for individual model). 
+-->
+
+<!--***DAN: Dan stopped here on 2019 10 27 until we resolve some of the issues that have come up-->
 
@@
 ```
+<!--***DAN: Shyamolina, in this table and in the sup mat (and anywhere else where this comment applies),
+please replace the notation P^i_t and \epsilon^i_t by P_i(t) and \epsilon_i(t), the reason being,
+putting an i in the superscript makes it look like and exponent. This is legitimately confusing because
+we use lots of other exponents. I want you to generally refrain from placing the abitat patch index in the
+superscript, putting in the subscript instead. This means you have to put the t in parens in front of the
+variable in question, instead of in the subscript. This will require several changes to this table. Please
+make those changes and then commit. Then please make the corresponding changes to the sup mat, whereever needed.-->
+<!--***DAN: Dan, once Shyamolina does the above, you will then have to go back and edit this table in various
+ways that you cannot do until after that change is made.-->
 

### 2019-10-28 SuppMat.Rmd
working on para intro-ing partial spearman

@@
 
+<!--yet more stuff from the main text-->
+
+For moderately left-tail associated 
+noises we drew points from a $n$ dimensional 
+Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
+any two out of $n$ patches and made 
+the marginals normal with zero mean and standard deviation $\sigma$ 
+(see the specified value $\sigma$ for each model in 3$^{rd}$ 
+column of Table \ref{tab_model_summary}). For moderately 
+right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
+from Clayton copula so that all the statistical properties remained 
+same except for the direction of tail association. We generated
+extreme left- or right-tail associated noises via algorithm stated in 
+Appendix \ref{SM-extr_tail_dep}.
+
 # Stability analysis of deterministic metapopulation models  \label{model_stability}

### 2019-10-28 SuppMat.Rmd
running with common spearman's rho

@@
 source("mtime.R")
-# Note : for moderate tail dep. we used Kendall's tau = 0.6 to get C or SC
-```
-
-```{r get_ktau_common, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("ExtremeTailDep.R"))}
+```
+
+```{r get_srho_common, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-np<-100000
-
-#for(i in 1:5){ # for loop to ensure the consistency of Kendall's tau with specified np
+np<-1000000
+srho_common<-0.875 #7/8
+
 xtc_r<-retd(n=np,d=2,rl=1,mn=0,sdev=1) # copula with extreme right tail
 
-  start_time <- Sys.time()
-  ktau_common<- cor(xtc_r[,1],xtc_r[,2],method="kendall") #get the kendall's tau
-  end_time <- Sys.time()
-
-#  cat("points = ",np, ", ktau_common = ", ktau_common, ", run time = ", (end_time -start_time),"\n")
-#}
-
-# see the change in 3rd decimal place
-#points =  1e+05 , ktau_common =  0.7512526 , run time =  1.837576 
-#points =  1e+05 , ktau_common =  0.7507594 , run time =  1.864947 
-#points =  1e+05 , ktau_common =  0.7497681 , run time =  1.862879 
-#points =  1e+05 , ktau_common =  0.7487744 , run time =  1.854317 
-#points =  1e+05 , ktau_common =  0.7516491 , run time =  1.840355
-
-saveRDS(ktau_common,"./Results/ktau_common.RDS")
+saveRDS(srho_common,"./Results/srho_common.RDS")
 
@@
 copC<-claytonCopula(3)
-parC<-iTau(copC,tau=ktau_common)
+parC<-iRho(copC,rho=srho_common)
 
@@
 
-# calculate tau with sampling variation
-kcor_el<-cor(xtc_l,xtc_l,method = "kendall")
-kcor_el<-kcor_el[1,2]
-
-kcor_er<-cor(xtc_r,xtc_r,method = "kendall")
-kcor_er<-kcor_er[1,2]
-
-kcor_ml<-cor(mtc_l,mtc_l,method = "kendall")
-kcor_ml<-kcor_ml[1,2]
-
-kcor_mr<-cor(mtc_r,mtc_r,method = "kendall")
-kcor_mr<-kcor_mr[1,2]
+# calculate rho with sampling variation
+scor_el<-cor(xtc_l,xtc_l,method = "spearman")
+scor_el<-scor_el[1,2]
+
+#scor_er<-cor(xtc_r,xtc_r,method = "spearman")
+scor_er<- -scor_el # Note it should be opposite to scor_el
+
+scor_ml<-cor(mtc_l,mtc_l,method = "spearman")
+scor_ml<- scor_ml[1,2]
+
+#scor_mr<-cor(mtc_r,mtc_r,method ="spearman")
+scor_mr<- -scor_ml # Note it should be opposite to scor_ml
 ```
@@
 text(x=0.1,y=0.9,"A",cex=2.5)
-text(x=0.8,y=0.1,round(kcor_el,4),cex=2)
+text(x=0.8,y=0.1,round(scor_el,4),cex=2)
 
@@
 text(x=0.1,y=0.9,"B",cex=2.5)
-text(x=0.8,y=0.1,round(kcor_ml,4),cex=2)
+text(x=0.8,y=0.1,round(scor_ml,4),cex=2)
 
@@
 text(x=0.1,y=0.9,"C",cex=2.5)
-text(x=0.8,y=0.1,round(kcor_mr,4),cex=2)
+text(x=0.8,y=0.1,round(scor_mr,4),cex=2)
 
@@
 text(x=0.1,y=0.9,"D",cex=2.5)
-text(x=0.8,y=0.1,round(kcor_er,4),cex=2)
-
-par(op)
-dev.off()
-```
-
-```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("ext_risk_cop_age_str.R"))}
+text(x=0.8,y=0.1,round(scor_er,4),cex=2)
+
+par(op)
+dev.off()
+```
+
+```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("ext_risk_cop_age_str.R"))}
 # extinction risk with moderate tail dep. for single patch linear age structure model
@@
 omg<-Inf
-ktau<-ktau_common  # kendall's tau
+srho<-srho_common  # Spearman's rho
 
@@
 copC<-claytonCopula(3)
-parC<-iTau(copC,tau=ktau)
+parC<-iRho(copC,rho=srho)
 par<-parC
@@
 
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_scor_",srho,"_linear",sep="")
 if (!dir.exists(tempo)){
@@
 
-# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same
+# Now for survival clayton : should be the same parameter as of clayton if Spearman's rho is the same
 
@@
 
-```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("ext_risk_cop_age_str.R"))}
+```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("ext_risk_cop_age_str.R"))}
 # extinction risk with moderate tail dep. for single patch non-linear age structure model
@@
 omg<-100
-ktau<-ktau_common  # kendall's tau
+srho<-srho_common  # Spearman's rho
 
@@
 
-tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_nonlinear",sep="")
+tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_scor_",srho,"_nonlinear",sep="")
 if (!dir.exists(tempo)){
@@
 
-```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-1
-ktau<-ktau_common
+srho<-srho_common
 ext_thrs<-K/10
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
-                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 
-```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-1
-ktau<-ktau_common
+srho<-srho_common
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
-                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 
-```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-0.8
-ktau<-ktau_common
+srho<-srho_common
 
@@
 
-```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,ktau_common,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,srho_common,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 
-```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 scl<-0.6
-ktau<-ktau_common
+srho<-srho_common
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
     ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 
-```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
   ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
+               scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)
 
@@
   ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
-               scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
+               scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)
 
@@
 moderate right-tail association in the vital rates. Both copulae had the same Kendall 
-correlation ($\tau =$ `r ktau_common`) but opposite association in their respective tails. 
+correlation ($\rho =$ `r srho_common`) but opposite association in their respective tails. 
 We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
@@
 We chose parameters for each copula family using 
-`iTau` function from `copula` package so that both copula had
-same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each column of
+`iRho` function from `copula` package so that both copula had
+same Kendall's correlation ($\rho =$ `r srho_common`). Entries within each column of
 matrix $M$, then became independent of each other but 
@@
 noises we drew points from a $n$ dimensional 
-Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
+Clayton copula with same Kendall correlation ($\rho =$ `r srho_common`) between
 any two out of $n$ patches and made 

### 2019-10-28 mydispersal.R

@@ -94,8 +94,21 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
       r<-params[1]
       K<-params[2]
       L<-params[3]
-      lam_malthus<-r*res[,,tct]*(K-res[,,tct]+(L*log(res[,,tct])))
+      
+      x1<-res[,,tct]
+      id0<-which(x1==0,arr.ind = T)
+      idp<-which(x1!=0,arr.ind = T)
+      
+      x2<-x1
+      
+      x2[idp]<-r*x1[idp]*(K-x1[idp]+(L*log(x1[idp])))
+      x2[id0]<-0 # just to ensure extinct value
+      
+      #lam_malthus<-r*res[,,tct]*(K-res[,,tct]+(L*log(res[,,tct])))
+      lam_malthus<-x2
+      
       res[,,tct+1]<-lam_malthus*lam_sto #numsims by numlocs matrix
+      
     }else if(model=="austinbrewer"){
       r<-params[1]
       K<-params[2]
@@ -173,13 +186,13 @@ get_avg_acf<-function(sims){
 #       6) params : a vector containing model parameters
 #       7) ext_thrs : extinction threshold below which populations go extinct
 #       8) scl : a scaling factor which is multiplied with noise generated
-#       9) ktau : given kendall's tau to generate moderate tail dep. copula
+#       9) srho : given Spearman's rho to generate moderate tail dep. copula
 #       10) model : a character specifying model name
 #       11) ploton : logical(T or F) to get optional plot
 #       12) resloc : location to save plots
 #       13) getacf : these are tags
 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,ktau,model,ploton,resloc,getacf){
+plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,srho,model,ploton,resloc,getacf){
   
   # ========================== when noises are extreme tail dep. ========================================
   ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1,mn=0,sdev=1)# a righttail dep matrix(numpoints by numlocs,     
@@ -210,11 +223,10 @@ plotter_ext_risk<-function(numsims,numsteps,numlocs,D,p0,params,ext_thrs,scl,kta
   }
   
   # ========================== when noises have moderate tail dep. ========================================
[…]

### 2019-10-29 MainText.Rmd
more edits

@@
 <!--ECEs have impacts in the spatial context as well-->
-ECEs also have impacts in ecological scenarios for which spatial dynamics are
+ECEs also have impacts in ecological contexts for which spatial dynamics are
 important.
 In the metapopulation context, spatial synchrony of population dynamics, which can
-be caused by spatial synchrony of climatic fluctuations, is linked to metapopulation
+be caused by spatial synchrony of climatic fluctuations [@Liebhold2004], is linked to metapopulation
 extinction risk [@hanski1998metapopulation; @earn2000coherence].
@@
 
+<!--***DAN: maybe you want to have a Discussion para on copulas. It could start with
+some text about ranks being the proper way to study associations, and then could go
+on to link to copulas and cite BIVAN-->
+
 <!--ECEs probably produce tail associations-->
-There are reasons to believe that ECEs may produce asymmetric tail 
+There are reasons to hypothesize that ECEs may produce asymmetric tail 
 associations between climatic variables
@@
 variable are associated with local weather rather than ECEs, the result could
-be stronger right-tail association between measurements of the climatic variable
-made in different locations, through time.
+be stronger right-tail association, through time, between measurements of the climatic variable
+made in different locations.
 For example, rainstorms which deposit exceptionally large amounts of rain
-may do so over large areas; but months with less rain may experience
+may do so over large areas; but periods with less rain may experience
 scattered showers only, which may be much less correlated across locales.
@@
 were the same, because in that case it would be very good years for the component 
-populations that would occur at the same time in many patches [@Ghosh_copula]. 
+populations that would occur at the same time in many patches. 
 @Ghosh_copula substantiated this intuition by showing that it held true for a 
@@
 if overall (non-tail-specific) levels of environmental synchrony are the same
-in both these cases. The present study extends tests of this idea to metapopulation
+in both these scenarios. The present study extends tests of this idea to metapopulation
 models incorporating density dependence.
@@
 depend on an environmental variable such as spring rainfall in similar ways, but 
-only when that variable takes very low values. When rainfall takes
+only when that variable takes low values. When rainfall takes
 typical or high values, fecundity and survival may be limited instead 
@@
 because our intention here is to provide an initial exploration only of 
-the potential mechanisms outlined above which suggest that
+the potential mechanisms, outlined above, which suggest that
 asymmetries of tail association may make a difference for extinction risks. 
-Some material in the Discussion outlines possible next steps.
+By so doing we hope to open these potentially practically important questions to 
+wider examination other researchers. Some material in the Discussion outlines some possible next steps.
 <!--***DAN: In Discussion need to talk about what analytic results are needed
@@
 Our questions are an important step for understanding
-the potential influence of ECEs on extinction risks if ECEs are associated with
-tail dependence in environmental variables.
+the potential influence of ECEs on extinction risks if ECEs are related to
+tail associations in environmental variables.
 
@@
 \includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
-\caption{Two stochastic quantities can have same Kendall 
+\caption{Two stochastic quantities can have same Spearman 
 correlation, but different degrees of tail association: 
@@
 moderate right-tail association, and (D) extreme right-tail association.
-In all cases, Kendall's $\tau$ was the same, $\tau = 0.75$,
+In all cases, Spearman's $\rho$ was the same, $\rho = 0.875$,
 up to sample variation (the sample correlation for each dataset is displayed
@@
 \end{figure*}
-<!--***DAN: Shyamolina, it occured to be there are simple and good analytic reasons
-why ktau_common is exactly 0.75 (up to sample variation). 
-I bet you will see them too if you look up the
-definition of kendall tau and think about the extreme tail dependent cases. 
-Currently, around line 50 of the sup mat Rmd, you calculate ktau_common
-by simulating 100000 points from the extreme-tail-denepdnece noise and then
-calculating the sample kendall. Instead, can you please replace those lines 
-with the simple line "ktau_common<-0.75"? I would do this myself but I
-don't want to risk messing up your code in some way I don't understand. You should not
-need the seed anymore to ensure you get the same ktau_common on each run
-since you won't be using the sample correlartion anymore to estimate ktau_common,
-you'll just be using the analytic result that the population correlation
-of data from the extreme-tail-dependent algorithm (in the limit of infinite
-sample sizes) is exactly 0.75. Please
-let me know if you do not understand something here.-->
-<!--***DAN: Shyamolina, currently the sample kendalls are the same on A and D, and
-also on B and C, and I know why. This looks weird, and makes the reader
-wonder why this is the case. Rather than plotting the reversed version of the
-data from A on D, and the reversed version of the data from B on C, can you plase
-generate separate, independent data for A and D, and also for B and C? That
-will have the effect of making the sample kendalls on the 4 panels slightly
-different, which will mean readers wont think anything funny is going on.-->
 
@@
 The stochastic quantities $s_e(t)$ and $\tilde{f}_a(t)$ were generated 
-to have the same overall association (Kendall's $\tau$ equal to $0.75$,
+to have the same overall association (Spearman's $\rho$ equal to $0.875$,
 up to sampling variation), but different degrees and types of 
@@
 generate $T$ independent pairs $(u(t),v(t))$ for $t=1,\ldots,T$ with
-$\tau=0.75$ (up to sample variation), but with either extreme left-,
+$\rho=0.875$ (up to sample variation), but with either extreme left-,
 moderate left-, moderate right-, or extreme right-tail association 
@@
 a beta distribution with parameters
-$\alpha=3$ and $\beta=2$ 
-and used those values for the $s_e(t)$. We also applied a (different)
+$\alpha=3$ and $\beta=2$, 
+and we used those values for the $s_e(t)$. We also applied a (different)
 monotonic transformation to the $v(t)$ so that the resulting numbers were
 distributed according to a gamma distribution with parameters $k=2$
-and $\theta=1$ and used those values for the $\tilde{f}_a(t)$. Because
+and $\theta=1$, and we used those values for the $\tilde{f}_a(t)$. Because
 the $u(t)$ and $v(t)$ were uniformly distributed, the inverse 
@@
 gamma distributions were the transformations. Because these
-were monotonic transformations, the Kendall correlation of 
+were monotonic transformations, the Spearman correlation of 
 $s_e(t)$ and $\tilde{f}_a(t)$ was the same as that of 
@@
 
-<!--***DAN: Insert para here talking about in what sense the variables
-$s_e(t)$ and $f_a(t)$ are asymmetrically associated, through time, in their 
-tails. This is where you intro the partial Spearman and may also make references
-to Genest and others indicating that, when you are interested in dependence,
-you want a rank-based notion of association - it's the only thing that
-separates the association from the marginals. I will insert some text now,
-but it won't fit perfectly until after we go through and replace Kendall with
-Spearman throughout.-->
 The variables $u(t)$ and $v(t)$ are visibly asymmetrically tail associated (Fig. \ref{fig_intro_cop}),
@@
 $v(t)$ in the set $\{v(t) | i=1,\ldots,T\}$ (smallest elements
-are defined to have rank 1),
+are here defined to have rank 1),
 consider the portions of the 
@@
 for which the ranks, $x$ and $y$, lie between the boundary lines
-we defined (Fig. \ref{PartialSpearmanFig}) in
-the unit square. The quantity $\cor_{0,0.5}(u,v)-\cor_{0.5,1}(u,v)$ 
+we defined in
+the unit square (Fig. \ref{PartialSpearmanFig}). 
+The quantity $\cor_{0,0.5}(u,v)-\cor_{0.5,1}(u,v)$ 
 is a way to measure asymmetry of tail associations between
@@
 for $u(t)$ and for $s_e(t)$ because these quantities are 
-related by a monotonic transformation (and likewise for 
+related by an increasing monotonic transformation (and likewise for 
 $v(t)$ and $\tilde{f}_a(t)$). Thus, in the sense captured
@@
 association as the values $u(t)$ and $v(t)$ from which they
-were constructed.
+were constructed. Rank-based approaches to understanding the nature
+of association between variables have been recommended, as such
+approaches are not influenced by the marginal distributions of
+the quantities (e.g., @Genest2007). Marginal distributions contain no information
+about association between variables, as each marginal pertains to
+only one of the variables.
 
@@
 of six density dependent population models, the Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, 
-and Austin-Brewer models (@cohen1995unexpected; Table \ref{tab_model_summary}).
+and Austin-Brewer models (Table \ref{tab_model_summary}).
 All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
@@
 The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{\text{th}}$ diagonal entry 
-$\lambda_i(t)$ depnding on the model used (see Table \ref{tab_model_summary}). The matrix $D$
+$\lambda_i(t)$ depending on the model used (Table \ref{tab_model_summary}). The matrix $D$
 was an $n \times n$ dispersal matrix. For simplicity, 
@@
 the two or one nearest neighbors. For global dispersal the same fraction $d$ dispersed
-equally to the other $(n-1)$ patches. When $d$ was $0$, dispersal did not occur and 
+equally to the other $n-1$ patches. When $d$ was $0$, dispersal did not occur and 
 rescue effects were not possible.
-Two models that @cohen1995unexpected used were excluded from our analyses because they were not
-readily adapted to our simulation setup: the Varley model,
-which is not continuous;
-<!--***DAN: Shyamolina, looking back at the Cohen paper, using the value C=1 as they used, the
-Varley model *is* continuous. It is not differentiable, but it's not clear to me that 
-non-differentiability means we cannot use the model. Perhaps it does mean that, but I don't see 
-it. Let's discuss. I'm not saying we have to go back and use it, but we may have to justify
-differently why we used 6 of Cohen's models and not the other two.-->
-and the Malthus-Condorcet-Mill model, which employs a logarithm of the population and is therefore
-inconvenient to analyze in the context of extinction.
-<!--***DAN: Unless I am mistaken, the limit of the right hand side of the M-C-M model as
-described by Cohen, as P approaches 0 from the right, is 0. That means the model has
-a removable singularity at 0, i.e., we could just define the model so P_t+1=0 when P_t
-is 0, and we would get a perfectly valid continuous model. This means that using logs is 
-not really a valid reason to not use the models. Let's discuss. I'm not saying we have to go 
-back and use it, but we may have to justify differently why we used 6 of Cohen's models and not 
-the other two.-->
+
+<!--***DAN: Also need to talk about how you selected parameters-->
+We obtained our list of classic population models (which we then spatialized and made 
+stochastic) from @cohen1995unexpected. 
+Model parameters were selected so that varying $r$ across a range, while keeping
+the other selected parameters fixed at selected values for a model, led to a 
+transition from undercompensatory to overcompensatory equilibrium dynamics
+for the deterministic, one-patch versions of the models. Whereas @cohen1995unexpected
+was interested in studying chaotic model behavior and therefore chose parameters
+for which the models exhibited chaotic dynamics, we were interested in non-chaotic model
+behaviors, which seem likely to be more common in real populations. 
+Two models that @cohen1995unexpected 
+used were excluded from our analyses because their functional forms complicated 
+the analyses we wished to perform. First, the equilibrium of the 
+Malthus-Condorcet-Mill model [@cohen1995unexpected] could not be determined 
+analytically, making downstream analysis difficult. Second, for
+the Varley model and for the portions of parameter space we explored, 
+varying $r$ produced no transition from under- to overcompensatory dynamics.
+
+<!--***DAN: got to here-->
 
@@
 
-<!--***DAN: Dan is starting to wonder if referees will wonder something like "what 
-happens to tail associations when you do these transformatons." We do not wonder 
-this because we have in mind a defn for tail associations that only depends on ranks,
-but we have not presented that defn. We have instead only given an intuitive description
-of tail association. This may be a problem. Referees will see uf generating data with tail associatons
-and uniform marginals, but then transforming it, and putting the transformed values into 
-our models. And we are confortable with the idea that that transformation does not
-alter the tail associations, but they are not comfortable with that.
-We may need to define tail associations
-using partial Spearman or something, to get this idea across. The problem with that is, 
-we use Kendall throughout
-the rest of the ms, and so if we define partial Spearman that will create an inconsistency
-(using Kendall in some places, Spearman in others) that may raise eyebrows in a short paper 
-like this. Not sure right now what to do anout this, we need to think about it. Can we add a new
-para at the beginnign of methods giving some kind of formal introduction to tail associations
-that makes clear it is something we consider to be for ranks? Without using partial Spearman?
-There is no partial Kendall, since Kendall is based on concordant and discordant pairs.
-An alterntive is to replace Kendall wth Spearman throughout the paper, then go ahead
-and intro partial Spearmn as a way of making it clear tail associaton is for ranks. 
-Is this possible? It would definitely be a pain, so therefore want to avoid unless
-necessary. One thing this would require would be calculating the populaton (not sample) 
-Spearman of the extreme tail associated data. At least we could estimate it very precisely
-using a huge sample (would not have the same problems doing that as we did for Kendall),
-and may even be able to calculate it analytically. Let's just use Spearman and calculate it very
-precisely using a large sample. Well, maybe we could calculate it analytically. 
--->
-
 To calculate extinction risk for a given model and a given choice of tail association,
@@
 -->
-
-<!--***DAN: Dan stopped here on 2019 10 27 until we resolve some of the issues that have come up-->
 
@@
 putting an i in the superscript makes it look like and exponent. This is legitimately confusing because
-we use lots of other exponents. I want you to generally refrain from placing the abitat patch index in the
+we use lots of other exponents. I want you to generally refrain from placing the habitat patch index in the
 superscript, putting in the subscript instead. This means you have to put the t in parens in front of the

### 2019-10-29 MainText.Rmd
edited notations

@@
 ```{r read_res, echo=F}
-ktau_common<-readRDS("./Results/ktau_common.RDS")
+srho_common<-readRDS("./Results/srho_common.RDS")
 ```
@@
 independent through time, that were generated to have the same overall associations 
-between all patches and in all simulations (Kendall's $\tau$ again $0.75$, up to sampling variation), 
+between all patches and in all simulations (Spearman's $\rho$ again `r srho_common`, up to sampling variation), 
 but with different degrees and types of asymmetry of tail associations 
@@
 we used methods that can generate $T$ independent $n$-tuples $(u_1(t),\ldots,u_n(t))$
-for $t=1,\ldots,T$ with $\tau=0.75$ between $u_i(t)$ and $u_j(t)$ for all $i \neq j$
+for $t=1,\ldots,T$ with $\rho=$ `r srho_common` between $u_i(t)$ and $u_j(t)$ for all $i \neq j$
 (up to sampling variation), but with either extreme left-, moderate left-, moderate right-,
@@
 with mean $0$ and standard deviation $\sigma$. Because this was a monotonic transformation,
-the Kendall correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
+the Spearman correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
 as that between $u_i(t)$ and $u_j(t)$. The quantities $(u_1(t),\ldots,u_n(t))$
@@
 rate $d= 0,0.1,0.2,..0.5$ and moderate to extreme left 
-tail- and right-tail associated environmental noises with same Kendall's $\tau$ (= `r ktau_common`)
+tail- and right-tail associated environmental noises with same Kendall's $\rho$ (= `r srho_common`)
 in different simulations (for code see [`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R)).
@@
 Fig.
-\ref{fig_extrisk_agestr} represents this scenario when vital rates have moderate to extreme tail association in either tail but with same Kendall's correlation ($\tau =$ `r ktau_common`).
+\ref{fig_extrisk_agestr} represents this scenario when vital rates have moderate to extreme tail association in either tail but with same Kendall's correlation ($\rho =$ `r srho_common`).
 Extinction risks for two different tail-dependent vital rates were not strinkingly different (e.g.
@@
              format="latex",digits=3, align="c",table.env='table*',
-             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
-Superscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
+             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_{e}$), critical growth rate ($r_{c}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
+Subscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"

### 2019-10-29 SuppMat.Rmd
edited notations

@@
 
-dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise $(\\epsilon_t^{i})$" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
-                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/(1+aP_t^{i})^b$",
-                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+(aP_t^{i})^b]$",
-                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_t^{i}/b)\\}]$",
-                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i}/K)]$",
-                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(K-P_t^{i})\\{1-\\exp(-sP_t^{i})\\}]$"
+dt2<-tibble("Density dependence $(\\lambda_i(t))$ with noise $(\\epsilon_i(t))$" = c("$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
+                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",
+                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+(aP_i(t))^b]$",
+                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_i(t)/b)\\}]$",
+                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(1-P_i(t)/K)]$",
+                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(K-P_i(t))\\{1-\\exp(-sP_i(t))\\}]$"
 ))
@@
 
-dt4<-tibble("$P_e^{i}$" = c("K",
+dt4<-tibble("$P_{e}$" = c("K",
                         "$(r^{1/b}-1)/a$",
@@
 
-dt5<-tibble("$r_c^{i}$" = c(rc_ricker,
+dt5<-tibble("$r_{c}$" = c(rc_ricker,
                         rc_hassell,
@@
 
-dt6<-tibble("$r_{bf}^{i}$" = c(r_bf_ricker,
+dt6<-tibble("$r_{bf}$" = c(r_bf_ricker,
                            r_bf_hassell,
@@
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of tail association in environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
@@
 \includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of tail association in environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
@@
 \includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
-standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations.. \label{fig_msmith_risk_local_disp_scl_0.8}}
+\caption[Effect of tail association in environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years) for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal;
+standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_msmith_risk_local_disp_scl_0.8}}
 \end{center}
@@
 \includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of tail association in environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
+\caption[Effect of tail association in environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
 for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
@@
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
+\caption[Effect of tail association in environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years) for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal;
 standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
@@
 copula to model moderate left-tail and Survival Clayton copula to model 
-moderate right-tail association in the vital rates. Both copulae had the same Kendall 
+moderate right-tail association in the vital rates. Both copulae had the same Spearman 
 correlation ($\rho =$ `r srho_common`) but opposite association in their respective tails. 
@@
 `iRho` function from `copula` package so that both copula had
-same Kendall's correlation ($\rho =$ `r srho_common`). Entries within each column of
+same Spearman correlation ($\rho =$ `r srho_common`). Entries within each column of
 matrix $M$, then became independent of each other but 
 entries between columns had a specific tail association
-with positive Kendall's correlation.
+with positive Spearman correlation.
 
@@
 noises we drew points from a $n$ dimensional 
-Clayton copula with same Kendall correlation ($\rho =$ `r srho_common`) between
+Clayton copula with same Spearman correlation ($\rho =$ `r srho_common`) between
 any two out of $n$ patches and made 
@@
 First, we will briefly discuss about the stability analysis for a difference 
-equation which is frequently used in population ecology model. Let's say, $P_{t+1} = F(P_t,r)$
+equation which is frequently used in population ecology model. Let's say, $P(t+1) = F(P(t),r)$
 represents the model dynamics of a single population $P$ which progress recursively via 
@@
 employ the concept that the system at fixed point (or equilibrium $P_e$) does not change with time 
-(i.e. $P_{t+1} = P_t$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
-returns to its original state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
+(i.e. $P(t+1) = P(t)$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
+returns to its initial state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
 Mathematically, from standard linear stability analysis we can say an equilibrium is locally 
-asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P_t=P_e}$.
+asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P(t)=P_e}$.
 If the approach towards its equilibrium is monotonic (when $0<{F}'<1$), the dynamics is said to be
@@
 is represented as 
-$P_{t+1}=F(P_t,r,K)=P_t \exp[r(1-P_t/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
+$P(t+1)=F(P(t),r,K)=P(t) \exp[r(1-P(t)/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
 growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
@@
 **Hassell model** [@hassell1975density; @may1976bifurcations], motivated for modelling insect populations,
-can be represented as $P_{t+1} = F(P_t,r,a,b)=rP_t/(1+aP_t)^b$, where $r>0$ is the intrinsic growth rate 
+can be represented as $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$, where $r>0$ is the intrinsic growth rate 
 of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
@@
 **Maynard Smith model** [@cohen1995unexpected; @maynard1978models; @may1976bifurcations] follows the dynamics 
-$P_{t+1}=F(P_t,r,a,b)=rP_t/[1+(aP_t)^b]$. $r$ is the growth rate, $a, b$ are the parameters with positive values.
+$P(t+1)=F(P(t),r,a,b)=rP(t)/[1+(aP(t))^b]$. $r$ is the growth rate, $a, b$ are the parameters with positive values.
 Equilibriums are: $P_e=0$(as expected extinction equilibrium) and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
@@
 in prey-predator perspective as indicated in [@pennycuick1968computer].
-In general, model dynamics is expressed as $P_{t+1}=F(P_t,r,a,b)=rP_t/[1+\exp \left\{-a(1-P_t/b)\right\}]$, where 
+In general, model dynamics is expressed as $P(t+1)=F(P(t),r,a,b)=rP(t)/[1+\exp \left\{-a(1-P(t)/b)\right\}]$, where 
 $r$ is the intrinsic growth rate and $a>0,b>0$ are the model parameters. Model
@@
 representation of logistic growth of non-human biological population and can be written 
-as $P_{t+1} = F(P_t,r,K) =P_t[1+r(1-P_t/K)]$, wirh $r>0$ and $K>0$ being growth rate and carrying capacity, respectively. 
+as $P(t+1) = F(P(t),r,K) =P(t)[1+r(1-P(t)/K)]$, with $r>0$ and $K>0$ being growth rate and carrying capacity, respectively. 
 Model equilibrium $P_e=0$ is always unstable whereas $P_e=K$ is stable below Hopf-bifurcation 
@@
 is used to model growth of human population. It follows the
-discrete dynamics : $P_{t+1} = F(P_t,r,K,s)= P_t[1+r(K-P_t) \left\{ 1-\exp (-sP_t)\right\}]$, where
+discrete dynamics : $P(t+1) = F(P(t),r,K,s)= P(t)[1+r(K-P(t)) \left\{ 1-\exp (-sP(t))\right\}]$, where
 $r, K, s$ are positive, representing growth rate, carrying capacity 

### 2019-10-30 MainText.Rmd
more edits, working on Methods

@@
 
-<!--***DAN: Also need to talk about how you selected parameters-->
 We obtained our list of classic population models (which we then spatialized and made 
-stochastic) from @cohen1995unexpected. 
-Model parameters were selected so that varying $r$ across a range, while keeping
-the other selected parameters fixed at selected values for a model, led to a 
-transition from undercompensatory to overcompensatory equilibrium dynamics
-for the deterministic, one-patch versions of the models. Whereas @cohen1995unexpected
+stochastic, as above) from @cohen1995unexpected. 
+Model parameters were selected so that varying the growth parameter, $r$, across a 
+range, while keeping the other parameters fixed at selected values, 
+caused the  deterministic, one-patch versions of the models to transition from dynamics
+showing an undercompensatory (i.e., monotonic) approach to an equilibrium; to
+dynamics showing an overcompensatory (i.e., oscillatory) approach to an equilibrium. 
+The transition value of $r$ was denoted $r_c$. Appendix \ref{SM-model_stability}
+summarizes stability analyses of the models.
+Whereas @cohen1995unexpected
 was interested in studying chaotic model behavior and therefore chose parameters
-for which the models exhibited chaotic dynamics, we were interested in non-chaotic model
+for which the models exhibited chaotic dynamics, we were interested in non-chaotic 
 behaviors, which seem likely to be more common in real populations. 
@@
 Malthus-Condorcet-Mill model [@cohen1995unexpected] could not be determined 
-analytically, making downstream analysis difficult. Second, for
+analytically, making downstream analyses difficult for that model. Second, for
 the Varley model and for the portions of parameter space we explored, 
@@
 
-<!--***DAN: got to here-->
+<!--***DAN: By the way, Shyamolina, these are not Hopf bifuractions - those cannot happen
+with univariate models like these. These are called period-doubling bifucations.
+If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
+model equilibrium $P_e$ became unstable. Beyond the Hopf-bifurcation point, the model showed sustained
+oscillation and eventually led to chaos with increasing $r$. Appendix \ref{SM-model_stability}
+summarized the stability analysis of each model in their deterministic form.
+-->
 
@@
 the Spearman correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
-as that between $u_i(t)$ and $u_j(t)$. The quantities $(u_1(t),\ldots,u_n(t))$
+as that between $u_i(t)$ and $u_j(t)$; likewise partial Spearman correlations were the same. 
+The quantities $(u_1(t),\ldots,u_n(t))$
 for $t=1,\ldots,T$ were generated as described in Appendix \ref{SM-makenoise}.
@@
 The initial condition $P_e$ was used in each habitat patch ($P_e$ is the equilibrium value of the 
-one-patch, deterministic version of the model, see Table \ref{tab_model_summary}).
+one-patch, deterministic version of the model, Table \ref{tab_model_summary}).
 After each time step, if $P_i(t)$ was less than $P_e/10$ then it was set to $0$.
@@
 
-The models stated in Table \ref{tab_model_summary} had one common 
-feature: they showed undercompensatory
-dynamics (monotonic approach towards stable steady state $P_e$) when
-growth parameter $r$ remained below a critical 
-level (say, $r_c$) and overcompensatory dynamics with oscillatory approach 
-towards stable equilibrium $P_e$ when $r$ was beyond $r_c$. 
-If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
-model equilibrium $P_e$ became unstable. Beyond the Hopf-bifurcation point, the model showed sustained
-oscillation and eventually led to chaos with increasing $r$. Appendix \ref{SM-model_stability}
-summarized the stability analysis of each model in their deterministic form.
+<!--**DAN: got to here 2019 10 29-->
 
 <!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
-To answer (Q2) of \nameref{Introduction} for each metapopulation model,
-we estimated extinction risk for a wide range of parameter $r$ 
-that covered the undercompensatory and overcompensatory regime. We 
-chose equispaced values of $r$ from the 95\% quantiles 
+To answer (Q2) of the \nameref{Introduction} for each metapopulation model,
+we estimated extinction risk of each model using each form of tail 
+association in the environmental noise,
+across the range of $r$ indicated above that spanned under- and overcompensatory dynamics.
+We chose equispaced values of $r$ from the 95\% quantiles 
 of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
@@
              format="latex",digits=3, align="c",table.env='table*',
-             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_{e}$), critical growth rate ($r_{c}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}$ : transition point from stable to unstable equilibrium of deterministic models), respectively.
-Subscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
+             caption = "Summary of the six density-dependent stochastic metapopulation models we used.
+The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, where $P_i(t)$ is the population in patch $i$ as time $t$.
+The growth rate depends on model parameters, $P_i(t)$, and the environment $\\epsilon_i(t)$ in patch $i$ at time $t$.
+$\\epsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times.
+$P_e$ is the equilibrium of the deterministic, one-patch version of the model.
+$r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text).
+$r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation).
+We use only $r<r_{bf}$.
+Dispersal between patches was also implemented as described in the main text.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"
@@
 ```
-<!--***DAN: Shyamolina, in this table and in the sup mat (and anywhere else where this comment applies),
-please replace the notation P^i_t and \epsilon^i_t by P_i(t) and \epsilon_i(t), the reason being,
-putting an i in the superscript makes it look like and exponent. This is legitimately confusing because
-we use lots of other exponents. I want you to generally refrain from placing the habitat patch index in the
-superscript, putting in the subscript instead. This means you have to put the t in parens in front of the
-variable in question, instead of in the subscript. This will require several changes to this table. Please
-make those changes and then commit. Then please make the corresponding changes to the sup mat, whereever needed.-->
-<!--***DAN: Dan, once Shyamolina does the above, you will then have to go back and edit this table in various
-ways that you cannot do until after that change is made.-->
+<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)?-->
 

### 2019-10-30 SuppMat.Rmd
table 1, second column heading fixed

@@
 
-dt2<-tibble("Density dependence $(\\lambda_i(t))$ with noise $(\\epsilon_i(t))$" = c("$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
+dt2<-tibble("$\\lambda_i(t)$" = c("$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
                           "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",

### 2019-10-30 MainText.Rmd
edited through methods

@@
 In all cases, Spearman's $\rho$ was the same, $\rho = 0.875$,
-up to sample variation (the sample correlation for each dataset is displayed
+up to sampling variance (the sample correlation for each dataset is displayed
 on its panel). The extreme left-tail (respectively, right-tail) dependent 
@@
 \end{figure*}
+<!--***DAN: Shyamolina, you put a negative sign in front of the Spearman correlations
+on the C and D panels, but that is erroneous. The Spearman correlations are positive
+in all four cases! Please fix!-->
 
@@
 to have the same overall association (Spearman's $\rho$ equal to $0.875$,
-up to sampling variation), but different degrees and types of 
+up to sampling variance), but different degrees and types of 
 asymmetry of tail association in different simulations, as follows.
@@
 generate $T$ independent pairs $(u(t),v(t))$ for $t=1,\ldots,T$ with
-$\rho=0.875$ (up to sample variation), but with either extreme left-,
+$\rho=0.875$ (up to sampling variance), but with either extreme left-,
 moderate left-, moderate right-, or extreme right-tail association 
@@
 was considered to have gone extinct if the total population $E(t)+A(t)$ ever 
-went below an extinction threshold (we used a threshold of $2$) within the 
+went below an extinction threshold (we used $2$) within the 
 $25$-time-step time horizon. Extinction risk was the fraction of $10000$
-simulations that went extinct.
+simulations that went extinct. Confidence intervals (95\%) of extinction
+risk were also computed based on a binomial distribution.
 
@@
 consider the portions of the 
-lines $x+y = 2l_b$ and $x+y=2u_b$ which meet with the unit square
+lines $x+y = 2l_b$ and $x+y=2u_b$ which lie within the unit square
 (Fig. \ref{PartialSpearmanFig}). 
-<!--***DAN: need to make this figure-->
+<!--***DAN: Shyamolina, please make this figure. You can start from the code that 
+make the similar figure in BIVAN (Fig. 7), but change it slightly so the figure is not
+*exactly* the same. E.g., you could just switch from Clayton to survival Clayton data,
+and maybe slightly change the locations of the bands. The caption can also be similar,
+but obviously should not mention D^2 or P, and wording should be changed slightly.
+The caption will also have to say something like "This figure was reproduced largely 
+following Fig. 7 of Ghosh et al".-->
 We can then define the *partial Spearman correlation*,
@@
 for $u(t)$ and for $s_e(t)$ because these quantities are 
-related by an increasing monotonic transformation (and likewise for 
-$v(t)$ and $\tilde{f}_a(t)$). Thus, in the sense captured
+related by an increasing monotonic transformation; and likewise for 
+$v(t)$ and $\tilde{f}_a(t)$. Thus, in the sense captured
 by the statistic $\cor_{0,0.5}-\cor_{0.5,1}$, $s_e(t)$
@@
 
-<!--***DAN: Shyamolina, the below commented text makes me wonder if I am misunderstanding something, please
-explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
-of that and what does it have to do with extinction riks? I would have thought extinction risk
-was just the fraction of 10000 sims that went extinct.
-
-after each time step, we simulated model (\ref{eq_age_str}) for $25$ years
-with $10000$ simulations. For each simulation, we started with initial population $(E,A)_{t=0} = (5,10)$
-and following Eq. \ref{eq_age_str} we checked whether after each time step the total population $(E+A)_{t}$
-was below a certain extinction threshold (we set it as $2$) or not. If total population was below that threshold then we 
-considered population went extinct for that simulation. We also calculated $2.5^{th}$ and $97.5^{th}$ quantiles
-of $10000$ simulations with the probability of computed extinction risk at each time step (see [`code`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R)).
--->
-
 <!--density-dependence metapopulation model-->
@@
 of six density dependent population models, the Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, 
-and Austin-Brewer models (Table \ref{tab_model_summary}).
+and Austin-Brewer models.
 All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
@@
 with univariate models like these. These are called period-doubling bifucations.
+
 If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
@@
 independent through time, that were generated to have the same overall associations 
-between all patches and in all simulations (Spearman's $\rho$ again `r srho_common`, up to sampling variation), 
+between all patches and in all simulations (Spearman's $\rho$ again `r srho_common`, up to sampling variance), 
 but with different degrees and types of asymmetry of tail associations 
@@
 for $t=1,\ldots,T$ with $\rho=$ `r srho_common` between $u_i(t)$ and $u_j(t)$ for all $i \neq j$
-(up to sampling variation), but with either extreme left-, moderate left-, moderate right-,
+(up to sampling variance), but with either extreme left-, moderate left-, moderate right-,
 or extreme right-tail association between all pairs of locations (as in Fig. \ref{fig_intro_cop}). 
@@
 
-To calculate extinction risk for a given model and a given choice of tail association,
+To calculate extinction risk for a given metapopulation model, given parameters,
+and a given choice of tail association,
 we simulated the model for $25$ time steps, $10000$ times, using $n=5$ habitat patches.
-The initial condition $P_e$ was used in each habitat patch ($P_e$ is the equilibrium value of the 
-one-patch, deterministic version of the model, Table \ref{tab_model_summary}).
+The equilibrium value of the 
+one-patch, deterministic version of the model, $P_e$ (Table \ref{tab_model_summary})
+was used as the initial condition in all patches and simulations.
 After each time step, if $P_i(t)$ was less than $P_e/10$ then it was set to $0$.
@@
 Extinction risk was the fraction of 10000 metapopulations that went extinct. 
+Estimates of extinction risk were plotted together with 95\% confidence intervals
+based on a binomial distribution.
 
@@
 
-<!--**DAN: got to here 2019 10 29-->
+<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
+To answer (Q2) of the \nameref{Introduction}, 
+we estimated extinction risk for each metapopulation model, using each form of tail 
+association for the environmental noise,
+using values of $r$ that included under- and overcompensatory dynamics,
+as indicated above. We chose values of $r$ within a range, the
+range specified as follows. The second part of Q2 (about how density dependence mediates the
+influence of asymmetric tail associations on extinction risk) was addressed by comparing
+results for under- and overcompensatory model regimes.
+Defining $r_{min}$ as the minimum value of $r$ 
+for which $P_e$ was non-negative; and defining
+$r_{bf}$ as the value of $r$ at which a bifurcation occurs, the range of
+$r$ used spanned 95\% of the interval ($r_{min},r_{bf}$). For details, see Appendix 
+\ref{SM-model_stability}. The range of $r$ encompassed $r_c$, so included 
+a transition from under- to overcompensatory dynamics. For each value of $r$
+used within the range, we calculated extinction risk as indicated above.
+Dispersal parameter values $d= 0,0.1,0.2,..0.5$ were also used in different 
+simulations. We plotted extinction risk as a function of $d$ for two
+fixed values of $r$, one producing under- and one overcompensatory dynamics.
+For $d=0$ we also plotted extinction risk as a function of $r$. All code for 
+this project is available at https://github.com/sghosh89/ERC.
 
-<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
-To answer (Q2) of the \nameref{Introduction} for each metapopulation model,
-we estimated extinction risk of each model using each form of tail 
-association in the environmental noise,
-across the range of $r$ indicated above that spanned under- and overcompensatory dynamics.
-We chose equispaced values of $r$ from the 95\% quantiles 
-of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
-to get a positive, finite equilibrium $P_e$ (for details, see Appendix 
-\ref{SM-model_stability}) to start with (at $t=0$). For each chosen $r$,
-we calculated extinction risk after $25$ years with each metapopulation model 
-where population could disperse locally (or globally) with varying dispersal 
-rate $d= 0,0.1,0.2,..0.5$ and moderate to extreme left 
-tail- and right-tail associated environmental noises with same Kendall's $\rho$ (= `r srho_common`)
-in different simulations (for code see [`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R)).
-
+<!--
 We plotted extinction risks along the y-axis
@@
 or overcompensatory.
+-->
 
@@
 
-<!--Q1 : age-structured results : linear and nonlinear model with asymmetric copula-->
-
-For single patch linear age-structured models, we found higher extinction risk for moderate left-tail 
-association (modeled by Clayton copula) in fluctuating vital 
-rates compared with the one when vital rates were right tail dependent (modeled by Survival Clayton copula).
-It was expected because an extreme "bad" year (e.g. with severe cold) can suppress 
-the fecundity of adult and survival/ maturation of eggs simultaneously 
-and thereby can increase the extinction risk of the population. For non-linear age-structured 
-models, we got a similar trend for extinction risks with the different tail associations (left and right).
-The non-linear model showed higher extinction risk than its corresponding 
-linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
-Fig.
-\ref{fig_extrisk_agestr} represents this scenario when vital rates have moderate to extreme tail association in either tail but with same Kendall's correlation ($\rho =$ `r srho_common`).
-Extinction risks for two different tail-dependent vital rates were not strinkingly different (e.g.
-see pair of red and
-blue lines in Fig. \ref{fig_extrisk_agestr}A or B) and we anticipated the difference being partly 
-dependent on the degree of tail association. Therefore we repeated the same analysis with 
-vital rates coming from extreme-tail dependent (left and right in two different simulations)
-copulas and got similar results but with much prominent difference between their extinction risks
-(Fig. \ref{fig_extrisk_agestr}C or D). This result answered (Q1) and showed extinction risk of 
-a population could be considered as a direct manifestation of tail association property 
-between vital rates of different age-class
-of that population. <!--discussion later: add Tuljapurkar's approximation-->
+<!--Q1 : age-structured model results-->
+For our age-structured, non-spatial model, extinction risks were higher when 
+associations between the vital rates $s_e$ and $f_a$ were principally for low values
+of these vital rates (left-tail association), and were lower when associations
+were principally for higher values; but the effect was minor for this model (Fig. \ref{fig_extrisk_agestr}).
+Though extinction risks were overall higher for the density-dependent version of our model compared
+to density-independent version, extinction risk differences across tail association 
+scenarios were similar for the two versions. Because the same Spearman correlation (up to sampling
+variance) was used in all simulations, the differences between lines in the panels
+of Fig. \ref{fig_extrisk_agestr} are attributable solely to the changes we made in the
+asymmetry of tail associations, and answer Q1. Though differences were significant (confidence
+bands in Fig. \ref{fig_extrisk_agestr}, which are very narrow, did not overlap),
+they were also small, corresponding to only a few percentage points of additional extinction
+risk after 25 years. A potential reason for this outcome is elaborated in the Discussion.
+<!--discussion later: add Tuljapurkar's approximation stuff-->
 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}
-\caption{Extinction risk against time from single patch age-structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(B, D)$ models for $10000$ simulations; shading represents 95\% confidence interval.
-$(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have a moderate left tail (LT) or a moderate right tail (RT) association, $(C, D)$ show the analogous scenario but with the extreme tail association between vital rates.
-\label{fig_extrisk_agestr}}
+\caption{Extinction risk versus time for our stochastic, age-structured model; risks
+were higher when associations between vital rates were stronger in the left compared to
+the right tails of the rates, but only slightly so. Results are for the linear
+version of the model ($\Omega = \infty$; A, C) and for the nonlinear model with density
+dependence in the fecundity rate ($\Omega = 100$; B, D). Results show cases of moderate
+(A, B) and extreme (C, D) asymmetry of tail association.
+The solid line is the estimated extinction risk and the area of shading, which is barely visible because it is so narrow, show 95\% confidence intervals for extinction risk.\label{fig_extrisk_agestr}}
 \end{center}
@@
              caption = "Summary of the six density-dependent stochastic metapopulation models we used.
-The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, where $P_i(t)$ is the population in patch $i$ as time $t$.
+The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, and defines how the model works within each patch.
+Here $P_i(t)$ is the population in patch $i$ at time $t$.
 The growth rate depends on model parameters, $P_i(t)$, and the environment $\\epsilon_i(t)$ in patch $i$ at time $t$.
@@
 $P_e$ is the equilibrium of the deterministic, one-patch version of the model.
+$r_{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative.
 $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text).
-$r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation).
-We use only $r<r_{bf}$.
-Dispersal between patches was also implemented as described in the main text.\\label{tab_model_summary}",
+$r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation; see main text).
+We used $r$ such that $r_{min}<r<r_{bf}$.
+Dispersal between patches was also implemented (see main text).\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"
@@
 ```
-<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)?-->
+<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)? Also, can you add a column for r_min, please?-->
 

### 2019-10-30 MainText.Rmd
response added

@@
 moderate right-tail association, and (D) extreme right-tail association.
-In all cases, Spearman's $\rho$ was the same, $\rho = 0.875$,
+In all cases, Spearman's $\rho$ was the same, $\rho =$ `r srho_common`,
 up to sampling variance (the sample correlation for each dataset is displayed
@@
 on the C and D panels, but that is erroneous. The Spearman correlations are positive
-in all four cases! Please fix!-->
+in all four cases! Please fix!
+***Shya: Yes, you are right, I fixed that.-->
 
@@
 The caption will also have to say something like "This figure was reproduced largely 
-following Fig. 7 of Ghosh et al".-->
+following Fig. 7 of Ghosh et al".
+***Shya: see below.-->
+
+<!--FIG : schematic diagram for partial spearman correlation, adapted from BIVAN-->
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=6 cm]{./Results/PartialSpearmanFig.pdf}
+\caption{The partial Spearman correlation, $\cor_{lb,ub}(u, v)$, within a band can be computed
+for any band to describe how the strength of association between $u$ and $v$
+varies from one part of the two distributions to another. Diagonal lines 
+show two bands, the data in the right/upper band showing
+stronger association than those in the left/lower band. This figure was reproduced
+largely following $Fig. 7$ of $Ghosh et al.(2019)$.
+\label{PartialSpearmanFig}}
+\end{center}
+\end{figure*}
+
 We can then define the *partial Spearman correlation*,
@@
 
+***Shya: I also mentioned Hopf bifurcation in Appendix S2. I leave as it is, when you checked that 
+we need to modify from hopf to period-doubling.
+
 If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
@@
 a time step $25$.
-<!--***DAN: Shyamolina, please check the above two sentences, is this what you did?-->
+<!--***DAN: Shyamolina, please check the above two sentences, is this what you did?
+***Shya: I agreed with you upto line 482, the last line "The whole metapopulation was considered to have gone extinct if all patches were $0$ a time step $25$" I could not understand, did you write that conceptually?
+(then it's ok),
+or you meant I apply that in my code?
+Similarly, the below line explains lines 149-150 of mydispersal.R, I am not fluent in using apply function on two dims at once e.g. c(1,3), you wrote that piece of code, I want to explain you 
+on Friday gow I interpret that portion, then I can tell you accurately what I did.-->
 Extinction risk was the fraction of 10000 metapopulations that went extinct. 
@@
              booktabs=T, linesep = "\\addlinespace"
-             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
+             )%>%column_spec(c(1,3,5,6,7),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
 ```
-<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)? Also, can you add a column for r_min, please?-->
+<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)? Also, can you add a column for r_min, please?
+***Shya: Done, I added r_min value, if a population starts with exactly r=r_min, then P_e remains at 0, it should start with r>r_min to get a positive equilibrium.-->
 

### 2019-10-30 SuppMat.Rmd
response added

@@
 
-#scor_er<-cor(xtc_r,xtc_r,method = "spearman")
-scor_er<- -scor_el # Note it should be opposite to scor_el
+scor_er<-cor(xtc_r,xtc_r,method = "spearman")
+scor_er<- scor_er[1,2]
 
@@
 
-#scor_mr<-cor(mtc_r,mtc_r,method ="spearman")
-scor_mr<- -scor_ml # Note it should be opposite to scor_ml
+scor_mr<-cor(mtc_r,mtc_r,method ="spearman")
+scor_mr<- scor_mr[1,2] 
 ```
@@
 
+par(op)
+dev.off()
+
+#========== Schematic fig for partial spearman correlation ==============
+pdf("./Results/PartialSpearmanFig.pdf", width=6, height=6)
+op<-par(mar=c(6,6,2,2),mgp=c(3,1,0))
+plot(mtc_r[1:150,1],mtc_r[1:150,2],type='p',pch=19,col=rgb(0,0,0,0.15),cex=1.5,
+     cex.axis=2,cex.lab=2.5,xlab="u",ylab="v")
+lines(c(0,0.3),c(0.3,0),type='l')
+lines(c(0,0.6),c(0.6,0),type='l')
+lines(c(0,1.4),c(1.4,0),type='l')
+lines(c(0,1.7),c(1.7,0),type='l')
 par(op)
@@
 
+dt4_2<-tibble("$r_{min}$" = c("0",
+                        "1",
+                        "1",
+                        "$1+\\exp(-a)$",
+                        "0",
+                        "0"
+))
+
 dt5<-tibble("$r_{c}$" = c(rc_ricker,
@@
 
-dt<-cbind(dt1,dt2,dt3,dt4,dt5,dt6)
+dt<-cbind(dt1,dt2,dt3,dt4,dt4_2,dt5,dt6)
 saveRDS(dt,"./Results/tab_model_summary.RDS")

### 2019-10-30 MainText.Rmd
working on results edits

--- 2019-10-30 MainText.Rmd
@@
 author: "Shyamolina Ghosh, Daniel Reuman"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: ""
 geometry: "left=1in,right=1in,top=1in,bottom=1in"
@@
 \noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
+
+\noindent \emph{Correspondence:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
 
@@
 
-<!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->
+<!--Q2 : metapopulation results-->
+For our metapopulation simulations, for low values of $r$, left-tail
+associations of environmental noises accentuated extinction risks, and right-tail associations
+mitigated them; but the reverse was true for high values of $r$ in the range
+$(r_{min},r_{bf})$. Thus 
+density dependent models can either produce similar results, regarding the influence of 
+tail associations on extinction risk, as the density-independent models of @Ghosh_copula, or
+opposite results, depending on the nature of density dependence. 
+For each model of Table \ref{tab_model_summary}, using a value of $r$ on
+the low end of the range $(r_{min},r_{bf})$, extinction risk was highest
+for extreme left-tail associated noise, lowest for extreme right-tail associated
+noise, and was intermediate for the moderate tail association scenarios, for
+both local (Fig. \ref{fig_extrisk_allmodel_localdisp}, left panels) and global 
+(Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}, left panels) dispersal, and for 
+essentially all the dispersal rates, $d$, that we considered. However, 
+using a value of $r$ on the high end of the range $(r_{min},r_{bf})$,
+extinction risks were ordered oppositely (right panels of Figs 
+\ref{fig_extrisk_allmodel_localdisp} and \ref{SM-fig_extrisk_allmodel_globaldisp}). 
+These results help address Q2 from the \nameref{Introduction}.
+Intermediate values of $r$ are shown in Figs. \ref{SM-fig_ricker_risk_local_disp_scl_1} - 
+\ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} for local dispersal and 
+Figs. \ref{SM-fig_ricker_risk_global_disp_scl_1} - 
+\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6} for global dispersal among 
+habitat patches.
 
-We found density-dependent metapopulation models in stochastic 
-environment showed similar patterns for extinction risk as density-independent 
-metapopulation model did in a recent study [@Ghosh_copula]. For each density-dependent 
-model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
-the metapopulations when spatially correlated noises were left-tail associated 
-than the one with the right-tail association. But that observation remained 
-valid when the model parameter - here considered as growth rate $(r)$ - was 
-sufficiently low so that under-compensatory dynamics (monotonous approach towards
-its stable equilibrium) for the deterministic analog of the model was ensured.
-We observed right tail associated noise caused higher extinction 
-risk for the metapopulations than left tail associated noise when the model 
-was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
-value of $r$. Fig. \ref{fig_extrisk_allmodel_localdisp} shows
-variation of extinction risk against local dispersal rate ($d$) for 
-six density dependent metapopulation models (see Table \ref{tab_model_summary}) with growth 
-rate $r$ considered either from under-compensatory or 
-from over-compensatory regime. This finding addresses (Q2) from \nameref{Introduction}. We 
-also got similar results 
-for global dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
-Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
-\ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6}) 
-demonstrate how extinction risk changed against local dispersal rate for a whole range
-of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
-represent the same but for global dispersal among patches. 
-Both case showed the signature of "rescue-effect" (i.e. reduced the extinction risk for 
-finite dispersal compared with $d=0$).
+Because the deterministic, one-patch versions of our models showed undercompensatory
+dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_{min},r_{bf})$),
+and showed overcompensatory dynamics for $r>r_c$, Figs \ref{fig_extrisk_allmodel_localdisp}
+and \ref{SM-fig_extrisk_allmodel_globaldisp} suggest the hypothesis that the
+opposite effects of tail association on extinction risk correspond to the domains
+in which models exhibit under- versus overcompensatory dynamics. This was 
+approximately, but not exactly, true. Careful inspection of Figs 
+\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
+and Table
 
@@
 \includegraphics[width=14.5 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
-\caption{Extinction risks are computed after $25$ years for six nonlinear density-dependence models (Table \ref{tab_model_summary}) with local dispersal among $n=5$ patches and for $10000$ simulations.
-Spatially correlated
-noises have varying degree of tail associations (extreme left tail, moderate left tail, moderate right tail and extreme right tail) as introduced in Fig.
-\ref{fig_intro_cop}.
-Last two columns represent variation of extinction risk against dispersal rate ($d$), calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
+\caption{Extinction risk versus dispersal rate for different scenarios of tail association
+for our six metapopulation models (Methods, Table \ref{tab_model_summary}). Dispersal was
+local, see Fig. \ref{SM-fig_extrisk_allmodel_globaldisp} for global dispersal. Values of $r$
+used are in the panel headers.
+Extinction risks are for a 25-year time horizon. \label{fig_extrisk_allmodel_localdisp}}
 \end{center}

### 2019-10-30 SuppMat.Rmd
working on results edits

--- 2019-10-30 SuppMat.Rmd
@@
 \includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_extreme_and_moderate_noise.pdf}
-\caption[Extinction risk vs.
-dispersal rate for six density dependent metapopulation model with global dispersal.]{Analogous figure as of Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp} in the main text but for global dispersal between metapopulation patches.\label{fig_extrisk_allmodel_globaldisp}}
+\caption[Extinction risk v.
+tail association and $d$: metapopulations, global dispersal]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but using global dispersal between habitat patches.\label{fig_extrisk_allmodel_globaldisp}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of tail association in environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme extreme tail dependent noises for 10000 simulations. \label{fig_ricker_risk_local_disp_scl_1}}
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, local dispersal, Ricker]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Ricker model.\label{fig_ricker_risk_local_disp_scl_1}}
 \end{center}
@@
 \includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of tail association in environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_local_disp_scl_1}}
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, local dispersal, Hassell]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Hassell model.
+\label{fig_hassell_risk_local_disp_scl_1}}
 \end{center}
@@
 \includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of tail association in environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years) for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal;
-standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_msmith_risk_local_disp_scl_0.8}}
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, local dispersal, Maynard Smith]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Maynard Smith model.
+\label{fig_msmith_risk_local_disp_scl_0.8}}
 \end{center}
@@
 \includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of tail association in environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
-standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, local dispersal, Pennycuick]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Pennycuick model.
+\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of tail association in environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
-for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
-standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, local dispersal, Velhurst]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Velhurst model.
+\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Effect of tail association in environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years) for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal;
-standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, local dispersal, Austin-Brewer]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Austin-Brewer model.
+\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Analogous figure as of Fig.
-\ref{fig_ricker_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
-\ref{fig_ricker_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, global dispersal, Ricker]{Analogous to Fig.
+\ref{fig_ricker_risk_local_disp_scl_1} but for global dispersal between habitat patches.
 \label{fig_ricker_risk_global_disp_scl_1}}
@@
 \includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Analogous figure as of Fig.
-\ref{fig_hassell_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
-\ref{fig_hassell_risk_local_disp_scl_1} but for global dispersal between metapopulation patches.
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, global dispersal, Hassell]{Analogous to Fig.
+\ref{fig_hassell_risk_local_disp_scl_1} but for global dispersal between habitat patches.
 \label{fig_hassell_risk_global_disp_scl_1}}
@@
 \includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Analogous figure as of Fig.
-\ref{fig_msmith_risk_local_disp_scl_0.8} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
-\ref{fig_msmith_risk_local_disp_scl_0.8} but for global dispersal between metapopulation patches.
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, global dispersal, Maynard Smith]{Analogous to Fig.
+\ref{fig_msmith_risk_local_disp_scl_0.8} but for global dispersal between habitat patches.
 \label{fig_msmith_risk_global_disp_scl_0.8}}
@@
 \includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Analogous figure as of Fig.
-\ref{fig_risk_local_disp_pennycuick_r_vary_scl_1} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
-\ref{fig_risk_local_disp_pennycuick_r_vary_scl_1} but for global dispersal between metapopulation patches.
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, global dispersal, Pennycuick]{Analogous to Fig.
+\ref{fig_risk_local_disp_pennycuick_r_vary_scl_1} but for global dispersal between habitat patches.
 \label{fig_risk_global_disp_pennycuick_r_vary_scl_1}}
@@
 \includegraphics[width=15 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Analogous figure as of Fig.
-\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
-\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} but for global dispersal between metapopulation patches.
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, global dispersal, Velhurst]{Analogous to Fig.
+\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} but for global dispersal between habitat patches.
 \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6}}
@@
 \includegraphics[width=15 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_extreme_and_moderate_noise.pdf}\\
-\caption[Analogous figure as of Fig.
-\ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} but for global dispersal between metapopulation patches.]{Analogous figure as of Fig.
-\ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} but for global dispersal between metapopulation patches.
+\caption[Extinction risk v.
+tail association, $d$ and $r$: metapopulations, global dispersal, Austin-Brewer]{Analogous to Fig.
+\ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} but for global dispersal between habitat patches.
 \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6}}

### 2019-10-30 MainText.Rmd
further comments on extrisk methods

@@
 \end{figure*}
-<!--***DAN: Shyamolina, you put a negative sign in front of the Spearman correlations
-on the C and D panels, but that is erroneous. The Spearman correlations are positive
-in all four cases! Please fix!
-***Shya: Yes, you are right, I fixed that.-->
 
@@
 (Fig. \ref{PartialSpearmanFig}). 
-<!--***DAN: Shyamolina, please make this figure. You can start from the code that 
-make the similar figure in BIVAN (Fig. 7), but change it slightly so the figure is not
-*exactly* the same. E.g., you could just switch from Clayton to survival Clayton data,
-and maybe slightly change the locations of the bands. The caption can also be similar,
-but obviously should not mention D^2 or P, and wording should be changed slightly.
-The caption will also have to say something like "This figure was reproduced largely 
-following Fig. 7 of Ghosh et al".
-***Shya: see below.-->
-
-<!--FIG : schematic diagram for partial spearman correlation, adapted from BIVAN-->
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=6 cm]{./Results/PartialSpearmanFig.pdf}
-\caption{The partial Spearman correlation, $\cor_{lb,ub}(u, v)$, within a band can be computed
-for any band to describe how the strength of association between $u$ and $v$
-varies from one part of the two distributions to another. Diagonal lines 
-show two bands, the data in the right/upper band showing
-stronger association than those in the left/lower band. This figure was reproduced
-largely following $Fig. 7$ of $Ghosh et al.(2019)$.
-\label{PartialSpearmanFig}}
-\end{center}
-\end{figure*}
-
 We can then define the *partial Spearman correlation*,
@@
 only one of the variables.
+
+<!--FIG : schematic diagram for partial spearman correlation, adapted from BIVAN-->
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=6 cm]{./Results/PartialSpearmanFig.pdf}
+\caption{The partial Spearman correlation, $\cor_{l_b,u_b}(u, v)$, within a band can be computed
+for any band to describe how the strength of association between $u$ and $v$
+varies from one part of the two distributions to another. Diagonal lines 
+show two bands, the data in the right/upper band showing
+stronger association than those in the left/lower band. This figure was reproduced
+largely following Fig. 7 of Ghosh \emph{et al.} (2019).
+\label{PartialSpearmanFig}}
+\end{center}
+\end{figure*}
+<!--***DAN: Shyamolina, please change the u and v axis labels on the figure
+to x and y, respectively, since we defined those as the ranks in the main text.-->
 
@@
 The whole metapopulation was considered to have gone extinct if all patches were $0$
-a time step $25$.
+at time step $25$.
 <!--***DAN: Shyamolina, please check the above two sentences, is this what you did?
@@
 or you meant I apply that in my code?
+***DAN: I wrote it guessing that that was what you did. If that is not what you did,
+when we should discuss, preferably as soon as possible, before Friday.
 Similarly, the below line explains lines 149-150 of mydispersal.R, I am not fluent in using apply function on two dims at once e.g. c(1,3), you wrote that piece of code, I want to explain you 
-on Friday gow I interpret that portion, then I can tell you accurately what I did.-->
+on Friday gow I interpret that portion, then I can tell you accurately what I did.
+***DAN: I added some comments above the function extrisk in mydispersal.R, after reading those
+can you tell me by adding another comment what you did and if what I wrote here was correct?
+-->
 Extinction risk was the fraction of 10000 metapopulations that went extinct. 

### 2019-10-30 mydispersal.R

@@ -145,6 +145,15 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
 }
 
 # function to calculate extinction risk
+#
+#Args
+#sims     A numeric array, assumed to be numsims by numlocs by numsteps
+#
+#Output
+#Produces a vector of length numsteps with the entry in location i being
+#the fraction of simulations which have gone extinct by time step i. 
+#"Extinct" here means all locations have population 0.
+#
 extrisk<-function(sims){
   totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
   return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])

### 2019-10-30 MainText.Rmd
added response

@@
 <!--***DAN: Shyamolina, please change the u and v axis labels on the figure
-to x and y, respectively, since we defined those as the ranks in the main text.-->
+to x and y, respectively, since we defined those as the ranks in the main text.
+***Shya: Done.-->
 
@@
 can you tell me by adding another comment what you did and if what I wrote here was correct?
+***Shya: Yes, you are correct. I also added commented text in mydispersal.R lines 153-165. Have a look 
+to check that you are agreed with me or not.
 -->

### 2019-10-30 SuppMat.Rmd
added response

@@
 plot(mtc_r[1:150,1],mtc_r[1:150,2],type='p',pch=19,col=rgb(0,0,0,0.15),cex=1.5,
-     cex.axis=2,cex.lab=2.5,xlab="u",ylab="v")
+     cex.axis=2,cex.lab=2.5,xlab="x",ylab="y")
 lines(c(0,0.3),c(0.3,0),type='l')

### 2019-10-30 MainText.Rmd
edits to results

--- 2019-10-30 MainText.Rmd
@@
 approximately, but not exactly, true. Careful inspection of Figs 
-\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
-and Table
+\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{fig_risk_global_disp_abrewer_r_vary_scl_0.6}
+and the $r_c$ values of 
+Table \ref{tab_model_summary} reveals that, for the Velhurst and Austin-Brewer models,
+the switch occurs at a value of $r$ less than $r_c$ for these models. For the example
+case of $d=0$, plots of extinction risk versus $r$ (Fig. \ref{fig_extrisk_d0}) seem to indicate that,
+for the Ricker, Hassell, and Pennycuick models, left-tail association accentuates and
+left-tail association mitigates extinction risk for undercompensatory dynamics ($r<r_c$),
+whereas the reverse occurs for overcompensatory dynamics ($r>r_c$). Fig. \ref{fig_extrisk_d0}
+reveals a similar reversal for the Maynard Smith, Velhurst and Austin-Brewer models; but 
+it occurs at a value of $r$ close to but distinguishable from $r_c$.
 
-<!--Q3 : metapopulation results : switching near rc for d=0-->
-Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) one can reasonably hypothesize
-that the switching between the extinction risk 
-generated by left-tail associated and right-tail associated noise 
-occurred near the critical parameter $r_c$. For simplicity, we only considered 
-$d=0$ case, i.e., when there was no dispersal between patches. 
-Fig. \ref{fig_extrisk_d0} demonstrates higher extinction risk 
-with left-tail associated noise for $r<r_c$ and with right-tail associated noise 
-for $r>r_c$. $r_c$ was the vertical line in Fig. \ref{fig_extrisk_d0} 
-indicating the value of $r$ for each model where their deterministic analog switches 
-from under- to over-compensatory dynamics. Though the switching visually appeared far away
-from the theoretical $r_c$ value for Verhulst and Austin-Brewer model, they were 
-consistent with our hypothesis as they had comparatively a smaller range of $r$ than the other models.
+<!--***DAN: stopped here, also need to do the caption of the last figure-->
 
@@
 We used $r$ such that $r_{min}<r<r_{bf}$.
-Dispersal between patches was also implemented (see main text).\\label{tab_model_summary}",
+Dispersal between patches was also implemented (see main text).
+Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_{min}$, $r_c$ and $r_{bf}$.
+See Appendix \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"
@@
 ```
-<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)? Also, can you add a column for r_min, please?
-***Shya: Done, I added r_min value, if a population starts with exactly r=r_min, then P_e remains at 0, it should start with r>r_min to get a positive equilibrium.-->
 

### 2019-10-30 mydispersal.R
changed a comment

@@ -153,11 +153,6 @@ popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
 # A vector of length numsteps+1, where i^th entry represent the fraction of simulations 
 # which have gone extinct by time step i. 
 #"Extinct" here means all locations have population 0.
-
-# DAN wrote this:
-#Produces a vector of length numsteps with the entry in location i being
-#the fraction of simulations which have gone extinct by time step i. 
-#"Extinct" here means all locations have population 0.
 #
 extrisk<-function(sims){
   totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3)) # this totpop is a matrix (dim= numsims by numsteps+1)

### 2019-11-01 MainText.Rmd
working on discussion

--- 2019-10-30 MainText.Rmd
@@
 
-<!--***DAN: maybe you want to have a Discussion para on copulas. It could start with
-some text about ranks being the proper way to study associations, and then could go
-on to link to copulas and cite BIVAN-->
-
 <!--ECEs probably produce tail associations-->
@@
 scattered showers only, which may be much less correlated across locales.
-
-<!--***DAN: May be scope for a discussion para linking what is known about
-how ECEs are changing to what effect that may have on how tail assiciations
-in environmental variables might be changing, and what influence that may have
-on extinction risk. If little is known, then this might be a good topic for 
-future work.-->
 
@@
 wider examination other researchers. Some material in the Discussion outlines some possible next steps.
-<!--***DAN: In Discussion need to talk about what analytic results are needed
-to take the next step of determining when tail depencence may make enough of
-a difference to think about. I think our results, together with Tulja's 
-approximation, suggest that tail associations probably don't make much difference
-for models without a spatial component. But our results illsutaret that it can
-make a big difference when there is a spatial component, apparently because of 
-the influence of tail associations on rescue effects. Could maybe even outline
-analytic approaches for non-dd models? Not sure what would be done with dd
-models. The point is you need to take the next step in the Discussion of 
-talking about how our results, which admitedly are just for specific models,
-could be turned into analytic results or general results. Interestingly, 
-Tulja's approximation may also apply to the spatial case, at least when 
-all locations can be reached from all other locations via dispersal. But the 
-approximation seems poor in that case. Need to think about it more, and maybe
-need more research.-->
 Though tail associations are used
@@
 \end{figure*}
-<!--***DAN: Shyamolina, please change the u and v axis labels on the figure
-to x and y, respectively, since we defined those as the ranks in the main text.-->
 
@@
 at time step $25$.
-<!--***DAN: Shyamolina, please check the above two sentences, is this what you did?
-***Shya: I agreed with you upto line 482, the last line "The whole metapopulation was considered to have gone extinct if all patches were $0$ a time step $25$" I could not understand, did you write that conceptually?
-(then it's ok),
-or you meant I apply that in my code?
-***DAN: I wrote it guessing that that was what you did. If that is not what you did,
-when we should discuss, preferably as soon as possible, before Friday.
-Similarly, the below line explains lines 149-150 of mydispersal.R, I am not fluent in using apply function on two dims at once e.g. c(1,3), you wrote that piece of code, I want to explain you 
-on Friday gow I interpret that portion, then I can tell you accurately what I did.
-***DAN: I added some comments above the function extrisk in mydispersal.R, after reading those
-can you tell me by adding another comment what you did and if what I wrote here was correct?
--->
 Extinction risk was the fraction of 10000 metapopulations that went extinct. 
@@
 based on a binomial distribution.
-
-<!--***DAN: Dan plans to delete this text as it has been superseded by the above. Keeping it
-around until I am sure the above is right:
-Our models (Table \ref{tab_model_summary}) 
-computed extinction risk after $25$ years for $10000$ simulations at $n = 5$ patches
-with spatially correlated noises having either moderate or extreme tail 
-association in left and right-tail. 
-We started with initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ for each patch, where 
-$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
-for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
-We assumed the population went extinct after each time step if it was below a 
-certain threshold (set to 10\% of the initial population $P_e$ for individual model). 
--->
 
@@
 
-<!--***DAN: stopped here, also need to do the caption of the last figure-->
+<!--***DAN: stopped here-->
 
@@
 <!--summary results:-->
-We showed the potential use of the "asymmetric" tail association in modeling the impact of 
-extreme ecological events on population-extinction risk. Our results demonstrated left and right-tail 
-association in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
-in years of extreme events ("catastrophes" and "bonanzas"). Though we could expect the left tail association 
-to show higher extinction risk than the one for right tail association with age-structured 
-and density-dependent metapopulation models, it was not always true. In the over-compensatory regime,
-right tail dependent noise caused higher extinction risk than the left tail dependent noise. We showed 
-that in the absence of dispersal the above-said switching took place near the critical parameter value where 
-the deterministic model showed the transition from under-compensatory to the over-compensatory regime.
+Our results combined with earlier results 
+show that asymmetries in tail associations between environmental variables in
+different locations can have a substantial influence on metapopulation extinction risk,
+for density-dependent as well as for density-independent 
+models. These results
+extend the result of @Ghosh_copula, which was for a density-independent model. For a
+density-independent model, @Ghosh_copula corroborated the intuition
+that, when "harmful" environmental events 
+(i.e., events which reduce the growth rates of local populations)
+are correlated across space,
+extinction risk of metapopulations will be accentuated compared 
+to when "beneficial" events
+(i.e., events which increase the growth of local populations)
+are spatially correlated, even if overall correlation is the same in 
+both scenarios. This was presumably because the first of these scenarios reduces the 
+potential for rescue effects. For the density-dependent metapopulation 
+models we considered, the same 
+result held, typically, but not always, for undercompensatory model dynamics. However,
+for overcompensatory dynamics, the oposite result held, again typically, but not always: 
+when "harmful" events were correlated across space,
+extinction risk of metapopulations was actually mitigated compared 
+to when "beneficial" environmental events were 
+spatially correlated by the same amount. This was probably because the correlated 
+beneficial events caused populations to exceed carrying capacity, and 
+to subsequently crash, across large areas. In contrast 
+to the metapopulation case, our results 
+suggest that asymmetry of tail associations between vital rates
+in a stage-structured, non-spatial stochastic model may be of limited 
+importance for extinction risk. However, our results provide suggestive, rather
+than definitive evidence, here, because we considered 
+only one model. In our stage-structured, non-spatial model, extinction 
+risk was only weakly influenced by asymmetry of tail associations
+between vital rates.
+
+<!--***DAN: May be scope for a discussion para linking what is known about
+how ECEs are changing to what effect that may have on how tail assiciations
+in environmental variables might be changing, and what influence that may have
+on extinction risk. If little is known, then this might be a good topic for 
+future work.-->
+Extensive evidence now exists that ECEs have become more frequent and more intense
+over the last several decades, and substantial progress has recently been made so that
+some event categories and specific events can now be attributed with confidence to
+human-induced climate change [@ummenhofer2017extreme]. We here provide a fuller 
+explanation of how such changes may modify the asymmetry of tail association of 
+weather variables measured in different places, and thereby influence metapopulation
+extinction risk through the mechansms that were explored in this paper.
 
 <!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
+<!--***DAN: Dan to elaborate and make more specific/precise-->
 We found a left tail association in vital rates (induced by ECEs) caused higher extinction risk than 
@@
 left-tail or right-tail association.
+
+<!--***DAN: Dan to maybe add another para on the fact that I thnk Tulja's approx should also
+work for our spatial, non-dd model (check first, Dan), but we see a strong effect here,
+so not clear why. And also suggests the possibilty that for some age-structured models,
+tail associations will make a bigger difference then they did in our case. Maybe somehow
+get to the suggestion that someone needs to revisit Tulja's approx and the analytic
+approximation of extinction risk of Lande and Orzack looking at tail associations? Maybe
+this and the previous para need to be combined, or broken up in a different way.-->
+
+<!--***DAN: maybe you want to have a Discussion para on copulas. It could start with
+some text about ranks being the proper way to study associations, and then could go
+on to link to copulas and cite BIVAN-->
+
+<!--***DAN: In Discussion need to talk about what analytic results are needed
+to take the next step of determining when tail depencence may make enough of
+a difference to think about. I think our results, together with Tulja's 
+approximation, suggest that tail associations probably don't make much difference
+for models without a spatial component. But our results illsutaret that it can
+make a big difference when there is a spatial component, apparently because of 
+the influence of tail associations on rescue effects. Could maybe even outline
+analytic approaches for non-dd models? Not sure what would be done with dd
+models. The point is you need to take the next step in the Discussion of 
+talking about how our results, which admitedly are just for specific models,
+could be turned into analytic results or general results. Interestingly, 
+Tulja's approximation may also apply to the spatial case, at least when 
+all locations can be reached from all other locations via dispersal. But the 
+approximation seems poor in that case. Need to think about it more, and maybe
+need more research.-->
 
@@
 or under-estimation of extinction risks [@ferson1995correlations]. We believe modeling 
-dependencies with copula could help in this case. Second, we did not consider multiple sites of 
+dependencies with copula could help in this case. 
+<!--***DAN: the ideas in this para up to here (taking out the wrong statements) might 
+follow on well from a para based on the notes I wrote above on Tulja's approx.-->
+Second, we did not consider multiple sites of 
 age-structured metapopulation which could be linked via dispersal. So, the possible future investigation 
@@
 survival of different stage class) across sites showed specific tail association 
-with spatially correlated extreme climatic events. Third, we considered the tail association 
+with spatially correlated extreme climatic events. 
+<!--***DAN: a reasonable point, above-->
+Third, we considered the tail association 
 between positively correlated vital rates, but what would be the extinction risk 
 and viability of a population if demographic rates were negatively correlated with the asymmetric tail association. 
+<!--***DAN: a good point, above-->
 We could consider this scenario to model extreme events that kill adults and allow more successful
@@
 to model more complex prey-predator model where asymmetric tail association could feature the
-interspecific-dependence. An earlier study [@carnicer2011widespread] showed extreme events
+interspecific-dependence. 
+<!--***DAN: mentioning interactions strikes me as a step too far-->
+An earlier study [@carnicer2011widespread] showed extreme events
 (i.e., heatwaves and droughts) increased mortality and defoliation in the southern European forest and its 
@@
 multiyear insect outbreak dynamics [@esper20061200].--> We believe copula (maybe a multidimensional one) has 
-a lot more potential to estimate the risk of cascading extinction in model food webs. <!--should I mention about 
+a lot more potential to estimate the risk of cascading extinction in model food webs. <!--should I 
+mention about 
 match-mismatch hypothesis leads to extinction? our phenology paper in prep.--> Fifth, we considered 
@@
 network with different clusters having various dispersal rates.
+<!--***DAN: might also mention how this relates with geog of sync of environmental vars, linking to
+Jon's paper-->
 
@@
 \includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}
-\caption{Extinction risk versus time for our stochastic, age-structured model; risks
+\caption{Extinction risk versus time for our stochastic, age-structured model. Risks
 were higher when associations between vital rates were stronger in the left compared to
@@
 (A, B) and extreme (C, D) asymmetry of tail association.
-The solid line is the estimated extinction risk and the area of shading, which is barely visible because it is so narrow, show 95\% confidence intervals for extinction risk.\label{fig_extrisk_agestr}}
+The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk.\label{fig_extrisk_agestr}}
 \end{center}
@@
 used are in the panel headers.
-Extinction risks are for a 25-year time horizon. \label{fig_extrisk_allmodel_localdisp}}
+Extinction risks are for a 25-year time horizon. Confidence intervals for extinction 
+risk are not plotted, to provide a clearer plot. But 
+because 10000 simulations were
+done for each extinction risk estimate, confidence intervals
+were very narrow, as in Fig. \ref{fig_extrisk_agestr}.\label{fig_extrisk_allmodel_localdisp}}
 \end{center}
@@
 \includegraphics[width=6 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extreme_and_moderate_noise.pdf}\\
-\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
-the metapopulations affected by extreme left tail dependent noise across
-$5$ multiple patches without dispersal ($d=0$), as intrinsic 
-growth rate ($r$) goes from under to over-compensatory regime. Extinction risk for 
-extreme right tail dependent noise follows the opposite trend.
-Error bars represent the standard errors associated with $10000$
-simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
-for each model where their deterministic analog switches from under- to over-compensatory dynamics.
+\caption{Extinction risk versus $r$, with $d=0$, for different scenarios of tail association for 
+our six metapopulation models (Methods, Table \ref{tab_model_summary}). 
+Extinction risks are for a 25-year time horizon. Vertical lines show $r_c$,
+the transition from under- to overcompensatory dynamics of the deterministic, one-patch
+version of the model. Error bars show 95\% confidence intervals of extinction risk.
 \label{fig_extrisk_d0}}

### 2019-11-01 MainText.Rmd
done with discussion for now

@@
 used population models. We intentionally use a simulation approach and simple models
-instead of attempting mathematical analysis and/or using models that are more realistic, 
+for both Q1 and Q2 instead of attempting mathematical 
+analysis and/or using models that are more realistic, 
 because our intention here is to provide an initial exploration only of 
@@
 By so doing we hope to open these potentially practically important questions to 
-wider examination other researchers. Some material in the Discussion outlines some possible next steps.
+wider examination other researchers. Some material in the Discussion 
+outlines some possible next steps.
+<!--***DAN: make sure you do this-->
 Though tail associations are used
@@
 
-<!--***DAN: stopped here-->
-
 # Discussion\label{Discussion}
@@
 in a stage-structured, non-spatial stochastic model may be of limited 
-importance for extinction risk. However, our results provide suggestive, rather
+importance for extinction risk. In our stage-structured, non-spatial model, extinction 
+risk was only weakly influenced by asymmetry of tail associations
+between vital rates. Our results provide suggestive, rather
 than definitive evidence, here, because we considered 
-only one model. In our stage-structured, non-spatial model, extinction 
-risk was only weakly influenced by asymmetry of tail associations
-between vital rates.
+only one model (but see below).
 
-<!--***DAN: May be scope for a discussion para linking what is known about
-how ECEs are changing to what effect that may have on how tail assiciations
-in environmental variables might be changing, and what influence that may have
-on extinction risk. If little is known, then this might be a good topic for 
-future work.-->
+<!--linking back to ECEs-->
 Extensive evidence now exists that ECEs have become more frequent and more intense
 over the last several decades, and substantial progress has recently been made so that
-some event categories and specific events can now be attributed with confidence to
+some event categories and even specific events can now be attributed to
 human-induced climate change [@ummenhofer2017extreme]. We here provide a fuller 
@@
 weather variables measured in different places, and thereby influence metapopulation
-extinction risk through the mechansms that were explored in this paper.
+extinction risk through the mechansms we explored in this paper. 
+If $X_i(t)$ denotes the value of an environmental variable (e.g., a summer maximum 
+temperature, or winter minimum temperature, spring rainfall, etc.) 
+in location $i$ in year $t$,
+@ummenhofer2017extreme provide a diagram of how ECEs may be increasing in frequency 
+and intensity due to changes in the distribution of $X_i(t)$ such as changes
+in the mean, variance, or skewness of this distribution. Although changes in 
+the correlation between $X_i(t)$ and $X_j(t)$ are increasingly documented, 
+and may also be a consequence of climate change (e.g., @Black2018),
+we are unaware of any studies examining the extent to which tail associations
+between $X_i(t)$ and $X_j(t)$, and asymmetries of tail association, may be 
+changing. Documented changes in correlation between $X_i(t)$ and $X_j(t)$
+may be due partly or wholly to changes in tail associations. If ECEs are becoming
+more spatially extensive, as seems likely given that they are becoming
+more intense [@ummenhofer2017extreme], then the conditional probability that both of
+$X_i(t)$ and $X_j(t)$ are influenced by an ECE given that one of them is so
+influenced should be increasing. This should contribute to increased association
+of the two variables in one of their tails (which tail depends on the type of 
+ECE and what the variable $X$ is). A study of changes through time in 
+tail associations of weather variables would be straightforward and 
+potentially important given the mechanisms we have explored.
+<!--***DAN: Shyamolina, another aspect of an exploration you may do on the 
+the geography of tail association may be to look at how tail association 
+may be changing. This relates to ECEs, so may help better justify the work
+to a funder. You had proposed to look at geography of tail association in 
+NDVI, and that makes good sense. You could also look at how tail association
+is changing in NDVI (the records are about 35 years long, so you could just
+about do this). Even better (because data are better), you could get weather
+station records (lots of these can be downloaded, over 1000 stations
+have been operating for over a century continuously in North America) and 
+look at how tail associations are changing in those data.-->
 
-<!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
-<!--***DAN: Dan to elaborate and make more specific/precise-->
-We found a left tail association in vital rates (induced by ECEs) caused higher extinction risk than 
-the right tail dependent ones, but the difference was not remarkable 
-unless the tail association was at its 
-extreme (Fig. \ref{fig_extrisk_agestr}). This finding could be explained 
-by the approximation @tuljapurkar1990lecture made for the 
-stochastic log(growth rate) in an iid environment.
-The approximation considered growth rate being 
-dependent on the most dominant eigenvalue of the mean projection matrix 
-and the temporal variance of the log(population growth rate) caused 
-by environmental stochasticity - which was based
-on the linear covariance measure between any 
-two matrix elements, not on their tail association properties. Therefore,
-following @tuljapurkar1990lecture, the population growth 
-rate in an iid environment could be approximated
-as same for our model (Eq. \ref{eq_age_str}) with 
-left-tail or right-tail association.
+<!--copulas and rank methods-->
+It is known that the distributions $X_i(t)$ are changing, both in
+mean and variance, and possibly in other ways, depending on what
+environmental variable $X$ represents.
+Studying associations between the values of an environmental variable in different
+locations, $X_i(t)$ and $X_j(t)$, using standard methods such as Pearson correlation
+may be complicated by these changes; rank-based and related 
+measures of association such as 
+we have used are recommended instead.
+If $f$ is a strictly monotonic, increasing transformation, then the Spearman
+(or Kendall) correlation of $f(X_i(t))$ and $f(X_j(t))$ is the same as the
+Spearman (or Kendall) correlation of $X_i(t)$ and $X_j(t)$. However, the Pearson
+correlation of $f(X_i(t))$ and $f(X_j(t))$ may differ substantially
+from the Pearson correlation of $X_i(t)$ and $X_j(t)$. 
+Relatedly, if the marginal distributions of $X_i(t)$ and $X_j(t)$ are
+changing through time, it may also cause changes through time in the 
+Pearson correlation of these quantities, even while the values
+of rank-based correlations such as Spearman and Kendall may be unchanging.
+Essentially, in such a case, associations between $X_i(t)$ and $X_j(t)$
+would be unchanging through time even through Pearson correlation
+appears to inducate a change in association, which is instead a change 
+in marginal distributions.
+Perhaps unfortunately, the Pearson correlation,
+which is the most commonly used measure of association of two variables,
+depends on both marginal information and association information.
+It is possible that recently observed changes in environmental
+(e.g., @Black2018) and biological (e.g., @Koenig2016) synchrony 
+are, at least in part, just another feature of 
+already-studied changes in distributional characteristics of  
+variables, rather than reflecting true changes in association.
+Rank-based methods are linked to a well-developed suite of statistical
+approaches related to the *copula* concept of statistics. Copulas and related ideas
+can be used to formally separate the information in the joint 
+distribution of $X_i$ and $X_j$ into the information that pertains
+to the marginal distributions of these two variables, and the information
+that pertains solely to their association. 
+@Ghosh_copula argued generally for the potential of copulas and related ideas
+to improve our understanding of ecology 
+(see also @Popovic2018, @Valpine2014, @Anderson2018).
+Rank-based methods such as the Spearman and partial Spearman correlations
+which we have used are closely related to copula approaches, and @Ghosh_copula,
+@Genest2007 and others have recommended copula and rank approaches when the 
+goal is to understand associations between variables, and how such 
+associations may be changing.
 
-<!--***DAN: Dan to maybe add another para on the fact that I thnk Tulja's approx should also
+<!--Tuljapurkar approximation-->
+Our results suggest that tail associations and asymmetries of tail associations
+may have limited influence on the extinction risk of 
+stage-structured, non-spatial models. Although we believe more research is 
+certainly still needed
+as to whether this conclusion holds more generally than the model we considered
+(see below), if it does hold 
+generally, it may be possible to obtain a better understanding of 
+the result using a well-known approximation
+of Tuljapurkar [@tulja1982; @tulja1990] for the long-term stochastic growth rate.
+The long-term stochastic growth rate is an important quantity in demography
+[@Caswell2000] and population viability analysis for conservation 
+[@MorrisDoak2002], and it factors prominently into
+extintion risk estimates for density-independent age- or 
+stage-structured population models 
+[@Lande1988], $\vec{P}(t+1)=M(t)\vec{P}(t)$. Here $\vec{P}(t)$ is the 
+age- or stage-structured population and $M$ is a matrix-valued stationary stochastic 
+process. The matrix entries of $M$ can correspond, for instance, to the fluctuating
+fecundity and survival rates of the life stages of the population. 
+For such a model, Tuljapurkar's approximation expresses the long-term stochastic 
+growth rate as a function of the variances and covariances of the entries
+of $M$ and of the eigenvalues and eigenvectors of the expected value of 
+$M$. None of these quantities depends on tail associations between matrix entries.
+Thus, whenever Tuljapurkar's approximation is a good
+approximation, the long-term stochastic growth rate should not depend sensitively
+on tail association. @Lande2003 provides reasoning supporting a claim that 
+Tuljapurkar's approximation
+is "likely to be accurate in many cases," (p. 60) though we are unaware of the extent
+to which accuracy of the approximation has been tested for scenarios with
+asymmetric tail associations.
+We also note that this reasoning pertains only to density-independent models,
+and says nothing about our results for the density-dependent version of our
+stage-structured model.
+
+<!--***DAN: decided the below was a step too far
+Finally, it seems likely that the 
+stochastic matrix modelling framework, and Tuljapurkar's approximation,
+should apply to the density-independent metapopulation model of 
+@Ghosh_copula, which was comprised of Lewontin-Cohen models linked 
+by dispersal. Nevertheless asymmetry of tail associations made a 
+big difference to extinction risk for that model [@Ghosh_copula].
+We have not examined whether Tuljapurkar's approximation is
+accurate in that spatially explicit case.
+
+***DAN: Dan to maybe add another para on the fact that I thnk Tulja's approx should also
 work for our spatial, non-dd model (check first, Dan), but we see a strong effect here,
@@
 get to the suggestion that someone needs to revisit Tulja's approx and the analytic
-approximation of extinction risk of Lande and Orzack looking at tail associations? Maybe
-this and the previous para need to be combined, or broken up in a different way.-->
+approximation of extinction risk of Lande and Orzack looking at tail associations?
 
-<!--***DAN: maybe you want to have a Discussion para on copulas. It could start with
-some text about ranks being the proper way to study associations, and then could go
-on to link to copulas and cite BIVAN-->
-
-<!--***DAN: In Discussion need to talk about what analytic results are needed
+***DAN: In Discussion need to talk about what analytic results are needed
 to take the next step of determining when tail depencence may make enough of
@@
 
-<!--Intution/ possible explanation for expected results/ hypo?-->
-
 <!--future study: worthy for PVA, IUCN red list:
@@
 -->
-Although we demonstrated that tail association between vital rates of an age-structured model can
-influence extinction risk, substantial work remains to determine the importance of this effect in the 
-following scenarios. First, we demonstrated the effect of the tail association on extinction risk 
-with a basic age-structured (two age stages) model where we considered the total semelparous 
-population would go extinct if there would be only one female adult and one egg. But in reality population 
-viability analysis (PVA) deals with a little bit higher threshold with several age, stage or size class and a variety 
-of populations including semelparous and iteroparous [@morris2002quantitative]. Future studies should include 
-tail association effect on extinction risk 
-for a multiple age-, size- and stage-structured model and estimate extinction risk 
-for each age, size and stage class separately 
-so that we can identify (via multisite-PVA) which life-stage or demographic process 
-to target for population management. An earlier study by @doak1994modeling showed PVA models with correlated 
-demographic rates indicated a marked difference in predicting confidence interval for 
-future population size compared with the model which did not include correlation. But when 
-such correlation cannot be estimated directly, more generally where the
-forms of statistical dependence among the model's parameters are unknown, making incorrect
-assumptions about correlations and dependencies may lead to severe over-estimation 
-or under-estimation of extinction risks [@ferson1995correlations]. We believe modeling 
-dependencies with copula could help in this case. 
-<!--***DAN: the ideas in this para up to here (taking out the wrong statements) might 
-follow on well from a para based on the notes I wrote above on Tulja's approx.-->
-Second, we did not consider multiple sites of 
-age-structured metapopulation which could be linked via dispersal. So, the possible future investigation 
-would be to explore the viability for the metapopulations if vital rates (e.g. fecundity or 
-survival of different stage class) across sites showed specific tail association 
-with spatially correlated extreme climatic events. 
-<!--***DAN: a reasonable point, above-->
-Third, we considered the tail association 
-between positively correlated vital rates, but what would be the extinction risk 
-and viability of a population if demographic rates were negatively correlated with the asymmetric tail association. 
-<!--***DAN: a good point, above-->
-We could consider this scenario to model extreme events that kill adults and allow more successful
-recruitment of seeds or other propagules. Fourth, following the aforementioned idea, it would be worthy 
-to model more complex prey-predator model where asymmetric tail association could feature the
-interspecific-dependence. 
-<!--***DAN: mentioning interactions strikes me as a step too far-->
-An earlier study [@carnicer2011widespread] showed extreme events
-(i.e., heatwaves and droughts) increased mortality and defoliation in the southern European forest and its 
-dependent herbivorous insects and fungi through trophic linkage. <!--ECEs can even truncate 
-multiyear insect outbreak dynamics [@esper20061200].--> We believe copula (maybe a multidimensional one) has 
-a lot more potential to estimate the risk of cascading extinction in model food webs. <!--should I 
-mention about 
-match-mismatch hypothesis leads to extinction? our phenology paper in prep.--> Fifth, we considered 
-our metapopulation patches were equispaced on a linear chain with equal global dispersal 
-rate between patches. But in realistic network models, as inter-patch distances 
-can be different, the dispersal rate should depend on distance: dispersal is possible at higher 
-rates between nearby patches than the patches which are far away. It is worthy to study 
-metapopulation viability with non-uniform spatial correlation (i.e. asymmetric tail association) in a 
-network with different clusters having various dispersal rates.
-<!--***DAN: might also mention how this relates with geog of sync of environmental vars, linking to
-Jon's paper-->
+
+<!--future ideas-->
+In our opinion, additional research would be worthwhile on the 
+impact of tail associations on growth rates and 
+extinction risk under additional modelling scenarios not considered
+in this paper.   
+First, we considered a model with positive association between 
+egg and adult vital rates,
+corresponding to similar impacts of an environmental variable on 
+eggs and adults. Negative associations between vital rates
+can occur as well, relating to
+life history trade-offs, e.g. between survival and reproduction 
+(e.g, @Lande2003, p. 59). Asymmetric tail associations may occur
+in such a context if, for instance, there is a three way trade-off
+among vital rates $r_1$, $r_2$, and $r_3$. Values of $r_1$ near
+an upper limit may be well be associated with,
+for instance, $r_2$, whereas lower values of $r_1$ may not be strongly
+associated with $r_2$ because reductions/increases in $r_1$ can
+then be traded off against increases/reductions in *either* 
+$r_2$ or $r_3$. Second, for our metapopulation models,
+we assumed habitat patches were equispaced on a transect or
+were all equally mutually accessible through dispersal.
+Furthermore, associations between environmental variables in different
+patches were assumed to be spatially homogeneous. In real metapopulations,
+neither of these assumptions holds. Connecting the ideas of tail association that
+we have developed in this study with the ideas on the geography of synchrony
+of @walteretal2017 seems like a promising future direction.
 

### 2019-11-03 MainText.Rmd
a bit of progress on the abstract

@@
 \noindent \textbf{ABSTRACT} 
-<!--200 words limit : J Am Nat-->
+<!--350 words limit : Ecosphere-->
 
-<!--***DAN: Dan has not edited this--> 
+\noindent Extreme climatic events are becoming more frequent and more intense
+due to climate change. Furthermore, there is reason to believe extreme 
+climatic events may modify "tail associations" between values of an environmental 
+variable measured in different locations, i.e., associations that 
+occur principally between values in the left or right tails of the distributions 
+of values of the variable that can occur in the individual locations. 
+Two positively associated variables can be "left-tail associated" (i.e., more 
+correlated when they take low than high values) or "right-tail associated"
+(more correlated when they take high than low values) even with the same
+overall correlation coefficient in both cases. We tested whether the nature of 
+spatial tail associations of environmental variables can influence metapopulation 
+extinction risk. For instance, if low values of a variable are bad for local
+populations, one may expect that left-tail associations increase
+metapopulation extinction risks because then environmental catastrophes are
+spatially synchronized, presumably reducing the potential for rescue effects.
+We also tested, in the context of (non-spatial) stage-structured matrix models, 
+whether tail associations between values of stage-specific vital rates may
+influence extinction risk.
+
+Tail associations and their asymmetries are taken into account in assessing risks
+in finance and other fields, but to our knowledge our study is one of the first
+to consider how tail associations influence extinction risk.
+
 With global climate change, extreme climatic events (ECEs) become more frequent.
@@
 
-\textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model
+\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model
 
@@
 By so doing we hope to open these potentially practically important questions to 
-wider examination other researchers. Some material in the Discussion 
-outlines some possible next steps.
-<!--***DAN: make sure you do this-->
-Though tail associations are used
+wider examination by other researchers. Though tail associations are used
 to assess risks in fields such as finance and hydrology 

### 2019-11-03 MainText.Rmd
edits

@@
 Extinction risk is also very commonly estimated for age-structured models 
-such as stochastic matrix models [@MorrisDoak2002; @Caswell2006] and 
+such as stochastic matrix models [@MorrisDoak2002; @Caswell2000] and 
 integral projection models
@@
 approximately, but not exactly, true. Careful inspection of Figs 
-\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{fig_risk_global_disp_abrewer_r_vary_scl_0.6}
+\ref{SM-fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{SM-fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
 and the $r_c$ values of 
@@
 It is possible that recently observed changes in environmental
-(e.g., @Black2018) and biological (e.g., @Koenig2016) synchrony 
+[e.g., @Black2018] and biological [e.g., @Koenig2016] synchrony 
 are, at least in part, just another feature of 
@@
 to improve our understanding of ecology 
-(see also @Popovic2018, @Valpine2014, @Anderson2018).
+(see also @Popovic2018; @Valpine2014; @Anderson2018).
 Rank-based methods such as the Spearman and partial Spearman correlations

### 2019-11-03 MainText.Rmd
edited the abstract

--- 2019-11-03 MainText.Rmd
@@
 variable measured in different locations, i.e., associations that 
-occur principally between values in the left or right tails of the distributions 
-of values of the variable that can occur in the individual locations. 
-Two positively associated variables can be "left-tail associated" (i.e., more 
+occur between values in the left or right tails of the distributions 
+of the variable that occur in the locations. 
+Two positively associated variables can be principally "left-tail associated" (i.e., more 
 correlated when they take low than high values) or "right-tail associated"
-(more correlated when they take high than low values) even with the same
+(more correlated when they take high than low values), even with the same
 overall correlation coefficient in both cases. We tested whether the nature of 
 spatial tail associations of environmental variables can influence metapopulation 
-extinction risk. For instance, if low values of a variable are bad for local
+extinction risk. For instance, if low values of a variable reduce the growth 
+rates of local
 populations, one may expect that left-tail associations increase
@@
 spatially synchronized, presumably reducing the potential for rescue effects.
-We also tested, in the context of (non-spatial) stage-structured matrix models, 
-whether tail associations between values of stage-specific vital rates may
+We also tested, in the context of non-spatial stage-structured matrix models, 
+whether tail associations between stage-specific vital rates may
 influence extinction risk. 
-
+We showed that density dependence interacts with the
+propensity for tail associations to influence metapopulation extinction risk: for 
+population models
+showing undercompensatory density dependence, left-tail associations in environmental 
+variables typically accentuated
+and right-tail associations mitigated extinction risk; whereas the 
+reverse was usually true for models showing overcompensatory density dependence.
+For the non-spatial, stage structured models we considered, left-tail associations 
+between vital rates did accentuate extinction risk compared to right-tail
+associations, but the effect was small. 
 Tail associations and their asymmetries are taken into account in assessing risks
 in finance and other fields, but to our knowledge our study is one of the first
-to consider how tail associations influence extinction risk.
+to consider how tail associations influence population extinction risk. 
+Our modelling results 
+provide an initial demonstration of a new mechanism influencing extinction risks,
+and, in our view, should help motivate more comprehensive
+study of the mechanism and its importance for real populations in future work.
 
-With global climate change, extreme climatic events (ECEs) become more frequent.
-These ECEs could pose a severe impact on the resistance or resilience of a population and/ or community level through life-history traits of the species.
-Therefore predicting the impact of extreme events on population extinction risk is necessary.
-We here introduce the potential use of "tail association" property of two correlated ecological variables in modeling the extinction risk of an age-structured single patch model.
-We also extend this idea to density-dependent metapopulation models.
-For metapopulation models, we show extinction risk not only depends on the nature of the tail association of spatially correlated noise but is also influenced by the inherent dynamics of the population-model.
-To elaborate, we demonstrate left tail dependent noise across spatially correlated patches cause higher extinction risk than the right tail dependent noise when the model show undercompensatory dynamics.
-But in the overcompensatory regime, the trend is the opposite.
-Our study paves the path for future research considering this tail association instead of a simple correlation in population viability analysis.
-
-\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model
+\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model, copula
 

### 2019-11-03 MainText.Rmd
working on sup mat

@@
 It is possible that recently observed changes in environmental
-[e.g., @Black2018] and biological [e.g., @Koenig2016] synchrony 
+(e.g., @Black2018) and biological (e.g., @Koenig2016) synchrony 
 are, at least in part, just another feature of 

### 2019-11-03 SuppMat.Rmd
edits to results
working on discussion
working on sup mat

@@
                         "1",
-                        "$1+\\exp(-a)$",
+                        "1.90",
                         "0",
@@
 
-<!--needs editing, I just dumped old text here-->
-
-Copula, a rank based measure of association, is much more informative 
-than linear correlation measure and a widely applied tool in dependence 
-modeling [@joe2014_dependence; @nelsen2006_copula]. Copulas separate 
-the information content of a bivariate dataset,
-$(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
-distributions (which is not about the association between the variables) and the rest of 
-the information (which is solely about the association). The association information is revealed by 
-the copula plot ($v_n$ against $u_n$) of the raw variables ($y_n$ vs. $x_n$) in Fig...., where 
-$u_n$ is the rank of $x_n$ in the set ${x_1, x_2, \ldots , x_n }$, divided by $N + 1$; 
-and $v_n$ is the rank of $y_n$ in the set ${y_1, y_2, \ldots , y_n }$, also divided
-by $N + 1$. Here the rank of the smallest element of each sets is considered to be 1 and the rank of
-the largest element is $N$. We refer to the $u_n$ and $v_n$ as normalized ranks of the $x_n$ and $y_n$. 
-Ranking makes the marginal distributions uniform, isolating only the information on 
-association between the variables.
-
 <!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
-
-For extreme tail association algorithm, a $N \times n$ matrix, M was generated of independent 
-standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->
-Then, each row of M was identified for a first set of modifications
-randomly and independently with a 50\% chance; for those rows identified, every
-element of the row was replaced with the absolute
-value of the first element of the row (including the first element itself).
-For other rows, every element in the row was replaced by the negative of 
-its own absolute value. This produces right-tail dependent
-matrix, normally distributed for each column. If left-tail 
-dependent noise is desired, take the negative of the entire matrix.
-Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
-generating extreme tail association matrix is flexible enough, so that 
-matrix M can be generated initially with independent random draws from a normal 
-distribution with a specified mean and standard deviation.
-
-<!--some stuff that used to be in the main text-->
-
-The moderate tail association was modeled by using the 
-concept of "copula", a statistical tool to model 
-dependencies. Appendix \ref{SM-intro_copula} summarized
-a brief introduction of "copulas" and their relation to tail associations, symmetries 
-with examples of available copula families 
-from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
-copula to model moderate left-tail and Survival Clayton copula to model 
-moderate right-tail association in the vital rates. Both copulae had the same Spearman 
-correlation ($\rho =$ `r srho_common`) but opposite association in their respective tails. 
-We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
-right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
-`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
-by number of simulations ($=10000$) we used. 
-We chose parameters for each copula family using 
-`iRho` function from `copula` package so that both copula had
-same Spearman correlation ($\rho =$ `r srho_common`). Entries within each column of
-matrix $M$, then became independent of each other but 
-entries between columns had a specific tail association
-with positive Spearman correlation.
-
-<!--some more stuff from the main text-->
-
-On a similar note, we generated extreme left- (or right-) tail dependent matrix (M)
-from standard iid (Appendix \ref{SM-extr_tail_dep}) and made the marginals uniform so that
-it could mimic an extreme tail dependent copula. We made the first column of matrix 
-$M$ as gamma ($\Gamma(k,\theta)$) distributed 
-with shape parameter $k = 2$ and scale parameter $\theta = 1$.
-This distribution represented fecundity of the adult population, $f_a \geq 0$.
-For a given $\theta$, increasing $k$ would shift the distribution towards
-higher fecundity and reduce the extinction risk if the other parameters
-remain unchanged. We took the second column of $M$ and applied a 
-beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
-$\beta = 2$ were the shape parameters that could control the distribution 
-of the survival rates of eggs ($s_e$) ranging in between 0 to 1. 
-For each $25$ timesteps from each of $10000$ simulations,
-we incorporated pairwise entries from each row of matrix $M$ in the expression for 
-vital rates (fecundity $f_a$ and survival rates $s_e$, Eq. \ref{eq_vitalrates}) of 
-the stochastic projrction matrix in model (\ref{eq_age_str}).
-
-<!--yet more stuff from the main text-->
-
-For moderately left-tail associated 
-noises we drew points from a $n$ dimensional 
-Clayton copula with same Spearman correlation ($\rho =$ `r srho_common`) between
-any two out of $n$ patches and made 
-the marginals normal with zero mean and standard deviation $\sigma$ 
-(see the specified value $\sigma$ for each model in 3$^{rd}$ 
-column of Table \ref{tab_model_summary}). For moderately 
-right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
-from Clayton copula so that all the statistical properties remained 
-same except for the direction of tail association. We generated
-extreme left- or right-tail associated noises via algorithm stated in 
-Appendix \ref{SM-extr_tail_dep}.
+Bivarite environmental noise with extreme left-tail association (as in Fig. 
+\ref{fig_intro_cop}A) was generated using an algorithm originally presented 
+by @Ghosh_copula (their Appendix S10), presented again here for convenience.
+If $T$ time steps of noise were desired, step 1 was to generate a $T$ by $2$
+matrix, $M$, of independent random numbers, each standard-normally distributed. 
+Then, each row of $M$ was identified for modification randomly and
+independently, with a 50\% chance. For those rows identified, every element
+of the row was replaced by the negative of the absolute value of the first element
+of the row, including the first element itself. For the other rows, every 
+element in the row was replaced by its own absolute value. For noise
+with extreme right-tail association (as in Fig. 
+\ref{fig_intro_cop}D), take the negative of the entire matrix. For
+both types of noise, every entry was transformed by the cummulative distribution
+function of a standard normal distribution, to make each column uniformly
+distributed.
+
+<!--Spearman correlation of the extreme-tail-associated noise-->
+The Spearman correlation of the extreme left- or right-tail associated 
+noise was $0.875$, up to sampling variation.
+<!--***DAN: fill this in-->
+
+<!--moderate tail association noise-->
+Bivarite environmental noise with moderate left-tail association (as in Fig. 
+\ref{fig_intro_cop}B) was generated using a 2-dimensional Clayton copula. The statistical
+theory of copulas has been introduced systematically in several texts 
+(e.g., @nelsen2006_copula; @joe2014_dependence), and was briefly 
+introduced by @Genest2007. Short, conceptual introductions aimed at ecological
+audiences were provided by sections within the papers of @Anderson2018 and  
+@Ghosh_copula. A 2-dimensional Clayton copula is equivalent to a particular bivariate 
+probability density function with marginal distributions which are 
+uniform on the interval $(0,1)$. The points on Fig. 
+\ref{fig_intro_cop}B were independent draws from the associated random variable.
+The functions `claytonCopula` and `rCopula` from the `copula` package for the 
+R programming language were used. The parameter of the Clayton family was
+chosen so that the Spearman correlation was $0.875$, the same as that of the
+extreme left-tail associated noise. This was accomplished with the `iRho`
+function of the `copula` package. Noise with moderate right-tail
+association was generated in a similar way but using the 2-dimensional 
+survival Clayton copula.
+
+<!--Multivariate extreme tail association-->
+Multivariate noise with extreme left-tail association was generated using a
+very simlar algorithm to the bivariate case,
+as follows. If $T$ time steps of noise of dimension $n$ were desired,
+step 1 was to generate a $T$ by $n$
+matrix, $M$, of independent random numbers, each standard-normally distributed.
+Each row of $M$ was identified for modification randomly and
+independently, with a 50\% chance. For those rows identified, every element
+of the row was replaced by the negative of the absolute value of the first element
+of the row, including the first element itself. For the other rows, every 
+element in the row was replaced by its own absolute value. For noise
+with extreme right-tail association, take the negative of the entire matrix. For
+both types of noise, every entry was transformed by the cummulative distribution
+function of a standard normal distribution, to make each column uniformly
+distributed. 
+Multivariate moderately tail associated noise of dimension $n$
+was generated using $n$-dimensional Clayton or survival Clayton copulas.
 
@@
 
-First, we will briefly discuss about the stability analysis for a difference 
-equation which is frequently used in population ecology model. Let's say, $P(t+1) = F(P(t),r)$
-represents the model dynamics of a single population $P$ which progress recursively via 
-a nonlinear density-dependence function $F$ with discrete time variable $t$. $r$ is the model parameter 
-which plays major role in assessing the model's dynamical stability. Fixed-point analysis can
-be used to get information about this type of nonlinear difference equation. The idea is to
-employ the concept that the system at fixed point (or equilibrium $P_e$) does not change with time 
-(i.e. $P(t+1) = P(t)$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
-returns to its initial state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
-Mathematically, from standard linear stability analysis we can say an equilibrium is locally 
-asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P(t)=P_e}$.
-If the approach towards its equilibrium is monotonic (when $0<{F}'<1$), the dynamics is said to be
-in under-compensatory regime. An over-compensatory dynamics means the approach towards its stable equlibrium
-is oscillatory ($-1<{F}'<0$). Following we will describe stability condition of the model dynamics considered for each 
-patch of metapopulation models in the maintext (see Table \ref{MT-tab_model_summary}).
+Let $P(t+1) = F(P(t),r)$ be a deterministic population-dynamic model, where 
+$P(t)$ is the population at time $t$ and $r$ is a model parameter. Let $P_e$
+be an equilibrium (which may depend on $r$), i.e., $F(P_e,r)=P_e$. $P_e$
+is called a stable equilibrium if, when the population is perturbed slightly
+from equilibrium, the system returns to the equilibrium value; $P_e$ is called
+an unstable equilibrium otherwise.
+From standard linear stability analysis, we say an equilibrium is locally 
+asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P=P_e}$.
+If $0<{F}'<1$, then the approach of the system to the equilibrium is monotonic,
+and dynamics are called undercompensatory. If $-1<{F}'<0$, then the approach of the
+system to equilibrium is oscillatory, and dynamics are called overcompensatory. 
+We will now describe stability analyses of the deterministic 1-habitat-patch version
+of each model of Table \ref{MT-tab_model_summary}.
+
+<!--***DAN: got to here-->
 

### 2019-11-04 SuppMat.Rmd
done with a round of edits for the whole thing

@@
 
-<!--***DAN: got to here-->
-
-**Ricker model** [@ricker1954stock; @may1976bifurcations], originally introduced in the context of modelling
-stock and recruitment of salmon populations, is a classic 
-discrete population model 
-and frequently used in fisheries management problem. The model dynamics
-is represented as 
-$P(t+1)=F(P(t),r,K)=P(t) \exp[r(1-P(t)/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
-growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
-The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ for a positive $r$ and considered 
-as the extinction equilibrium 
-in population ecology context. Depending on parameter $r$, 
-the other equilibrium $P_e = K$ can be stable or unstable as ${F}' = (1-r)$. Following generalized
-concept, the model can show monotonic approach to its stable equilibrium $P_e = K$ if $0<r<1$ 
-or can approach $P_e = K$ in an oscillating fashion if $1<r<2$. When 
-$r>2$ (i.e., ${F}'<-1$), system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
-as $r_c$, the critical point where system goes from under- to over-compensatory regime and 
-$r=2$ as $r_{bf}$ where bifurcation (transition from stable to unstable equilibrium) occurs.
-
-**Hassell model** [@hassell1975density; @may1976bifurcations], motivated for modelling insect populations,
-can be represented as $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$, where $r>0$ is the intrinsic growth rate 
-of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
-Beverton-Holt model [@beverton2012dynamics] used in fisheries. In general, if $b \sim 1$,
-then we say that there is contest and $F$ will saturate eventually at a finite
-level for large populations, which indicates that the population stabilizes by
-rejecting too many newborns. For scramble, $b$ is large and $F$ tends to
-zero for large populations which indicates that the resources are over-utilized
-leading to eventual extinction. Like Ricker, Hassel model too has two equilibrim:
-$P_e =0$ (trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
-For $P_e = P^\star$, model remains in under-compensatory regime if $1<r<[b/(b-1)]^b$ 
-and in over-compensatory regime if $[b/(b-1)]^b<r<[b/(b-2)]^b$ with $r_c = [b/(b-1)]^b$.
-Hopf bifurcation occurs at $r_{bf} = [b/(b-2)]^b$, so if $r>r_{bf}$, $P_e=P^\star$
-becomes unstable equilibrium.
-
-**Maynard Smith model** [@cohen1995unexpected; @maynard1978models; @may1976bifurcations] follows the dynamics 
-$P(t+1)=F(P(t),r,a,b)=rP(t)/[1+(aP(t))^b]$. $r$ is the growth rate, $a, b$ are the parameters with positive values.
-Equilibriums are: $P_e=0$(as expected extinction equilibrium) and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
-For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$. This $P^\star$ will be stable
-in under-compensatory regime for $1<r<r_c(=b/(b-1))$ and in over-compensatory regime for $r_c<r<b/(b-2)$.
-$P^\star$ will become unstable when $r>r_{bf}(=b/(b-2))$.
-
-**Pennycuick model** [@pennycuick1968computer; @may1976bifurcations] was 
-initially developed
-to follow change in age-structured total population
-level which assumed
-to have a constant sex 
-ratio and therefore change in female population was 
-expected through the change 
-in total population density. 
-A two population variant could also be possible
-in prey-predator perspective as indicated in [@pennycuick1968computer].
-In general, model dynamics is expressed as $P(t+1)=F(P(t),r,a,b)=rP(t)/[1+\exp \left\{-a(1-P(t)/b)\right\}]$, where 
-$r$ is the intrinsic growth rate and $a>0,b>0$ are the model parameters. Model
-equilibriums are 
-$P_e=0$ (extinction equilibrium) and $P_e=P^\star=\frac{b}{a}[a+\ln{(r-1)}]$. For 
-a finite equilibrium population 
-$r$ must be greater than $r_{min}$, where 
-$r_{min} = 1+\exp(-a)$. The critical transition point $r_c$ (from under- to 
-over-compensatory dynamics) and Hopf-bifurcation point $r_{bf}$ can be obtained
+**The Ricker model.** The Ricker model was originally introduced in the context of modelling
+the stock and recruitment of salmon populations, and has frequently been used in fisheries 
+management. The model is 
+$P(t+1)=F(P(t),r,K)=P(t) \exp[r(1-P(t)/K)]$, where $r>0$ is the intrinsic 
+growth rate and $K>0$ is the carrying capacity of the population. There are two equilibrium 
+states, $P_e = 0$ and $P_e = K$.
+The equilibrium $P_e = 0$ is unstable, since ${F}' = \exp(r)$ for that equilibtrium, and 
+$\exp(r) > 1$ for a positive $r$. Depending on $r$, 
+the other equilibrium, $P_e = K$, can be stable or unstable, since ${F}' = (1-r)$. 
+Specifically, $P_e=K$ is stable if $|1-r|<1$, i.e., $0<r<2$. 
+The model shows undercompesatory dynamics if $0<r<1$ and overcompensatory dynamics
+if $1<r<2$. Thus $r_{min}=0$, $r_c=1$, and $r_{bf}=2$.
+
+**The Hassell model.** The Hassell model [@hassell1975density; @may1976bifurcations] was 
+developed to help understand the dynamics of insect populations, and as a model of contest versus scramble 
+instraspecific competition. The model
+can be written $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$. Here $r>0$ is the intrinsic growth rate and
+$a$ and $b$ are positive parameters. If $b=1$, the Hassell model reduces to the
+Beverton-Holt model [@beverton2012dynamics] used in fisheries. The Hassell model is a representation
+of contest competition if $b \sim 1$, and scramble competition if $b$ is large.
+Like the Ricker model, the Hassell model also has two equilibria, $P_e =0$ 
+(the trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
+For $P_e = P^\star$, the model shows undercompensatory dymamics if $1<r<[b/(b-1)]^b$ 
+and overcompensatory dynamics if $[b/(b-1)]^b<r<[b/(b-2)]^b$, so $r_{min}=1$, 
+$r_c = [b/(b-1)]^b$, and $r_{bf}=[b/(b-2)]^b$.
+
+**The Maynard Smith model.** The Maynard Smith 
+model [@cohen1995unexpected; @maynard1978models; @may1976bifurcations] is
+$P(t+1)=F(P(t),r,a,b)=rP(t)/[1+(aP(t))^b]$. Here $r>0$ is the growth rate and $a, b$ are 
+positive parameters. The equilibria are $P_e=0$ (the extinction equilibrium) 
+and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
+For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$, so $r_{min}=1$. 
+Then $P^\star$ will 
+be a stable equilibrium and the model will show undercompensatory dynamics if $1<r<r_c=b/(b-1)$;
+the model will show overcompensatory dymamics $r_c<r<b/(b-2)$.
+$P^\star$ will become unstable for $r>r_{bf}=b/(b-2)$.
+
+**The Pennycuick model.** The Pennycuick model [@pennycuick1968computer; @may1976bifurcations] 
+is $P(t+1)=F(P(t),r,a,b)=rP(t)/[1+\exp \left\{-a(1-P(t)/b)\right\}]$, where 
+$r>0$ is the intrinsic growth rate and $a$ and $b$ are positive parameters. Model
+equilibria are 
+$P_e=0$ (the extinction equilibrium) and $P_e=P^\star=\frac{b}{a}[a+\ln{(r-1)}]$. For 
+a positive equilibrium population 
+$r$ must be greater than $r_{min} = 1+\exp(-a)$. The transition point $r_c$ (from under- to 
+overcompensatory dynamics) and the bifurcation point $r_{bf}$ were obtained
 by numerical solution of ${F}'=0$ and ${F}'=-1$ at $P_e=P^{\star}$, respectively.
 
-**Verhulst model** [@cohen1995unexpected; @may1976bifurcations] is a discrete time difference 
-representation of logistic growth of non-human biological population and can be written 
-as $P(t+1) = F(P(t),r,K) =P(t)[1+r(1-P(t)/K)]$, with $r>0$ and $K>0$ being growth rate and carrying capacity, respectively. 
-Model equilibrium $P_e=0$ is always unstable whereas $P_e=K$ is stable below Hopf-bifurcation 
-point $r_{bf}(=2)$. Model at stable equilibrium ($P_e=K$) lies in the under-compensatory regime when $0<r<r_c$
-and in over-compensatory regime when $rc<r<r_{bf}$ with $r_c$ being equal to 1.
-
-**Austin-Brewer model** [@austin1971world; @cohen1995unexpected], unlike from the first
-five models we considered,
-is used to model growth of human population. It follows the
-discrete dynamics : $P(t+1) = F(P(t),r,K,s)= P(t)[1+r(K-P(t)) \left\{ 1-\exp (-sP(t))\right\}]$, where
-$r, K, s$ are positive, representing growth rate, carrying capacity 
-and model parameter, respectively.
-Usual extinction equilibrium is $P_e=0$. The other
-finite, positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. The equilibrium $P^\star$
-is stable when $r<r_{bf}$ and unstable beyond the bifurcatio point $r_{bf} = \frac{2}{K[1-\exp (-sK)]}$.
-After perturbation, model can return back monotonically to its stable equilibrium if parameter $r$ lies in the 
-under-compensatory regime ($0<r<r_c$), where $r_c = \frac{1}{K[1-\exp (-sK)]}$. In the 
-over-compensatory regime ($r_c<r<r_{bf}$), model approaches to stable equilibrium $P^\star$
-in oscillatory fashion.
+**The Verhulst model.** The Velhurst model 
+[@cohen1995unexpected; @may1976bifurcations] is a discrete-time 
+representation of logistic growth, and can be written 
+as $P(t+1) = F(P(t),r,K) =P(t)[1+r(1-P(t)/K)]$, with $r>0$ and $K>0$ being growth rate and 
+carrying capacity parameters, respectively. 
+The model equilibrium $P_e=0$ is always unstable whereas the equilibrium 
+$P_e=K$ is stable below the bifurcation 
+point $r_{bf}=2$. The model shows undercompensatory dynamics for $0<r<r_c=1$
+and overcompensatory dynamics for $r_c<r<r_{bf}$.
+
+**The Austin-Brewer model.** The Austin-Brewer model 
+[@austin1971world; @cohen1995unexpected] was originally used 
+to model the growth of human populations. The model is $P(t+1) = F(P(t),r,K,s)= P(t)[1+r(K-P(t)) \left\{ 1-\exp (-sP(t))\right\}]$, where
+$r, K, s$ are positive parameters. $r$ and $K$ represent the growth rate and carrying 
+capacity, respectively.
+the usual extinction equilibrium is $P_e=0$. The 
+positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. $P^\star$
+is stable when $r<r_{bf}$ and unstable beyond than bifurcation point, where 
+$r_{bf} = \frac{2}{K[1-\exp (-sK)]}$. The model shows undercompensatory dynamics when
+$0<r<r_c$, where $r_c = \frac{1}{K[1-\exp (-sK)]}$. It shows 
+overcompensatory dynamics when $r_c<r<r_{bf}$.
 

### 2019-11-05 MainText.Rmd
edits

@@
 the $u(t)$ and $v(t)$ were uniformly distributed, the inverse 
-cummulative distribution functions of the desired beta and
+cumulative distribution functions of the desired beta and
 gamma distributions were the transformations. Because these
@@
 for all $i$. This was 
-accomplished with the inverse cummulative distribution function of the normal distribution
+accomplished with the inverse cumulative distribution function of the normal distribution
 with mean $0$ and standard deviation $\sigma$. Because this was a monotonic transformation,
@@
 \ref{SM-fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{SM-fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
-and the $r_c$ values of 
-Table \ref{tab_model_summary} reveals that, for the Velhurst and Austin-Brewer models,
+and the $r_c$ values of Table \ref{tab_model_summary} reveals that, for the Velhurst and Austin-Brewer models,
 the switch occurs at a value of $r$ less than $r_c$ for these models. For the example
@@
 for the Ricker, Hassell, and Pennycuick models, left-tail association accentuates and
-left-tail association mitigates extinction risk for undercompensatory dynamics ($r<r_c$),
+right-tail association mitigates extinction risk for undercompensatory dynamics ($r<r_c$),
 whereas the reverse occurs for overcompensatory dynamics ($r>r_c$). Fig. \ref{fig_extrisk_d0}
@@
 weather variables measured in different places, and thereby influence metapopulation
-extinction risk through the mechansms we explored in this paper. 
+extinction risk through the mechanisms we explored in this paper. 
 If $X_i(t)$ denotes the value of an environmental variable (e.g., a summer maximum 
@@
 Essentially, in such a case, associations between $X_i(t)$ and $X_j(t)$
-would be unchanging through time even through Pearson correlation
-appears to inducate a change in association, which is instead a change 
+would be unchanging through time even though Pearson correlation
+appears to indicate a change in association, which is instead a change 
 in marginal distributions.
@@
 extinction risk under additional modelling scenarios not considered
-in this paper.   
-First, we considered a model with positive association between 
+in this paper. First, we considered a model with positive association between 
 egg and adult vital rates,

### 2019-11-05 SuppMat.Rmd
edits

@@
 Bivarite environmental noise with extreme left-tail association (as in Fig. 
-\ref{fig_intro_cop}A) was generated using an algorithm originally presented 
+\ref{MT-fig_intro_cop}A) was generated using an algorithm originally presented 
 by @Ghosh_copula (their Appendix S10), presented again here for convenience.
@@
 with extreme right-tail association (as in Fig. 
-\ref{fig_intro_cop}D), take the negative of the entire matrix. For
-both types of noise, every entry was transformed by the cummulative distribution
+\ref{MT-fig_intro_cop}D), take the negative of the entire matrix. For
+both types of noise, every entry was transformed by the cumulative distribution
 function of a standard normal distribution, to make each column uniformly
@@
 The Spearman correlation of the extreme left- or right-tail associated 
-noise was $0.875$, up to sampling variation.
-<!--***DAN: fill this in-->
-
+noise was `r srho_common`, up to sampling variation.
 <!--moderate tail association noise-->
-Bivarite environmental noise with moderate left-tail association (as in Fig. 
-\ref{fig_intro_cop}B) was generated using a 2-dimensional Clayton copula. The statistical
+Bivarite environmental noise with moderate left-tail association (as in Fig. \ref{MT-fig_intro_cop}B) was 
+generated using a 2-dimensional Clayton copula. The statistical
 theory of copulas has been introduced systematically in several texts 
@@
 introduced by @Genest2007. Short, conceptual introductions aimed at ecological
-audiences were provided by sections within the papers of @Anderson2018 and  
-@Ghosh_copula. A 2-dimensional Clayton copula is equivalent to a particular bivariate 
+audiences were provided by sections within the papers of @Anderson2018 and @Ghosh_copula. A
+2-dimensional Clayton copula is equivalent to a particular bivariate 
 probability density function with marginal distributions which are 
 uniform on the interval $(0,1)$. The points on Fig. 
-\ref{fig_intro_cop}B were independent draws from the associated random variable.
+\ref{MT-fig_intro_cop}B were independent draws from the associated random variable.
 The functions `claytonCopula` and `rCopula` from the `copula` package for the 
 R programming language were used. The parameter of the Clayton family was
-chosen so that the Spearman correlation was $0.875$, the same as that of the
+chosen so that the Spearman correlation was `r srho_common`, the same as that of the
 extreme left-tail associated noise. This was accomplished with the `iRho`
@@
 with extreme right-tail association, take the negative of the entire matrix. For
-both types of noise, every entry was transformed by the cummulative distribution
+both types of noise, every entry was transformed by the cumulative distribution
 function of a standard normal distribution, to make each column uniformly
@@
 positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. $P^\star$
-is stable when $r<r_{bf}$ and unstable beyond than bifurcation point, where 
+is stable when $r<r_{bf}$ and unstable beyond the bifurcation point, where 
 $r_{bf} = \frac{2}{K[1-\exp (-sK)]}$. The model shows undercompensatory dynamics when

### 2019-11-05 MainText.Rmd
responses to Shyamolina's commnts

@@
 association is given in Methods.
+<!--**DAN: You suggested we replace this with \nameref{Methods}. I do not care
+one way or the other. You decide, we either do this for all such links or 
+for none. Please decide yourself and then search through and make it that
+way systematically across the ms. I point out though, links between pdfs will
+not work.-->
 Figure \ref{fig_intro_cop} shows that two variables can have a wide 
@@
 associations between climatic variables
-measured in different locations. For instance, if a single ECE is associated
+measured in different locations. For instance, suppose a single ECE is associated
 with spatially widespread, large values 
-of some climatic variable, but small values of the
-variable are associated with local weather rather than ECEs, the result could
+of some climatic variable. Also suppose that small values of the
+variable are associated, instead, with local weather, which can differ from location
+to location, rather than with ECEs. The result could
 be stronger right-tail association, through time, between measurements of the climatic variable
@@
 For example, rainstorms which deposit exceptionally large amounts of rain
-may do so over large areas; but periods with less rain may experience
+may do so over large areas. But periods with less rain may experience
 scattered showers only, which may be much less correlated across locales.
@@
 cumulative distribution functions of the desired beta and
-gamma distributions were the transformations. Because these
+gamma distributions were the transformations. 
+<!--***DAN: Shyamolina, you asked if I was talking about qbeta and qgamma here. 
+Yes, but we do not need to say that - these are very standard functions in R. Anyone
+who would be using R will know them, anyone not using R won't care that these
+are the names of the functions in R. We just have to give what the mathematical
+name is, which is "inverse 
+cumulative distribution functions". Thanks for correcting my spelling of cumulative. 
+Please delete this comment once you have read it.-->
+Because these
 were monotonic transformations, the Spearman correlation of 
@@
 accomplished with the inverse cumulative distribution function of the normal distribution
-with mean $0$ and standard deviation $\sigma$. Because this was a monotonic transformation,
+with mean $0$ and standard deviation $\sigma$. 
+<!--***DAN: Shyamolina, you made a comment here in the pdf. I understand what you are saying.
+What you did was mathematically equivalent to what I have written here, so it is OK we described it this way.-->
+<!--***DAN: Shyamolina, in future please make comments in the Rmd, not attached to a pdf - that 
+keeps everything in one place and makes it possible to reply to your comments more easily.-->
+Because this was a monotonic transformation,
 the Spearman correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
@@
 of $M$ and of the eigenvalues and eigenvectors of the expected value of 
-$M$. None of these quantities depends on tail associations between matrix entries.
+$M$. 
+<!--***DAN: Shyamolina, I did not put it as a formula because that would have 
+required a lot more text - we would have to define all the terms. We don't
+actually need to know the specific dependence, just the fact that the growth
+rate depends on the variances and covariances. -->
+None of these quantities depends on tail associations between matrix entries.
 Thus, whenever Tuljapurkar's approximation is a good

### 2019-11-05 SuppMat.Rmd
responses to Shyamolina's commnts

@@
 noise was `r srho_common`, up to sampling variation.
+<!--***DAN: Dan to fill this in with a derivation of this fact-->
+
 <!--moderate tail association noise-->
@@
 survival Clayton copula.
+<!--***DAN: Shyamolina, in response to your comment in the pdf, what you
+did is mathematically equivalent to this, so OK to say it this way. I added a 
+bit of explanation of what that survival Clayon is. Please delete this
+comment once you have read it.-->
+The survival Clayton copula is a rotation of the Clayton copula by 180 degrees.
 

### 2019-11-05 MainText.Rmd
deleting all \nameref{} link

--- 2019-11-05 MainText.Rmd
@@
 (Figure \ref{fig_intro_cop}C, D). A precise definition of a measure of tail 
-association is given in Methods.
-Figure \ref{fig_intro_cop} shows that two variables can have a wide 
+association is given in Methods. Figure \ref{fig_intro_cop} shows 
+that two variables can have a wide 
 range of patterns of tail association, from extreme left- to extreme right-tail
@@
 <!--stochastic matrix model : age-structured-->
-To investigate Q1 from the \nameref{Introduction},
+To investigate Q1 from the Introduction,
 we considered a simple stochastic age-structured model
@@
 <!--density-dependence metapopulation model-->
-To investigate Q2 from the \nameref{Introduction}, we considered spatialized, stochastic versions
+To investigate Q2 from the Introduction, we considered spatialized, stochastic versions
 of six density dependent population models, the Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, 
@@
 <!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
-To answer (Q2) of the \nameref{Introduction}, 
+To answer (Q2) of the Introduction, 
 we estimated extinction risk for each metapopulation model, using each form of tail 
@@
 \ref{fig_extrisk_allmodel_localdisp} and \ref{SM-fig_extrisk_allmodel_globaldisp}). 
-These results help address Q2 from the \nameref{Introduction}.
+These results help address Q2 from the Introduction.
 Intermediate values of $r$ are shown in Figs. \ref{SM-fig_ricker_risk_local_disp_scl_1} - 

### 2019-11-06 MainText.Rmd
added a derivation of the Spearman for extreme tail dependent noise

--- 2019-11-05 MainText.Rmd
@@
 
-<!--***DAN: By the way, Shyamolina, these are not Hopf bifuractions - those cannot happen
-with univariate models like these. These are called period-doubling bifucations.
-
-***Shya: I also mentioned Hopf bifurcation in Appendix S2. I leave as it is, when you checked that 
-we need to modify from hopf to period-doubling.
-
-If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
-model equilibrium $P_e$ became unstable. Beyond the Hopf-bifurcation point, the model showed sustained
-oscillation and eventually led to chaos with increasing $r$. Appendix \ref{SM-model_stability}
-summarized the stability analysis of each model in their deterministic form.
--->
-
 We simulated our metapopulation models using environmental noises $\epsilon_i(t)$ (Table \ref{tab_model_summary}), 

### 2019-11-06 SuppMat.Rmd
deleting all \nameref{} link
added a derivation of the Spearman for extreme tail dependent noise

@@
 distributed.
-
-<!--Spearman correlation of the extreme-tail-associated noise-->
 The Spearman correlation of the extreme left- or right-tail associated 
-noise was `r srho_common`, up to sampling variation.
-<!--***DAN: Dan to fill this in with a derivation of this fact-->
+noise was `r srho_common`, up to sampling variation. A derivation
+of this fact is presented below.
 
@@
 
+<!--derivation of the fact that Spearman is 0.875 for the extreme tail dependent noise-->
+We now present a derivation of the fact that, up to sampling variance,
+the Spearman correlation of our extreme tail associated noise was 
+$0.875$. Let $u_i$, $v_i$ be the ranks of a sample of noise, $i=1,\ldots,N$.
+Assuming that $N$ is divisible by $4$ makes the presentation of the 
+derivation simpler, so we adopt that assumption. We show the derivation for the left-tail
+associated noise, the right-tail case being similar. Then 
+\vspace{-10pt}
+\begin{equation}
+\rho=\frac{\sum_{i=1}^N (u_i - \mean(u_i))(v_i-\mean(v_i))}{(N-1)\sqrt{\var(u_i)\var(v_i)}}.
+\end{equation}
+\noindent However, it is straightforward to see that $\mean(u_i) = \mean(v_i)=(N+1)/2$,
+which we henceforth call $\mu$. It is also straightforward to see
+that $\var(u_i)=\var(v_i)=\frac{\sum_{i=1}^N (i-\mu)^2}{N-1}$, since 
+the set consisting of the $u_i$ is the same as the set consisting of the 
+$v_i$, and is also the same as the set consisting of the numbers $1,\ldots,N$. 
+Therefore,
+\vspace{-10pt}
+\begin{equation}
+\rho=\frac{\sum_{i=1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2}.
+\end{equation}
+\noindent We can reindex the $u_i$ and $v_i$ so that $u_i=v_i=i$ for $i=1,\ldots,N/2$.
+This uses the fact that we are working with the left-tail associated noise.
+Then we have 
+\vspace{-20pt}
+\begin{align}
+\rho &= \frac{\sum_{i=1}^{N/2} (i-\mu)^2 +\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2} \\
+&= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2}.\label{eq:rho1}
+\end{align}
+\noindent We define $\mu_2 = \mean_{i=N/2+1}^N(u_i)=\mean_{i=N/2+1}^N(i)$. It is straightforward
+to show that this is $3N/4+1/2$. Then the numerator of the right term in 
+(\ref{eq:rho1}) can be rearranged as
+\vspace{-20pt}
+\begin{align}
+\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu) &= \sum_{i=N/2+1}^N (u_i-\mu_2+\mu_2-\mu)(v_i-\mu_2+\mu_2-\mu) \\
+&= \sum_{i=N/2+1}^N \left( (u_i-\mu_2)(v_i-\mu_2)+(u_i-\mu_2)(\mu_2-\mu)
+  +(v_i-\mu_2)(\mu_2-\mu)+(\mu_2-\mu)^2 \right) \\
+&= \sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2) + \frac{N}{2}(\mu_2-\mu)^2 \\
+&= \sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2) + \frac{N^3}{32}.
+\end{align}
+\noindent The denominator of the right term in (\ref{eq:rho1}) can be rearranged as
+\vspace{-20pt}
+\begin{align}
+\sum_{i=1}^N (i-\mu)^2 &= \sum_{i=1}^N (i^2-2\mu i +\mu^2) \\
+&= \sum_{i=1}^N i^2 - 2\mu\sum_{i=1}^N i +N\mu^2.
+\end{align}
+\noindent Applying Faulhaber's formula to simplify the first term, this is
+\vspace{-20pt}
+\begin{align}
+\sum_{i=1}^N (i-\mu)^2 &= \frac{N(N+1)(2N+1)}{6}-2\mu\frac{N(N+1)}{2}+N\mu^2 \label{eq:rho2}\\
+&= \frac{N(N+1)(N-1)}{12},\label{eq:rho3}
+\end{align}
+\noindent where we have omitted the simple algebra leading from (\ref{eq:rho2}) 
+to (\ref{eq:rho3}). Plugging in our expressions for the numerator and denominator,
+(\ref{eq:rho1}) becomes
+\vspace{-20pt}
+\begin{align}
+\rho &= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2} \\
+&= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2) + \frac{N^3}{32}}{\frac{N(N+1)(N-1)}{12}} \\
+&= \frac{1}{2}+\frac{12\sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2)}{N(N+1)(N-1)}+
+\frac{3N^3}{8N(N+1)(N-1)} \label{eq:rho4}
+\end{align}
+\noindent The second term of (\ref{eq:rho4}) is proportional to the covariance 
+$\cov_{i=N/2+1}^N (u_i,v_i)$. To see this, note that
+\vspace{-20pt}
+\begin{align}
+\cov_{i=N/2+1}^N (u_i,v_i) &= \frac{\sum_{i=N/2+1}^N(u_i-\mu_2)(v_i-\mu_2)}{(N/2-1)
+\frac{\sum_{i=N/2+1}^N (i-\mu_2)^2}{N/2-1}} \\
+&= \frac{\sum_{i=N/2+1}^N(u_i-\mu_2)(v_i-\mu_2)}{\sum_{i=N/2+1}^N (i-\mu_2)^2}.\label{eq:rho5}
+\end{align}
+\noindent But it is straightforward to show that
+\vspace{-10pt}
+\begin{equation}
+\sum_{i=N/2+1}^N (i-\mu_2)^2 = \frac{N(N+2)(N-2)}{96}.
+\end{equation}
+\noindent Plugging this into (\ref{eq:rho5}), we get
+\vspace{-20pt}
+\begin{align}
+\cov_{i=N/2+1}^N (u_i,v_i) &= \frac{96\sum_{i=N/2+1}^N(u_i-\mu_2)(v_i-\mu_2)}{N(N+2)(N-2)}.
+\end{align}
+\noindent Therefore the middle term of (\ref{eq:rho4}) equals
+\vspace{-10pt}
+\begin{equation}
+\frac{12\sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2)}{N(N+1)(N-1)} = \cov_{i=N/2+1}^N (u_i,v_i) \left( \frac{N(N+2)(N-2)}{8N(N+1)(N-1)} \right).
+\end{equation}
+\noindent The limit of this as $N \rightarrow \infty$ is $0$ because the covariance
+goes to 0. Thus the limit of (\ref{eq:rho4}) as $N \rightarrow \infty$
+is $7/8=0.875$, as originally claimed.
+
 # Stability analysis of deterministic metapopulation models  \label{model_stability}

### 2019-11-06 SuppMat.Rmd

@@
 association was generated in a similar way but using the 2-dimensional 
-survival Clayton copula.
-<!--***DAN: Shyamolina, in response to your comment in the pdf, what you
-did is mathematically equivalent to this, so OK to say it this way. I added a 
-bit of explanation of what that survival Clayon is. Please delete this
-comment once you have read it.-->
-The survival Clayton copula is a rotation of the Clayton copula by 180 degrees.
+survival Clayton copula. The survival Clayton copula is a rotation
+of the Clayton copula by 180 degrees.
 

### 2019-11-06 MainText.Rmd
small edits on final read through

@@
 title: "Tail associations and their impact on extinction risk"
-author: "Shyamolina Ghosh, Daniel Reuman"
+author: "Shyamolina Ghosh, Daniel C. Reuman"
 date: ""
@@
 Two positively associated variables can be principally "left-tail associated" (i.e., more 
-correlated when they take low than high values) or "right-tail associated"
+correlated when they take low values than when they take high values) or "right-tail associated"
 (more correlated when they take high than low values), even with the same
@@
 spatial tail associations of environmental variables can influence metapopulation 
-extinction risk. For instance, if low values of a variable reduce the growth 
+extinction risk. For instance, if low values of an environmental variable reduce the growth 
 rates of local
 populations, one may expect that left-tail associations increase
-metapopulation extinction risks because then environmental catastrophes are
+metapopulation extinction risks because then environmental "catastrophes" are
 spatially synchronized, presumably reducing the potential for rescue effects.
@@
 reverse was usually true for models showing overcompensatory density dependence.
-For the non-spatial, stage structured models we considered, left-tail associations 
+For the non-spatial, stage-structured models we considered, left-tail associations 
 between vital rates did accentuate extinction risk compared to right-tail
@@
 
+\vspace{12pt}
 \noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model, copula
@@
 [@hansen2012perception]. All such changes have the potential to influence 
-populations, including their risk of extirpation from an area 
+populations, including through risks of species extinction or extirpation from an area 
 [@GarciaCarerras2013; @Vasseuretal2014].
@@
 because our intention here is to provide an initial exploration only of 
-the potential mechanisms, outlined above, which suggest that
-asymmetries of tail association may make a difference for extinction risks. 
+the potential mechanisms outlined above. 
 By so doing we hope to open these potentially practically important questions to 
-wider examination by other researchers. Though tail associations are used
+wider examination by other researchers, using a variety of models and perspectives. 
+Though tail associations are used
 to assess risks in fields such as finance and hydrology 
@@
 knowledge asymmetric tail associations have rarely been assessed for their
-importance for extinction risks of populations and species (but see @Ghosh_copula). 
-Our questions are an important step for understanding
+importance for extinction risks of populations and species (but see @Ghosh_copula),
+and our essential goal is to advance the idea that tail associations may
+also be important in this context. 
+Our questions are also an important step for understanding
 the potential influence of ECEs on extinction risks if ECEs are related to
@@
 \includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
-\caption{Two stochastic quantities can have same Spearman 
+\caption{Two stochastic quantities can have the same Spearman 
 correlation, but different degrees of tail association: 
@@
 on its panel). The extreme left-tail (respectively, right-tail) dependent 
-case is perfectly correlated below (respectively, above) the threshold of 0.
+case is perfectly correlated below (respectively, above) the threshold of $1/2$.
 See Methods for information on how bivariate noise was generated with
@@
 about association between variables, as each marginal pertains to
-only one of the variables.
+only one of the variables. These ideas are revisited in the Discussion.
 
@@
 show two bands, the data in the right/upper band showing
-stronger association than those in the left/lower band. This figure was reproduced
+stronger association than those in the left/lower band. Each band is bounded by
+the lines $x+y=2l_b$ and $x+y=2u_b$ described in Methods (for values of $l_b$ and $u_b$
+that depend on the band to be used). This figure was reproduced
 largely following Fig. 7 of Ghosh \emph{et al.} (2019).
@@
 
-<!--
-We plotted extinction risks along the y-axis
-for both left tail and right tail associated noises 
-against the varying range of $r$ along x-axis when no dispersal ($d=0$) was considered.
-We excluded the possibility of dispersal to avoid any rescue effect among patches. 
-For each model with extreme (or moderate) left- and right-tail associated noises, we calculated 
-2.5$^{th}$ and 97.5$^{th}$ quantiles
-of 10000 simulations with the probability of computed
-extinction risks at each $r$. We compared $r_c$ and the value of $r$ at which 
-95\% CI of extinction risks from both noises overlapped
-with each other. This could give us an idea of how extinction risks affected by the model parameter 
-$r$ that controlled the underlying dynamics of the model, i.e., undercompensatory 
-or overcompensatory.
--->
-
 # Results\label{Results}
@@
 risk after 25 years. A potential reason for this outcome is elaborated in the Discussion.
-<!--discussion later: add Tuljapurkar's approximation stuff-->
 
 <!--Q2 : metapopulation results-->
-For our metapopulation simulations, for low values of $r$, left-tail
+For our metapopulation simulations, for low values of $r$ (close to $r_{min}$), left-tail
 associations of environmental noises accentuated extinction risks, and right-tail associations
-mitigated them; but the reverse was true for high values of $r$ in the range
-$(r_{min},r_{bf})$. Thus 
+mitigated them; but the reverse was true for high values of $r$ (close to
+$r_{bf}$). Thus 
 density dependent models can either produce similar results, regarding the influence of 
@@
 beneficial events caused populations to exceed carrying capacity, and 
-to subsequently crash, across large areas. In contrast 
+to subsequently crash, across all or most habitat patches. In contrast 
 to the metapopulation case, our results 
@@
 @ummenhofer2017extreme provide a diagram of how ECEs may be increasing in frequency 
-and intensity due to changes in the distribution of $X_i(t)$ such as changes
+and intensity, linked to changes in the distribution of $X_i(t)$ such as changes
 in the mean, variance, or skewness of this distribution. Although changes in 
@@
 between $X_i(t)$ and $X_j(t)$, and asymmetries of tail association, may be 
-changing. Documented changes in correlation between $X_i(t)$ and $X_j(t)$
+changing. However, documented changes in correlation between $X_i(t)$ and $X_j(t)$
 may be due partly or wholly to changes in tail associations. If ECEs are becoming
@@
 more intense [@ummenhofer2017extreme], then the conditional probability that both of
-$X_i(t)$ and $X_j(t)$ are influenced by an ECE given that one of them is so
-influenced should be increasing. This should contribute to increased association
+$X_i(t)$ and $X_j(t)$ are influenced by an ECE, given that one of them is so
+influenced, should be increasing. This should contribute to increased association
 of the two variables in one of their tails (which tail depends on the type of 
@@
 which is the most commonly used measure of association of two variables,
-depends on both marginal information and association information.
-It is possible that recently observed changes in environmental
+depends not only on association information, but also on the marginals.
+For this reason, it is possible that recently observed changes in environmental
 (e.g., @Black2018) and biological (e.g., @Koenig2016) synchrony 
@@
 already-studied changes in distributional characteristics of 
-variables, rather than reflecting true changes in association.
+variables, rather than reflecting true changes in association (these studies
+used Pearson correlations).
 Rank-based methods are linked to a well-developed suite of statistical
@@
 @Genest2007 and others have recommended copula and rank approaches when the 
-goal is to understand associations between variables, and how such 
+goal is to understand associations between variables, including how such 
 associations may be changing.
@@
 
+# Acknowledgments
+
+\noindent The authors were partly supported by US National Science Foundation
+grant 1714195 and the James S. McDonnell Foundation.
+
  <!--FIG : age structure model results : linear, nonlinear-->
@@
 (A, B) and extreme (C, D) asymmetry of tail association.
-The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk.\label{fig_extrisk_agestr}}
+The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk.
+LT=left tail, RT=right tail.\label{fig_extrisk_agestr}}
 \end{center}
@@
 done for each extinction risk estimate, confidence intervals
-were very narrow, as in Fig. \ref{fig_extrisk_agestr}.\label{fig_extrisk_allmodel_localdisp}}
+were very narrow, as in Fig. \ref{fig_extrisk_agestr}. LT=left tail, RT=right 
+tail. \label{fig_extrisk_allmodel_localdisp}}
 \end{center}
@@
 version of the model. Error bars show 95\% confidence intervals of extinction risk.
-\label{fig_extrisk_d0}}
+LT=left tail, RT=right tail. \label{fig_extrisk_d0}}
 \end{center}
@@
 $r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation; see main text).
-We used $r$ such that $r_{min}<r<r_{bf}$.
+For extinction-risk simulations, we used $r$ such that $r_{min}<r<r_{bf}$.
 Dispersal between patches was also implemented (see main text).

### 2019-11-06 SuppMat.Rmd
small edits on final read through

@@
 chosen so that the Spearman correlation was `r srho_common`, the same as that of the
-extreme left-tail associated noise. This was accomplished with the `iRho`
+extreme left- and right-tail associated noise. This was accomplished with the `iRho`
 function of the `copula` package. Noise with moderate right-tail
@@
 the Spearman correlation of our extreme tail associated noise was 
-$0.875$. Let $u_i$, $v_i$ be the ranks of a sample of noise, $i=1,\ldots,N$.
+$0.875$. Let $x_i$, $y_i$ be the ranks of a sample of noise, $i=1,\ldots,N$.
 Assuming that $N$ is divisible by $4$ makes the presentation of the 
 derivation simpler, so we adopt that assumption. We show the derivation for the left-tail
-associated noise, the right-tail case being similar. Then 
+associated noise, the right-tail case being similar. Then we have
 \vspace{-10pt}
 \begin{equation}
-\rho=\frac{\sum_{i=1}^N (u_i - \mean(u_i))(v_i-\mean(v_i))}{(N-1)\sqrt{\var(u_i)\var(v_i)}}.
+\rho=\frac{\sum_{i=1}^N (x_i - \mean(x_i))(y_i-\mean(y_i))}{(N-1)\sqrt{\var(x_i)\var(y_i)}}.
 \end{equation}
-\noindent However, it is straightforward to see that $\mean(u_i) = \mean(v_i)=(N+1)/2$,
+\noindent However, it is straightforward to see that $\mean(x_i) = \mean(y_i)=(N+1)/2$,
 which we henceforth call $\mu$. It is also straightforward to see
-that $\var(u_i)=\var(v_i)=\frac{\sum_{i=1}^N (i-\mu)^2}{N-1}$, since 
-the set consisting of the $u_i$ is the same as the set consisting of the 
-$v_i$, and is also the same as the set consisting of the numbers $1,\ldots,N$. 
+that $\var(x_i)=\var(y_i)=\frac{\sum_{i=1}^N (i-\mu)^2}{N-1}$, since 
+the set consisting of the $x_i$ is the same as the set consisting of the 
+$y_i$, and is also the same as the set consisting of the numbers $1,\ldots,N$. 
 Therefore,
@@
 \begin{equation}
-\rho=\frac{\sum_{i=1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2}.
+\rho=\frac{\sum_{i=1}^N (x_i-\mu)(y_i-\mu)}{\sum_{i=1}^N (i-\mu)^2}.
 \end{equation}
-\noindent We can reindex the $u_i$ and $v_i$ so that $u_i=v_i=i$ for $i=1,\ldots,N/2$.
-This uses the fact that we are working with the left-tail associated noise.
+\noindent We can reindex the $x_i$ and $y_i$ so that $x_i=y_i=i$ for $i=1,\ldots,N/2$.
+This uses the fact that we are working with the extreme left-tail associated noise.
 Then we have 
@@
 \begin{align}
-\rho &= \frac{\sum_{i=1}^{N/2} (i-\mu)^2 +\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2} \\
-&= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2}.\label{eq:rho1}
+\rho &= \frac{\sum_{i=1}^{N/2} (i-\mu)^2 +\sum_{i=N/2+1}^N (x_i-\mu)(y_i-\mu)}{\sum_{i=1}^N (i-\mu)^2} \\
+&= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (x_i-\mu)(y_i-\mu)}{\sum_{i=1}^N (i-\mu)^2}.\label{eq:rho1}
 \end{align}
-\noindent We define $\mu_2 = \mean_{i=N/2+1}^N(u_i)=\mean_{i=N/2+1}^N(i)$. It is straightforward
+\noindent We define $\mu_2 = \mean_{i=N/2+1}^N(x_i)=\mean_{i=N/2+1}^N(i)$. It is straightforward
 to show that this is $3N/4+1/2$. Then the numerator of the right term in 
@@
 \begin{align}
-\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu) &= \sum_{i=N/2+1}^N (u_i-\mu_2+\mu_2-\mu)(v_i-\mu_2+\mu_2-\mu) \\
-&= \sum_{i=N/2+1}^N \left( (u_i-\mu_2)(v_i-\mu_2)+(u_i-\mu_2)(\mu_2-\mu)
-  +(v_i-\mu_2)(\mu_2-\mu)+(\mu_2-\mu)^2 \right) \\
-&= \sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2) + \frac{N}{2}(\mu_2-\mu)^2 \\
-&= \sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2) + \frac{N^3}{32}.
+\sum_{i=N/2+1}^N (x_i-\mu)(y_i-\mu) &= \sum_{i=N/2+1}^N (x_i-\mu_2+\mu_2-\mu)(y_i-\mu_2+\mu_2-\mu) \\
+&= \sum_{i=N/2+1}^N \left( (x_i-\mu_2)(y_i-\mu_2)+(x_i-\mu_2)(\mu_2-\mu)
+  +(y_i-\mu_2)(\mu_2-\mu)+(\mu_2-\mu)^2 \right) \\
+&= \sum_{i=N/2+1}^N (x_i-\mu_2)(y_i-\mu_2) + \frac{N}{2}(\mu_2-\mu)^2 \\
+&= \sum_{i=N/2+1}^N (x_i-\mu_2)(y_i-\mu_2) + \frac{N^3}{32}.
 \end{align}
@@
 \begin{align}
-\rho &= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (u_i-\mu)(v_i-\mu)}{\sum_{i=1}^N (i-\mu)^2} \\
-&= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2) + \frac{N^3}{32}}{\frac{N(N+1)(N-1)}{12}} \\
-&= \frac{1}{2}+\frac{12\sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2)}{N(N+1)(N-1)}+
+\rho &= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (x_i-\mu)(y_i-\mu)}{\sum_{i=1}^N (i-\mu)^2} \\
+&= \frac{1}{2}+\frac{\sum_{i=N/2+1}^N (x_i-\mu_2)(y_i-\mu_2) + \frac{N^3}{32}}{\frac{N(N+1)(N-1)}{12}} \\
+&= \frac{1}{2}+\frac{12\sum_{i=N/2+1}^N (x_i-\mu_2)(y_i-\mu_2)}{N(N+1)(N-1)}+
 \frac{3N^3}{8N(N+1)(N-1)} \label{eq:rho4}
@@
 \noindent The second term of (\ref{eq:rho4}) is proportional to the covariance 
-$\cov_{i=N/2+1}^N (u_i,v_i)$. To see this, note that
+$\cov_{i=N/2+1}^N (x_i,y_i)$. To see this, note that
 \vspace{-20pt}
 \begin{align}
-\cov_{i=N/2+1}^N (u_i,v_i) &= \frac{\sum_{i=N/2+1}^N(u_i-\mu_2)(v_i-\mu_2)}{(N/2-1)
+\cov_{i=N/2+1}^N (x_i,y_i) &= \frac{\sum_{i=N/2+1}^N(x_i-\mu_2)(y_i-\mu_2)}{(N/2-1)
 \frac{\sum_{i=N/2+1}^N (i-\mu_2)^2}{N/2-1}} \\
-&= \frac{\sum_{i=N/2+1}^N(u_i-\mu_2)(v_i-\mu_2)}{\sum_{i=N/2+1}^N (i-\mu_2)^2}.\label{eq:rho5}
+&= \frac{\sum_{i=N/2+1}^N(x_i-\mu_2)(y_i-\mu_2)}{\sum_{i=N/2+1}^N (i-\mu_2)^2}.\label{eq:rho5}
 \end{align}
@@
 \begin{align}
-\cov_{i=N/2+1}^N (u_i,v_i) &= \frac{96\sum_{i=N/2+1}^N(u_i-\mu_2)(v_i-\mu_2)}{N(N+2)(N-2)}.
+\cov_{i=N/2+1}^N (x_i,y_i) &= \frac{96\sum_{i=N/2+1}^N(x_i-\mu_2)(y_i-\mu_2)}{N(N+2)(N-2)}.
 \end{align}
@@
 \begin{equation}
-\frac{12\sum_{i=N/2+1}^N (u_i-\mu_2)(v_i-\mu_2)}{N(N+1)(N-1)} = \cov_{i=N/2+1}^N (u_i,v_i) \left( \frac{N(N+2)(N-2)}{8N(N+1)(N-1)} \right).
+\frac{12\sum_{i=N/2+1}^N (x_i-\mu_2)(y_i-\mu_2)}{N(N+1)(N-1)} = \cov_{i=N/2+1}^N (x_i,y_i) \left( \frac{N(N+2)(N-2)}{8N(N+1)(N-1)} \right).
 \end{equation}

### 2019-11-08 MainText.Rmd
adding new csl
finished incorporating Lawrence comments

@@
 title: "Tail associations and their impact on extinction risk"
-author: "Shyamolina Ghosh, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
 date: ""
@@
 \noindent Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
+
+\noindent Sheppard: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
 
@@
 are computed using all the data but the sum is computed only
-over the $t$ such that $x(t)+y(t) > 2l_b$ and $x(t)+y(t)<2u_b$. 
+over the $t$ such that $x(t)+y(t) > 2l_b$ and $x(t)+y(t)<2u_b$. The variance, $\var$,
+here and throughout, is the sample variance, with Bessel's correction (denominator $T-1$).
 The partial Spearman correlation is the component of the Spearman
@@
 \ref{SM-fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{SM-fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
-and the $r_c$ values of Table \ref{tab_model_summary} reveals that, for the Velhurst and Austin-Brewer models,
+and the $r_c$ values of Table \ref{tab_model_summary} reveals that, for the Verhulst and Austin-Brewer models,
 the switch occurs at a value of $r$ less than $r_c$ for these models. For the example
@@
 whereas the reverse occurs for overcompensatory dynamics ($r>r_c$). Fig. \ref{fig_extrisk_d0}
-reveals a similar reversal for the Maynard Smith, Velhurst and Austin-Brewer models; but 
+reveals a similar reversal for the Maynard Smith, Verhulst and Austin-Brewer models; but 
 it occurs at a value of $r$ close to but distinguishable from $r_c$.
@@
 extinction risk through the mechanisms we explored in this paper. 
-If $X_i(t)$ denotes the value of an environmental variable (e.g., a summer maximum 
-temperature, or winter minimum temperature, spring rainfall, etc.) 
-in location $i$ in year $t$,
-@ummenhofer2017extreme provide a diagram of how ECEs may be increasing in frequency 
-and intensity, linked to changes in the distribution of $X_i(t)$ such as changes
-in the mean, variance, or skewness of this distribution. Although changes in 
+Although changes in 
 the correlation between $X_i(t)$ and $X_j(t)$ are increasingly documented, 
@@
 environmental variable $X$ represents.
-Studying associations between the values of an environmental variable in different
-locations, $X_i(t)$ and $X_j(t)$, using standard methods such as Pearson correlation
+Studying associations between $X_i(t)$ and $X_j(t)$ using standard 
+methods such as Pearson correlation
 may be complicated by these changes; rank-based and related 
 measures of association such as 
-we have used are recommended instead.
+we have used may be more appropriate for some applications.
 If $f$ is a strictly monotonic, increasing transformation, then the Spearman
@@
 already-studied changes in distributional characteristics of 
-variables, rather than reflecting true changes in association (these studies
-used Pearson correlations).
+variables, rather than reflecting true changes in association as would be measured
+using rank-based methods (these studies used Pearson correlations).
 Rank-based methods are linked to a well-developed suite of statistical
@@
 goal is to understand associations between variables, including how such 
-associations may be changing.
+associations may be changing. We recognize that the Pearson correlation
+of $X_i(t)$ and $X_j(t)$, and changes through time in that correlation, 
+will still be and appropriate and important quantity to study 
+for many applications. For instance, in cases for which $X_i$ has a linear 
+or approximately linear influence on some aspect of the ecosystem
+at location $i$, Pearson correlations of the ecological impacts at locations 
+$i$ and $j$, which may be of fundamental interest, will relate to 
+Pearson correlations between $X_i$ and $X_j$. But in many cases the
+ecological impact of $X_i$ and $X_j$ will be a nonlinear or unknown monotonic functions of 
+the values of $X_i$ and $X_j$. So studying rank-based correlations
+may be more informative, ecologically as well as providing the mathematical advantages
+listed above.
+
+<!--future ideas start here-->
+Our results suggest some new lines of research on the 
+impact of tail associations on growth rates and 
+extinction risks. First, we considered a model with positive association between 
+egg and adult vital rates,
+corresponding to similar impacts of an environmental variable on 
+eggs and adults. Negative associations between vital rates
+can occur as well, and the impacts of tail associations should be considered
+in that context as well. Negative associations relate to
+life history trade-offs, e.g. between survival and reproduction 
+(e.g, @Lande2003, p. 59). Asymmetric tail associations may occur
+in such a context if, for instance, there is a three way trade-off
+among vital rates $r_1$, $r_2$, and $r_3$. Values of $r_1$ near
+an upper limit may be strongly associated with,
+for instance, $r_2$, whereas lower values of $r_1$ may not be strongly
+associated with $r_2$ because reductions/increases in $r_1$ can
+then be traded off against increases/reductions (respectively) in either 
+$r_2$ or $r_3$. It is possible that tail associations in negatively 
+related vital rates produce different outcomes for extinction risk
+from what we have observed by studying positively associated vital rates.
 
 <!--Tuljapurkar approximation-->
+A second topic of research that our work suggests has to do with Tuljapurkar's
+approximation of the long-term stochastic growth rate.
 Our results suggest that tail associations and asymmetries of tail associations
@@
 certainly still needed
-as to whether this conclusion holds more generally than the model we considered
-(see below), if it does hold 
+as to whether this conclusion holds more generally than the model we considered, 
+if it does hold 
 generally, it may be possible to obtain a better understanding of 
-the result using a well-known approximation
+the result using the well-known approximation
 of Tuljapurkar [@tulja1982; @tulja1990] for the long-term stochastic growth rate.
@@
 to which accuracy of the approximation has been tested for scenarios with
-asymmetric tail associations.
-We also note that this reasoning pertains only to density-independent models,
+asymmetric tail associations. It seems worth explicitly exploring the accuracy of 
+Tuljapurkar's approximation under scenarios of tail association in the vital rates
+and relating results to the effects of tail associations on the long-term
+stochastic growth rate and extinction risks.
+We note that the above reasoning pertains only to density-independent models,
 and says nothing about our results for the density-dependent version of our
@@
 
-<!--***DAN: decided the below was a step too far
-Finally, it seems likely that the 
-stochastic matrix modelling framework, and Tuljapurkar's approximation,
-should apply to the density-independent metapopulation model of 
-@Ghosh_copula, which was comprised of Lewontin-Cohen models linked 
-by dispersal. Nevertheless asymmetry of tail associations made a 
-big difference to extinction risk for that model [@Ghosh_copula].
-We have not examined whether Tuljapurkar's approximation is
-accurate in that spatially explicit case.
-
-***DAN: Dan to maybe add another para on the fact that I thnk Tulja's approx should also
-work for our spatial, non-dd model (check first, Dan), but we see a strong effect here,
-so not clear why. And also suggests the possibilty that for some age-structured models,
-tail associations will make a bigger difference then they did in our case. Maybe somehow
-get to the suggestion that someone needs to revisit Tulja's approx and the analytic
-approximation of extinction risk of Lande and Orzack looking at tail associations?
-
-***DAN: In Discussion need to talk about what analytic results are needed
-to take the next step of determining when tail depencence may make enough of
-a difference to think about. I think our results, together with Tulja's 
-approximation, suggest that tail associations probably don't make much difference
-for models without a spatial component. But our results illsutaret that it can
-make a big difference when there is a spatial component, apparently because of 
-the influence of tail associations on rescue effects. Could maybe even outline
-analytic approaches for non-dd models? Not sure what would be done with dd
-models. The point is you need to take the next step in the Discussion of 
-talking about how our results, which admitedly are just for specific models,
-could be turned into analytic results or general results. Interestingly, 
-Tulja's approximation may also apply to the spatial case, at least when 
-all locations can be reached from all other locations via dispersal. But the 
-approximation seems poor in that case. Need to think about it more, and maybe
-need more research.-->
+<!--relate things to the geography of synchrony-->
+Finally, our results suggest that combining
+ideas about tail association with recently developed ideas about the geography 
+of synchrony may provide fertile territory for future research. 
+Our metapopulation models assumed that habitat patches were equispaced on a transect or
+were all equally mutually accessible through dispersal.
+Furthermore, associations between environmental variables in different
+patches were assumed to be spatially homogeneous. In real metapopulations,
+neither of these assumptions holds. Several recent studies, reviewed and synthesized
+by @walteretal2017, have explored geographic variation in the synchrony of environmental
+and population variables; @walteretal2017 explored the mechanisms, consequences
+and ecological importance of the geography of synchrony. For instance, synchrony was found to 
+have a prominent geography which can help provide inferences about the causes
+of synchrony and about organism ecology. Two among many research questions that 
+could be explored that blend the geography of synchrony with tail associations
+include: 1) what is the nature of the geography of tail associations of 
+environmental and ecological variables? and 2) do asymmetries of tail
+associations of environmental variables interact with geographies of synchrony
+to influence metapopulation extinction risk?
+Connecting the ideas of tail association that
+we have developed with the ideas on the geography of synchrony
+of @walteretal2017 seems like a promising future direction.
 
@@
 -->
-
-<!--future ideas-->
-In our opinion, additional research would be worthwhile on the 
-impact of tail associations on growth rates and 
-extinction risk under additional modelling scenarios not considered
-in this paper. First, we considered a model with positive association between 
-egg and adult vital rates,
-corresponding to similar impacts of an environmental variable on 
-eggs and adults. Negative associations between vital rates
-can occur as well, relating to
-life history trade-offs, e.g. between survival and reproduction 
-(e.g, @Lande2003, p. 59). Asymmetric tail associations may occur
-in such a context if, for instance, there is a three way trade-off
-among vital rates $r_1$, $r_2$, and $r_3$. Values of $r_1$ near
-an upper limit may be well be associated with,
-for instance, $r_2$, whereas lower values of $r_1$ may not be strongly
-associated with $r_2$ because reductions/increases in $r_1$ can
-then be traded off against increases/reductions in *either* 
-$r_2$ or $r_3$. Second, for our metapopulation models,
-we assumed habitat patches were equispaced on a transect or
-were all equally mutually accessible through dispersal.
-Furthermore, associations between environmental variables in different
-patches were assumed to be spatially homogeneous. In real metapopulations,
-neither of these assumptions holds. Connecting the ideas of tail association that
-we have developed in this study with the ideas on the geography of synchrony
-of @walteretal2017 seems like a promising future direction.
 

### 2019-11-08 SuppMat.Rmd
adding new csl
finished incorporating Lawrence comments

@@
 title: "Supporting information: Tail associations and their impact on extinction risk"
-author: "Shyamolina Ghosh, Daniel Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
 date: ""
@@
 \caption[Extinction risk v.
-tail association, $d$ and $r$: metapopulations, local dispersal, Velhurst]{Analogous to Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Velhurst model.
+tail association, $d$ and $r$: metapopulations, local dispersal, Verhulst]{Analogous to Fig.
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Verhulst model.
 \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
@@
 \caption[Extinction risk v.
-tail association, $d$ and $r$: metapopulations, global dispersal, Velhurst]{Analogous to Fig.
+tail association, $d$ and $r$: metapopulations, global dispersal, Verhulst]{Analogous to Fig.
 \ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} but for global dispersal between habitat patches.
@@
 \begin{equation}
-\rho=\frac{\sum_{i=1}^N (x_i - \mean(x_i))(y_i-\mean(y_i))}{(N-1)\sqrt{\var(x_i)\var(y_i)}}.
+\rho=\frac{\sum_{i=1}^N (x_i - \mean(x_i))(y_i-\mean(y_i))}{(N-1)\sqrt{\var(x_i)\var(y_i)}},
 \end{equation}
-\noindent However, it is straightforward to see that $\mean(x_i) = \mean(y_i)=(N+1)/2$,
+\noindent where $\var$, here and throughout, is the sample variance, with Bessel's correction 
+(denominator $N-1$). However, it is straightforward to see that $\mean(x_i) = \mean(y_i)=(N+1)/2$,
 which we henceforth call $\mu$. It is also straightforward to see
@@
 \noindent The second term of (\ref{eq:rho4}) is proportional to the covariance 
-$\cov_{i=N/2+1}^N (x_i,y_i)$. To see this, note that
+$\cov_{i=N/2+1}^N (x_i,y_i)$ (again using $N-1$ in the denominator, here and throughout). 
+To see this, note that
 \vspace{-20pt}
@@
 
-**The Verhulst model.** The Velhurst model 
+**The Verhulst model.** The Verhulst model 
 [@cohen1995unexpected; @may1976bifurcations] is a discrete-time 

### 2019-11-09 MainText.Rmd
added 'mathematical ecology' as a keywrd

@@
 
-csl: TheAmericanNaturalist.csl
+csl: ecosphere.csl
 bibliography: Ref_extrisk.bib
@@
 \vspace{12pt}
-\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model, copula
+\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model, copula, mathematical ecology
 

### 2019-11-09 MainText.Rmd
fixed typos

@@
 result held, typically, but not always, for undercompensatory model dynamics. However,
-for overcompensatory dynamics, the oposite result held, again typically, but not always: 
+for overcompensatory dynamics, the opposite result held, again typically, but not always: 
 when "harmful" events were correlated across space,
@@
 of $X_i(t)$ and $X_j(t)$, and changes through time in that correlation, 
-will still be and appropriate and important quantity to study 
+will still be an appropriate and important quantity to study 
 for many applications. For instance, in cases for which $X_i$ has a linear 
@@
 $i$ and $j$, which may be of fundamental interest, will relate to 
-Pearson correlations between $X_i$ and $X_j$. But in many cases the
-ecological impact of $X_i$ and $X_j$ will be a nonlinear or unknown monotonic functions of 
+Pearson correlations between $X_i$ and $X_j$. But in many cases, the
+ecological impact of $X_i$ and $X_j$ will be a nonlinear or unknown monotonic function of 
 the values of $X_i$ and $X_j$. So studying rank-based correlations
@@
 [@MorrisDoak2002], and it factors prominently into
-extintion risk estimates for density-independent age- or 
+extinction risk estimates for density-independent age- or 
 stage-structured population models 

### 2019-11-09 SuppMat.Rmd
fixed typos

@@
 <!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
-Bivarite environmental noise with extreme left-tail association (as in Fig. 
+Bivariate environmental noise with extreme left-tail association (as in Fig. 
 \ref{MT-fig_intro_cop}A) was generated using an algorithm originally presented 
@@
 <!--moderate tail association noise-->
-Bivarite environmental noise with moderate left-tail association (as in Fig. \ref{MT-fig_intro_cop}B) was 
+Bivariate environmental noise with moderate left-tail association (as in Fig. \ref{MT-fig_intro_cop}B) was 
 generated using a 2-dimensional Clayton copula. The statistical
@@
 probability density function with marginal distributions which are 
-uniform on the interval $(0,1)$. The points on Fig. 
+uniform on the interval $(0,1)$. The points in Fig. 
 \ref{MT-fig_intro_cop}B were independent draws from the associated random variable.
@@
 Multivariate noise with extreme left-tail association was generated using a
-very simlar algorithm to the bivariate case,
+very similar algorithm to the bivariate case,
 as follows. If $T$ time steps of noise of dimension $n$ were desired,

### 2019-11-10 MainText.Rmd
formatting for journal

@@
 date: ""
+fontsize: 12pt
 geometry: "left=1in,right=1in,top=1in,bottom=1in"
@@
 
+mainfont: Times New Roman      
 tables: True
@@
 \vspace{12pt}
-\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model, copula, mathematical ecology
+\noindent \textbf{\textit{Keywords:}} copula; density-dependent model; extinction risk; extreme events; mathematical ecology; tail association.
 
@@
 \begin{center}
-\includegraphics[width=14.5 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
+\includegraphics[width=14 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
 \caption{Extinction risk versus dispersal rate for different scenarios of tail association

### 2019-11-10 SuppMat.Rmd
formatting for journal

@@
 
+mainfont: Times New Roman       
 tables: True
@@
 
-# Generating noise with asymmetric tail associations  \label{makenoise}
-
-<!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
+# Generating noise with asymmetric tail associations\label{makenoise}
+
 Bivariate environmental noise with extreme left-tail association (as in Fig. 

### 2020-01-15 Response.txt
started adding response to reviewer's comment...to be continued

********************************
Reviewer #1 (Comments to the Author):

Summary of article
This manuscript investigates the effect of asymmetries in tail associations (increased correlation at either extreme low or high values of a population or environmental variable) on extinction risk for populations in stage-structured demographic models and in spatial metapopulations. It extends the little previous ecological work on asymmetric tail associations in spatial metapopulations by including density-dependence and introduces them in stage structured models. The authors find that spatial left-tail associations substantially increase extinction risk in metapopulations, possibly due to lower chance of rescue for populations with low growth rates. They find these asymmetries in vital rate tail associations impact extinction risk very little in stage-structured non-spatial models.

Overarching thoughts
The overall concepts of what asymmetric tail associations are and why they are important ecologically are clearly communicated, with the use of conceptual figures such as Fig 1.

The inclusion of two kinds of asymmetric tail associations – those between vital rates within a population and spatial tail associations in a metapopulation – illustrate that the type of asymmetric tail association matters to determining their role in extinction risk.

The use of several forms of density dependence in the spatial metapopulation model demonstrates the degree to which effect sizes are not dependent on a specific density dependence formulation, but a broader feature of asymmetric tail associations across metapopulation model structures.

***Shya: Thanks for the positive feedback! We feel happy that our communication has made it clear about the usefulness of asymmetric tail associations in estimating extinction risk for the stage-structured demographic model and meta-populations.

I suggest only a few clarifications:

Specific comments
L164-175: Has this particular formulation been analyzed in detail in a classic text? Cite it to guide the reader to more resources on the behavior of this model more generally.

***Shya: The formulation [Eq. 1] is quite common in population-demography modeling [Caswell; Moris & Doak] where vital rates are found to be time-dependent and contains stochasticity (due to random environmental noise) in population projection matrix. (see L168-171)

L222-224: Cite introduction or thorough analyses of these models to direct the reader to more resources. Or add those citations to Table 1 and reference it here for brevity.

***Shya: These models are adapted from Cohen(1995) - now mentioned in L227. Details of each model are discussed in Appendix S2.

********************************
Reviewer #2 (Comments to the Author):

Brief summary: This paper examined the effects of “tail associations” of ecological variables on the extinction risk of metapopulations and stage-structured populations using nonlinear population models. The results show that the effect of tail associations between fecundity and juvenile survival is minor, while that between patches that a metapopulation occupies increases or decreases extinction risk, depending on the domain of different compensatory dynamics. This work is one of the earliest studies that examine how copula structure in ecological variables influence population persistence. The authors investigated this problem using a variety of models, showing the general importance of tail associations of ecological variables on population persistence. Because copula structure in ecological variables should be ubiquitous in nature and ongoing climate change is likely altering correlation structure between variables, the problem investigated in this paper is of interest to the readers of ESA journals.

***Shya: Many thanks for the positive feedback and constructive suggestions! We are happy that we have got the attention in estimating extinction risk using copula to model "tail-associations" between ecological/ environmental variables. For specific response, please see below.

General comments
1.      The title seems a bit cryptic to readers who are not a statistician. I would suggest that title be modified to clarify it is about the tails of the distributions of ecological variables.

***Shya: I could think of a few alternative titles:
(i) Tail associations between ecological variables and their impact on extinction risk
(ii) Tail associations between ecological variables could affect the extinction risk
(iii) Tail associations between ecological variables, extreme events and their impact on extinction risk

2.      A conceptual schematic diagram early in the introduction will greatly help readers understand the main idea quickly and smoothly follow the text. In the current form, it is relatively hard to follow. Figure 1 does help, but a figure showing the idea in a more concrete way using tangible examples would make it much easier for readers to absorb the idea early on in the paper. It also would be helpful to see an example from finance or other fields that are somewhat analogous to ecological situations.

***Shya: I think Fig. 1 is the most simplified pedagogical fig. we've ever had in our all copula-related papers. 
So, as of now I have no idea what extra info does the 2nd reviewer ask for? Copula related paper from finance or other fields are already cited in the paper in L156-157.

3.      The amount of analyses for metapopulations and age-structured populations seems quite unbalanced. Why is this so? Obviously, the paper would greatly increase its value by including more analyses for stage-structured populations, but then the paper may become too long to fit in the required length for Ecosphere. Would it be more beneficial to focus only on metapopulations in this paper and expand on the stage-structured population part in a separate paper to more deeply examine Tuljapurkar’s approximation?

***Shya: This is our first attempt to introduce the general impact of asymmetric tail-association (using copula) on the extinction risk - be it on the stage-structured model or on the spatial metapopulation models. We agree that further studies are needed as mentioned in the Discussion to examine why those effects are less severe in stage-structured population model compared to the metapopulation - probably we will address that in detail in our 
subsequent papers (I'm not sure about this statement - but I could not think of a better response :P).

4.      The order of appearance of the metapopulation problem and the age-structured population problem keeps switching throughout the paper. For readability, the order should be consistent.

***Shya: Ok, I will read the paper again, and comment on this later.

5.      Whether variables have tail associations can be altered by transformation; e.g., variables may have no tail association in one way but show any of the tail associations discussed in this paper in another. How should researchers who want to use the approached introduced in this paper deal with this? Should they look for a transformation that can give one of such tail associations as long as it is ecologically meaningful in a given system? I suppose so, but this is not mentioned in the paper. Also, the interpretation that certain variables are “bad” or “good” depends on how it is formulated or transformed. It could be that, if a “bad” variable is associated with a “good” variable in their right tails, for example, their effects cancel each other and leave no overall effect, which would be also an interesting result. This may be related to the discussion about negative associations between Line 411 and 423, but it does not clearly come across. Maybe these are discussed in the Ghosh et al. paper in BioRxiv, but it is not yet published in a peer reviewed journal. These aspects should be worth mentioning in this paper for clarification of the method.

***Shya: I have several comments about this criticism:
(i) Reviewer2 has misunderstood us here. Transformation (Linear) does not alter the nature of tail-association, i.e., stronger left tail association remains left-tail dominated, so as for right tail.
(ii) After reading his own story of "good" or "bad" variable - now I'm confused. I think "good" or "bad" notion of variable has nothing to do with the transformation - it is their underlying nature that how they affect the population-size with time. Also, I don't agree that a right tail (or left-tail) association between two contradictory variable (i.e. negatively correlated) nullify the effect on population-extinction risk.
(iii) Our BIVAN paper is now accepted and in press, though not updated in BioRxiv, I am not sure whether I should upload the accepted version in BioRxiv or not because of copyright agreement with Elsevier.

6.      Figure 5 presents cases with d = 0 only and in the appendix different values of d are considered only for global dispersal. Would the results qualitatively hold when different values of d for local dispersal is considered?

***Shya: I personally believe the results should hold, in general, even when d>0 (because d=0 is more stringent case). My goal is to show the crossover happened solely because of the change in dynamical regime, so setting d=0 means there will be no rescue-effect on extinction risk due to metapopulation dispersal. (But still I will check my first statement - I will write a toy code using my saved data)

7.      In general, the text in certain parts can be much tightened for readability and clarity.

Minor comment
•       Line 97-105. It is too abstract to follow in the current form until the reader read the last sentence.
•       Line 100, insert “same” in between “the” and “variable” for clarity.

***Shya: okay, I've rephrased this para.

•       As I mentioned above, Q1 and Q2 appear in an opposite order from what I expected from the introduction so far.

***Shya: see point 4.

•       Line 299. Because simulation studies can increase the number of replicates relatively easily and usually use a large number of replicates, statistical significance does not mean much in most of cases, including this one.

***Shya: Here Reviewer2 is talking about Fig. 3. I've no response for this :(

•       Line 357-9. Rewrite the sentence. The paper did not show “how such (climatic) changes modify the asymmetry of tail association of weather variables measured in different places.” It showed only whether the latter part of the sentence happens.

***Shya: I am more or less agreed with the reviewer here, but I don't rewrite the sentence. I leave it for DAN.

•       In the second paragraph in Discussion, X’s are used to refer to generic random variables, but it appeared rather abruptly as the letter was not used to denote variables in the paper up to that point in the main text. It would be helpful to give the definition parenthetically in the text. Do i and j refer to different weather variables, to different ecological variable, or also to different patches?

***Shya: Agreed. As far as I understand, X's could be time series of two environmental/ ecological variables (i.e. for two vital rates  i, j for stage-structured model) from same location or metapopulation time series from two sites i, j (in case of metapopulation model). I'm not sure though, so I did not edit that para in text, I will do that after DAN comments.

•       The paragraph starting from Line 372 is very long and winding. This needs to be tightened substantially (to 2/3 to ½ of the current length, which is 1.5 pages).

***Shya: I leave it for DAN. I agree that the para is too long but the contents are useful, may be we could split the para into multiple ones.

•       Line 419-421. This sentence is very confusing. Using “/” and “(respectively)” makes it almost intractable.

***Shya: Agreed and modified in the text.

•       Line 424-450. I suggest that the sentence starting with “Although” should come after “…with asymmetric tail associations” in Line 445 for a better flow. As a connection, add a sentence like “This result is partly consistent with the prediction from Tuljapurkar’s approximation” before “The long-term stochastic growth rate… (Line 431)”.

***Shya: I leave it for DAN's judgement.

•       Line 459-460. Please give a brief definition of “the geography of synchrony,” and clarify what “synchrony was found to have a prominent geography…” means here.

***Shya: modified in the text but please check.

Please see also the attached file.

********************************

### 2020-01-15 MainText.Rmd
started adding response to reviewer's comment...to be continued

@@
 because our goal was to carry out an initial evaluation, only, of 
-the realism of a potential mechanism. The stochastic vital rates $s_e(t)$ and $f_a(t)$
-were selected, independently across times, as follows. 
-The egg survival rate $s_e(t)$ was drawn from a beta distribution with parameters
-$\alpha=3$ and $\beta=2$. 
+the realism of a potential mechanism. The formulation (\ref{eq_age_str}) is quite common 
+in population-demography modeling [@Caswell2000; @MorrisDoak2002] where vital rates 
+are found to be time-dependent and contains stochasticity (due to random 
+environmental noise) in population projection matrix. The stochastic vital 
+rates $s_e(t)$ and $f_a(t)$ in Eq. \ref{eq_age_str} were selected, independently 
+across times, as follows. The egg survival rate $s_e(t)$ was 
+drawn from a beta distribution with parameters $\alpha=3$ and $\beta=2$. 
 Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
@@
 of six density dependent population models, the Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, 
-and Austin-Brewer models.
+and Austin-Brewer models - as adapted from @cohen1995unexpected.
 All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
@@
 associated with $r_2$ because reductions/increases in $r_1$ can
-then be traded off against increases/reductions (respectively) in either 
+then be traded off against increases/reductions in either 
 $r_2$ or $r_3$. It is possible that tail associations in negatively 
@@
 neither of these assumptions holds. Several recent studies, reviewed and synthesized
-by @walteretal2017, have explored geographic variation in the synchrony of environmental
-and population variables; @walteretal2017 explored the mechanisms, consequences
-and ecological importance of the geography of synchrony. For instance, synchrony was found to 
-have a prominent geography which can help provide inferences about the causes
+by @walteretal2017, have explored geographic variation in the spatial synchrony of environmental
+and population variables (known as "geography of synchrony"); @walteretal2017 explored the 
+mechanisms, consequences and ecological importance of this geography of synchrony. For instance, 
+synchrony was found to have a prominent geography (i.e. a distinctive spatial pattern) which can 
+help provide inferences about the causes
 of synchrony and about organism ecology. Two among many research questions that 

### 2020-01-16 Response.txt
response edited as of now

@@
 
-***Shya: Ok, I will read the paper again, and comment on this later.
+***Shya: Ok, If we assume the accepted order would be first Stage-structured model (Addressing Q1...) and then metapopulation model (addressing Q2...), then I found the reverse order as follows:
+
+In the Introduction: we first explained metapop (para started with "There are also reasons to hypothesize that asymmetry of tail association in environmental...") and then stage structured model (para started with "Extinction risk is also very commonly estimated for age-structured models such as....")
+
+In the Discussion: we first explained the first topic as metapop results and 
+then as a second topic stage-structured results (in brief compared to first topic).
+Currently, I do not change anything in the maintext, I can rearrange the para but I am afraid in doing so, may be I 
+would break the continuation. So, I am leaving for DAN.
 
@@
 My goal is to show the crossover happened solely because of the change in dynamical regime, so setting d=0 means there will be no rescue-effect on extinction risk due to metapopulation dispersal.
-(But still I will check my first statement - I will write a toy code using my saved data)
+(ok, I checked the simulation result with d=0.3 for local and global dispersal for ricker and hassell, the crossover is around the same r_c, I am attaching the pdfs in my mail).
 

### 2020-01-16 MainText.Rmd
response edited as of now

@@
 <!--ECEs probably produce tail associations-->
-There are reasons to hypothesize that ECEs may produce asymmetric tail 
+
+<!--***DAN wrote this-->
+<!--There are reasons to hypothesize that ECEs may produce asymmetric tail 
 associations between climatic variables
@@
 may do so over large areas. But periods with less rain may experience
-scattered showers only, which may be much less correlated across locales.
+scattered showers only, which may be much less correlated across locales.-->
+
+<!--***Shya: I rephrased the following para-->
+There are reasons to hypothesize that different locations when influenced by single ECEs may 
+show asymmetric tail associations through time. This asymmetry arises because a single 
+ECE is associated with spatially widespread, large values 
+of some climatic variable whereas small values of the same
+variable are associated, instead, with local weather, which can differ from location
+to location, rather than with ECEs. For example, rainstorms which deposit 
+exceptionally large amounts of rain may do so over large areas. But periods with 
+less rain may experience scattered showers only, which may be much 
+less correlated across locales. The result could
+be stronger right-tail association, through time, between measurements of rainfall
+occurred in different locations.
 

### 2020-01-19 Response.txt
worked on comments and edits

@@
+Dear editors,
+
+Thank you for the attention and expertise you used in assessing, for Ecosphere, our paper "Tail associations and their impact on extinction risk" (this was the old title, we will now be changing it in response to a referee request - see below).
+Thank you also for the generally very positive reviews.
+We have responded to all referee comments through changes to the manuscript, and below.
+We find the paper is improved and we hope you now find it ready for formal acceptance to Ecosphere.
+Our responses to comments are preceeded by ">>>" to make them easy to find.
+Thank you again for an enjoyable and helpful review process.
+
+Yours,
+Shyamolina Ghosh, Lawrence Sheppard and Dan Reuman
+
 ********************************
@@
 
+>>>Thank you for the positive feedback!
+
 The inclusion of two kinds of asymmetric tail associations – those between vital rates within a population and spatial tail associations in a metapopulation – illustrate that the type of asymmetric tail association matters to determining their role in extinction risk.
@@
 
-***Shya: Thanks for the positive feedback!
-We feel happy that our communication has made it clear about the usefulness of asymmetric tail associations in estimating extinction risk for the stage-structured demographic model and meta-populations.
+>>>Correct. Thank you for the positive feedback.
 
@@
 
-***Shya: The formulation [Eq.
-1] is quite common in population-demography modeling [Caswell; Moris & Doak] where vital rates are found to be time-dependent and contains stochasticity (due to random environmental noise) in population projection matrix.
-(see L168-171)
+>>>Models of this form (i.e., stochastic matrix models, of which this is an example) are very common in demography, evolutionary ecology, and quantitative conservation biology.
+As the referee requests, we have added references to the classic books of Caswell and Morris & Doak.
 
@@
 
-***Shya: These models are adapted from Cohen(1995) - now mentioned in L227. Details of each model are discussed in Appendix S2.
+>>>Thank you for raising this point, as it spurred us to add clarity about where these models come from.
+These models are all described by Cohen (1995), who also cites the original sources for each model.
+Details of each model are also given in Appendix S2.
+These facts are now mentioned at the point in the text to which the referee refers.
+Thanks again.
 
@@
 
-***Shya: Many thanks for the positive feedback and constructive suggestions!
-We are happy that we have got the attention in estimating extinction risk using copula to model "tail-associations" between ecological/ environmental variables.
-For specific response, please see below.
+>>>Thank you for the positive feedback!
 
@@
 
-***Shya: I could think of a few alternative titles:
-(i) Tail associations between ecological variables and their impact on extinction risk
-(ii) Tail associations between ecological variables could affect the extinction risk
-(iii) Tail associations between ecological variables, extreme events and their impact on extinction risk
+>>>Thank you for the suggestion. We changed the title as suggested.
 
@@
 So, as of now I have no idea what extra info does the 2nd reviewer ask for? Copula related paper from finance or other fields are already cited in the paper in L156-157.
+***DAN: I do not think we can dodge this - we have to do SOME or even MOST of what the refere says, or he/she will simply say, if the paper is sent back to him/her, "they did not do what I asked for, so REJECT".
+Of the different suggestions, I think we should make some effort at this one and at 4 below.
+I will write a justification below for not doing 3.
+Please think more about such a figure, ASAP.
+Please think of an *additional* figure, to come right after the curent Fig.
+1 The current Fig.
+1 explains what tail dependence is.
+But I think what the referee wants is a figure explaining, conceptually, how tail dependence could cause accentuated extinction risk.
+I think he/she wants a conceptual figure showing the main idea.
+I think the figure should be associated with, and cited from, the Intro paras beginning "There are also reasons to hypothesize..." and "Extinction risk is also very commonly estimated..." I agree with the referee that we can produce a good conceptual figure illustrating this main idea.
+I know you can do this.
+Please think about it Monday and come to me Tuesday with a hand-drawn sketch of a figure that we can discuss before you then create the figure on Tuesday.
+Please do not come back to me saying you do not know how to do this - you can do this, I know you can.
+And time is getting a little tight, and I have other things I have to do on Monday, so this has to be done by you!
+I am sure you can do a good job.
+***DAN: Part of the reason we need to move efficiently on this is they only gave us three weeks, and we may also have to convert to Word, which you know can be labor intensive.
+Can you please look on the Ecosphere website and find out if they need final submissions to be in Word or if they can take latex?
+Even if they can take latex, we may want to convery to Word.
+If there is any evidence that they prefer Word, we will want to convert to Word to avoid them introducing a million errors trying to do latex when they do not really know how.
 
@@
 
-***Shya: This is our first attempt to introduce the general impact of asymmetric tail-association (using copula) on the extinction risk - be it on the stage-structured model or on the spatial metapopulation models.
-We agree that further studies are needed as mentioned in the Discussion to examine why those effects are less severe in stage-structured population model compared to the metapopulation - probably we will address that in detail in our
-subsequent papers (I'm not sure about this statement - but I could not think of a better response :P).
+>>>This is a first attempt to understand the impact of asymmetric tail associations on extinction risk, and as such we intentionally used a simple approach, using simulations instead of attempting mathematical analyses, and using simple models, in order to assess if the ideas have merit.
+We believe this could turn into a broad topic, which many researchers should investigate.
+Our intention was to provide an initial exploration of some potential mechanisms, and in so doing, hopefully to open what could be a practically important topic (e.g., for conservation) to exploration by other researchers.
+With regard to mathematical analysis of stochastic matrix models, we are journeymen only, not true experts.
+We believe attempting to carry out an analytic approach ourselves, e.g., using Tuljapurkars approximation, as the referee suggests, would actually slow down the scientfic progress of the field relative to what progress may be possible if stochastic matrix model experts see our curent results and decide to start from the ideas we have posited and then to carry out an appropriate mathematical analysis that would test the generality of our matrix model results (and related models such as integral projection models, which are becoming more and more common and which require even more sophistication to mathematically analyze).
+We agree with the referee that a systematic, analytic treatment of stochastic matrix models with asymmetric tail associations between vital rates is a worthy goal, but we think the best/fastest way to achieve this is to publish our current result, in hopes of stimulating some appropriate experts to carry out such an analysis.
+Carving our matrix models results out of the current paper and then undertaking a more systematic explortion ourselves, for possible later publication, does not get these potentially practically important ideas out there as fast.
+Finally, we do not see such a serious imbalance - to us the two components somewhat, but not dramatically different in the extent of their coverage, not so much for us to want to slow the transmission to the research community of what could be practically important ideas.
+Certainly our matrix modelling results are enough to raise the possibility that tail associations in vital rates may influence extinction risk, but to suggest that this effect may be minor; and transmitting this idea was our goal.
+A truly systematic exploration of that idea, using analysis, will require a real expert in the field.
+
+***DAN: Dan will edit and shorten the above.
 
@@
 
+***DAN: OK I will do this
+
 5.
@@
 I suppose so, but this is not mentioned in the paper.
+
 Also, the interpretation that certain variables are “bad” or “good” depends on how it is formulated or transformed.

### 2020-01-19 MainText.Rmd
worked on comments and edits

@@
 ---
-title: "Tail associations and their impact on extinction risk"
+title: "Tail associations in ecological variables and their impact on extinction risk"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
@@
 because our goal was to carry out an initial evaluation, only, of 
-the realism of a potential mechanism. The formulation (\ref{eq_age_str}) is quite common 
-in population-demography modeling [@Caswell2000; @MorrisDoak2002] where vital rates 
-are found to be time-dependent and contains stochasticity (due to random 
-environmental noise) in population projection matrix. The stochastic vital 
+the realism of a potential mechanism. The stochastic vital 
 rates $s_e(t)$ and $f_a(t)$ in Eq. \ref{eq_age_str} were selected, independently 
@@
 $\Omega$ made the model nonlinear, with density dependence in the adult
-fecundity rate.
+fecundity rate. Stochastic matrix models are very commonly used in
+evolutionary ecology and quantitative conservation biology, so our model
+is an example of a very widely used category of models. See, for instance,
+the classic books of @Caswell2000, and @MorrisDoak2002.
 
@@
 of six density dependent population models, the Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, 
-and Austin-Brewer models - as adapted from @cohen1995unexpected.
+and Austin-Brewer models. These models were all described by @cohen1995unexpected. 
+The models are also listed in Table \ref{tab_model_summary}, and 
+are analyzed in detail in Appendix \ref{SM-model_stability}, where 
+references for the models are also provided.
 All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,

### 2020-01-20 Response.txt
getting close to being done addressing Ecosphere referee comments

@@
 
->>>This is a first attempt to understand the impact of asymmetric tail associations on extinction risk, and as such we intentionally used a simple approach, using simulations instead of attempting mathematical analyses, and using simple models, in order to assess if the ideas have merit.
-We believe this could turn into a broad topic, which many researchers should investigate.
-Our intention was to provide an initial exploration of some potential mechanisms, and in so doing, hopefully to open what could be a practically important topic (e.g., for conservation) to exploration by other researchers.
-With regard to mathematical analysis of stochastic matrix models, we are journeymen only, not true experts.
-We believe attempting to carry out an analytic approach ourselves, e.g., using Tuljapurkars approximation, as the referee suggests, would actually slow down the scientfic progress of the field relative to what progress may be possible if stochastic matrix model experts see our curent results and decide to start from the ideas we have posited and then to carry out an appropriate mathematical analysis that would test the generality of our matrix model results (and related models such as integral projection models, which are becoming more and more common and which require even more sophistication to mathematically analyze).
-We agree with the referee that a systematic, analytic treatment of stochastic matrix models with asymmetric tail associations between vital rates is a worthy goal, but we think the best/fastest way to achieve this is to publish our current result, in hopes of stimulating some appropriate experts to carry out such an analysis.
-Carving our matrix models results out of the current paper and then undertaking a more systematic explortion ourselves, for possible later publication, does not get these potentially practically important ideas out there as fast.
-Finally, we do not see such a serious imbalance - to us the two components somewhat, but not dramatically different in the extent of their coverage, not so much for us to want to slow the transmission to the research community of what could be practically important ideas.
-Certainly our matrix modelling results are enough to raise the possibility that tail associations in vital rates may influence extinction risk, but to suggest that this effect may be minor; and transmitting this idea was our goal.
-A truly systematic exploration of that idea, using analysis, will require a real expert in the field.
-
-***DAN: Dan will edit and shorten the above.
+>>>Thank you for asking these questions.
+First, this is an initial attempt to understand the impact of asymmetric tail associations on extinction risk, and as such we intentionally used a simple approach for a rapid assessment of whether the basic ideas have merit.
+Based on our results, we believe this could turn into a broad topic, which many researchers should investigate.
+Our intention was to provide an initial exploration of the mechanisms rapidly, and in so doing, hopefully to open what could be a practically important topic (for conservation) to consideration by other researchers, as quickly and broadly as possible.
+With regard to extensive mathematical analyses (e.g.
+using Tuljapurkar's approaximation) of stochastic matrix models, which the referee suggests, we are journeymen only, not true experts.
+Mathematical analysis of stochastic matrix models is a deep field.
+So we believe carving our matrix model results out of the current paper and then undertaking a more systematic explortion ourselves, for possible later publication, does not get these potentially practically important ideas out there as fast and risks the possibility that we will not be able to do the analyses as well as some of our readers.
+  Second, for stochastic matrix models, our initial explorations suggested that probably the effects of tail association in that context are minor, so we went no further.
+In contrast, for metapopulation models, our initial explorations (with the Ricker model) showed a likely strong effect, so we went further with additional models.
+So what the referee perceives as imbalance is actually just a rational attempt to explore promising ideas further and to simply identify other ideas as less promising, to help future researchers choose what to work on.
+  Finally, we do not see as much of an imbalance as the referee suggests.
+To us, the two components are treated with slightly different throroughness, but not markedly so, and for good reasons (see above).
+Our matrix modelling results are certainly enough to make the points we wanted to make, i.e., an initial assessment that the influence of tail associations on extinction risk in this context seems likely to be minor.
 
@@
 
-***Shya: Ok, If we assume the accepted order would be first Stage-structured model (Addressing Q1...) and then metapopulation model (addressing Q2...), then I found the reverse order as follows:
-
-In the Introduction: we first explained metapop (para started with "There are also reasons to hypothesize that asymmetry of tail association in environmental...") and then stage structured model (para started with "Extinction risk is also very commonly estimated for age-structured models such as....")
-
-In the Discussion: we first explained the first topic as metapop results and 
-then as a second topic stage-structured results (in brief compared to first topic).
-Currently, I do not change anything in the maintext, I can rearrange the para but I am afraid in doing so, may be I 
-would break the continuation. So, I am leaving for DAN.
-
-***DAN: OK I will do this
+>>>Thanks for raising this point. We have edited to enforce consistency of this ordering throughout the manuscript.
 
@@
 I suppose so, but this is not mentioned in the paper.
+
+>>>The referee is mistaken here, transformation does not alter tail associations as we have defined them based on data ranks.
+This is a standard way of defining tail associations and copula approaches to the study of dependencies, justified by the Ghosh et al paper (which, by the way, has now been formally accepted at Advances in Ecological Research, and is in press) and by Genest & Favre, 2007, J Hydol Eng, "Everything you always wanted to know about copula modelling but were afraid to ask".
+We have added an explanation of this fact in the Discussion, and made some related points here that will help readers in future applications of our ideas.
+
+***DAN: Don't forget to add the explanation.
 
@@
 
-***Shya: I have several comments about this criticism:
-(i) Reviewer2 has misunderstood us here. Transformation (Linear) does not alter the nature of tail-association, i.e., stronger left tail association remains left-tail dominated, so as for right tail.
-(ii) After reading his own story of "good" or "bad" variable - now I'm confused.
-I think "good" or "bad" notion of variable has nothing to do with the transformation - it is their underlying nature that how they affect the population-size with time.
-Also, I don't agree that a right tail (or left-tail) association between two contradictory variable (i.e.
-negatively correlated) nullify the effect on population-extinction risk.
-(iii) Our BIVAN paper is now accepted and in press, though not updated in BioRxiv, I am not sure whether I should upload the accepted version in BioRxiv or not because of copyright agreement with Elsevier.
+>>>The Ghosh et al paper is now formally accepted.
+The version on the BioRxiv is essentially the same as the final, accepted version.
+In spite of thinking for a quite a long time about what the referee may have meant by the above comment, we confess we are not sure, so we do not know what to do about this comment.
+In our model formulations, variables are never "good" or "bad" but are instead assumed to fluctuate, taking values in some years which are "good" for the population (by which we mean they promote increased reproduction or increased survival in that location, in that year), and values in other years which are "bad" (in the opposite sense).
+Aside from an apparent desire to have us make some additional point (I'm afriad we cannot understand what point), the referee does not appear to be criticizing anything we actually do say.
+So we suggest proceeding with this text as it is.
+As we said, the Ghosh et al paper is now accepted, and is already available online on the BioRxiv, and many important related points are made there.
+So readers looking to base some of their own research on the points made in the current manuscript can also use the Ghosh et al paper as an additional background resource before getting started.
 
@@
 
-***Shya: I personally believe the results should hold, in general, even when d>0 (because d=0 is more stringent case).
-My goal is to show the crossover happened solely because of the change in dynamical regime, so setting d=0 means there will be no rescue-effect on extinction risk due to metapopulation dispersal.
-(ok, I checked the simulation result with d=0.3 for local and global dispersal for ricker and hassell, the crossover is around the same r_c, I am attaching the pdfs in my mail).
+>>>If we understand correctly what the referee is saying here, he/she seems to be mistaken.
+Figs S2-S7 show results for each model for a whole range of d, for local dispersal.
+These figures are referenced at the end of the second-to-last paragraph of Results, and in the last paragraph of Results.
+It can be determined through examination of Figs S2-S13 that yes, results will qualitatively hold for nonzero d for the types of dispersal we considered (local and global).
+A statement to this effect has been added to the last paragraph of Results.
 
@@
 
+>>>In the course of doing other edits we went through the entire text again and tightened where possible.
+
+***DAN: make sure to do this
+
 Minor comment
 •       Line 97-105. It is too abstract to follow in the current form until the reader read the last sentence.
+
+>>>We reworded the paragraph.
+
 •       Line 100, insert “same” in between “the” and “variable” for clarity.
 
-***Shya: okay, I've rephrased this para.
+>>>Done.
 
@@
 
-***Shya: see point 4.
+>>>We have fixed this ordering issue throughout the manuscript, see our response to point 4 above.
 
@@
 
-***Shya: Here Reviewer2 is talking about Fig. 3. I've no response for this :(
+>>>Thank you, we take the referee's point here and have eliminated the statement about significance.
 
@@
 
-***Shya: I am more or less agreed with the reviewer here, but I don't rewrite the sentence. I leave it for DAN.
+>>>The sentence the referee mentions was the topic sentence of the paragraph.
+The first part of that sentence was what the argument of the paragraph is about.
+So yes, we did do this, it was accomplished not in the Results of the paper (which is what the referee seems to have expected), but instead was accomplished at this point in the manuscript, in this very paragraph.
+We apologize, however, that this was unclear!
+So we reworded to make it clearer.
+We also worked on this whole paragraph, and the following one, to make them clearer.
 
@@
 
-***Shya: Agreed.
-As far as I understand, X's could be time series of two environmental/ ecological variables (i.e.
-for two vital rates  i, j for stage-structured model) from same location or metapopulation time series from two sites i, j (in case of metapopulation model).
-I'm not sure though, so I did not edit that para in text, I will do that after DAN comments.
+>>>We apologize for this omission! We have now added the definition. We have also clarified the rest of this paragraph.
 
@@
 
-***Shya: I leave it for DAN. I agree that the para is too long but the contents are useful, may be we could split the para into multiple ones.
+>>>We split the paragraph into two and shortened, as the referee requesrs.
 
@@
 
-***Shya: modified in the text but please check.
+>>>Done, and the clarity is improved. Thanks.
 
@@
 
+***DAN: Shyamolina, did you look carefully through the attached file mentioned above to see if it had any additional comments, not represented above?
+
 ********************************

### 2020-01-20 MainText.Rmd
getting close to being done addressing Ecosphere referee comments

@@
 
-<!--
-Alternative titles
-1) Tail associations in environmental variables and extinction risk
-2) The influence of correlations in the extremes on extinction risk
-3) Tail associations in environmental variables and their impact on extinction risk
-4) Asymmetric tail association in extreme climatic events affects population-extinction risk
--->
-
 \newpage
@@
 due to climate change. Furthermore, there is reason to believe extreme 
-climatic events may modify "tail associations" between values of an environmental 
-variable measured in different locations, i.e., associations that 
+climatic events may modify "tail associations" between distinct population vital rates,
+or between values of an environmental 
+variable measured in different locations. "Tail associations" between two variables are associations that 
 occur between values in the left or right tails of the distributions 
-of the variable that occur in the locations. 
+of the variables. 
 Two positively associated variables can be principally "left-tail associated" (i.e., more 
@@
 (more correlated when they take high than low values), even with the same
-overall correlation coefficient in both cases. We tested whether the nature of 
+overall correlation coefficient in both cases. 
+We tested, in the context of non-spatial stage-structured matrix models, 
+whether tail associations between stage-specific vital rates may
+influence extinction risk.
+We also tested whether the nature of 
 spatial tail associations of environmental variables can influence metapopulation 
@@
 spatially synchronized, presumably reducing the potential for rescue effects.
-We also tested, in the context of non-spatial stage-structured matrix models, 
-whether tail associations between stage-specific vital rates may
-influence extinction risk. 
-We showed that density dependence interacts with the
-propensity for tail associations to influence metapopulation extinction risk: for 
-population models
-showing undercompensatory density dependence, left-tail associations in environmental 
-variables typically accentuated
-and right-tail associations mitigated extinction risk; whereas the 
-reverse was usually true for models showing overcompensatory density dependence.
 For the non-spatial, stage-structured models we considered, left-tail associations 
@@
 associations, but the effect was small. 
+In contrast, we showed that density dependence interacts with tail associations 
+to influence metapopulation extinction risk substantially: for 
+population models
+showing undercompensatory density dependence, left-tail associations in environmental 
+variables often strongly accentuated
+and right-tail associations mitigated extinction risk; whereas the 
+reverse was usually true for models showing overcompensatory density dependence.
 Tail associations and their asymmetries are taken into account in assessing risks
@@
 
+Extinction risk is very commonly estimated for age-structured models 
+such as stochastic matrix models [@MorrisDoak2002; @Caswell2000] and 
+integral projection models
+[@Merowetal2014]. For such models, extinction risk depends, in part, on the long-term
+population growth rate, a population parameter which in turn 
+depends on the means, variances and covariances of population vital rates such 
+as the survival and fecundity rates of age classes. Extinction risks may,
+*a priori*, also depend on possible asymmetries of tail association between
+population vital rates. For instance, it seems reasonable to imagine
+a case in which both adult fecundity and juvenile survival rates in a population 
+depend on an environmental variable such as spring rainfall in similar ways, but 
+only when that variable takes low values. When rainfall takes
+typical or high values, fecundity and survival may be limited instead 
+by different, unrelated factors. This could produce left-tail dependence between these 
+vital rates, possibly resulting in different extinction risk for the model
+compared to a case in which vital rates were symmetrically associated in 
+their tails but otherwise statistically similar.
+
 <!--ECEs probably produce tail associations-->
-
-<!--***DAN wrote this-->
-<!--There are reasons to hypothesize that ECEs may produce asymmetric tail 
+There are reasons to hypothesize that ECEs may produce asymmetric tail 
 associations between climatic variables
-measured in different locations. For instance, suppose a single ECE is associated
-with spatially widespread, large values 
-of some climatic variable. Also suppose that small values of the
-variable are associated, instead, with local weather, which can differ from location
-to location, rather than with ECEs. The result could
-be stronger right-tail association, through time, between measurements of the climatic variable
-made in different locations.
-For example, rainstorms which deposit exceptionally large amounts of rain
-may do so over large areas. But periods with less rain may experience
-scattered showers only, which may be much less correlated across locales.-->
-
-<!--***Shya: I rephrased the following para-->
-There are reasons to hypothesize that different locations when influenced by single ECEs may 
-show asymmetric tail associations through time. This asymmetry arises because a single 
-ECE is associated with spatially widespread, large values 
-of some climatic variable whereas small values of the same
-variable are associated, instead, with local weather, which can differ from location
-to location, rather than with ECEs. For example, rainstorms which deposit 
-exceptionally large amounts of rain may do so over large areas. But periods with 
-less rain may experience scattered showers only, which may be much 
-less correlated across locales. The result could
-be stronger right-tail association, through time, between measurements of rainfall
-occurred in different locations.
+measured in different locations. The idea, here, is based on the reasonable assumption 
+that ECEs are both more intense and spatially more extensive than non-extreme 
+climatic events. For clarity, we explain the idea using the example case
+of extreme rain events. Suppose a single extreme rain event is associated
+with spatially widespread large precipitation values, i.e., not only does it rain a lot
+in each locale, but it does so for many locales across a large area. Also suppose that smaller local
+precipitation measurements are associated, instead, with local, non-extreme weather 
+which can differ from location to location, rather than with a single extreme rain event.
+The result could be stronger right-tail association, through time, between measurements of precipitation
+made in different locations: large precipation measurements are coherent across space, since 
+they are all attributable to a single event; whereas small values are less spatially correlated
+because they come from a multiplicity of distinct local events.
 
@@
 models incorporating density dependence.
-
-Extinction risk is also very commonly estimated for age-structured models 
-such as stochastic matrix models [@MorrisDoak2002; @Caswell2000] and 
-integral projection models
-[@Merowetal2014]. For such models, extinction risk depends, in part, on the long-term
-population growth rate, a population parameter which in turn 
-depends on the means, variances and covariances of population vital rates such 
-as the survival and fecundity rates of age classes. Extinction risks may,
-*a priori*, also depend on possible asymmetries of tail association between
-population vital rates. For instance, it seems reasonable to imagine
-a case in which both adult fecundity and juvenile survival rates 
-depend on an environmental variable such as spring rainfall in similar ways, but 
-only when that variable takes low values. When rainfall takes
-typical or high values, fecundity and survival may be limited instead 
-by different factors. This could produce left-tail dependence between these 
-vital rates, possibly resulting in different extinction risk for the model
-compared to a case in which vital rates were symmetrically associated in 
-their tails but otherwise statistically similar.
 
@@
 of Fig. \ref{fig_extrisk_agestr} are attributable solely to the changes we made in the
-asymmetry of tail associations, and answer Q1. Though differences were significant (confidence
-bands in Fig. \ref{fig_extrisk_agestr}, which are very narrow, did not overlap),
-they were also small, corresponding to only a few percentage points of additional extinction
-risk after 25 years. A potential reason for this outcome is elaborated in the Discussion.
+asymmetry of tail associations, and answer Q1. Differences were small, corresponding 
+only to a few percentage points of additional extinction risk after 25 years. A 
+potential reason for this outcome is elaborated in the Discussion.
 
@@
 and the $r_c$ values of Table \ref{tab_model_summary} reveals that, for the Verhulst and Austin-Brewer models,
-the switch occurs at a value of $r$ less than $r_c$ for these models. For the example
+the switch occurs at a value of $r$ less than $r_c$. For the example
 case of $d=0$, plots of extinction risk versus $r$ (Fig. \ref{fig_extrisk_d0}) seem to indicate that,
@@
 reveals a similar reversal for the Maynard Smith, Verhulst and Austin-Brewer models; but 
-it occurs at a value of $r$ close to but distinguishable from $r_c$.
+it occurs at a value of $r$ close to but distinguishable from $r_c$. Inspection of Figs
+\ref{SM-fig_ricker_risk_local_disp_scl_1}-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
+shows that a qualitatively similar conclusion holds for nonzero $d$, for both local and global
+dispersal.
 
@@
 <!--summary results:-->
-Our results combined with earlier results 
+Our results 
+suggest that asymmetry of tail associations between vital rates
+in a stage-structured, non-spatial stochastic model may be of limited 
+importance for extinction risk. In our stage-structured, non-spatial model, extinction 
+risk was only weakly influenced by asymmetry of tail associations
+between vital rates. Our results provide suggestive, rather
+than definitive evidence, here, because we considered 
+only one model (but see below). 
+In contrast, our metapopulation results 
 show that asymmetries in tail associations between environmental variables in
@@
 beneficial events caused populations to exceed carrying capacity, and 
-to subsequently crash, across all or most habitat patches. In contrast 
-to the metapopulation case, our results 
-suggest that asymmetry of tail associations between vital rates
-in a stage-structured, non-spatial stochastic model may be of limited 
-importance for extinction risk. In our stage-structured, non-spatial model, extinction 
-risk was only weakly influenced by asymmetry of tail associations
-between vital rates. Our results provide suggestive, rather
-than definitive evidence, here, because we considered 
-only one model (but see below).
-
-<!--linking back to ECEs-->
-Extensive evidence now exists that ECEs have become more frequent and more intense
-over the last several decades, and substantial progress has recently been made so that
-some event categories and even specific events can now be attributed to
-human-induced climate change [@ummenhofer2017extreme]. We here provide a fuller 
-explanation of how such changes may modify the asymmetry of tail association of 
-weather variables measured in different places, and thereby influence metapopulation
-extinction risk through the mechanisms we explored in this paper. 
-Although changes in 
-the correlation between $X_i(t)$ and $X_j(t)$ are increasingly documented, 
-and may also be a consequence of climate change (e.g., @Black2018),
-we are unaware of any studies examining the extent to which tail associations
-between $X_i(t)$ and $X_j(t)$, and asymmetries of tail association, may be 
-changing. However, documented changes in correlation between $X_i(t)$ and $X_j(t)$
-may be due partly or wholly to changes in tail associations. If ECEs are becoming
-more spatially extensive, as seems likely given that they are becoming
-more intense [@ummenhofer2017extreme], then the conditional probability that both of
-$X_i(t)$ and $X_j(t)$ are influenced by an ECE, given that one of them is so
-influenced, should be increasing. This should contribute to increased association
-of the two variables in one of their tails (which tail depends on the type of 
-ECE and what the variable $X$ is). A study of changes through time in 
-tail associations of weather variables would be straightforward and 
-potentially important given the mechanisms we have explored.
-<!--***DAN: Shyamolina, another aspect of an exploration you may do on the 
-the geography of tail association may be to look at how tail association 
-may be changing. This relates to ECEs, so may help better justify the work
-to a funder. You had proposed to look at geography of tail association in 
-NDVI, and that makes good sense. You could also look at how tail association
-is changing in NDVI (the records are about 35 years long, so you could just
-about do this). Even better (because data are better), you could get weather
-station records (lots of these can be downloaded, over 1000 stations
-have been operating for over a century continuously in North America) and 
-look at how tail associations are changing in those data.-->
-
-<!--copulas and rank methods-->
-It is known that the distributions $X_i(t)$ are changing, both in
-mean and variance, and possibly in other ways, depending on what
-environmental variable $X$ represents.
-Studying associations between $X_i(t)$ and $X_j(t)$ using standard 
-methods such as Pearson correlation
-may be complicated by these changes; rank-based and related 
-measures of association such as 
-we have used may be more appropriate for some applications.
-If $f$ is a strictly monotonic, increasing transformation, then the Spearman
-(or Kendall) correlation of $f(X_i(t))$ and $f(X_j(t))$ is the same as the
-Spearman (or Kendall) correlation of $X_i(t)$ and $X_j(t)$. However, the Pearson
-correlation of $f(X_i(t))$ and $f(X_j(t))$ may differ substantially
-from the Pearson correlation of $X_i(t)$ and $X_j(t)$. 
-Relatedly, if the marginal distributions of $X_i(t)$ and $X_j(t)$ are
-changing through time, it may also cause changes through time in the 
-Pearson correlation of these quantities, even while the values
-of rank-based correlations such as Spearman and Kendall may be unchanging.
-Essentially, in such a case, associations between $X_i(t)$ and $X_j(t)$
-would be unchanging through time even though Pearson correlation
-appears to indicate a change in association, which is instead a change 
-in marginal distributions.
-Perhaps unfortunately, the Pearson correlation,
-which is the most commonly used measure of association of two variables,
-depends not only on association information, but also on the marginals.
-For this reason, it is possible that recently observed changes in environmental
-(e.g., @Black2018) and biological (e.g., @Koenig2016) synchrony 
-are, at least in part, just another feature of 
-already-studied changes in distributional characteristics of 
-variables, rather than reflecting true changes in association as would be measured
-using rank-based methods (these studies used Pearson correlations).
-Rank-based methods are linked to a well-developed suite of statistical
-approaches related to the *copula* concept of statistics. Copulas and related ideas
-can be used to formally separate the information in the joint 
-distribution of $X_i$ and $X_j$ into the information that pertains
-to the marginal distributions of these two variables, and the information
-that pertains solely to their association. 
-@Ghosh_copula argued generally for the potential of copulas and related ideas
-to improve our understanding of ecology 
-(see also @Popovic2018; @Valpine2014; @Anderson2018).
-Rank-based methods such as the Spearman and partial Spearman correlations
-which we have used are closely related to copula approaches, and @Ghosh_copula,
-@Genest2007 and others have recommended copula and rank approaches when the 
-goal is to understand associations between variables, including how such 
-associations may be changing. We recognize that the Pearson correlation
-of $X_i(t)$ and $X_j(t)$, and changes through time in that correlation, 
-will still be an appropriate and important quantity to study 
-for many applications. For instance, in cases for which $X_i$ has a linear 
-or approximately linear influence on some aspect of the ecosystem
-at location $i$, Pearson correlations of the ecological impacts at locations 
-$i$ and $j$, which may be of fundamental interest, will relate to 
-Pearson correlations between $X_i$ and $X_j$. But in many cases, the
-ecological impact of $X_i$ and $X_j$ will be a nonlinear or unknown monotonic function of 
-the values of $X_i$ and $X_j$. So studying rank-based correlations
-may be more informative, ecologically as well as providing the mathematical advantages
-listed above.
+to subsequently crash, across all or most habitat patches.
 
 <!--future ideas start here-->
-Our results suggest some new lines of research on the 
-impact of tail associations on growth rates and 
-extinction risks. First, we considered a model with positive association between 
+Our initial exploration for stage-structured models suggested limited influence of 
+tail associations between vital rates on extinction risk, but some
+additional lines of inquiry might be worth exploring as part of an effort to
+understand how general our results are.
+First, we considered a model with positive association between 
 egg and adult vital rates,
@@
 related vital rates produce different outcomes for extinction risk
-from what we have observed by studying positively associated vital rates.
+from what we observed by studying positively associated vital rates.
 
@@
 A second topic of research that our work suggests has to do with Tuljapurkar's
-approximation of the long-term stochastic growth rate.
-Our results suggest that tail associations and asymmetries of tail associations
+well-known approximation of the long-term stochastic growth rate [@tulja1982; @tulja1990].
+It may be possible to obtain a better understanding, using Tuljapurkar's
+approximation, of our result that asymmetries of tail association
 may have limited influence on the extinction risk of 
-stage-structured, non-spatial models. Although we believe more research is 
-certainly still needed
-as to whether this conclusion holds more generally than the model we considered, 
-if it does hold 
-generally, it may be possible to obtain a better understanding of 
-the result using the well-known approximation
-of Tuljapurkar [@tulja1982; @tulja1990] for the long-term stochastic growth rate.
+stage-structured, non-spatial models.
 The long-term stochastic growth rate is an important quantity in demography
@@
 to which accuracy of the approximation has been tested for scenarios with
-asymmetric tail associations. It seems worth explicitly exploring the accuracy of 
+asymmetric tail associations. It seems potentially worth exploring the accuracy of 
 Tuljapurkar's approximation under scenarios of tail association in the vital rates
@@
 stage-structured model.
+
+<!--linking back to ECEs - this para about the metapop case-->
+Extensive evidence now exists that ECEs have become more frequent and more intense
+over the last several decades, and substantial progress has recently been made so that
+some event categories and even specific events can now be attributed to
+human-induced climate change [@ummenhofer2017extreme]. We now provide an  
+explanation of how such changes may modify the asymmetry of tail association of 
+weather variables measured in different places. Such modification may thereby influence metapopulation
+extinction risk through the mechanisms explored earlier in this paper. 
+Let $X_i(t)$ and $X_j(t)$ denote the values of some weather variable as measured in fixed 
+locations $i$ and $j$ at time $t$. Although changes in 
+the correlation, through time, between $X_i(t)$ and $X_j(t)$ are increasingly documented, 
+and may also be a consequence of climate change (e.g., @Black2018),
+we are unaware of any studies examining the extent to which tail associations
+between $X_i(t)$ and $X_j(t)$, and asymmetries of tail association, may be 
+changing. If ECEs are becoming
+more spatially extensive, as seems likely given that they are becoming
+more intense [@ummenhofer2017extreme], then the conditional probability that both of
+$X_i(t)$ and $X_j(t)$ are influenced by an ECE, given that one of them is so
+influenced, should be increasing. That is, since ECEs are spatially bigger, it is now
+more likely that a single weather event
+influences both locations $i$ and $j$ given that it influences one of them. 
+This should contribute to increased association
+of the two variables in one of their tails (which tail depends on the type of 
+ECE and what the variable $X$ is). A study of changes through time in 
+tail associations of weather variables would be straightforward and 
+potentially important given the mechanisms we have explored. 
+<!--***DAN: Shyamolina, another aspect of an exploration you may do on the 
+the geography of tail association may be to look at how tail association 
+may be changing. This relates to ECEs, so may help better justify the work
+to a funder. You had proposed to look at geography of tail association in 
+NDVI, and that makes good sense. You could also look at how tail association
+is changing in NDVI (the records are about 35 years long, so you could just
+about do this). Even better (because data are better), you could get weather
+station records (lots of these can be downloaded, over 1000 stations
+have been operating for over a century continuously in North America) and 
+look at how tail associations are changing in those data.-->
+
+<!--copulas and rank methods-->
+It is known that the distributions $X_i(t)$ are changing, in
+mean, variance, and possibly other ways, depending on what
+environmental variable $X$ represents.
+Studying associations between $X_i(t)$ and $X_j(t)$ using standard 
+methods such as Pearson correlation
+may be complicated by these changes; rank-based 
+measures of association such as 
+we have used may be more appropriate for some applications.
+If $f$ is a strictly monotonic, increasing transformation, then the Spearman
+(or Kendall) correlation of $f(X_i(t))$ and $f(X_j(t))$ is the same as the
+Spearman (or Kendall) correlation of $X_i(t)$ and $X_j(t)$. However, the Pearson
+correlation of $f(X_i(t))$ and $f(X_j(t))$ may differ substantially
+from the Pearson correlation of $X_i(t)$ and $X_j(t)$. 
+Relatedly, if the marginal distributions of $X_i(t)$ and $X_j(t)$ are
+changing through time, it may also cause changes through time in the 
+Pearson correlation of these quantities, even while the values
+of rank-based correlations such as Spearman and Kendall may be unchanging.
+Essentially, in such a case, associations between $X_i(t)$ and $X_j(t)$
+would be unchanging through time even though Pearson correlation
+appears to indicate a change, which is instead a change 
+in marginal distributions. Perhaps unfortunately, the Pearson correlation,
+which is the most commonly used measure of association of two variables,
+depends not only on association information, but also on the marginals.
+For this reason, it is possible that recently observed changes in environmental
+(e.g., @Black2018) and biological (e.g., @Koenig2016) synchrony 
+are, at least in part, just another feature of 
+already-studied changes in distributional characteristics of 
+variables, rather than reflecting true changes in association as would be measured
+using rank-based methods.
+
+Rank-based methods are linked to a well-developed suite of statistical
+approaches related to the *copula* concept of statistics. Copulas and related ideas
+can be used to formally separate the information in the joint 
+distribution of $X_i$ and $X_j$ into the information that pertains
+to the marginal distributions of these variables, and the information
+that pertains solely to their association. 
+@Ghosh_copula argued generally for the potential of copulas 
+to improve our understanding of ecology 
+(see also @Popovic2018; @Valpine2014; @Anderson2018).
+Rank-based methods such as the Spearman and partial Spearman correlations
+we used are closely related to copulas, and @Ghosh_copula,
+@Genest2007 and others have recommended copula and rank approaches when the 
+goal is to understand associations between variables. We recognize that the Pearson correlation
+of $X_i(t)$ and $X_j(t)$, and changes through time in that correlation, 
+will still be useful for many applications. But to understand potential 
+changes in the association 
+of $X_i(t)$ and $X_j(t)$, unconfounded by changes in their marginal distributions,
+it is necessary to use rank-based or copula methods [@Ghosh_copula,
+@Genest2007].
 
@@
 by @walteretal2017, have explored geographic variation in the spatial synchrony of environmental
-and population variables (known as "geography of synchrony"); @walteretal2017 explored the 
-mechanisms, consequences and ecological importance of this geography of synchrony. For instance, 
-synchrony was found to have a prominent geography (i.e. a distinctive spatial pattern) which can 
+and population variables. @walteretal2017 coined such variation the 
+"geography of synchrony", and explored its 
+mechanisms, consequences and ecological importance. For instance, 
+synchrony in several systems was found to have prominent geographic variation, which can 
 help provide inferences about the causes
@@
 (A, B) and extreme (C, D) asymmetry of tail association.
-The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk.
+The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk based on the number of simulations done.
 LT=left tail, RT=right tail.\label{fig_extrisk_agestr}}

### 2020-01-21 Response.txt
even closer, just one thing to add plus the new figure and related text

@@
 
->>>We split the paragraph into two and shortened, as the referee requesrs.
+>>>We split the paragraph into two and shortened, as the referee requests.
 
@@
 
-***Shya: Agreed and modified in the text.
+>>>We revised, as requested.
 
@@
 
-***Shya: I leave it for DAN's judgement.
+>>>In the course of performing other edits, we revised this paragraph so the comment no longer applies.
 

### 2020-01-21 MainText.Rmd
even closer, just one thing to add plus the new figure and related text

@@
 among vital rates $r_1$, $r_2$, and $r_3$. Values of $r_1$ near
-an upper limit may be strongly associated with,
-for instance, $r_2$, whereas lower values of $r_1$ may not be strongly
-associated with $r_2$ because reductions/increases in $r_1$ can
-then be traded off against increases/reductions in either 
-$r_2$ or $r_3$. It is possible that tail associations in negatively 
+an upper limit may be strongly associated with $r_2$, whereas lower 
+values of $r_1$ may not be strongly
+associated with $r_2$ because $r_1$ trades off against both 
+$r_2$ and $r_3$. It is possible that tail associations in negatively 
 related vital rates produce different outcomes for extinction risk
@@
 asymmetric tail associations. It seems potentially worth exploring the accuracy of 
-Tuljapurkar's approximation under scenarios of tail association in the vital rates
-and relating results to the effects of tail associations on the long-term
+Tuljapurkar's approximation under scenarios of tail association in the vital rates;
+and then relating results to the effects of tail associations on the long-term
 stochastic growth rate and extinction risks.

### 2020-01-21 MainText.Rmd
minor stuff during first half of a read through

@@
 extinction risk for the model, and to what extent? Q2) Does the nature
-of tail associations between measurements of an environmental variable in different
+of tail associations between measurements of an environmental variable made in different
 locations influence the extinction risk of a density-dependent metapopulation model
@@
 for both Q1 and Q2 instead of attempting mathematical 
-analysis and/or using models that are more realistic, 
+analysis and/or using models that are more realistic and complex, 
 because our intention here is to provide an initial exploration only of 

### 2020-01-21 MainText.Rmd
minr changes on read through

@@
 knowledge asymmetric tail associations have rarely been assessed for their
-importance for extinction risks of populations and species (but see @Ghosh_copula),
-and our essential goal is to advance the idea that tail associations may
+importance for extinction risks of populations and species (but see @Ghosh_copula).
+Our essential goal is to advance the idea that tail associations may
 also be important in this context. 
@@
 that $0 \leq l_b < u_b \leq 1$, and letting $x(t)$ be the rank of 
-$u(t)$ in the set $\{u(t) | t=1,\ldots,T\}$ and letting 
+$u(t)$ in the set $\{u(t) | t=1,\ldots,T\}$, divided by $T+1$, and letting 
 $y(t)$ be the rank of 
-$v(t)$ in the set $\{v(t) | i=1,\ldots,T\}$ (smallest elements
+$v(t)$ in the set $\{v(t) | i=1,\ldots,T\}$, divided by $T+1$ (smallest elements
 are here defined to have rank 1),
@@
 correlation of $u$ and $v$ which can be attributed to the points
-for which the ranks, $x$ and $y$, lie between the boundary lines
+for which the normalized ranks, $x$ and $y$, lie between the boundary lines
 we defined in
@@
 that depend on the band to be used). This figure was reproduced
-largely following Fig. 7 of Ghosh \emph{et al.} (2019).
+largely following Fig. 7 of Ghosh \emph{et al.} (2020).
 \label{PartialSpearmanFig}}
@@
 are analyzed in detail in Appendix \ref{SM-model_stability}, where 
-references for the models are also provided.
+original references for the models are also provided.
 All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
@@
 Estimates of extinction risk were plotted together with 95\% confidence intervals
-based on a binomial distribution.
+based on a binomial distribution. Though these confidence intervals have limited
+utility because they can be made arbitrarily narrow by increasing the number of simulations,
+they are nonetheless presented to illustrate the confidence achieved with the number of 
+simulations we performed.
 
@@
 used within the range, we calculated extinction risk as indicated above.
-Dispersal parameter values $d= 0,0.1,0.2,..0.5$ were also used in different 
-simulations. We plotted extinction risk as a function of $d$ for two
-fixed values of $r$, one producing under- and one overcompensatory dynamics.
+Dispersal parameter values $d= 0,0.1,0.2,\ldots,0.5$ were also used in different 
+simulations. We plotted extinction risk as a function of $d$ for several
+fixed values of $r$, some producing under- and some overcompensatory dynamics.
 For $d=0$ we also plotted extinction risk as a function of $r$. All code for 
@@
 Though extinction risks were overall higher for the density-dependent version of our model compared
-to density-independent version, extinction risk differences across tail association 
+to the density-independent version, extinction risk differences across tail association 
 scenarios were similar for the two versions. Because the same Spearman correlation (up to sampling
@@
 <!--future ideas start here-->
-Our initial exploration for stage-structured models suggested limited influence of 
+Our exploration for stage-structured models suggested limited influence of 
 tail associations between vital rates on extinction risk, but some
@@
 variables, rather than reflecting true changes in association as would be measured
-using rank-based methods.
+using rank-based correlation methods.
 
@@
 of $X_i(t)$ and $X_j(t)$, unconfounded by changes in their marginal distributions,
-it is necessary to use rank-based or copula methods [@Ghosh_copula,
+it is necessary to use rank-based or copula methods [@Ghosh_copula;
 @Genest2007].

### 2020-01-22 MainText.Rmd
new figs added, captions written, and cited from text. just need typo-free versions from Shyamolina

@@
 vital rates, possibly resulting in different extinction risk for the model
-compared to a case in which vital rates were symmetrically associated in 
-their tails but otherwise statistically similar.
+compared to a case in which vital rates are symmetrically associated in 
+their tails or right-tail associated but otherwise statistically similar (Fig. \ref{VitalRatePedagogFig}).
 
@@
 if overall (non-tail-specific) levels of environmental synchrony are the same
-in both these scenarios. The present study extends tests of this idea to metapopulation
-models incorporating density dependence.
+in both these scenarios (Fig. \ref{MetapopPedagogFig}). The present study extends tests 
+of this idea to metapopulation models incorporating density dependence.
 
@@
 \end{figure*}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=25cm]{vitalrate_pedagog_fig.pdf}
+\caption{Left-tail association between vital rates makes it more likely that low values
+of fecundity and survival will happen simultaneously (A), causing both life stages to crash 
+simultaneously. This seems likely to increase extinction risk. In contrast, right-tail association
+between vital rates instead makes it more likely that high values of fecundity and survival will 
+happen simultaneously (B). \label{VitalRatePedagogFig}}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=25cm]{metapop_pedagog_fig.pdf}
+\caption{Under extreme left-tail association of environments between two locations, catastrophes happen 
+simultaneously in both locations (A). Whereas under extreme right-tail association, catastrophes do not happen
+simultaneously, even when overall environmental association is the same in both scenarios (B). \label{MetapopPedagogFig}}
+\end{center}
+\end{figure}
 

### 2020-01-24 MainText.Rmd
cleaned upto Methods

@@
 The result could be stronger right-tail association, through time, between measurements of precipitation
-made in different locations: large precipation measurements are coherent across space, since 
+made in different locations: large precipitation measurements are coherent across space, since 
 they are all attributable to a single event; whereas small values are less spatially correlated
@@
 $y(t)$ be the rank of 
-$v(t)$ in the set $\{v(t) | i=1,\ldots,T\}$, divided by $T+1$ (smallest elements
+$v(t)$ in the set $\{v(t) | t=1,\ldots,T\}$, divided by $T+1$ (smallest elements
 are here defined to have rank 1),
@@
 
-We simulated our metapopulation models using environmental noises $\epsilon_i(t)$ (Table \ref{tab_model_summary}), 
+We simulated our metapopulation models using environmental noises $\varepsilon_i(t)$ (Table \ref{tab_model_summary}), 
 independent through time, that were generated to have the same overall associations 
@@
 simulations described above. For any $i$, $u_i(t)$ was again uniformly distributed on 
-the interval $(0,1)$. To obtain the values $\epsilon_i(t)$, the $u_i(t)$ were 
+the interval $(0,1)$. To obtain the values $\varepsilon_i(t)$, the $u_i(t)$ were 
 transformed so that the resulting numbers were
@@
 with mean $0$ and standard deviation $\sigma$. Because this was a monotonic transformation,
-the Spearman correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
+the Spearman correlation between $\varepsilon_i(t)$ and $\varepsilon_j(t)$ was the same 
 as that between $u_i(t)$ and $u_j(t)$; likewise partial Spearman correlations were the same. 
@@
 Here $P_i(t)$ is the population in patch $i$ at time $t$.
-The growth rate depends on model parameters, $P_i(t)$, and the environment $\\epsilon_i(t)$ in patch $i$ at time $t$.
-$\\epsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times.
+The growth rate depends on model parameters, $P_i(t)$, and the environment $\\varepsilon_i(t)$ in patch $i$ at time $t$.
+$\\varepsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times.
 $P_e$ is the equilibrium of the deterministic, one-patch version of the model.

### 2020-01-25 SuppMat.Rmd
worked on comments and edits
table cleaned
typo

@@
 ---
-title: "Supporting information: Tail associations and their impact on extinction risk"
+title: "Supporting information: Tail associations in ecological variables and their impact on extinction risk"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
@@
 
-dt2<-tibble("$\\lambda_i(t)$" = c("$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
-                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",
-                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+(aP_i(t))^b]$",
-                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_i(t)/b)\\}]$",
-                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(1-P_i(t)/K)]$",
-                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(K-P_i(t))\\{1-\\exp(-sP_i(t))\\}]$"
+dt2<-tibble("$\\lambda_i(t)$" = c("$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
+                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",
+                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} r/[1+(aP_i(t))^b]$",
+                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_i(t)/b)\\}]$",
+                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} [1+r(1-P_i(t)/K)]$",
+                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} [1+r(K-P_i(t))\\{1-\\exp(-sP_i(t))\\}]$"
 ))
@@
 developed to help understand the dynamics of insect populations, and as a model of contest versus scramble 
-instraspecific competition. The model
+intraspecific competition. The model
 can be written $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$. Here $r>0$ is the intrinsic growth rate and

### 2020-01-25 Response.txt
done with edits, subject to Shya review, except for new fig
minor edits to responses to referees

@@
 
-***Shya: I think Fig. 1 is the most simplified pedagogical fig. we've ever had in our all copula-related papers. 
-So, as of now I have no idea what extra info does the 2nd reviewer ask for? Copula related paper from finance or other fields are already cited in the paper in L156-157.
-***DAN: I do not think we can dodge this - we have to do SOME or even MOST of what the refere says, or he/she will simply say, if the paper is sent back to him/her, "they did not do what I asked for, so REJECT".
-Of the different suggestions, I think we should make some effort at this one and at 4 below.
-I will write a justification below for not doing 3.
-Please think more about such a figure, ASAP.
-Please think of an *additional* figure, to come right after the curent Fig.
-1 The current Fig.
-1 explains what tail dependence is.
-But I think what the referee wants is a figure explaining, conceptually, how tail dependence could cause accentuated extinction risk.
-I think he/she wants a conceptual figure showing the main idea.
-I think the figure should be associated with, and cited from, the Intro paras beginning "There are also reasons to hypothesize..." and "Extinction risk is also very commonly estimated..." I agree with the referee that we can produce a good conceptual figure illustrating this main idea.
-I know you can do this.
-Please think about it Monday and come to me Tuesday with a hand-drawn sketch of a figure that we can discuss before you then create the figure on Tuesday.
-Please do not come back to me saying you do not know how to do this - you can do this, I know you can.
-And time is getting a little tight, and I have other things I have to do on Monday, so this has to be done by you!
-I am sure you can do a good job.
-***DAN: Part of the reason we need to move efficiently on this is they only gave us three weeks, and we may also have to convert to Word, which you know can be labor intensive.
-Can you please look on the Ecosphere website and find out if they need final submissions to be in Word or if they can take latex?
-Even if they can take latex, we may want to convery to Word.
-If there is any evidence that they prefer Word, we will want to convert to Word to avoid them introducing a million errors trying to do latex when they do not really know how.
+>>>Thank you for this suggestion, which spurred a lot of thought about how to best communicate the main ideas of the manuscript in the intial parts of the manuscript.
+The referee's suggestion led to two new figures.
 
@@
 So we believe carving our matrix model results out of the current paper and then undertaking a more systematic explortion ourselves, for possible later publication, does not get these potentially practically important ideas out there as fast and risks the possibility that we will not be able to do the analyses as well as some of our readers.
+So to us it is important to get these results out there now.
   Second, for stochastic matrix models, our initial explorations suggested that probably the effects of tail association in that context are minor, so we went no further.
-In contrast, for metapopulation models, our initial explorations (with the Ricker model) showed a likely strong effect, so we went further with additional models.
+In contrast, for metapopulation models, our initial explorations (initially our explorations were with the Ricker model only) showed a likely strong effect, so we went further with additional models.
 So what the referee perceives as imbalance is actually just a rational attempt to explore promising ideas further and to simply identify other ideas as less promising, to help future researchers choose what to work on.
@@
 This is a standard way of defining tail associations and copula approaches to the study of dependencies, justified by the Ghosh et al paper (which, by the way, has now been formally accepted at Advances in Ecological Research, and is in press) and by Genest & Favre, 2007, J Hydol Eng, "Everything you always wanted to know about copula modelling but were afraid to ask".
-We have added an explanation of this fact in the Discussion, and made some related points here that will help readers in future applications of our ideas.
-
-***DAN: Don't forget to add the explanation.
+We have clarified and amended some text, in the Discussion, which helps explain these facts.
 
@@
 The version on the BioRxiv is essentially the same as the final, accepted version.
-In spite of thinking for a quite a long time about what the referee may have meant by the above comment, we confess we are not sure, so we do not know what to do about this comment.
-In our model formulations, variables are never "good" or "bad" but are instead assumed to fluctuate, taking values in some years which are "good" for the population (by which we mean they promote increased reproduction or increased survival in that location, in that year), and values in other years which are "bad" (in the opposite sense).
-Aside from an apparent desire to have us make some additional point (I'm afriad we cannot understand what point), the referee does not appear to be criticizing anything we actually do say.
-So we suggest proceeding with this text as it is.
-As we said, the Ghosh et al paper is now accepted, and is already available online on the BioRxiv, and many important related points are made there.
+In spite of thinking for a quite a long time about what the referee may have meant by the above comment, we confess we are not sure, so we do not know quite what to do about this comment.
+In our model formulations, variables themselves are never "good" or "bad" but are instead assumed to fluctuate, taking values in some years which are "good" for the population (by which we mean they promote increased reproduction or increased survival in that location, in that year), and values in other years which are "bad" (in the opposite sense).
+Aside from an apparent desire to have us make some additional point (I'm afraid we cannot understand what point), the referee does not appear to be criticizing anything we actually do say.
+So we suggest proceeding with this part of the manuscript as it is.
+As we said, the Ghosh et al paper is now accepted, and is already available online on the BioRxiv, and many important  points which seem likely to be related to the referee's point are made there.
 So readers looking to base some of their own research on the points made in the current manuscript can also use the Ghosh et al paper as an additional background resource before getting started.
@@
 
->>>In the course of doing other edits we went through the entire text again and tightened where possible.
-
-***DAN: make sure to do this
+>>>In the course of doing other edits we went through the entire text again and tightened where possible. Thank you for spurring us to do so.
 
@@
 
->>>Thank you, we take the referee's point here and have eliminated the statement about significance.
+>>>Thank you, we take the referee's point, here, and have eliminated the statement about significance.
 

### 2020-01-27 Response.txt
adding my comments
delete last line

@@
 We find the paper is improved and we hope you now find it ready for formal acceptance to Ecosphere.
-Our responses to comments are preceeded by ">>>" to make them easy to find.
+Our responses to comments are preceded by ">>>" to make them easy to find.
 Thank you again for an enjoyable and helpful review process.
@@
 
->>>Thank you for this suggestion, which spurred a lot of thought about how to best communicate the main ideas of the manuscript in the intial parts of the manuscript.
+>>>Thank you for this suggestion, which spurred a lot of thought about how to best communicate the main ideas of the manuscript in the initial parts of the manuscript.
 The referee's suggestion led to two new figures.
@@
 Mathematical analysis of stochastic matrix models is a deep field.
-So we believe carving our matrix model results out of the current paper and then undertaking a more systematic explortion ourselves, for possible later publication, does not get these potentially practically important ideas out there as fast and risks the possibility that we will not be able to do the analyses as well as some of our readers.
+So we believe carving our matrix model results out of the current paper and then undertaking a more systematic exploration ourselves, for possible later publication, does not get these potentially practically important ideas out there as fast and risks the possibility that we will not be able to do the analyses as well as some of our readers.
 So to us it is important to get these results out there now.
@@
   Finally, we do not see as much of an imbalance as the referee suggests.
-To us, the two components are treated with slightly different throroughness, but not markedly so, and for good reasons (see above).
+To us, the two components are treated with slightly different thoroughness, but not markedly so, and for good reasons (see above).
 Our matrix modelling results are certainly enough to make the points we wanted to make, i.e., an initial assessment that the influence of tail associations on extinction risk in this context seems likely to be minor.
@@
 >>>Done, and the clarity is improved. Thanks.
-
-Please see also the attached file.
-
-***DAN: Shyamolina, did you look carefully through the attached file mentioned above to see if it had any additional comments, not represented above?
-
-********************************

### 2020-01-31 MainText.Rmd
cleaned pto fig legend
table cleaned
Popovic Ref updated

@@
 life history trade-offs, e.g. between survival and reproduction 
-(e.g, @Lande2003, p. 59). Asymmetric tail associations may occur
+(e.g., @Lande2003, p. 59). Asymmetric tail associations may occur
 in such a context if, for instance, there is a three way trade-off
@@
 to improve our understanding of ecology 
-(see also @Popovic2018; @Valpine2014; @Anderson2018).
+(see also @Popovic2019; @Valpine2014; @Anderson2018).
 Rank-based methods such as the Spearman and partial Spearman correlations
@@
 $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text).
-$r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation; see main text).
+$r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilibrium (a bifurcation; see main text).
 For extinction-risk simulations, we used $r$ such that $r_{min}<r<r_{bf}$.

### 2020-03-23 MainText.Rmd
made changes in Rmds as DAN said

@@
 and $\theta=1$, respectively, 
-and $\Omega$ is a density dependence parameter. Taking $\Omega=\infty$ made
+and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$
 $\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
@@
 \begin{equation}
-\cor_{l_b,u_b}(u,v) = \frac{\sum (x(t)-\mean(x))(y(t)-\mean(y))}{(n-1)\sqrt{\var(x) \var(y)}},\label{PartialSpearmanEq}
+\cor_{l_b,u_b}(u,v) = \frac{\sum (x(t)-\mean(x))(y(t)-\mean(y))}{(T-1)\sqrt{\var(x) \var(y)}},\label{PartialSpearmanEq}
 \end{equation}
@@
 results for under- and overcompensatory model regimes.
-Defining $r_{min}$ as the minimum value of $r$ 
+Defining $r_\text{min}$ as the minimum value of $r$ 
 for which $P_e$ was non-negative; and defining
-$r_{bf}$ as the value of $r$ at which a bifurcation occurs, the range of
-$r$ used spanned 95\% of the interval ($r_{min},r_{bf}$). For details, see Appendix 
+$r_\text{bf}$ as the value of $r$ at which a bifurcation occurs, the range of
+$r$ used spanned 95\% of the interval ($r_\text{min},r_\text{bf}$). For details, see Appendix 
 \ref{SM-model_stability}. The range of $r$ encompassed $r_c$, so included 
@@
 <!--Q2 : metapopulation results-->
-For our metapopulation simulations, for low values of $r$ (close to $r_{min}$), left-tail
+For our metapopulation simulations, for low values of $r$ (close to $r_\text{min}$), left-tail
 associations of environmental noises accentuated extinction risks, and right-tail associations
 mitigated them; but the reverse was true for high values of $r$ (close to
-$r_{bf}$). Thus 
+$r_\text{bf}$). Thus 
 density dependent models can either produce similar results, regarding the influence of 
@@
 For each model of Table \ref{tab_model_summary}, using a value of $r$ on
-the low end of the range $(r_{min},r_{bf})$, extinction risk was highest
+the low end of the range $(r_\text{min},r_\text{bf})$, extinction risk was highest
 for extreme left-tail associated noise, lowest for extreme right-tail associated
@@
 essentially all the dispersal rates, $d$, that we considered. However, 
-using a value of $r$ on the high end of the range $(r_{min},r_{bf})$,
+using a value of $r$ on the high end of the range $(r_\text{min},r_\text{bf})$,
 extinction risks were ordered oppositely (right panels of Figs 
@@
 Because the deterministic, one-patch versions of our models showed undercompensatory
-dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_{min},r_{bf})$),
+dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_\text{min},r_\text{bf})$),
 and showed overcompensatory dynamics for $r>r_c$, Figs \ref{fig_extrisk_allmodel_localdisp}
@@
 the right tails of the rates, but only slightly so. Results are for the linear
-version of the model ($\Omega = \infty$; A, C) and for the nonlinear model with density
+version of the model ($\Omega \rightarrow \infty$; A, C) and for the nonlinear model with density
 dependence in the fecundity rate ($\Omega = 100$; B, D). Results show cases of moderate
@@
 $P_e$ is the equilibrium of the deterministic, one-patch version of the model.
-$r_{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative.
+$r_\\text{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative.
 $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text).
-$r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilibrium (a bifurcation; see main text).
-For extinction-risk simulations, we used $r$ such that $r_{min}<r<r_{bf}$.
+$r_\\text{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilibrium (a bifurcation; see main text).
+For extinction-risk simulations, we used $r$ such that $r_\\text{min}<r<r_\\text{bf}$.
 Dispersal between patches was also implemented (see main text).
-Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_{min}$, $r_c$ and $r_{bf}$.
+Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_\\text{min}$, $r_c$ and $r_\\text{bf}$.
 See Appendix \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",

### 2020-03-23 SuppMat.Rmd
starting to change appendices in trivial ways to accommodate journal style
made changes in Rmds as DAN said

@@
 lines(t,x$erA,col="blue",type="l")
-legend(x=-2,y=1.05, c("moderate LT","moderate RT", expression(paste(Omega,"= ", infinity))), lty=c(1,1, NA),  
+legend(x=-2,y=1.05, c("moderate LT","moderate RT", 
+                      #expression(paste(Omega,"%->%", infinity))
+                      expression(symbol('\127') *symbol('\256')* symbol('\245'))
+                      ), 
+       lty=c(1,1, NA),  
        col=c('red', 'blue', NA), horiz=F, bty='n', cex=1.5, x.intersp = 0.2
@@
 lines(t,x$erA,col="blue",type="l")
-legend(x=-2,y=1.05, c("extreme LT","extreme RT", expression(paste(Omega,"= ", infinity))), lty=c(1,1, NA),  
+legend(x=-2,y=1.05, c("extreme LT","extreme RT", expression(symbol('\127') *symbol('\256')* symbol('\245'))), 
+       lty=c(1,1, NA),  
        col=c('red', 'blue', NA), horiz=F, bty='n', cex=1.5, x.intersp = 0.2
@@
 
-dt4_2<-tibble("$r_{min}$" = c("0",
+dt4_2<-tibble("$r_\\text{min}$" = c("0",
                         "1",
@@
 
-dt6<-tibble("$r_{bf}$" = c(r_bf_ricker,
+dt6<-tibble("$r_\\text{bf}$" = c(r_bf_ricker,
                            r_bf_hassell,
@@
 tail association, $d$ and $r$: metapopulations, local dispersal, Ricker]{Analogous to Fig.
-\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Ricker model.\label{fig_ricker_risk_local_disp_scl_1}}
+\ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Ricker model.
+Local dispersal.\label{fig_ricker_risk_local_disp_scl_1}}
 \end{center}
@@
 \ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Hassell model.
-\label{fig_hassell_risk_local_disp_scl_1}}
+Local dispersal.\label{fig_hassell_risk_local_disp_scl_1}}
 \end{center}
@@
 \ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Maynard Smith model.
-\label{fig_msmith_risk_local_disp_scl_0.8}}
+Local dispersal.\label{fig_msmith_risk_local_disp_scl_0.8}}
 \end{center}
@@
 \ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Pennycuick model.
-\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
+Local dispersal.\label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
 \end{center}
@@
 \ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Verhulst model.
-\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
+Local dispersal.\label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
 \end{center}
@@
 \ref{MT-fig_extrisk_allmodel_localdisp} in the main text, but considering more values of $r$ and only the Austin-Brewer model.
-\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
+Local dispersal.\label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
 \end{center}
@@
 The model shows undercompesatory dynamics if $0<r<1$ and overcompensatory dynamics
-if $1<r<2$. Thus $r_{min}=0$, $r_c=1$, and $r_{bf}=2$.
+if $1<r<2$. Thus $r_\text{min}=0$, $r_c=1$, and $r_\text{bf}=2$.
 
@@
 For $P_e = P^\star$, the model shows undercompensatory dymamics if $1<r<[b/(b-1)]^b$ 
-and overcompensatory dynamics if $[b/(b-1)]^b<r<[b/(b-2)]^b$, so $r_{min}=1$, 
-$r_c = [b/(b-1)]^b$, and $r_{bf}=[b/(b-2)]^b$.
+and overcompensatory dynamics if $[b/(b-1)]^b<r<[b/(b-2)]^b$, so $r_\text{min}=1$, 
+$r_c = [b/(b-1)]^b$, and $r_\text{bf}=[b/(b-2)]^b$.
 
@@
 and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
-For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$, so $r_{min}=1$. 
+For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$, so $r_\text{min}=1$. 
 Then $P^\star$ will 
@@
 the model will show overcompensatory dymamics $r_c<r<b/(b-2)$.
-$P^\star$ will become unstable for $r>r_{bf}=b/(b-2)$.
+$P^\star$ will become unstable for $r>r_\text{bf}=b/(b-2)$.
 
@@
 a positive equilibrium population 
-$r$ must be greater than $r_{min} = 1+\exp(-a)$. The transition point $r_c$ (from under- to 
-overcompensatory dynamics) and the bifurcation point $r_{bf}$ were obtained
+$r$ must be greater than $r_\text{min} = 1+\exp(-a)$. The transition point $r_c$ (from under- to 
+overcompensatory dynamics) and the bifurcation point $r_\text{bf}$ were obtained
 by numerical solution of ${F}'=0$ and ${F}'=-1$ at $P_e=P^{\star}$, respectively.
@@
 $P_e=K$ is stable below the bifurcation 
-point $r_{bf}=2$. The model shows undercompensatory dynamics for $0<r<r_c=1$
-and overcompensatory dynamics for $r_c<r<r_{bf}$.
+point $r_\text{bf}=2$. The model shows undercompensatory dynamics for $0<r<r_c=1$
+and overcompensatory dynamics for $r_c<r<r_\text{bf}$.
 
@@
 positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. $P^\star$
-is stable when $r<r_{bf}$ and unstable beyond the bifurcation point, where 
-$r_{bf} = \frac{2}{K[1-\exp (-sK)]}$. The model shows undercompensatory dynamics when
+is stable when $r<r_\text{bf}$ and unstable beyond the bifurcation point, where 
+$r_\text{bf} = \frac{2}{K[1-\exp (-sK)]}$. The model shows undercompensatory dynamics when
 $0<r<r_c$, where $r_c = \frac{1}{K[1-\exp (-sK)]}$. It shows 
-overcompensatory dynamics when $r_c<r<r_{bf}$.
+overcompensatory dynamics when $r_c<r<r_\text{bf}$.
 

### 2020-03-23 MainText.Rmd
appendix fixed

@@
 $u(t)$ and $v(t)$. The quantities $(u(t),v(t))$ for $t=1,\ldots,T$ were
-generated as described in Appendix \ref{SM-makenoise}.
+generated as described in Appendix, section \ref{SM-makenoise}.
 
@@
 The models are also listed in Table \ref{tab_model_summary}, and 
-are analyzed in detail in Appendix \ref{SM-model_stability}, where 
+are analyzed in detail in Appendix, section \ref{SM-model_stability}, where 
 original references for the models are also provided.
@@
 dynamics showing an overcompensatory (i.e., oscillatory) approach to an equilibrium. 
-The transition value of $r$ was denoted $r_c$. Appendix \ref{SM-model_stability}
+The transition value of $r$ was denoted $r_c$. Appendix, section \ref{SM-model_stability}
 summarizes stability analyses of the models.
@@
 The quantities $(u_1(t),\ldots,u_n(t))$
-for $t=1,\ldots,T$ were generated as described in Appendix \ref{SM-makenoise}.
+for $t=1,\ldots,T$ were generated as described in Appendix, section \ref{SM-makenoise}.
 
@@
 $r_\text{bf}$ as the value of $r$ at which a bifurcation occurs, the range of
-$r$ used spanned 95\% of the interval ($r_\text{min},r_\text{bf}$). For details, see Appendix 
+$r$ used spanned 95\% of the interval ($r_\text{min},r_\text{bf}$). For details, see Appendix, section 
 \ref{SM-model_stability}. The range of $r$ encompassed $r_c$, so included 
@@
 Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_\\text{min}$, $r_c$ and $r_\\text{bf}$.
-See Appendix \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",
+See Appendix, section \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",
              booktabs=T, linesep = "\\addlinespace"

### 2020-03-25 MainText.Rmd
minor edit

@@
 and $\theta=1$, respectively, 
-and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$
+and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$,
 $\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence

### 2020-03-27 MainText.Rmd
minor edit

--- 2020-03-23 MainText.Rmd
@@
 and $\theta=1$, respectively, 
-and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$
-$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
+and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$,
+$\exp(-A(t)/\Omega) \rightarrow 1$, so that, in that limit, the model did not include density dependence
 (it was then a standard stochastic matrix model). Finite, positive values of 

### 2020-03-27 MainText.Rmd
minor edits

@@
 title: "Tail associations in ecological variables and their impact on extinction risk"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
-date: ""
+author:
+ - Shyamolina Ghosh\textsuperscript{1}, Lawrence W. Sheppard\textsuperscript{1}, Daniel C. Reuman\textsuperscript{1,2,\dag}
+ - \textsuperscript{1}Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA; \textsuperscript{2}Laboratory of Populations, Rockefeller University, 1230 York Ave., New York, NY 10065, USA
+date: '\textsuperscript{\dag}Correspondence: Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084'
 fontsize: 12pt
@@
 
-\noindent \emph{Affiliations:}
+<!--\noindent \emph{Affiliations:}
 
@@
 
-\noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
+\noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
 
 \noindent \emph{Correspondence:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084. 
+-->
 
@@
 \noindent The authors were partly supported by US National Science Foundation
-grant 1714195 and the James S. McDonnell Foundation.
+grant 1714195 and the James S. McDonnell Foundation, and the California Department of Fish 
+and Wildlife Delta Science Program.
 

### 2020-04-27 SuppMat.Rmd
typo fixed

@@
 ---
-title: "Supporting information: Tail associations in ecological variables and their impact on extinction risk"
+title: "Appendix: Tail associations in ecological variables and their impact on extinction risk"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
@@
 states, $P_e = 0$ and $P_e = K$.
-The equilibrium $P_e = 0$ is unstable, since ${F}' = \exp(r)$ for that equilibtrium, and 
+The equilibrium $P_e = 0$ is unstable, since ${F}' = \exp(r)$ for that equilibrium, and 
 $\exp(r) > 1$ for a positive $r$. Depending on $r$, 
